### 2016-04-26 paper/mscov.tex
Added paper

\documentclass{article}

\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{textcomp}
\usepackage{longtable}
\usepackage{multirow, booktabs}
\usepackage{amsfonts,amsmath,amssymb}
\usepackage{natbib}
\usepackage{url}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{chemarrow}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{rotating}
\usepackage{multirow}
\usepackage[margin=.75in]{geometry}

\renewcommand{\baselinestretch}{1.5}
\newcommand{\bl}[1]{{\mathbf #1}}
\newcommand{\bs}[1]{{\boldsymbol #1}}
\newcommand{\tr}{\text{tr}}
\newcommand{\etr}{\text{etr}}
\newcommand{\Exp}[1]{{\text{E}}[ \ensuremath{ #1 } ]  }

\begin{document}

\title{Shared Subspace Model}

\date{\today}

\bibliographystyle{plain}
\title{Shared Subspace Covariance Model}
\author{Alexander Franks and Peter Hoff}
\date{\today}
\maketitle

\begin{abstract}
  We develop model-based methods for evaluating similarities and
  differences among several $p\times p$ covariance matrices in the
  $p>n$ setting.  This is done by assuming a spiked covariance model
  for each group, and sharing information about the subspace spanned
  by the group-level eigenvectors.  We propose an EM algorithm for
  estimating the shared subspace and covariance matrix of the data
  projected onto this subspace.  In addition, we provide an Gibbs
  sampling algorithm for estimating the posterior uncertainty of
  group-level eigenvectors and eigenvalues on this subspace.

\end{abstract}

% \section{NOTES}

% \begin{itemize}
% \item For MSCov to work in no-pooeld case need $R < S$ (doens't work if $S=P$).
% \item Explore the effect on estimation when R=2 two eigenvalues converge/diverge.
% \item TODO: Experiments: 
% \begin{itemize}
% \item S=20, R=2, data generated under no-pooled model.
% \item Fit with R=2 to R=20, Fix S=20
% \item Fit with R=2, S=5 to 20
% \item Fix $1/(1-Y^2)$ bug for Y=1
% \item debeta_un = 0 for large sample sizes (underflows)
% \item Check that CP models yield same results using fitBayesianSpike and fit-subspace.
% \end{itemize}
% \end{itemize}

\section{Introduction}

%\citet{donoho_gavish_johnstone_2013} 
%\citet{paul_2007}

\section{The Spiked Covariance Model}  
Suppose $\bl S$ has a possibly degenerate Wishart$(\Sigma,n$) 
distribution with density 
given by 
\begin{equation} 
p(\bl S | \Sigma, n) \propto_{\Sigma} l(\Sigma:\bl S) =  |\Sigma|^{-n/2} \text{etr}( - \Sigma^{-1} \bl S/2 ) ,  
\label{eqn:lik}
\end{equation}
where $\Sigma \in \mathcal S_p^+$ and $n$ may be less than $p$.

% However, when the population size is large and comparable with the
% sample size, as is in many contemporary data, it is known that the
% sample covariance matrix is no longer a good approximation to the
% covari- ance matrix. Baik

Such a likelihood results from $\bl S$ being, for example, a residual 
sum of squares matrix from a multivariate regression analysis. In  this case,
$n$ is the number of independent observations minus the 
rank of the design matrix. 
The spiked principle 
components model (spiked PCA) introduced by \citet{Johnstone2001} assumes that 
%\begin{equation} 
$\Sigma = \sigma^2 (  \bl U  \Lambda  \bl U^T  + \bl I )$  ,  
%\end{equation}
where $\Lambda$ is an $r\times r$ diagonal matrix and 
$\bl U \in \mathcal V_{r,p}$, with 
%$r< ( n \wedge p)$. 
$r<p$. Such a formulation is appealing because it explicitly
partitions the covariance into a tractable low rank ``signal'' and
homscedastic ``noise''.

For such a covariance matrix, we have
$\Sigma^{-1} = (\bl U \Lambda \bl U^T + \bl I )^{-1} = \bl I - \bl U
\Omega \bl U^T $
where $\Omega = \text{diag}(\omega_1,\ldots, \omega_r)$ with
$\omega_s = \lambda_j/(\lambda_s+1)$.  The likelihood (\ref{eqn:lik})
in terms of $(\sigma^2,\bl U, \Omega)$ is then
\begin{equation}
L(\sigma^2,\bl U , \Omega : \bl Y)  = 
(\sigma^2)^{-np/2} \etr(-\bl S/[2\sigma^2]) \times 
   \etr( \Omega \bl U^T [\bl S/(2\sigma^2)] \bl U ) \times  
 \left\{ \prod_{s=1}^r (1-\omega_s)^{n/2} \right \} .  
\label{eqn:rplik}
\end{equation}

Classical results for parametric models 
(e.g., \citet{schwartz_1965})  
imply that 
asymptotically in $n$ for fixed $p$, a Bayes estimator
will be consistent for a spiked  population covariance 
as long as the assumed number of spikes is greater than or equal to 
the true number. 
However, the situation is more difficult 
when both $n$ and $p$ are large.  
%Since a Bayes estimator for an invariant loss function 
%is orthogonally equivariant under this prior, 
%it can be expressed as 
%$\hat \Sigma(\bl S) = \bl E \bl D(\bl L) \bl E^T$, 
%where $\bl E$ and $\bl L$ are the matrices of eigenvectors and 
%eigenvalues of $\bl S$, and $\bl D$ denotes a diagonal matrix 
%that is a function of $\bl L$. 
%As a result 
Under the spiked covariance model, 
results of 
\citet{Baik2006, Paul2007} indicate that if 
$p/n \rightarrow \gamma >0$ as $n\rightarrow \infty$,   
the top $r$ eigenvalues of $\bl S/(n\sigma^2)$ will converge 
to upwardly  biased versions of $(\lambda_1+1,\ldots, \lambda_r+1)$, 
if each $\lambda_s $ is greater than $\sqrt{\gamma}$.   
This has led several authors to suggest estimating 
$\Sigma$ via shrinkage of the eigenvalues of the sample covariance 
matrix. In particular, 
in the setting that $\sigma^2$ is known, 
\citet{Donoho2013} propose  
setting all sample eigenvalues that fall below 
the value $(1+\sqrt{\gamma})^2$ to be one, and shrinking the larger 
eigenvalues in a way that depends on the particular 
loss function being used. 
These shrinkage functions are shown to be asymptotically 
optimal in the $p/n\rightarrow \gamma$ setting.

As demonstrated by Gavish et al \citep{Gavish2014}, even when the
true rank is large, if blah it is still asymptotically preferable to fit a lower rank
blah .  it is always better than hard thresholding at any other value, and is always better than ideal truncated singular value decomposition (TSVD)

Bayes estimators under the prior distribution described above 
can also be viewed as eigenvalue shrinkage estimators: 
Since a Bayes estimator for an invariant loss function 
is orthogonally equivariant under this prior, 
it can be expressed as 
$\hat \Sigma(\bl S) = \bl E  \hat \Sigma (\bl L) \bl E^T$, 
where $\bl E$ and $\bl L$ are the matrices of eigenvectors and 
eigenvalues of $\bl S$. Furthermore, it is straightforward to 
show that $\hat \Sigma (\bl L)$ must be a diagonal matrix, 
and so the Bayes estimator is obtained by applying the 
shrinkage function $\hat\Sigma$ to the sample eigenvalues $\bl L$. 
Understanding of this shrinkage function could in principle 
be pursued by investigation of the posterior distribution of the 
eigenvalue parameters $\Lambda$, or equivalently their  
transformed values $\Omega$.

However, in many modern applications, the primary objective is to
compare covariances across multiple groups of observations.  For
instance, Cite
Daniela Witten.  Hoff eigenpooling, no longer equivariance.  In suhc a
situation, equivariance blah blah.  In the following section, we
propose a model for estimating and exploring differences between
covariance matrices from multiple groups of related data.

\section{The Shared Subspace Spiked Covariance Model}

Assume we have $k=1, ..., K$ groups of matrices $\bl S_k$ which follow
the spiked covariance model with density as given in
\ref{eqn:rplik}.    Each group consists of of $n_k$ measurements of $p$ features, typically with $n_k << p$.  However, we assume that the eigenvectors
are related across groups in a particular way.  \citet{Hoff2009}
introduced a model for eigenvector shrinkage based on the matrix Bingham
distribution.

In contrast, in this paper, we assume a priori that the eigenvectors
are uniformally distributed on a low rank common subspace.  That is,

\begin{equation}
\Sigma_k = V\psi_kV^T + \sigma^2_kI
\end{equation}

With $V \in \mathcal{V}_{s,p}$ with $p
\geq s \geq r$.  Here, $VV^T \in \mathcal{G}_{s, p}$ is the plane of
variation shared by all groups and $\psi_k$ is the rank-r covariance matrix
on that subspace.  \textbf{Rationale}.

In this paper, as in blah we consider a model on eigendecomposition of
$\psi_k$ so that Equation \ref{} can be written as

\begin{equation}
\Sigma_k = VO_k\Lambda_KO_k^TV^T + \sigma^2_kI
\end{equation}

Where, $\Lambda_k$ is and $r \times r$ diagonal matrix of eigenvalues,
and $O_k \in \mathcal{V}_{s,r}$ is the matrix of eigenvectors of
$\psi_k$.  Note that for any single group we can recover the original
spiked covariance formuation (Equation
\ref{}) by letting $U_k = VO_k$.

The likelihood for $\Sigma_k$ given the sufficient statistic $\mathbf{S}_k = Y_kY_k^T$ is simply

\begin{equation}
p(S | \Sigma_k,n_k) \propto \prod_{k=1}^K |\Sigma_k|^{-n/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)
\end{equation}

First note that by the Woodbury matrix identity we can show that 
\begin{align}
\Sigma^{-1}_k &=  (\sigma_k^2(U_k\Lambda_kU_k^T+I))^{-1}\\
&= \frac{1}{\sigma_k^2}(U_k\Lambda_kU_k^T+I)^{-1}\\
&= \frac{1}{\sigma_k^2}(I-U_k\Omega_kU_k^T)\\
\end{align}

Where the diagonal matrix $\Omega = \Lambda(I+\Lambda)^{-1}$, e.g. $\omega_i = \frac{\lambda_i}{\lambda_{i}+1}$.

 Further,

\begin{align}
|\Sigma_k| &= (\sigma_k^2)^{p}|U_k\Lambda_kU_k^T+I|\\
&= (\sigma_k^2)^{p}|\Lambda_k+I| \\
&= (\sigma_k^2)^{p}\prod_{i=1}^r(\lambda_i+1)\\
&= (\sigma_k^2)^{p}\prod_{i=1}^r(1-\omega_i)
\end{align}

where the second line is due to Sylvester's determinant theorem.

Substituting for $\Sigma^{-1}_k$ and $|\Sigma_k|$ we have the likelihood of $V$, $O_k$, $\Lambda_k$ and $\sigma_k^2$

We can write the likelihood of $V$, $O_k$, $\Lambda_k$ and
$\sigma_k^2$ given $Y_k$ using Equation \ref{eqn:rplik} by replacing
$U_k$ with $VO_k$: 
\begin{equation}
 L(\sigma_k^2,\bl V , O_k \Omega_k : \bl Y_k) \propto \prod_{k=1}^K\left[ (\sigma_k^2)^{-n_kp/2}etr(-\frac{1}{2\sigma_k^2}\mathbf{S}_k)\left(\prod_{i=1}^r(1-\omega_{ki}) \right) ^{n_k/2}  etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) \right]
\end{equation}

\subsection*{An Expectation-Maximization Algorithm For Estimating the Shared Subspace}

We start by outlining a computationally efficient algorithm for
estimating the shared subspace, $VV^T$.

The full likelihood can be written as

\begin{align}
p(\bl S | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)  \\
&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-(\sigma_k^2(V\psi_kV^T +
  I))^{-1}\mathbf{S}_k/2) \\
&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-\left[V(\psi_k +
  I)^{-1}/\sigma_k^2 V^T + (I-VV^T)/\sigma^2_k\right]\mathbf{S}_k/2)
  \\
&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-r)/2}|M_k|^{-n_k/2}etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]\mathbf{S}_k/2) 
\end{align}

Where $M_k^{-1} = (\psi_k + I) ^{-1}/\sigma_k^2$.  The log-likelihood in
$V$ (up to additive constant) is

\begin{align}
l(V) &= \sum_k tr\left(-VM_k^{-1}V^T +
       VV^T/\sigma^2_k\right)\mathbf{S}_k/2)\\
&= \sum_k \frac{1}{2}tr\left(-M_k^{-1}V^T \mathbf{S}_kV\right) + \frac{1}{2\sigma_k^2}tr\left(VV^T \mathbf{S}_k\right)
\end{align}

% \begin{align}
% p(V|...) &\propto etr(\sum_{k=1}^K[O_k\Omega_kO^T_k]V^TS_k/(2\sigma^2_k)]V)\\
% \end{align}

To find the maximum marginal likelihood of $V$, we use an EM approach,
treating the $M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as the ``missing''
parameters to be imputed. %%We assume a Wishart prior on $M_k^{-1}$ and an Gamma prior
on $\frac{1}{\sigma_k^2}$.  If we take these prior's to be
``non-informative'' degenerate priors, we have

We put a Jeffreys prior on both $M_k^{-1}$ and $\frac{1}{\sigma_k^2}$.

$$p(M_k^{-1} | V) \propto |M_k^{-1}|^{n_k/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) $$

\noindent which is a Wishart($V^T\mathbf{S}_kV$, $n_k + r + 1$).

\noindent For $\frac{1}{\sigma_k^2}$ we have

$$p\left(\frac{1}{\sigma^2} | V\right) \propto \left(\frac{1}{\sigma_k^2}\right)^{n_k(p-r)/2}etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)\mathbf{S}_k\right)  $$

\noindent which is a Gamma($n_k(p-r)/2+1$, $\frac{1}{2}tr((I-VV^T)\mathbf{S}_k)$)

\noindent Thus, for each iteration, $t$, of the EM algorithm:

\begin{enumerate}
\item For each $k$, compute relevant conditional expectations:
\begin{enumerate}
\item $E[M_k^{-1} | V_{t-1}] = (n_k + r  + 1)(V^T \mathbf{S}_kV)^{-1}$
\item $E[\frac{1}{2\sigma_k^2}|V_{t-1}] = \frac{n_k(p-r) + 2}{tr((I-VV^T)S_k)}$
\end{enumerate}
\item Compute $V_{t}$ = $\underset{V}{argmax}  \sum_k tr\left(-VE[M^{-1}|V_{t-1}]V^T +
       E[\frac{1}{2\sigma_k^2}|V_{t-1}]VV^T\right)\mathbf{S}_k/2)$ (optimization
  approach from Wen et al, 2013)
\end{enumerate}

\subsection{Alternative formulation}
$M_k = \sigma_k^2(\Phi_k + I)$ where 
$$ \Phi_k =\left( \begin{array}{cc}
\Psi_k & 0  \\
0 & D  \end{array} \right)$$

Letting $V = [V_1\, V_2]$

\begin{eqnarray*}
p(M_k^{-1} | V) &\propto \prod_k |M_k^{-1}|^{n_k/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) \\
                & = (|\psi_k + I|^{-1}|D +
                  I|^{-1})^{n_k/2}etr(-1/2((\psi_k+I)^{-1}V_1^T\mathbf{S}_kV_1
                  -1/2(D+I)^{-1}V_2^T\mathbf{S}_kV_2)) \\
\end{eqnarray*}

$$p(\frac{1/d_i} | V) \propto \prod_k (d_i + 1)^{-n_k/2}exp\left(\frac{1}{-1/2(d_i+1)}(V_2^T\mathbf{S}_kV_2)_{i,i}\right) $$

\section{Bayesian Inference}

\subsection{Prior and posterior distributions}

\paragraph{Conditional estimation of $(\sigma^2, \bl U)$:}
Bayesian inference is facilitated with the use of semiconjugate 
prior distributions  for $(\sigma^2, \bl U)$.
These include the inverse-gamma class of priors for $\sigma^2$, 
and the von Mises-Fisher-Bingham  
distributions for $\bl U$ \citep{hoff_2009a,hoff_2009b}.  
In the absence of real prior information,  invariance 
considerations suggest the use of priors on $\sigma^2$ and 
$\bl U$ that lead to equivariant estimators. 
In particular, an (improper) prior having density 
$p(\sigma^2,\bl U) = 1/\sigma^2 $, with respect to the 
product of Lebesgue measure on $\mathbb R^+$ and the 
uniform probability measure on $\mathcal V_{r,p}$, 
leads to Bayes estimators $\hat \Sigma(\bl S)$ that 
are equivariant with respect to 
rotations and scale changes, so that 
$\hat \Sigma(a \bl W \bl S \bl W^T) = 
   a \bl W \hat\Sigma(\bl S) \bl W$ for all 
$a>0$ and $\bl W\in \mathcal O_{p}$  
(assuming an invariant loss function).  
Under this invariant prior, straightforward calculations
show that the full conditional distribution of
$\sigma^2$ is
inverse-gamma$( n p/2 , \tr(\bl S(\bl I-\bl U\Omega\bl U)/2)$.
%Under the uniform prior on $\mathcal V_{r,p}$,
The full conditional distribution of $\bl U$ has a density proportional
to the likelihood, so
\[
 p(\bl U | \bl S, \sigma^2, \Omega) \propto 
  \etr( \Omega \bl U^T [\bl S/(2\sigma^2) \bl U). 
\]
This is a Bingham$(\Omega, \bl S/(2\sigma^2)$ distribution on
$\mathcal V_{r,p}$ \citep{khatri_mardia_1977}. A Gibbs sampler to
simulate from this distribution is given in
\citet{hoff_2009a}.
Interestingly, if $\Omega$ were known (which it is not), then 
for a given invariant loss function
the Bayes estimator under this prior 
minimizes the (frequentist) risk among all 
equivariant estimators \citep{eaton_1989}.

\paragraph{Estimation of $\Omega$:} 
We first consider a uniform(0,1) prior distribution  for
each element of $\Omega$, or equivalently, an 
$F_{2,2}$ prior distribution for the elements of $\Lambda$. 
As a result, the full conditional distribution of 
an element  $\omega_s$ of $\Omega$ is proportional to the 
likelihood function, so 
%an inverse-gamma$(\nu_0/2 , \nu_0 \sigma_0^2/2)$  prior distribution for  
%$\sigma^2$  gives an inverse-gamma$(\nu_n/2 , \nu_n \sigma_n^2/2)$
%full conditional distribution, 
%where  $\nu_n = \nu_0+np$ and 
%  $\nu_n \sigma_n^2  = \nu_0\sigma_0^2 + 
%   \tr(\bl S(\bl I-\bl U\Omega\bl U))$. 
\begin{equation} 
p(\omega_s | \bl U, \bl S, \sigma^2)   \propto_{\omega_s} 
 l(\sigma^2,\bl U, \Omega:\bl S)  \propto_{\omega_s} 
  (1-\omega_s)^{n/2} e^{c \omega_s  n/2},    
\label{eqn:wpost}
\end{equation}
where $c= \bl u_s^T \bl S \bl u_s/(n \sigma^2)$ and 
$\bl u_s$ is column $s$ of $\bl U$. 
While not proportional to a density belonging to a standard class 
of distributions, 
the corresponding 
univariate distribution is easy to sample from numerically, and 
its 
behavior is straightforward to 
understand:
%First note that the likelihood 
%\ref{eqn:lik} is maximized in 
%if $\bl S \approx n \Sigma$, 
%and $\bl U$ and $\sigma^2$ are approximately close to thier 
%correct values, then $c\approx  $
If $c\leq 1$, then the the function has a maximum 
at $\omega_s =0$, and decays monotonically to zero as $\omega \rightarrow 1$. 
%If $c<n/2$ then the derivative is negative at zero, 
%whereas 
If $c>1$ then the function is uniquely  maximized at 
$(c-1)/c \in (0,1)$. 
To see why this makes sense, note that the likelihood is maximized when 
the columns of 
$\bl U$ are equal to the eigenvectors of $\bl S$ corresponding to its top 
$r$ eigenvalues \citep{tipping_bishop_1999}. At this value of $\bl U$, 
$c$ will then equal one of the top $r$ eigenvalues of $\bl S/(n\sigma^2)$. 
In the case  that $n\gg p$, we expect  $\bl S/(n\sigma^2)\approx \Sigma_0/\sigma^2$, 
the true (scaled) population covariance, and so we expect $c$ to be near one of the 
top $r$ eigenvalues of $\Sigma_0/\sigma^2$, say $\lambda_0+1$. 
If indeed $\Sigma_0$ has 
$r$ spikes, then  $\lambda_0>0$, $c \approx \lambda_0 +1 > 1$, and so the conditional 
mode of $w_s$ is approximately $(c-1)/c = \lambda_0/(\lambda_0+1)$, 
the correct value. 
On the other hand, if we have assumed the existence of a spike when 
there is none, then $\lambda_0=0$, $c\approx 1$ and the Bayes estimate 
of $w_s$ will be shrunk towards zero, as it should be.

\subsection*{Conditional distribution of $\sigma_k^2$}
TODO: Inverse-Gamma conjugacy

\subsection*{Conditional distribution of $\Lambda$}
Here, we specify the conditional distribution of $\Omega = \Lambda(I+\Lambda)^{-1}$.

\begin{align}
p(\Omega_k|V,O_k,\Lambda_k,\Sigma_k) \propto \left(\prod_{i=1}^r(1-\omega_{ki})^{n_k/2}  \right) etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) 
\end{align}

TODO: Finish

\subsection*{Conditional distribution of $O_k$}

\begin{align}
\label{lik_vo}
 p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto etr(\Omega_kO^T_kV^T[S_k/(2\sigma^2_k)]VO_k)
\end{align}

Assuming a uniform distribution on $O_k$, the conditional distribution of $O_k$ is clearly a matrix Bingham distribution with parameters $\Omega_k$ and $V^T[S_k/(2\sigma^2_k)]V$.  It may be desirable to specify a conjugate prior for $O_k$:

\begin{equation}
P(O_k|\Omega_k) = etr(\Omega_kO_kHO_k^T)
\end{equation}

in which case the conditional distribution is simply
\begin{equation}
P(O_k|...) \propto etr(\Omega_kO^T_k(V^T[S_k/(2\sigma^2_k)]V + H)O_k)
\end{equation}
which is still a matrix Bingham with parameters $\Omega_k$ and $V^T[S_k/(2\sigma^2_k)]V+H$
\subsection*{Conditional distribution of $V$}
\begin{align}
p(V|...) &\propto etr(\sum_{k=1}^K[S_k/(2\sigma^2_k)]V[O_k\Omega_kO^T_k]V^T)\\
\end{align}

Let, $A_k = S_k/(2\sigma^2_k)$ and $B_k = O_k\Omega_kO^T_k$.  We can write the conditional distribution of $V$ (again assuming a uniform prior) given $A_k$ and $B_k$ as
\begin{align}
P(V|A_k,B_k) &\propto etr(\sum_{k=1}^KA_kVB_kV^T)\\
&= etr(\sum_{k=1}^K\sum_{i,j}^sA_kv_iv_j^Tb_{ijk})\\
&= etr(\sum_{i,j}^s \sum_{k=1}^K v_j^T[b_{ijk}A_k]v_i)
\end{align}

From here, we specify the conditional distribution of a single column of $V$ given $A_k$, $B_k$ and all other columns of V:

\begin{align}
P(v_i|A_k,B_kV_{-i})  &\propto etr(2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])v_i+v_i^T[\sum_{k=1}^K b_{ijk}A_k]v_i))\\
&= etr(Cv_i+v_i^TMv_i)
\end{align}

This is a vector BMF(C,M) with $C = 2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])$ and $M= \sum_{k=1}^K b_{ijk}A_k$.

% \begin{align}
% P(V|O_k,\Omega_k,\sigma^2_k,\mathbf{S}_k) = (\sigma^2_k)^{-n_kp}etr(\sum_{k=1}^K[S_k/(2\sigma^2_k)]V[O_k\Omega_kO^T_k]V^T)\\
% \end{align}

\pagebreak

\section{Simulation Study}

First, we present the results of a simulation which demonstrates the
relative performance of shared subspace group covariance estimation
under different data generating models.  We consider three different
spiked covariance data models: 1) ``complete pooling'', in which the
covariance matrices from all groups are identical, e.g.
$\Sigma_1 = \Sigma_2 = ... \Sigma_K$; 2) ``shared subspace'', where
the signal lies on the same low rank subspace (Equation \ref{}); and 3) ``no
pooling'' assuming little overlap in the eigenvectors of the
covariance matrices (Equation \ref{}).  For each of these data models, we
simulated 30 datasets from a spiked covariance model with rank $r=2$
and eigenvalues $(\lambda_1, \lambda_2) = (250, 25)$.

We evalute the three corresponding inferential models,
corresponding to the complete pooling blah and blah.

For each group, $k$, we compute Stein's loss
$L_S( \Sigma_k , \hat\Sigma_k) = \text{tr}( \Sigma_k^{-1} \hat
\Sigma_k ) - \log |\Sigma_k^{-1} \Sigma_k | - p$.
Under Stein's loss, the Bayes estimator is the inverse of the
posterior mean of the precision matrix,
$\hat \Sigma_{BS} = \Exp{ \Sigma^{-1} | \bl S}^{-1}$ which can easily
be esimated using posterior samples.  In Table \ref{table:groupLoss}
we report the average group loss
$L(\Sigma_1, ..., \Sigma_K; \hat\Sigma_1, ..., \hat\Sigma_K ) =
\frac{1}{K} \sum_k L_S( \Sigma_k , \hat\Sigma_k)$.

\citep{Paul2007, Donoho2013}.

The results demonstrate that

\begin{table}
\begin{center}
  \begin{tabular}{ l  l | c | c | c |}
    \multicolumn{2}{c}{} & \multicolumn{3}{c}{\textbf{Inferential Model}} \\
  \multicolumn{2}{c|}{}  & Shared Subspace & Complete Pooling & No pooling \\  \cline{2-5}
    \multirow{3}{*}{\rotatebox[origin=c]{90}{\textbf{Data Model}}} & 
   Shared Subspace & 0.8 & 2.1 & 3.0 \\ \cline{2-5}
   & Complete Pooling & 0.8 & 0.7 & 3.0 \\ \cline{2-5}
   & No pooling & 8.3 & 136.3 & 3.0 \\ \cline{2-5}
  \end{tabular}
  \caption[Table caption text]{Stein's loss for different inferential
    and data generating models.  For each of the $K=10$ groups, we
    simulate a $p=200$ dimensional spiked covariance matrix for rank
    $r=2$.  For data from a shared subspace model the eigenvectors
    of the covariance matrix share a subspace dimension $s=2$.  All results are
    based on $n_k = 50$ observations per group.  Shared subspace
    inference blah between fitting each group covariance independently
  and assuming that they are all the same.  }
\label{table:groupLoss}
\end{center}
\end{table}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{Figs/LossVsDimension}
  \caption{Stein's loss for data generated under the ``no pooling''
    model and fit using the shared subspace model.  As the shared
    subspace dimension increases,  $s \rightarrow p$, the loss converges
    to the loss from the independent spiked covariance estimates.
    Thus, the shared subspace estimator is a generalization of blah.  }
\end{figure}

% n <- 50
% S <- 2
% R <- 2
% P <- 200
% ngroups <- 10

% evals <- c(250, 25)
% niters <- 1000
% nwarmup <- niters/2

\subsection{Evaluating the Subspace}

$$L_SS(\hat{V}^T\Sigma_K\hat{V} , \hat{V}^T\hat{\Sigma}_K\hat{V}) $$

We propose a simple metric  to evaluate how much of the variation in $Y_k$ lies on the
subspace spanned by $\hat{V}$:

$$ \frac{||Y_k^T\hat{V}||_F}{\underset{V \in \mathcal{V}_{p, s}}{sup}
  ||Y_k^TV||_F - \hat{\sigma}_k^2ps}$$

This quantity is  as an estimate of the ratio of the first $s$
sum of the eigenvalues of $V^T\bl \Sigma_k V$ to the sum of the first
$s$ eigenvalues of $\Sigma_k$.  To see this, note that $\underset{V \in \mathcal{V}_{p, s}}{sup}\frac{||Y_k^TV||_F}{n_k}$
is equivalent to the sum of the first $s$ eigenvalues of the sample
covariance matrix $\frac{\bl S_k}{n_k}$.  Letting
$\hat{\lambda}^{(k)}_i$ = the $i$-th eigenvalue of $\frac{\bl S_k}{n_k}$.  Asymptotically as $p, n_k \rightarrow \infty$ with
$\frac{p}{n} = \gamma$

$ \hat{\lambda}^{(k)}_i\rightarrow \lambda^{(k)}_i
(1 + \frac{\sigma_k^2\gamma}{\lambda^{(k)}_i - 1} \approx \lambda^{(k)}
+ \sigma^2_k\frac{p}{n_k}$ \citep{Baik2006}.

$\sum_i^s \hat{\lambda}^{(k)}_i \approx \sum_i^s \lambda^{(k)}_i + \sigma_k^2ps$

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{Figs/lossBoxplot}
  \caption{}
\end{figure}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.75\textwidth]{Figs/evalRatios}
  \caption{}
\label{fig:evalRatios}
\end{figure}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{Figs/posteriorRegions}
  \caption{}
\label{fig:posteriorRegions}
\end{figure}

\section{An Analysis of Gene Expression Data}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{Figs/leukemiaRatio}
  \caption{}
\label{fig:leukemiaRatio}
\end{figure}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{Figs/leukemiaPosterior}
  \caption{}
\label{fig:leukemiaRatio}
\end{figure}

\bibliographystyle{natbib}
\bibliography{refs.bib}

\end{document}

### 2016-05-10 paper/mscov.tex
Updated paper.

@@
 \usepackage{multirow}
+\usepackage{tikz}
 \usepackage[margin=.75in]{geometry}
@@
 
-\bibliographystyle{plain}
 \title{Shared Subspace Covariance Model}
@@
 \begin{abstract}
+
+% From Hoff Although the covariance matrices corresponding to different
+% populations are unlikely to be exactly equal they can still exhibit a
+% high degree of similarity. For example, some pairs of variables may be
+% positively correlated across most groups, whereas the correlation
+% between other pairs may be consistently negative. In such cases much
+% of the similarity across covari- ance matrices can be described by
+% similarities in their principal axes, which are the axes that are
+% defined by the eigenvectors of the covariance matrices. Estimating the
+% degree of across-popu- lation eigenvector heterogeneity can be helpful
+% for a variety of estimation tasks.
+
   We develop model-based methods for evaluating similarities and
   differences among several $p\times p$ covariance matrices in the
-  $p>n$ setting.  This is done by assuming a spiked covariance model
+  $p > n$ setting.  This is done by assuming a spiked covariance model
   for each group, and sharing information about the subspace spanned
   by the group-level eigenvectors.  We propose an EM algorithm for
-  estimating the shared subspace and covariance matrix of the data
-  projected onto this subspace.  In addition, we provide an Gibbs
-  sampling algorithm for estimating the posterior uncertainty of
-  group-level eigenvectors and eigenvalues on this subspace.
+  estimating the shared subspace and the covariance matrix of the data
+  projected onto this subspace.  We also provide a Gibbs sampling
+  algorithm for estimating the posterior uncertainty of group-level
+  eigenvectors and eigenvalues.  Finally, we demonstrate the utility
+  of our method in analyses exploring covariability in biological data.
 
@@
 
-%\citet{donoho_gavish_johnstone_2013} 
-%\citet{paul_2007}
+Introduction to covariance estimation and spiked covariance model.
 
-\section{The Spiked Covariance Model}  
 Suppose $\bl S$ has a possibly degenerate Wishart$(\Sigma,n$) 
@@
 components model (spiked PCA) introduced by \citet{Johnstone2001} assumes that 
-%\begin{equation} 
-$\Sigma = \sigma^2 (  \bl U  \Lambda  \bl U^T  + \bl I )$  ,  
-%\end{equation}
+\begin{equation} 
+\Sigma = \sigma^2 (  \bl U  \Lambda  \bl U^T  + \bl I )
+\label{eqn:spiked}
+\end{equation}
+
 where $\Lambda$ is an $r\times r$ diagonal matrix and 
@@
 %$r< ( n \wedge p)$. 
-$r<p$. Such a formulation is appealing because it explicitly
-partitions the covariance into a tractable low rank ``signal'' and
-homscedastic ``noise''.
+$r<p$. Thus, the spiked covariance model is a low rank factor model
+with homoscedastic error \citep{}.  Such a formulation is appealing because it explicitly
+partitions the covariance into a tractable low rank ``signal'' and homscedastic ``noise''.
 
@@
 
-However, in many modern applications, the primary objective is to
-compare covariances across multiple groups of observations.  For
-instance, Cite
-Daniela Witten.  Hoff eigenpooling, no longer equivariance.  In suhc a
-situation, equivariance blah blah.  In the following section, we
+However, in many modern applications, it is not enough to
+independently estimate an individual covariance matrix.  Rather, a
+comparative analysis of covariance patterns across multiple groups of
+observations would be most informative.  For example, in case
+control-studies, mean level effects may be small relative to subject
+variability.  Differences in the covariance
+matrices may lead to better insights about the differences between groups.
+Alternatively, when mean level differences are detectable, better
+estimates of the covariability of features across groups may lead to improved
+understanding of the mechanisms underlying these apparent mean level
+differences.
+
+With this situation in mind, Flury (1984) developed estimation and
+testing pro-cedures for the ‘common principal components’ model, in
+which a set of covariance matrices
+\citep{Flury1987} and \citep{Schott1991, Schott1999} considered
+cases in which only certain columns or subspaces of U are shared across populations, and \citep{Boik2002} described a very general model in which eigenspaces can be shared between all or some of the populations.
+
+Continuously parameterized covariance matrices. \citep{Hoff2011}, \citep{Chiu1996}, \citep{Yin2010}
+
+In such models, it is no longer reasonable to assume orthogonal equivariance .
+To address these issues, \citet{Hoff2009} introduced a model for
+eigenvector shrinkage based on the matrix Bingham distribution.
+
+Partial least squares discriminant analysis refs?
+\citep{Westerhuis2008, Szymanska2012, Eriksson2001, Barker2003, Saebo2008}
+
+Recently there has been significant interest gaussian graphical
+models, .   In gaussian graphical models, zeros in the
+precision matrix correspond to conditional independences between pairs
+of features given the remaining features \citep{}.  \citet{Witten2014} developed a
+method for pooled estimation inverse covariance matrices across
+multiple groups.  Identifying a shared pattern of zero's in the inverse covariance
+matrices for multiple classes of observations.
+
+In the following section, we
 propose a model for estimating and exploring differences between
@@
 
+Below we describe our approach for comparaing covariance matrices on
+a common low dimensional subspace.
+
 \section{The Shared Subspace Spiked Covariance Model}
+\label{sec:shared}
 
-Assume we have $k=1, ..., K$ groups of matrices $\bl S_k$ which follow
-the spiked covariance model with density as given in
-\ref{eqn:rplik}.    Each group consists of of $n_k$ measurements of $p$ features, typically with $n_k << p$.  However, we assume that the eigenvectors
-are related across groups in a particular way.  \citet{Hoff2009}
-introduced a model for eigenvector shrinkage based on the matrix Bingham
-distribution.
-
-In contrast, in this paper, we assume a priori that the eigenvectors
-are uniformally distributed on a low rank common subspace.  That is,
+Assume we have $k=1, ..., K$ groups of matrices $\bl S_k$ each of
+which follow a (degenerate) Wishart distribution, as in Equation
+\ref{eqn:lik}.  Each group consists of of $n_k$ measurements of $p$
+features, typically with $n_k << p$.  We pool information about the
+covariance matrices by assuming that most of the relevant variation
+for each group shares a low rank subspace spanned by
+the columns of a matrix $V$.  That is,
 
@@
 \Sigma_k = V\psi_kV^T + \sigma^2_kI
+\label{eqn:sspsi}
 \end{equation}
 
-With $V \in \mathcal{V}_{s,p}$ with $p
-\geq s \geq r$.  Here, $VV^T \in \mathcal{G}_{s, p}$ is the plane of
-variation shared by all groups and $\psi_k$ is the rank-r covariance matrix
-on that subspace.  \textbf{Rationale}.
+Here, $V \in \mathcal{V}_{s,p}$ and $p \geq s$.  $V$ is only
+identifiable up to right rotations.  Consequently,
+$VV^T \in \mathcal{G}_{s, p}$, which defines the plane of variation
+shared by all groups, is identifiable.  $\psi_k$ is a rank-$r$
+covariance matrix on the $s$-dimensional subspace.  Such a model is
+reasonable for many applications involving groups of highly structure
+data on the same set of $p$ features.  For instance, in biological
+applications, only certain subsets of molecules can physically
+interact which constrains the space of possible correlation patterns.
+Our model exploits this notion, by estimating differences between
+groups on a common lower dimensional subpace.
 
-In this paper, as in blah we consider a model on eigendecomposition of
-$\psi_k$ so that Equation \ref{} can be written as
+\section{Inference}
+\subsection{An Expectation-Maximization Algorithm}
+\label{sec:em}
+
+We start by outlining a computationally efficient algorithm for
+estimating the shared subspace, $VV^T$, noise $\sigma_k^2$ and
+projected covariance matrix $\psi_k$.
+
+First, note that the full likelihood can be written as
+
+\begin{align}
+p(S_1, ... S_k | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)  \\
+&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-(\sigma_k^2(V\psi_kV^T +
+  I))^{-1}\mathbf{S}_k/2) \\
+&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-\left[V(\psi_k +
+  I)^{-1}/\sigma_k^2 V^T + (I-VV^T)/\sigma^2_k\right]\mathbf{S}_k/2)
+  \\
+&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-r)/2}|M_k|^{-n_k/2}etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]\mathbf{S}_k/2) 
+\end{align}
+
+Where we define $M_k^{-1} = (\psi_k + I) ^{-1}/\sigma_k^2$.  The log-likelihood in
+$V$ (up to additive constant) is
+
+\begin{align}
+l(V) &= \sum_k tr\left(-VM_k^{-1}V^T +
+       VV^T/\sigma^2_k\right)\mathbf{S}_k/2)\\
+&= \sum_k \frac{1}{2}tr\left(-M_k^{-1}V^T \mathbf{S}_kV\right) + \frac{1}{2\sigma_k^2}tr\left(VV^T \mathbf{S}_k\right)
+\end{align}
+
+% \begin{align}
+% p(V|...) &\propto etr(\sum_{k=1}^K[O_k\Omega_kO^T_k]V^TS_k/(2\sigma^2_k)]V)\\
+% \end{align}
+
+We propose a computationally efficient expectation-maximizatino (EM)
+algorithm, to estimate the relevant quantities.  In particular we
+maximize the marginal posterior distribution of $V$, treating the
+$M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as the ``missing'' parameters to
+be imputed.  We assume independent Jeffreys' prior distributions for
+both $\frac{1}{\sigma_k^2}$ and $M_k^{-1}$ and a uniform prior for $V
+\in \mathcal{V}_{s, p}$.  In particular we let
+$p(\frac{1}{\sigma_k^2}) \propto \sigma_k^2$ and
+$p(M_k^{-1}) \propto |M_k^{-1}|^{-(r+1)/2}$.  From the likelihood it
+can easily be shown that the conditional posterior for $M_k^{-1}$ is
+
+$$p(M_k^{-1} | V) \propto |M_k^{-1}|^{(n_k - r -1)/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) $$
+
+\noindent which is a Wishart($V^T\mathbf{S}_kV$, $n_k$).  The
+conditional posterior distribution of $\frac{1}{\sigma_k^2}$ is simply
+
+$$p\left(\frac{1}{\sigma^2} | V\right) \propto \left(\frac{1}{\sigma_k^2}\right)^{n_k(p-r)/2-1}etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)\mathbf{S}_k\right)  $$
+
+\noindent which is a Gamma($n_k(p-r)/2$,
+$\frac{1}{2}tr((I-VV^T)\mathbf{S}_k)$).  Thus, for each iteration,
+$t$, of the EM algorithm:
+
+\begin{enumerate}
+\item For each $k$, compute relevant conditional expectations:
+\begin{enumerate}
+\item $E[M_k^{-1} | V_{(t-1)}] = n_k(V_{(t-1)}^T \mathbf{S}_kV_{(t-1)})^{-1}$
+\item $E[\frac{1}{2\sigma_k^2}|V_{(t-1)}] = \frac{n_k(p-r)}{tr((I-V_{(t-1)}V_{(t-1)}^T)S_k)}$
+\end{enumerate}
+\item Compute $V_{t}$ = $\underset{V}{argmax}  \sum_k tr\left(-VE[M^{-1}|V_{(t-1)}]^T +
+       E[\frac{1}{2\sigma_k^2}|V_{(t-1)}]VV^T\right)\mathbf{S}_k/2)$ 
+\end{enumerate}
+
+For the ``M-step'', we use a numerical optimization algorithm based on
+the Cayley transform to preserve the orthogonality constraints in $V$
+\citep{Wen2013}.  Importantly, the complexity of this algorithm is
+dominated by dimension of the shared subspace, $s$, not the number of
+features $p$.  Thus, our approach is computationally efficient for
+relatively small values of $s$, even when $p$ is large.
+
+% \subsection{Alternative formulation}
+% $M_k = \sigma_k^2(\Phi_k + I)$ where 
+% $$ \Phi_k =\left( \begin{array}{cc}
+% \Psi_k & 0  \\
+% 0 & D  \end{array} \right)$$
+
+% Letting $V = [V_1\, V_2]$
+
+% \begin{eqnarray*}
+% p(M_k^{-1} | V) &\propto \prod_k |M_k^{-1}|^{n_k/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) \\
+%                 & = (|\psi_k + I|^{-1}|D +
+%                   I|^{-1})^{n_k/2}etr(-1/2((\psi_k+I)^{-1}V_1^T\mathbf{S}_kV_1
+%                   -1/2(D+I)^{-1}V_2^T\mathbf{S}_kV_2)) \\
+% \end{eqnarray*}
+
+% $$p(\frac{1/d_i} | V) \propto \prod_k (d_i + 1)^{-n_k/2}exp\left(\frac{1}{-1/2(d_i+1)}(V_2^T\mathbf{S}_kV_2)_{i,i}\right) $$
+
+\subsection{Bayesian Inference}
+\label{sec:bayes}
+
+Although the EM algorithm presented in the previous section yields
+point estimates of $\psi_k$, it does not lead to natural uncertainty
+bounds for these estimates.  To quantify the uncertainty, we consider Bayesian
+inference for $\psi_k$, given the
+subspace $VV^T$.  Note that when the subspace is known, the posterior
+distribution of $\Sigma_k$ is conditionally independent from the other
+groups, i.e.  $P(\Sigma_{k} | V, \bl S_k, \Sigma_{-k}) = P(\Sigma_{k} | V, \bl S_k)$.
+Thus, we can independently estimate the conditional posterior
+distribution for each group.  In this paper, we focus on models based
+on the eigendecomposition of $\psi_k$.  In this setting, each covariance
+$\Sigma_k$ is a spiked covariance matrix where Equation \ref{eqn:sspsi}
+can be written as
 
@@
 \Sigma_k = VO_k\Lambda_KO_k^TV^T + \sigma^2_kI
+\label{eqn:ss}
 \end{equation}
 
-Where, $\Lambda_k$ is and $r \times r$ diagonal matrix of eigenvalues,
+Here, $\Lambda_k$ is an $r \times r$ diagonal matrix of eigenvalues,
 and $O_k \in \mathcal{V}_{s,r}$ is the matrix of eigenvectors of
-$\psi_k$.  Note that for any single group we can recover the original
-spiked covariance formuation (Equation
-\ref{}) by letting $U_k = VO_k$.
+$\psi_k$.  For any individual group, this corresponds to the original
+spiked covariance model (Equation \ref{eqn:spiked}) with $U_k = VO_k$.
 
@@
 \begin{equation}
-p(S | \Sigma_k,n_k) \propto \prod_{k=1}^K |\Sigma_k|^{-n/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)
+p(S | \Sigma_k,n_k) \propto |\Sigma_k|^{-n/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)
 \end{equation}
@@
 \begin{equation}
- L(\sigma_k^2,\bl V , O_k \Omega_k : \bl Y_k) \propto \prod_{k=1}^K\left[ (\sigma_k^2)^{-n_kp/2}etr(-\frac{1}{2\sigma_k^2}\mathbf{S}_k)\left(\prod_{i=1}^r(1-\omega_{ki}) \right) ^{n_k/2}  etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) \right]
+ L(\sigma_k^2,\bl V , O_k \Omega_k : \bl Y_k) \propto
+    (\sigma_k^2)^{-n_kp/2}etr(-\frac{1}{2\sigma_k^2}\mathbf{S}_k)\left(\prod_{i=1}^r(1-\omega_{ki})
+   \right) ^{n_k/2}
+   etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k)
+\label{eqn:sslik}
 \end{equation}
 
-\subsection*{An Expectation-Maximization Algorithm For Estimating the Shared Subspace}
+We use conjugate and semiconjugate priors for the parameters $O_k$,
+$\sigma^2_k$ and $\Omega_k$ to facilitate inference.  In the absence
+of real prior information, invariance considerations suggest the use
+of priors that lead to equivariant estimators.  Below we describe our
+choices for the prior distributions of each parameter and the
+resultant posterior distributions.
 
-We start by outlining a computationally efficient algorithm for
-estimating the shared subspace, $VV^T$.
+\paragraph{Conditional distribution of $\sigma_k^2$:}
 
-The full likelihood can be written as
+From Equation \ref{eqn:sslik} it is clear that the the inverse-gamma
+class of prior distributions is conjugate for$\sigma_k^2$.  We chose a
+default prior distriubtion for $\sigma^2$ that is equivariant with
+respect to scale changes.  Specifically, we use Jeffreys' prior, an
+improper prior with density $p(\sigma^2) \propto 1/\sigma^2 $.  Under
+this prior, straightforward calculations show that the full
+conditional distribution of $\sigma_k^2$ is
+inverse-gamma$( n_k p/2 , \tr(\bl S_k(\bl I -\bl U_k\Omega_k\bl
+U_k)/2)$, where $U_k = \hat{V}O_k$
 
-\begin{align}
-p(\bl S | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)  \\
-&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-(\sigma_k^2(V\psi_kV^T +
-  I))^{-1}\mathbf{S}_k/2) \\
-&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-\left[V(\psi_k +
-  I)^{-1}/\sigma_k^2 V^T + (I-VV^T)/\sigma^2_k\right]\mathbf{S}_k/2)
-  \\
-&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-r)/2}|M_k|^{-n_k/2}etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]\mathbf{S}_k/2) 
-\end{align}
-
-Where $M_k^{-1} = (\psi_k + I) ^{-1}/\sigma_k^2$.  The log-likelihood in
-$V$ (up to additive constant) is
-
-\begin{align}
-l(V) &= \sum_k tr\left(-VM_k^{-1}V^T +
-       VV^T/\sigma^2_k\right)\mathbf{S}_k/2)\\
-&= \sum_k \frac{1}{2}tr\left(-M_k^{-1}V^T \mathbf{S}_kV\right) + \frac{1}{2\sigma_k^2}tr\left(VV^T \mathbf{S}_k\right)
-\end{align}
-
-% \begin{align}
-% p(V|...) &\propto etr(\sum_{k=1}^K[O_k\Omega_kO^T_k]V^TS_k/(2\sigma^2_k)]V)\\
-% \end{align}
-
-To find the maximum marginal likelihood of $V$, we use an EM approach,
-treating the $M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as the ``missing''
-parameters to be imputed. %%We assume a Wishart prior on $M_k^{-1}$ and an Gamma prior
-on $\frac{1}{\sigma_k^2}$.  If we take these prior's to be
-``non-informative'' degenerate priors, we have
-
-We put a Jeffreys prior on both $M_k^{-1}$ and $\frac{1}{\sigma_k^2}$.
-
-$$p(M_k^{-1} | V) \propto |M_k^{-1}|^{n_k/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) $$
-
-\noindent which is a Wishart($V^T\mathbf{S}_kV$, $n_k + r + 1$).
-
-\noindent For $\frac{1}{\sigma_k^2}$ we have
-
-$$p\left(\frac{1}{\sigma^2} | V\right) \propto \left(\frac{1}{\sigma_k^2}\right)^{n_k(p-r)/2}etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)\mathbf{S}_k\right)  $$
-
-\noindent which is a Gamma($n_k(p-r)/2+1$, $\frac{1}{2}tr((I-VV^T)\mathbf{S}_k)$)
-
-\noindent Thus, for each iteration, $t$, of the EM algorithm:
-
-\begin{enumerate}
-\item For each $k$, compute relevant conditional expectations:
-\begin{enumerate}
-\item $E[M_k^{-1} | V_{t-1}] = (n_k + r  + 1)(V^T \mathbf{S}_kV)^{-1}$
-\item $E[\frac{1}{2\sigma_k^2}|V_{t-1}] = \frac{n_k(p-r) + 2}{tr((I-VV^T)S_k)}$
-\end{enumerate}
-\item Compute $V_{t}$ = $\underset{V}{argmax}  \sum_k tr\left(-VE[M^{-1}|V_{t-1}]V^T +
-       E[\frac{1}{2\sigma_k^2}|V_{t-1}]VV^T\right)\mathbf{S}_k/2)$ (optimization
-  approach from Wen et al, 2013)
-\end{enumerate}
-
-\subsection{Alternative formulation}
-$M_k = \sigma_k^2(\Phi_k + I)$ where 
-$$ \Phi_k =\left( \begin{array}{cc}
-\Psi_k & 0  \\
-0 & D  \end{array} \right)$$
-
-Letting $V = [V_1\, V_2]$
-
-\begin{eqnarray*}
-p(M_k^{-1} | V) &\propto \prod_k |M_k^{-1}|^{n_k/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) \\
-                & = (|\psi_k + I|^{-1}|D +
-                  I|^{-1})^{n_k/2}etr(-1/2((\psi_k+I)^{-1}V_1^T\mathbf{S}_kV_1
-                  -1/2(D+I)^{-1}V_2^T\mathbf{S}_kV_2)) \\
-\end{eqnarray*}
-
-$$p(\frac{1/d_i} | V) \propto \prod_k (d_i + 1)^{-n_k/2}exp\left(\frac{1}{-1/2(d_i+1)}(V_2^T\mathbf{S}_kV_2)_{i,i}\right) $$
-
-\section{Bayesian Inference}
-
-\subsection{Prior and posterior distributions}
-
-\paragraph{Conditional estimation of $(\sigma^2, \bl U)$:}
-Bayesian inference is facilitated with the use of semiconjugate 
-prior distributions  for $(\sigma^2, \bl U)$.
-These include the inverse-gamma class of priors for $\sigma^2$, 
-and the von Mises-Fisher-Bingham  
-distributions for $\bl U$ \citep{hoff_2009a,hoff_2009b}.  
-In the absence of real prior information,  invariance 
-considerations suggest the use of priors on $\sigma^2$ and 
-$\bl U$ that lead to equivariant estimators. 
-In particular, an (improper) prior having density 
-$p(\sigma^2,\bl U) = 1/\sigma^2 $, with respect to the 
-product of Lebesgue measure on $\mathbb R^+$ and the 
-uniform probability measure on $\mathcal V_{r,p}$, 
-leads to Bayes estimators $\hat \Sigma(\bl S)$ that 
-are equivariant with respect to 
-rotations and scale changes, so that 
-$\hat \Sigma(a \bl W \bl S \bl W^T) = 
-   a \bl W \hat\Sigma(\bl S) \bl W$ for all 
-$a>0$ and $\bl W\in \mathcal O_{p}$  
-(assuming an invariant loss function).  
-Under this invariant prior, straightforward calculations
-show that the full conditional distribution of
-$\sigma^2$ is
-inverse-gamma$( n p/2 , \tr(\bl S(\bl I-\bl U\Omega\bl U)/2)$.
-%Under the uniform prior on $\mathcal V_{r,p}$,
-The full conditional distribution of $\bl U$ has a density proportional
-to the likelihood, so
-\[
- p(\bl U | \bl S, \sigma^2, \Omega) \propto 
-  \etr( \Omega \bl U^T [\bl S/(2\sigma^2) \bl U). 
-\]
-This is a Bingham$(\Omega, \bl S/(2\sigma^2)$ distribution on
-$\mathcal V_{r,p}$ \citep{khatri_mardia_1977}. A Gibbs sampler to
-simulate from this distribution is given in
-\citet{hoff_2009a}.
-Interestingly, if $\Omega$ were known (which it is not), then 
-for a given invariant loss function
-the Bayes estimator under this prior 
-minimizes the (frequentist) risk among all 
-equivariant estimators \citep{eaton_1989}.
-
-\paragraph{Estimation of $\Omega$:} 
-We first consider a uniform(0,1) prior distribution  for
-each element of $\Omega$, or equivalently, an 
-$F_{2,2}$ prior distribution for the elements of $\Lambda$. 
-As a result, the full conditional distribution of 
-an element  $\omega_s$ of $\Omega$ is proportional to the 
-likelihood function, so 
-%an inverse-gamma$(\nu_0/2 , \nu_0 \sigma_0^2/2)$  prior distribution for  
-%$\sigma^2$  gives an inverse-gamma$(\nu_n/2 , \nu_n \sigma_n^2/2)$
-%full conditional distribution, 
-%where  $\nu_n = \nu_0+np$ and 
-%  $\nu_n \sigma_n^2  = \nu_0\sigma_0^2 + 
-%   \tr(\bl S(\bl I-\bl U\Omega\bl U))$. 
-\begin{equation} 
-p(\omega_s | \bl U, \bl S, \sigma^2)   \propto_{\omega_s} 
- l(\sigma^2,\bl U, \Omega:\bl S)  \propto_{\omega_s} 
-  (1-\omega_s)^{n/2} e^{c \omega_s  n/2},    
-\label{eqn:wpost}
-\end{equation}
-where $c= \bl u_s^T \bl S \bl u_s/(n \sigma^2)$ and 
-$\bl u_s$ is column $s$ of $\bl U$. 
-While not proportional to a density belonging to a standard class 
-of distributions, 
-the corresponding 
-univariate distribution is easy to sample from numerically, and 
-its 
-behavior is straightforward to 
-understand:
-%First note that the likelihood 
-%\ref{eqn:lik} is maximized in 
-%if $\bl S \approx n \Sigma$, 
-%and $\bl U$ and $\sigma^2$ are approximately close to thier 
-%correct values, then $c\approx  $
-If $c\leq 1$, then the the function has a maximum 
-at $\omega_s =0$, and decays monotonically to zero as $\omega \rightarrow 1$. 
-%If $c<n/2$ then the derivative is negative at zero, 
-%whereas 
-If $c>1$ then the function is uniquely  maximized at 
-$(c-1)/c \in (0,1)$. 
-To see why this makes sense, note that the likelihood is maximized when 
-the columns of 
-$\bl U$ are equal to the eigenvectors of $\bl S$ corresponding to its top 
-$r$ eigenvalues \citep{tipping_bishop_1999}. At this value of $\bl U$, 
-$c$ will then equal one of the top $r$ eigenvalues of $\bl S/(n\sigma^2)$. 
-In the case  that $n\gg p$, we expect  $\bl S/(n\sigma^2)\approx \Sigma_0/\sigma^2$, 
-the true (scaled) population covariance, and so we expect $c$ to be near one of the 
-top $r$ eigenvalues of $\Sigma_0/\sigma^2$, say $\lambda_0+1$. 
-If indeed $\Sigma_0$ has 
-$r$ spikes, then  $\lambda_0>0$, $c \approx \lambda_0 +1 > 1$, and so the conditional 
-mode of $w_s$ is approximately $(c-1)/c = \lambda_0/(\lambda_0+1)$, 
-the correct value. 
-On the other hand, if we have assumed the existence of a spike when 
-there is none, then $\lambda_0=0$, $c\approx 1$ and the Bayes estimate 
-of $w_s$ will be shrunk towards zero, as it should be.
-
-\subsection*{Conditional distribution of $\sigma_k^2$}
-TODO: Inverse-Gamma conjugacy
-
-\subsection*{Conditional distribution of $\Lambda$}
-Here, we specify the conditional distribution of $\Omega = \Lambda(I+\Lambda)^{-1}$.
-
-\begin{align}
-p(\Omega_k|V,O_k,\Lambda_k,\Sigma_k) \propto \left(\prod_{i=1}^r(1-\omega_{ki})^{n_k/2}  \right) etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) 
-\end{align}
-
-TODO: Finish
-
-\subsection*{Conditional distribution of $O_k$}
+\paragraph{Conditional distribution of $O_k$:} Given the likelihood
+from Equation \ref{eqn:sslik}, it is easy to show that the class of
+von Mises-Fisher-Bingham distributions are conjugate for $O_k$
+\citep{Hoff2009, Hoff2012}.  Again, invariance considerations
+lead us to use a rotationally invariant uniform probability measure on
+$\mathcal V_{s,p}$.  Under this uniform prior, the full conditional
+distribution of $\bl O_k$ has a density proportional to the
+likelihood, so
 
@@
 \label{lik_vo}
- p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto etr(\Omega_kO^T_kV^T[S_k/(2\sigma^2_k)]VO_k)
+ p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto etr(\Omega_kO^T_k\hat{V}^T[S_k/(2\sigma^2_k)]\hat{V}O_k)
 \end{align}
 
-Assuming a uniform distribution on $O_k$, the conditional distribution of $O_k$ is clearly a matrix Bingham distribution with parameters $\Omega_k$ and $V^T[S_k/(2\sigma^2_k)]V$.
-It may be desirable to specify a conjugate prior for $O_k$:
+\noindent This is a Bingham$(\Omega, \hat{V}^T \bl S_k \hat{V}/(2\sigma^2)$
+distribution on $\mathcal V_{s, p}$ \citep{Khatri1977}. A
+Gibbs sampler to simulate from this distribution is given in
+\citet{Hoff2012}.
 
-\begin{equation}
-P(O_k|\Omega_k) = etr(\Omega_kO_kHO_k^T)
-\end{equation}
+Together, the prior for $\sigma_k^2$ and $O_k$ leads to conditional
+(on $V$) Bayes estimators $\hat \Sigma(V^T \bl S V)$ that are
+equivariant with respect to scale changes and rotations on the
+subspace spanned by $V$, so
+that $\hat \Sigma(a \bl W V^T \bl S V \bl W^T) = a \bl W \hat\Sigma(V^T
+\bl S V) \bl W$
+for all $a>0$ and $\bl W\in \mathcal O_{s}$ (assuming an invariant
+loss function). Interestingly, if $\Omega$ were known (which it is
+not), then for a given invariant loss function the Bayes estimator
+under this prior minimizes the (frequentist) risk among all
+equivariant estimators \citep{Eaton1989}.
 
-in which case the conditional distribution is simply
-\begin{equation}
-P(O_k|...) \propto etr(\Omega_kO^T_k(V^T[S_k/(2\sigma^2_k)]V + H)O_k)
-\end{equation}
-which is still a matrix Bingham with parameters $\Omega_k$ and $V^T[S_k/(2\sigma^2_k)]V+H$
-\subsection*{Conditional distribution of $V$}
+\paragraph{Conditional distribution for $\Omega$:} Here we specify the conditional
+distribution of the diagonal matrix $\Omega_k =
+\Lambda_k(I+\Lambda_k)^{-1} = \text{diag}(\omega_{k1}, ... \omega_{kr})$.  We first consider a uniform(0,1) prior distribution for each element of $\Omega$, or
+equivalently, an $F_{2,2}$ prior distribution for the elements of
+$\Lambda$.  The full conditional distribution of an
+element $\omega_i$ of $\Omega$ is proportional to the likelihood
+function, so
+
 \begin{align}
-p(V|...) &\propto etr(\sum_{k=1}^K[S_k/(2\sigma^2_k)]V[O_k\Omega_kO^T_k]V^T)\\
+p(\omega_{ki}|V, O_k, \bl S_k) \propto_{\omega_{ki}}
+  \left(\prod_{i=1}^r(1-\omega_{ki})^{n_k/2}  \right)
+  etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) \\
+&  \propto & (1-\omega_{ki})^{n/2} e^{c \omega_s  n/2},    
+\label{eqn:wpost}
 \end{align}
 
-Let, $A_k = S_k/(2\sigma^2_k)$ and $B_k = O_k\Omega_kO^T_k$.  We can write the conditional distribution of $V$ (again assuming a uniform prior) given $A_k$ and $B_k$ as
-\begin{align}
-P(V|A_k,B_k) &\propto etr(\sum_{k=1}^KA_kVB_kV^T)\\
-&= etr(\sum_{k=1}^K\sum_{i,j}^sA_kv_iv_j^Tb_{ijk})\\
-&= etr(\sum_{i,j}^s \sum_{k=1}^K v_j^T[b_{ijk}A_k]v_i)
-\end{align}
+where $c = \bl u_{ki}^T S_k \bl u_{ki}/(n \sigma^2)$ and $\bl u_{ki}$ is
+column $i$ of $U_k = VO_k$.  While not proportional to a density
+belonging to a standard class of distributions, we can sample from the
+corresponding univariate distribution numerically.  The behavior of
+this distribution is straightforward to understand: if $c\leq 1$, then
+the the function has a maximum at $\omega_i =0$, and decays
+monotonically to zero as $\omega \rightarrow 1$.  If $c>1$ then the
+function is uniquely maximized at $(c-1)/c \in (0,1)$.  To see why
+this makes sense, note that the likelihood is maximized when the
+columns of $\bl U_k$ are equal to the eigenvectors of $\bl S_k$
+corresponding to its top $r$ eigenvalues
+\citep{Tipping1999}. At this value of $\bl U_k$, $c$ will then
+equal one of the top $r$ eigenvalues of $\bl S_k/(n\sigma_k^2)$.  In the
+case that $n\gg p$, we expect
+$\bl S_k/(n_k\sigma_k^2)\approx \Sigma_k/\sigma_k^2$, the true (scaled)
+population covariance, and so we expect $c$ to be near one of the top
+$r$ eigenvalues of $\Sigma_0/\sigma^2$, say $\lambda_0+1$.  If indeed
+$\Sigma_k$ has $r$ spikes, then $\lambda_k>0$,
+$c \approx \lambda_0 +1 > 1$, and so the conditional mode of $w_i$ is
+approximately $(c-1)/c = \lambda_i/(\lambda_i+1)$, the correct value.
+On the other hand, if we have assumed the existence of a spike when
+there is none, then $\lambda_0=0$, $c\approx 1$ and the Bayes estimate
+of $w_i$ will be shrunk towards zero, as it should be.
 
-From here, we specify the conditional distribution of a single column of $V$ given $A_k$, $B_k$ and all other columns of V:
+% \paragraph{Conditional distribution of $V$}
+% Need to finish / should we include?
 
-\begin{align}
-P(v_i|A_k,B_kV_{-i})  &\propto etr(2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])v_i+v_i^T[\sum_{k=1}^K b_{ijk}A_k]v_i))\\
-&= etr(Cv_i+v_i^TMv_i)
-\end{align}
-
-This is a vector BMF(C,M) with $C = 2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])$ and $M= \sum_{k=1}^K b_{ijk}A_k$.
+% Although we prefer EM... it is also posible to estimate $V$ using
+% Bayesian inference.  We assume a uniform prior on $V \in \mathcal{V}_{p, s}$ so that the
+% conditional posterior of $V$ is proportional to the likelihood:
 
 % \begin{align}
-% P(V|O_k,\Omega_k,\sigma^2_k,\mathbf{S}_k) = (\sigma^2_k)^{-n_kp}etr(\sum_{k=1}^K[S_k/(2\sigma^2_k)]V[O_k\Omega_kO^T_k]V^T)\\
+% p(V|\bl O, \bl S, \bl \sigma^2, \bl \Omega) &\propto etr(\sum_{k=1}^K[S_k/(2\sigma^2_k)]V[O_k\Omega_kO^T_k]V^T)\\
 % \end{align}
 
-\pagebreak
+% Although close in appearance, this conditional distribution is not a matrix
+% Bingham-von Mises-Fisher distribution.  However, the conditional distribution
+% of a column of $V$ given all other columns is a vector Bingham-von
+% Mises-Fisher distribution.  To see this, let
+% $A_k = S_k/(2\sigma^2_k)$ and $B_k = O_k\Omega_kO^T_k$.  We can write
+% the conditional distribution of $V$ given $A_k$ and $B_k$ as
+% \begin{align}
+% P(V|A_k,B_k) &\propto etr(\sum_{k=1}^KA_kVB_kV^T)\\
+% &= etr(\sum_{k=1}^K\sum_{i,j}^sA_kv_iv_j^Tb_{ijk})\\
+% &= etr(\sum_{i,j}^s \sum_{k=1}^K v_j^T[b_{ijk}A_k]v_i)
+% \end{align}
+
+% From here, we specify the conditional distribution of a single column of $V$ given $A_k$, $B_k$ and all other columns of V:
+
+% \begin{align}
+% P(v_i|A_k,B_kV_{-i})  &\propto etr(2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])v_i+v_i^T[\sum_{k=1}^K b_{ijk}A_k]v_i))\\
+% &= etr(Cv_i+v_i^TMv_i)
+% \label{eqn:vbmf}
+% \end{align}
+
+% This is a vector Bingham-von Mises-Fisher Distribution, BMF(C, M)
+% with $C = 2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])$ and $M=
+% \sum_{k=1}^K b_{ijk}A_k$.  Following \citet{Hoff2012},  we can derive
+% a Gibbs sampler for $V$ based on conditional draws of the columns of
+% $V$.
 
@@
 
-First, we present the results of a simulation which demonstrates the
+We present results of a simulation which demonstrates the
 relative performance of shared subspace group covariance estimation
-under different data generating models.  We consider three different
+under different data generating models.  In this simulation, we consider three different
 spiked covariance data models: 1) ``complete pooling'', in which the
-covariance matrices from all groups are identical, e.g.
-$\Sigma_1 = \Sigma_2 = ... \Sigma_K$; 2) ``shared subspace'', where
-the signal lies on the same low rank subspace (Equation \ref{}); and 3) ``no
-pooling'' assuming little overlap in the eigenvectors of the
-covariance matrices (Equation \ref{}).  For each of these data models, we
-simulated 30 datasets from a spiked covariance model with rank $r=2$
-and eigenvalues $(\lambda_1, \lambda_2) = (250, 25)$.
+spiked covariance matrices from all groups are identical, e.g.
+$\Sigma_1 = \Sigma_2 = ... \Sigma_K = U\Lambda U^T + \sigma^2I$; 2)
+the ``shared subspace'' model, outlined in Section \ref{sec:shared};
+and 3) ``no pooling'', e.g. the data from each group is generated from
+a spiked covariance matrix with each $U_k$ drawn as an independent
+uniform sample from the Stiefel manifold $\mathcal{V}_{r, p}$.  For
+each of these data generating models, we simulate datasets
+with $p=200$ features, rank $r=2$ and eigenvalues
+$(\lambda_1, \lambda_2) = (250, 25)$.  For data generated under the
+shared subspace model, we additionally assume that the first two
+eigenvectors span a common two dimensional subspace, e.g. we fix
+$s=2$.
 
-We evalute the three corresponding inferential models,
-corresponding to the complete pooling blah and blah.
+We evaluate the performance of the three corresponding inferential
+models on thirty simulated datasets for each data generating
+mechanism.  In practice, the true rank of the covariance matrix is not
+known, and thus, prior to fitting each of these inferential models we
+must first estimate the rank.  \citet{Gavish2014} provide an
+asymptotically optimal (in mean squared error) singular value
+threshold for low rank matrix recovery with noisy data.  We use their
+rank estimator, which is a function of the median singular value of
+the data matrix and $\frac{p}{n_k}$, to independently estimate the
+rank, $r$, of the spiked covariance matrices.  For data generated
+under the shared subspace model we must also estimate the dimension of
+the shared subspace.  As a useful heuristic, we use Gavish and
+Donoho's estimator on the pooled data to estimate the shared subspace
+dimension, $s$.
 
-For each group, $k$, we compute Stein's loss
+Given estimates of $s$ and $r$, we then proceed with Bayesian
+inference of the eigenvectors, eigenvalues, and noise variance using
+each of the three inferential models.  When applying the shared
+subspace estimator, we use the EM algorithm presented in Section
+\ref{sec:em} first, to estimate $VV^T$ and then use Bayesian inference
+to estimate the posterior distribution of all remaining quantities
+conditional on $V$.  For all covariance estimates, we compute Stein's
+loss
 $L_S( \Sigma_k , \hat\Sigma_k) = \text{tr}( \Sigma_k^{-1} \hat
-\Sigma_k ) - \log |\Sigma_k^{-1} \Sigma_k | - p$.
-Under Stein's loss, the Bayes estimator is the inverse of the
-posterior mean of the precision matrix,
-$\hat \Sigma_{BS} = \Exp{ \Sigma^{-1} | \bl S}^{-1}$ which can easily
-be esimated using posterior samples.  In Table \ref{table:groupLoss}
-we report the average group loss
+\Sigma_k ) - \log |\Sigma_k^{-1} \Sigma_k | - p$
+for each group.  Under Stein's loss, the Bayes estimator is the
+inverse of the posterior mean of the precision matrix,
+$\hat \Sigma_{k} = \Exp{ \Sigma_k^{-1} | \bl S_k}^{-1}$ which we estimate
+using MCMC samples.  In Table \ref{table:groupLoss} we report the
+average group loss
 $L(\Sigma_1, ..., \Sigma_K; \hat\Sigma_1, ..., \hat\Sigma_K ) =
-\frac{1}{K} \sum_k L_S( \Sigma_k , \hat\Sigma_k)$.
+\frac{1}{K} \sum_k L_S( \Sigma_k , \hat\Sigma_k)$
+and 95\% intervals (TODO) for the estimates derived using each
+inferential model.
 
-\citep{Paul2007, Donoho2013}.
+Naturally, the estimates with the lowest loss are derived from the
+inferential model that matches the data generating model. However, the
+shared subspace estimator has small loss under model misspecification
+relative to the alternative.  For example, when the data is
+generated using the complete pooling model, the shared subspace
+estimator has almost four times smaller loss than the no pooling
+estimator.  When the data is generated under the no pooling model, the
+shared subspace estimator is over an order of magnitude better than
+the complete pooling estimator.
 
-The results demonstrate that
+Thus, shared subspace estimation can be an appropriate choice when the
+amount of similarity between groups is not known a priori.  The shared
+subspace model combined with the rank estimation procedure suggested
+by \citet{Gavish2014} leads to an estimator which adapts to the amount
+of similarity between groups.  Importantly, the shared subspace model
+can be viewed as a generalization of the ``no pooling'' spiked
+covariance model.  The two are equivalent when we let
+the shared subspace span the full space, e.g. $V \in \mathcal{O}(p)$.
+Figure \ref{fig:sdimension} depicts the loss for data generated under
+the no pooling model and fit with shared subspace estimators with
+different dimension.  As the shared subspace dimension increases to
+the dimension of the feature space, $s \rightarrow p$, the loss for
+the shared subspace estimator quickly coverges to the loss for the
+individual spiked covariance estimator.  Note that
+span($[U_1, ..., U_K]) \leq rK$, thus even when there is little
+similarity between the eigenvectors from each group, the shared
+subspace estimator with $s = rK$ will perform comparably to the ``no
+pooling'' estimator, provided that we can identify a subspace where
+span($\hat{V})$ is close to $\text{span}([U_1, ..., U_K])$.
+
+In many real world examples, we are more interested in finding a
+useful space to interpret and compare differences between covariance
+matrices than we are in finding the lowest-loss estimator for the full
+covariance matrices.  In this context, the relevant quantity is often
+an estimate of the projected covariance, $V^T\Sigma V$.  In Figure
+\ref{fig:lossBoxplot} we depict boxplots for the distribution of
+losses,
+$L_S(\hat{V}^T\Sigma_k\hat{V}, \hat{V}^T\hat{\Sigma}_k\hat{V})$, on
+simulated data.  In the left boxplot the data come from a
+2-dimensional shared subspace model and on the right plot the data
+come from a 10-dimensional shared subspace model.  For both boxplots,
+estimates
+$\hat{V}^T\hat{\Sigma}_k\hat{V}$ are derived from assumed 2-dimensional
+model. This figure demonstrates that the loss for estimating the
+projected covariance matrix is not significantly larger when we
+underestimate the dimension of the true shared subspace.
 
@@
     of the covariance matrix share a subspace dimension $s=2$.  All results are
-    based on $n_k = 50$ observations per group.  Shared subspace
-    inference blah between fitting each group covariance independently
-  and assuming that they are all the same.  }
+    based on $n_k = 50$ observations per group.  }
 \label{table:groupLoss}
@@
 
-\begin{figure}[!ht]
+\begin{figure}[t]
     \centering
-    \includegraphics[width=0.5\textwidth]{Figs/LossVsDimension}
-  \caption{Stein's loss for data generated under the ``no pooling''
-    model and fit using the shared subspace model.  As the shared
-    subspace dimension increases,  $s \rightarrow p$, the loss converges
-    to the loss from the independent spiked covariance estimates.
-    Thus, the shared subspace estimator is a generalization of blah.  }
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{Figs/LossVsDimension}
+        \caption{Ratio}
+        \label{fig:sdimension}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{Figs/lossBoxplot}
+        \caption{Posterior}
+        \label{fig:lossBoxplot}
+    \end{subfigure}
+    \caption{Left) Stein's loss for data generated under the ``no
+      pooling'' model and fit using the shared subspace model.  As the
+      shared subspace dimension increases, $s \rightarrow p$, the loss
+      converges to the loss from the independent spiked covariance
+      estimates. Right) Distribution of Stein's loss
+      $L_S(\hat{V}^T\Sigma_k\hat{V}, \hat{V}^T\hat{\Sigma_k}\hat{V})$
+      for data generated from a two dimensional shared subspace (left
+      box) and data generated with a ten dimensional shared subspace
+      (right).  In both cases the estimates were computed assuming a
+      two-dimensional subspace, $s$ and matrix rank $r=2$.  Even when
+      we underestimate the dimension of the shared subspace, the
+      average loss of the estimator for the projected covariance matrices is
+      similar to the loss under the correctly specified model. }
 \end{figure}
 
-% n <- 50
-% S <- 2
-% R <- 2
-% P <- 200
-% ngroups <- 10
+\subsection{Model Checking and Uncertainty Quantification}
 
-% evals <- c(250, 25)
-% niters <- 1000
-% nwarmup <- niters/2
+If the shared subspace model is reasonable and we succeed in
+identifying an appropriate subspace, $\hat{V}$, then for each $k$,
+most of the variation in $\Sigma_k$ should lie on the subspace spanned
+by $\hat{V}$.  To quantify the extent to which this is true for
+different groups, we propose a simple estimator for the proportion of
+``signal'' variance that lies on any given subspace.  Specifically, we
+apply the following estimator for ratio of the sum of the first $s$ eigenvalues
+of $\hat{V}^T\bl \Sigma_k \hat{V}$ to the sum of the first $s$
+eigenvalues of $\Sigma_k$:
 
-\subsection{Evaluating the Subspace}
+\begin{equation}
+ \frac{||Y_k^T\hat{V}||_F}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{sup}
+  ||Y_k^T\widetilde{V}||_F - \hat{\sigma}_k^2ps}
+\label{eqn:ratio}
+\end{equation}
 
-$$L_SS(\hat{V}^T\Sigma_K\hat{V} , \hat{V}^T\hat{\Sigma}_K\hat{V}) $$
+If the inferred subspace is the true subspace, e.g. $\hat{V}\hat{V}^T = VV^T$,
+then this estimator will asymptotically approach one.  To see this, note that
+$\underset{V \in \mathcal{V}_{p, s}}{sup}\frac{||Y_k^TV||_F}{n_k}$ is
+equivalent to the sum of the first $s$ eigenvalues of the sample
+covariance matrix $\frac{\bl S_k}{n_k}$.  With $\hat{\lambda}^{(k)}_i$
+the $i$-th eigenvalue of $\frac{\bl S_k}{n_k}$, \citet{Baik2006}
+demonstrated that asymptotically as $p, n_k \rightarrow \infty$ with
+$\frac{p}{n_k} = \gamma_k$ fixed
 
-We propose a simple metric  to evaluate how much of the variation in $Y_k$ lies on the
-subspace spanned by $\hat{V}$:
+\begin{eqnarray}
+\hat{\lambda}^{(k)}_i &\rightarrow& \lambda^{(k)}_i\left(1 +
+                                    \frac{\sigma_k^2\gamma}{\lambda^{(k)}_i
+                                    - 1}\right)\\
+& \approx& \lambda^{(k)} + \sigma^2_k\frac{p}{n_k}
+\end{eqnarray}
 
-$$ \frac{||Y_k^T\hat{V}||_F}{\underset{V \in \mathcal{V}_{p, s}}{sup}
-  ||Y_k^TV||_F - \hat{\sigma}_k^2ps}$$
-
-This quantity is  as an estimate of the ratio of the first $s$
-sum of the eigenvalues of $V^T\bl \Sigma_k V$ to the sum of the first
-$s$ eigenvalues of $\Sigma_k$.  To see this, note that $\underset{V \in \mathcal{V}_{p, s}}{sup}\frac{||Y_k^TV||_F}{n_k}$
-is equivalent to the sum of the first $s$ eigenvalues of the sample
-covariance matrix $\frac{\bl S_k}{n_k}$.  Letting
-$\hat{\lambda}^{(k)}_i$ = the $i$-th eigenvalue of $\frac{\bl S_k}{n_k}$.  Asymptotically as $p, n_k \rightarrow \infty$ with
-$\frac{p}{n} = \gamma$
-
-$ \hat{\lambda}^{(k)}_i\rightarrow \lambda^{(k)}_i
-(1 + \frac{\sigma_k^2\gamma}{\lambda^{(k)}_i - 1} \approx \lambda^{(k)}
-+ \sigma^2_k\frac{p}{n_k}$ \citep{Baik2006}.
-
-$\sum_i^s \hat{\lambda}^{(k)}_i \approx \sum_i^s \lambda^{(k)}_i + \sigma_k^2ps$
-
-\begin{figure}[!ht]
-  \centering
-    \includegraphics[width=0.5\textwidth]{Figs/lossBoxplot}
-  \caption{}
-\end{figure}
+Thus,
+$\sum_i^s \hat{\lambda}^{(k)}_i \approx \sum_i^s \lambda^{(k)}_i +
+\sigma_k^2ps$.
+If we assume that $s$ is fixed in this asymptotic regime, then
+$||Y_k^TV||_F/n_k$ is a consistent estimator for
+$\sum_i^s \lambda^{(k)}_i$, and thus the above estimator will be close
+to one when $\hat{V}\hat{V}^T = VV^T$ and smaller if not.  Figure
+\ref{fig:evalRatios} depicts this estimate for each group in two
+different simulated datasets.  When the shared subspace model is
+appropriate, the subspace spanned by $\hat{V}$ explains a large
+proportion of variance in $\Sigma_k$ in all groups.  When the
+estimated subspace is too small, the proportion of variance will be
+small for at least some of the groups.  The metric thus provides a
+useful indicator for which groups can be reasonably compared on a
+given subspace and which groups cannot.
 
@@
     \includegraphics[width=0.75\textwidth]{Figs/evalRatios}
-  \caption{}
+    \caption{Proportion of variance explained by the estimated
+      subspace, $\hat{V}\hat{V}^T$, for each of 10 groups.  Left) Data
+      generated and fit under the shared subspace with $s=r=2$.  The
+      estimated two dimensional subspace explains nearly all of the
+      variance in each of the ten groups.  Right) Data generated from
+      the shared subspace model with $s=10$ and $r=2$ but estimated
+      assuming $s=r=2$.  In this case, a two dimensional subspace does
+      not capture the proportion of variance for all groups.  Here,
+      the inferred subspace explains over 50\% of the variance for
+      groups 2, 5, 7, 9 and 10 but almost none of the variance for
+      groups 6 and 8. }
 \label{fig:evalRatios}
 \end{figure}
+
+\paragraph{Uncertainty Quantification}
+
+The EM algorithm presented in Section \ref{sec:em} provides a fast way
+to estimate the shared subspace $VV^T$, the rank $r$ covariance,
+$U_k\Lambda_kU_k^T$, and the homoscedastic noise $\sigma_k^2$.
+Although the EM algorithm is useful for deriving point estimates, it
+does not provide associated uncertainty bounds.  Uncertainty
+quantification is particularly important in the $n << p$ setting where
+seemingly large differences in the low rank covariance matrices across groups may not
+actually be statistically significant.
 
@@
     \includegraphics[width=0.5\textwidth]{Figs/posteriorRegions}
-  \caption{}
+    \caption{95\% posterior regions for the ratio of the eigenvalues,
+      $\frac{\lambda_1}{\lambda_2}$ and the angle on the shared
+      subspace determined spanned by $V$.  The horizontal dashed line
+      represents the true ratio of eigenvalues
+      $\frac{\lambda_1}{\lambda_2} = 10$.  Colored tick marks at the
+      bottom of the figure represent ``true'' orientation of the
+      principle axis for each group relative to axes determined by
+      $\hat{V}R$.  Since $V$ is only identifiable up to rtation, we
+      fix R using the Procrustes rotation that maximizes the
+      similarity of $\hat{V}R$ to the true data generating $V$. }
 \label{fig:posteriorRegions}
@@
 
-\section{An Analysis of Gene Expression Data}
+In order to quantify the uncertainty associated with the relevant
+parameter estimates, we use the Bayesian inference algorithm outlined
+in Section \ref{sec:bayes}. In Figure \ref{fig:posteriorRegions} we
+depict the 95\% posterior regions for statistics based on the
+eigendecomposition of $\hat{V}^T\hat{\Sigma}\hat{V}$.  Specifically,
+for the two dimensional shared subspace model, we plot the eigenvalue
+ratio $\frac{\lambda_1}{\lambda_2}$ and the angle of the first
+eigenvector on the $\text{arctan}(\frac{O_1}{O_2})$, where
+$\lambda_1 > \lambda_2$.  Non-overlapping 95\% posterior regions are
+indicative of significant differences in the covariance structure
+between groups.
+
+In this example, the posterior regions for all groups cover the true
+eigenvalue ratio, $\frac{\lambda_1}{\lambda_2} = 10$, even though the
+posterior mean for this quantity varies by a factor of 2 across groups.  We also
+estimate how well we recover the true orientation of the covariance
+matrix.  Since $V$ is only identifiable up to rotations, we estimate
+the Procustes transformation, R, which maximizes the similarity
+between $\hat{V}R$ and $V$.  The posterior regions for each group also
+cover the correct orientation.
+
+\section{Analysis of Biological Data}
+
+As in many natural systems, high dimensional biological data is often
+very structured and thus can be well understood on a relatively low
+dimensional subspace. For example, with gene expression
+data, the effective dimensionality is thought to scale with the number
+of gene regulatory modules, not the number of genes themselves
+\citep{Heimberg2016}.  As such, differences in gene expression across
+groups should be expressed in terms of differences between these
+regulatory modules rather than strict differences between expression
+levels.  Such differences can be examined on a subspace that reflects
+the correlations resulting from these modules.
+
+With this in mind, we apply the shared subspace spiked covariance model to data in two
+different biological examples.  In the first, we compare gene
+expression data from juveniles with different subtypes of leukemia.  Even
+after removing mean effects, the data indicate significant differences
+in covariance matrices between groups.  In the second example, we
+compare covariance matrices in a metabolomic analysis of fly data.  In
+this example, after controlling for mean effects of fly age and
+gender, we find little evidence of differences between metabolite
+covariances across groups.
+
+\paragraph{Analysis of Gene Expression Data}
+
+Here, we demonstrate the utility of the shared subspace
+covariance estimator for exploring differences in the covariability of
+gene expression levels in young adults with different subtypes of
+pediatric acute lymphoblastic leukemia \citep{Yeoh2002}.  The raw data
+consists of gene expression levels for over 12,000 genes in seven
+different subtypes of leukemia: BCR-ABL, E2A-PBX1, Hyperdip50, MLL,
+T-ALL, TEL-AML1 and a seventh group for unidentified subtypes.  The
+groups have corresponding sample sizes of
+$n = (15, 27, 64, 20, 43, 79, 79)$.  Although there are over 12,000
+genes, the vast majority of expression levels are missing.
+Thus, we restrict our attention to the genes for which less than half
+of the values are missing and use Amelia, a software package for
+missing value imputation, to fill in the remaining missing values
+\citep{Amelia}.  After restricting the data in this way, $p=3124$
+genes remain.  Prior to analysis, we demean both the rows and columns of
+the gene expression levels in each group.
+
+In Figure \ref{fig:leukemia} we plot the results of our analysis.
+Panel \ref{fig:leukemiaRatio} shows that over $40\%$ of the variance
+can be explained by the estimated two-dimensional subspace for all groups, with 4 of
+the 6 groups over 60\%.  Panel \ref{fig:leukemiaPosterior} reflects some significant
+differences in the posterior distribution of eigenvalues and
+eigenvectors between groups.  The $x$-axis corresponds to the
+orientation of the principle axis of $\psi_k$ on $V$ and the $y$-axis
+corresponds to the ratio of the larger eigenvalue over the smaller
+eigenvalue.  Several groups show significantly different orientations
+for the first principle component of the subspace $\hat{V}$.  Further,
+the subtype E2A-PBX1 has a larger eigenvalue ratio which reflects a
+more correlated distribution on the subspace spanned by $V$.  The
+TEL-AML1 and Hyperdip50 appear to cluster together which suggests
+there is little detectable difference between their covariance
+structure.  Note that when the eigenvalue ratio is close to one (as is
+the case for the ``other'' subtype), the distribution of expression
+values on the subspace is nearly spherical and thus the orientation of
+the primary eigenvector is at best weakly identifiable.  This is
+reflected in the wide posterior range of orientations of the first
+principle component.
+
+Further intuition about the differences in covariances between groups
+can be understood in the biplot in Figure
+\ref{fig:leukemiaBiplot}.  Here, we plot the contours of the
+covariance matrices for three of subtypes of Leukemia.  We also plot
+points for the 1\% of genes with the largest loadings on the first two columns
+of $V$.  The genes in each cluster are
+listed in the corresponding table.  Even though we remove all
+mean-level differences in gene expressions between groups, we are
+still able to identify genes with known connections to cancer and
+leukemia.  As one example, MYC (top right) is an oncogene with well
+established association to many cancers \citep{Dang2012}.
+
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{Figs/leukemiaRatio-04-27}
+        \caption{Proportion of Variance on $\hat{V}$}
+        \label{fig:leukemiaRatio}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{Figs/leukemiaPosterior-04-27}
+        \caption{95\% Posterior Regions }
+        \label{fig:leukemiaPosterior}
+    \end{subfigure}
+\caption{}
+\label{fig:leukemia}
+\end{figure}
+
+In this figure, genes that fall in the upper right quadrant
+($\blacksquare$) have positive loadings in both columns of $V$. On the
+other hand, genes in the upper left corner have positive loadings on
+the first column of $V$ but negative loadings on the second column of
+$V$.  When the principle axis has an orientation close to
+$\frac{\pi}{4}$, genes that fall in the upper right or lower left
+quadrant will exhibit larger variability, whereas genes in the upper
+left or lower right exhibit little variability since their loadings
+effectively cancel out.  On the other hand, when the principle axis is
+closer to $-\frac{\pi}{4}$, the genes in the upper left
+($\blacktriangle$) and lower right exhibit more variability.  In this case,
+the genes in the upper left portion of the plot show large correlated
+variability for the Hyperdip50 group but not for E2A-PBX1 group.
+Similarly, genes in the upper right tend to vary together more for
+E2A-PBX1 but not for the Hyperdip50 group.  Both of these sets of
+genes have large variability for the BCR-ABL group because the
+principle axis of variation aligns predominantly with $V2$.
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{Figs/leukemiaRatio}
-  \caption{}
-\label{fig:leukemiaRatio}
+    \includegraphics[width=0.5\textwidth]{Figs/leukemia-biplot-05-06}
+    \qquad
+\raisebox{\height}{
+\footnotesize
+\input{Figs/biplot.tab}
+}
+\caption{Left) Variant of a biplot with contours for three leukemia
+  subtypes and the loadings of the f Right) Gene's with }
+\label{fig:leukemiaBiplot}
   \end{figure}
 
-\begin{figure}[!ht]
+\paragraph{Metabolomic Analysis}
+
+Next, we apply shared subspace group covariance estimation to
+a metabolomic data analysis on fly aging \citep{Hoffman2014}.  We bin the data to include
+groups of flies less than 10 days (young), between 10 and 51 days (middle) and greater
+than 51 days (old), further split by gender.   We analyze metabolomic
+data corresponding to metabolites with 3714 mass-charge ratios.  After removing mean
+effects due to age and gender we fit the shared subspace estimator and
+compare covariances between groups.  Interestingly, we identify a
+subspace that explains almost all of the variability in the first few
+components of $\Sigma_k$ for all groups (Figure
+\ref{fig:dmelanRatio}). However, we see little evidence for
+differences in metabolite covariances on this subspace (Figure
+\ref{dmelanPosterior}). This is indicative of large amount of
+variation which is common across all groups. In this example, each
+group contains flies from multiple genotypes.  As such, much of the
+variability is probably due differences in genotype rather than age or
+gender.
+
+\begin{figure}[t]
     \centering
-    \includegraphics[width=0.5\textwidth]{Figs/leukemiaPosterior}
-  \caption{}
-\label{fig:leukemiaRatio}
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{Figs/dmelanRatio-04-28}
+        \caption{Ratio}
+        \label{fig:dmelanRatio}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.45\textwidth}r
+        \includegraphics[width=\textwidth]{Figs/dmelanPosterior-04-28}
+        \caption{Posterior}
+        \label{fig:dmelanPosterior}
+    \end{subfigure}
+    \caption{a) Proportion of variance on $\hat{V}$ is close to one
+      for all groups suggesting that blah. However, there is
+      significant overlap in the posterior distribution of $\Lambda_k$
+      and $U_k$. This suggests that there is little difference in the
+      covariance matrices between groups, on this particular subspace.  }
 \end{figure}
 
-\bibliographystyle{natbib}
+\section{Discussion}
+
+In this paper, we present a method for comparing differences between
+covariance matrices on a low dimensional subspace.  We provide an EM
+algorithm for estimating the subspace which explains the largest
+proportion of variance .
+
+Variation that differs across groups versus variation which is the
+same across groups.
+
+Importantly, our approach can be combined with existing hierarchical
+models for multi-group covariance estimation.  For instance, we can
+incorporate additional shrinkage as in \citet{Hoff2009} by utilizing
+non-uniform Bingham prior distributions for the subspace eigenvectors.
+Alternatively, we can summarize the estimated group covariance
+matrices by thresholding entries of the precision matrix to visualize
+differences between groups using a graphical model \citep{}.
+
+Full Bayesian approach using Riemmanian manifold sampler?
+
+\bibliographystyle{plainnat}
 \bibliography{refs.bib}
 
+\section{Appendix}
+
 \end{document}

### 2016-05-19 paper/mscov.tex
Updated paper

@@
 
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}[b]{0.25\textwidth}
+        \includegraphics[width=\textwidth]{Figs/3dplot}
+        \caption{Ratio}
+        \label{fig:dmelanRatio}
+    \end{subfigure}
+    \qquad     \qquad     \qquad 
+ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.25\textwidth}r
+        \includegraphics[width=\textwidth]{Figs/2d-scatter}
+        \caption{Posterior}
+        \label{fig:dmelanPosterior}
+    \end{subfigure}
+    \caption{ }
+\end{figure}
+
 \section{Inference}
+
+In this section we outline Bayesian inference algorithms including
+both an optimization algorithm and an MCMC algorithm for estimating
+covariance parameters and their associated uncertainty.  MCMC
+inference is particularly appealing because of the ease with which we
+can quantify the uncertainty associated with each estimate.
+Uncertainty quantification is particularly important in the $n << p$
+setting where seemingly large differences in the point estimates of
+the low rank covariance matrices across groups may not actually
+reflect significant differences.  Although, uncertainty quantification
+for all parameters is desirable, for the size of the problems we are
+generally interested in, we found MCMC inference for the subspace
+$VV^T$ to be intractable.  Our optimziation based approached is
+signifcantly faster and works well for large large dimensional
+datasets.  As such, we start, in Section \ref{sec:em}, by outlining an
+expectation-maximization algorithm for estimating the shared subspace
+and derving point estimates of the projected covariance $\psi_k$. In
+Section \ref{sec:bayes} we describe a Gibbs sampler for posterior
+inference of $\psi_k$ given the common subspace.  These two algorithms
+can be used in conjuction to first identify and evaluate a shared
+subspace and then to quantify the uncertainty associated with the
+projected covariance matrices on this space.
+
 \subsection{An Expectation-Maximization Algorithm}
@@
 
-We start by outlining a computationally efficient algorithm for
-estimating the shared subspace, $VV^T$, noise $\sigma_k^2$ and
-projected covariance matrix $\psi_k$.
-
-First, note that the full likelihood can be written as
+In this section we describe a computationally efficient algorithm for
+estimating the shared subspace, $VV^T$, projected covariance matrix
+$\psi_k$ and noise parameter $\sigma_k^2$.  First, note that the full
+likelihood for the shared subspace model can be written as
 
 \begin{align}
-p(S_1, ... S_k | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)  \\
-&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-(\sigma_k^2(V\psi_kV^T +
+p(S_1, ... S_k | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}\etr(-\Sigma_k^{-1}\mathbf{S}_k/2)  \\
+&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-(\sigma_k^2(V\psi_kV^T +
   I))^{-1}\mathbf{S}_k/2) \\
-&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-\left[V(\psi_k +
+&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-\left[V(\psi_k +
   I)^{-1}/\sigma_k^2 V^T + (I-VV^T)/\sigma^2_k\right]\mathbf{S}_k/2)
   \\
-&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-r)/2}|M_k|^{-n_k/2}etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]\mathbf{S}_k/2) 
+&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-s)/2}|M_k|^{-n_k/2}\etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]\mathbf{S}_k/2) 
 \end{align}
@@
 % \begin{align}
-% p(V|...) &\propto etr(\sum_{k=1}^K[O_k\Omega_kO^T_k]V^TS_k/(2\sigma^2_k)]V)\\
+% p(V|...) &\propto \etr(\sum_{k=1}^K[O_k\Omega_kO^T_k]V^TS_k/(2\sigma^2_k)]V)\\
 % \end{align}
 
-We propose a computationally efficient expectation-maximizatino (EM)
+We propose an expectation-maximization (EM)
 algorithm, to estimate the relevant quantities.  In particular we
-maximize the marginal posterior distribution of $V$, treating the
+maximize the marginal posterior distribution of $V$, treating 
 $M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as the ``missing'' parameters to
@@
 $p(\frac{1}{\sigma_k^2}) \propto \sigma_k^2$ and
-$p(M_k^{-1}) \propto |M_k^{-1}|^{-(r+1)/2}$.  From the likelihood it
+$p(M_k^{-1}) \propto |M_k^{-1}|^{-(s+1)/2}$.  From the likelihood it
 can easily be shown that the conditional posterior for $M_k^{-1}$ is
 
-$$p(M_k^{-1} | V) \propto |M_k^{-1}|^{(n_k - r -1)/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) $$
+$$p(M_k^{-1} | V) \propto |M_k^{-1}|^{(n_k - s -1)/2}\etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) $$
 
@@
 
-$$p\left(\frac{1}{\sigma^2} | V\right) \propto \left(\frac{1}{\sigma_k^2}\right)^{n_k(p-r)/2-1}etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)\mathbf{S}_k\right)  $$
+$$p\left(\frac{1}{\sigma^2} | V\right) \propto \left(\frac{1}{\sigma_k^2}\right)^{n_k(p-s)/2-1}\etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)\mathbf{S}_k\right)  $$
 
@@
 
-% \subsection{Alternative formulation}
-% $M_k = \sigma_k^2(\Phi_k + I)$ where 
-% $$ \Phi_k =\left( \begin{array}{cc}
-% \Psi_k & 0  \\
-% 0 & D  \end{array} \right)$$
-
-% Letting $V = [V_1\, V_2]$
-
-% \begin{eqnarray*}
-% p(M_k^{-1} | V) &\propto \prod_k |M_k^{-1}|^{n_k/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) \\
-%                 & = (|\psi_k + I|^{-1}|D +
-%                   I|^{-1})^{n_k/2}etr(-1/2((\psi_k+I)^{-1}V_1^T\mathbf{S}_kV_1
-%                   -1/2(D+I)^{-1}V_2^T\mathbf{S}_kV_2)) \\
-% \end{eqnarray*}
-
-% $$p(\frac{1/d_i} | V) \propto \prod_k (d_i + 1)^{-n_k/2}exp\left(\frac{1}{-1/2(d_i+1)}(V_2^T\mathbf{S}_kV_2)_{i,i}\right) $$
-
-\subsection{Bayesian Inference}
+\paragraph{Evaluating a Shared Subspace}
+
+The above EM algorithm provides an estimate of a shared subspace, but
+in practice there are no guarantees that the inferred subspace is
+equally appropriate for all groups.  Moreover, the shared subspace can
+be chosen in other ways, including more ad hoc methods
+involving the singular value dcomposition or simply by incorporating
+subject matter considerations.  Regardless of how a particular
+subspace is identified, it is usually important to quantify how useful
+this subspace is for explaining the variability in each group.  In
+particular, if $\hat{V}$ is a basis for a subspace shared across
+groups, then for each $k$, most of the variation in $\Sigma_k$ should
+lie on the subspace spanned by $\hat{V}$.  To quantify the extent to
+which this is true for different groups, we propose a simple estimator
+for the proportion of ``signal'' variance that lies on any given
+subspace.  Specifically, we apply the following estimator for ratio of
+the sum of the first $s$ eigenvalues of
+$\hat{V}^T\bl \Sigma_k \hat{V}$ to the sum of the first $s$
+eigenvalues of $\Sigma_k$:
+
+\begin{equation}
+ \frac{||Y_k^T\hat{V}||_F}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{sup}
+  ||Y_k^T\widetilde{V}||_F - \hat{\sigma}_k^2ps}
+\label{eqn:ratio}
+\end{equation}
+
+where $||.||_F$ is the Frobenius norm. If the inferred subspace is the
+true subspace, e.g. $\hat{V}\hat{V}^T = VV^T$, then this estimator
+will asymptotically approach one.  To see this, note that
+$\underset{V \in \mathcal{V}_{p, s}}{sup}\frac{||Y_k^TV||_F}{n_k}$ is
+equivalent to the sum of the first $s$ eigenvalues of the sample
+covariance matrix $\frac{\bl S_k}{n_k}$.  With $\hat{\lambda}^{(k)}_i$
+the $i$-th eigenvalue of $\frac{\bl S_k}{n_k}$, \citet{Baik2006}
+demonstrated that asymptotically as $p, n_k \rightarrow \infty$ with
+$\frac{p}{n_k} = \gamma_k$ fixed
+
+\begin{eqnarray}
+\hat{\lambda}^{(k)}_i &\rightarrow& \lambda^{(k)}_i\left(1 +
+                                    \frac{\sigma_k^2\gamma}{\lambda^{(k)}_i
+                                    - 1}\right)\\
+& \approx& \lambda^{(k)} + \sigma^2_k\frac{p}{n_k}
+\end{eqnarray}
+
+Thus,
+$\sum_i^s \hat{\lambda}^{(k)}_i \approx \sum_i^s \lambda^{(k)}_i +
+\sigma_k^2ps$.
+If we assume that $s$ is fixed in this asymptotic regime, then
+$||Y_k^TV||_F/n_k$ is a consistent estimator for
+$\sum_i^s \lambda^{(k)}_i$, and thus the above estimator will be close
+to one when $\hat{V}\hat{V}^T = VV^T$ and smaller if not.  The metric provides a
+useful indicator for which groups can be reasonably compared on a
+given subspace and which groups cannot.  In Section \ref{sec:simulation} we plot
+these estimates for different simulated datasets.  In Section \ref{sec:bio}
+we plot the estimates for inferred subspaces for real biological data.
+
+\subsection{Uncertainty Quantification}
 \label{sec:bayes}
 
-Although the EM algorithm presented in the previous section yields
-point estimates of $\psi_k$, it does not lead to natural uncertainty
-bounds for these estimates.  To quantify the uncertainty, we consider Bayesian
-inference for $\psi_k$, given the
-subspace $VV^T$.  Note that when the subspace is known, the posterior
+Although the EM algorithm presented in the previous section yields a
+fast way to derive point estimates for $VV^T$ and $\psi_k$, it does
+not lead to natural uncertainty bounds for these estimates. Full
+uncertainty quantification fo $VV^T$ is desirable but we found this
+challenging for the dimensions of the problems we would like to
+consider.  As such, in this section, we assume that the subspace
+$VV^T$ is fixed and known and focus on uncertainty quantificatoin for
+$\psi_k$ only.  Further, by focusing on group differences on a fixed
+subspace, it is simpler to interpret differences across covariance
+matrices.  The shared subspace can be an estimated using the EM algorithm
+presented in Section \ref{sec:em} or chosen based on subejct matter
+considerations.  Note that when the subspace is known, the posterior
 distribution of $\Sigma_k$ is conditionally independent from the other
-groups, i.e.  $P(\Sigma_{k} | V, \bl S_k, \Sigma_{-k}) = P(\Sigma_{k} | V, \bl S_k)$.
-Thus, we can independently estimate the conditional posterior
-distribution for each group.  In this paper, we focus on models based
-on the eigendecomposition of $\psi_k$.  In this setting, each covariance
-$\Sigma_k$ is a spiked covariance matrix where Equation \ref{eqn:sspsi}
-can be written as
+groups, i.e.
+$P(\Sigma_{k} | V, \bl S_k, \Sigma_{-k}) = P(\Sigma_{k} | V, \bl
+S_k)$,
+so that we can independently estimate the conditional posterior
+distribution for each group.
+
+There are many different prior specifications for $\psi_k$ and
+$\sigma_k^2$ that we could consider for estimating posterior
+uncertainty.  One simple approach would be to use the same prior
+specification used in Section \ref{sec:em}.  That is, we can use
+Jeffreys' prior for $\frac{1}{\sigma_k^2}$ and
+$M_k^{-1} = (\psi_k + I) ^{-1}/\sigma_k^2$.  In fact, it is easy to
+show that the Gamma distribution for $\frac{1}{\sigma_k^2}$ and the
+Wishart distribution for $M_k^{-1}$ are conjugate priors for the
+respective quantities.  From posterior samples of $M_k$ and
+$\sigma_k^2$ we can easily recover samples of $\psi_k$.
+Unfortunately, if $M_k^{-1}$ has a Wishart distribution, the implied
+support of $\psi_k$ is not constrained to the cone of positive
+semi-definite matrices.
+
+Direct prior specifications for $\psi_k$ are more appealing, but many
+obvious choices, like the inverse-Wishart prior distribution for
+$\psi_k$ do not lead to simple posterior distributions.  As such,
+building on recent interest in spiked covariance models
+\citep{Hoff2009, Donoho2013, Paul2007} we focus on tractable models
+based on the eigendecomposition of $\psi_k$.  Such models are
+especially appealing because it is simple to summarize and compare
+$\psi_k$ across groups in terms their eigenstructure (See Section
+\ref{sec:simulation} and \ref{sec:bio}).
+
+By modeling the eigenstructure, we can now view each covariance
+$\Sigma_k$ as a spiked covariance matrix.  Equation \ref{eqn:sspsi}, written as a function $V$, becomes
 
@@
 
-Here, $\Lambda_k$ is an $r \times r$ diagonal matrix of eigenvalues,
-and $O_k \in \mathcal{V}_{s,r}$ is the matrix of eigenvectors of
-$\psi_k$.  For any individual group, this corresponds to the original
-spiked covariance model (Equation \ref{eqn:spiked}) with $U_k = VO_k$.
+Here, we make one further generalization to allow
+$\psi_k = O_k\Lambda_KO_k^T$ be a rank $r \leq s$ dimensional covariance
+matrix on the $s$-dimensional subspace.  Thus, $\Lambda_k$ is an $r \times
+r$ diagonal matrix of eigenvalues, and $O_k \in
+\mathcal{V}_{s,r}$ is the matrix of eigenvectors of
+$\psi_k$.  For any individual group, this corresponds to the original spiked
+covariance model (Equation \ref{eqn:spiked}) with $U_k = VO_k \in
+\mathcal{V}_{p, r}$.
 
@@
 \begin{equation}
-p(S | \Sigma_k,n_k) \propto |\Sigma_k|^{-n/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)
+p(S | \Sigma_k,n_k) \propto |\Sigma_k|^{-n/2}\etr(-\Sigma_k^{-1}\mathbf{S}_k/2)
 \end{equation}
@@
 
-Where the diagonal matrix $\Omega = \Lambda(I+\Lambda)^{-1}$, e.g. $\omega_i = \frac{\lambda_i}{\lambda_{i}+1}$.
-
- Further,
+Where the diagonal matrix $\Omega = \Lambda(I+\Lambda)^{-1}$, e.g. $\omega_i = \frac{\lambda_i}{\lambda_{i}+1}$.  Further,
 
@@
 
-where the second line is due to Sylvester's determinant theorem.
-
-Substituting for $\Sigma^{-1}_k$ and $|\Sigma_k|$ we have the likelihood of $V$, $O_k$, $\Lambda_k$ and $\sigma_k^2$
-
-We can write the likelihood of $V$, $O_k$, $\Lambda_k$ and
-$\sigma_k^2$ given $Y_k$ using Equation \ref{eqn:rplik} by replacing
-$U_k$ with $VO_k$: 
+where the second line is due to Sylvester's determinant theorem.  The
+likelihood of $V$, $O_k$, $\Lambda_k$ and $\sigma_k^2$ is available
+from Equation\ref{eqn:rplik} by substituting in the appropriate
+quantities in for $\Sigma^{-1}_k$ and $|\Sigma_k|$ and replacing $U_k$
+with $VO_k$:
+
 \begin{equation}
  L(\sigma_k^2,\bl V , O_k \Omega_k : \bl Y_k) \propto
-    (\sigma_k^2)^{-n_kp/2}etr(-\frac{1}{2\sigma_k^2}\mathbf{S}_k)\left(\prod_{i=1}^r(1-\omega_{ki})
+    (\sigma_k^2)^{-n_kp/2}\etr(-\frac{1}{2\sigma_k^2}\mathbf{S}_k)\left(\prod_{i=1}^r(1-\omega_{ki})
    \right) ^{n_k/2}
-   etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k)
+   \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k)
 \label{eqn:sslik}
@@
 We use conjugate and semiconjugate priors for the parameters $O_k$,
-$\sigma^2_k$ and $\Omega_k$ to facilitate inference.  In the absence
-of real prior information, invariance considerations suggest the use
-of priors that lead to equivariant estimators.  Below we describe our
-choices for the prior distributions of each parameter and the
-resultant posterior distributions.
+$\sigma^2_k$ and $\Omega_k$ to faciliate inference via a Gibbs
+sampling algorithm.  In the absence of real prior information,
+invariance considerations suggest the use of priors that lead to
+equivariant estimators.  Below we describe our choices for the prior
+distributions of each parameter and the resultant conditional posterior
+distributions.
 
@@
 \label{lik_vo}
- p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto etr(\Omega_kO^T_k\hat{V}^T[S_k/(2\sigma^2_k)]\hat{V}O_k)
+ p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto \etr(\Omega_kO^T_k\hat{V}^T[S_k/(2\sigma^2_k)]\hat{V}O_k)
 \end{align}
@@
 \noindent This is a Bingham$(\Omega, \hat{V}^T \bl S_k \hat{V}/(2\sigma^2)$
-distribution on $\mathcal V_{s, p}$ \citep{Khatri1977}. A
+distribution on $\mathcal V_{s, r}$ \citep{Khatri1977}. A
 Gibbs sampler to simulate from this distribution is given in
@@
   \left(\prod_{i=1}^r(1-\omega_{ki})^{n_k/2}  \right)
-  etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) \\
+  \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) \\
 &  \propto & (1-\omega_{ki})^{n/2} e^{c \omega_s  n/2},    
@@
 % \begin{align}
-% p(V|\bl O, \bl S, \bl \sigma^2, \bl \Omega) &\propto etr(\sum_{k=1}^K[S_k/(2\sigma^2_k)]V[O_k\Omega_kO^T_k]V^T)\\
+% p(V|\bl O, \bl S, \bl \sigma^2, \bl \Omega) &\propto \etr(\sum_{k=1}^K[S_k/(2\sigma^2_k)]V[O_k\Omega_kO^T_k]V^T)\\
 % \end{align}
@@
 % \begin{align}
-% P(V|A_k,B_k) &\propto etr(\sum_{k=1}^KA_kVB_kV^T)\\
-% &= etr(\sum_{k=1}^K\sum_{i,j}^sA_kv_iv_j^Tb_{ijk})\\
-% &= etr(\sum_{i,j}^s \sum_{k=1}^K v_j^T[b_{ijk}A_k]v_i)
+% P(V|A_k,B_k) &\propto \etr(\sum_{k=1}^KA_kVB_kV^T)\\
+% &= \etr(\sum_{k=1}^K\sum_{i,j}^sA_kv_iv_j^Tb_{ijk})\\
+% &= \etr(\sum_{i,j}^s \sum_{k=1}^K v_j^T[b_{ijk}A_k]v_i)
 % \end{align}
@@
 % \begin{align}
-% P(v_i|A_k,B_kV_{-i})  &\propto etr(2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])v_i+v_i^T[\sum_{k=1}^K b_{ijk}A_k]v_i))\\
-% &= etr(Cv_i+v_i^TMv_i)
+% P(v_i|A_k,B_kV_{-i})  &\propto \etr(2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])v_i+v_i^T[\sum_{k=1}^K b_{ijk}A_k]v_i))\\
+% &= \etr(Cv_i+v_i^TMv_i)
 % \label{eqn:vbmf}
@@
 \section{Simulation Study}
+\label{sec:simulation}
+
+\subsection{Model Works}
+First, we do a single simulation that shows that we can recover the
+true parameters using our inferential algorithm when the assumed model
+is correct.
+
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{Figs/posteriorRegions}
+        \caption{Posterior}
+        \label{fig:simPosterior}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{Figs/simRatio-05-19}
+        \caption{Ratio}
+        \label{fig:simRatio}
+    \end{subfigure}
+    \caption{a) 95\% posterior regions for the ratio of the eigenvalues,
+      $\frac{\lambda_1}{\lambda_2}$, of $\psi_k$ and the orientation
+      of the principle axis on space spanned by $\hat{V}$ cover the
+      truth.  True parameter values on $\hat{V}^T\Sigma_k\hat{V}$ are
+      indicated by the crosses.  Since $V$ is only identifiable up to
+      rotation we find the Procrustes rotation that maximizes the
+      similarity of $\hat{V}$ to the true data generating basis. True
+      eigenvalue ratios were 10 (red and black), 3 (green and blue)
+      and 1 (light blue).  True orientations were $\pi/4$ (black), $-\pi/4$
+      (red and green) and 0 (blue and cyan).  b)  Estimated proportion
+    of variance explained by the subspace (Equation \ref{}).  As expected, for each
+    group, all of the variance can be explained on the inferred subspace.  }
+\end{figure}
+
+In order to quantify the uncertainty associated with the relevant
+parameters conditional on a shared subspace, we use the Bayesian
+inference algorithm outlined in Section \ref{sec:bayes}. In Figure
+\ref{fig:posteriorRegions} we depict the 95\% posterior regions for
+statistics based on the eigendecomposition of
+$\psi_k$.  Specifically, for the two dimensional
+shared subspace model, we plot the eigenvalue ratio
+$\frac{\lambda_1}{\lambda_2}$ and the angle of the first eigenvector
+on the $\text{arctan}(\frac{O_2}{O_1})$, where
+$\lambda_1 > \lambda_2$.  Non-overlapping 95\% posterior regions are
+indicative of significant differences in the covariance structure
+between groups.
+
+In this example, the posterior regions for all groups cover the true
+eigenvalue ratio and orienntation of the primary axis for the matrix
+$\hat{V}^T\Sigma_k\hat{V}$.  Note that for smaller eigenvalue ratios,
+there is larger uncertainty about the orientation of the primary axis.
+When the ratio is one (light blue) there is no information about
+orientation of the primary axis since the contours are spherical.
+
+\subsection{Model Misspecification}
 
@@
 inverse of the posterior mean of the precision matrix,
-$\hat \Sigma_{k} = \Exp{ \Sigma_k^{-1} | \bl S_k}^{-1}$ which we estimate
-using MCMC samples.  In Table \ref{table:groupLoss} we report the
-average group loss
+$\hat \Sigma_{k} = \Exp{ \Sigma_k^{-1} | \bl S_k}^{-1}$ which we
+estimate using MCMC samples.  In Table \ref{table:groupLoss} we report
+the average group loss
 $L(\Sigma_1, ..., \Sigma_K; \hat\Sigma_1, ..., \hat\Sigma_K ) =
 \frac{1}{K} \sum_k L_S( \Sigma_k , \hat\Sigma_k)$
-and 95\% intervals (TODO) for the estimates derived using each
-inferential model.
+and 95\% intervals for the estimates derived using each inferential
+model.
 
@@
 shared subspace estimator has small loss under model misspecification
-relative to the alternative.  For example, when the data is
-generated using the complete pooling model, the shared subspace
-estimator has almost four times smaller loss than the no pooling
-estimator.  When the data is generated under the no pooling model, the
-shared subspace estimator is over an order of magnitude better than
-the complete pooling estimator.
-
-Thus, shared subspace estimation can be an appropriate choice when the
-amount of similarity between groups is not known a priori.  The shared
-subspace model combined with the rank estimation procedure suggested
-by \citet{Gavish2014} leads to an estimator which adapts to the amount
-of similarity between groups.  Importantly, the shared subspace model
-can be viewed as a generalization of the ``no pooling'' spiked
-covariance model.  The two are equivalent when we let
-the shared subspace span the full space, e.g. $V \in \mathcal{O}(p)$.
-Figure \ref{fig:sdimension} depicts the loss for data generated under
-the no pooling model and fit with shared subspace estimators with
-different dimension.  As the shared subspace dimension increases to
-the dimension of the feature space, $s \rightarrow p$, the loss for
-the shared subspace estimator quickly coverges to the loss for the
-individual spiked covariance estimator.  Note that
-span($[U_1, ..., U_K]) \leq rK$, thus even when there is little
-similarity between the eigenvectors from each group, the shared
+relative to the alternative.  For example, when the data is generated
+using the complete pooling model, the shared subspace estimator has
+almost four times smaller loss than the no pooling estimator.  When
+the data is generated under the no pooling model, the shared subspace
+estimator is over an order of magnitude better than the complete
+pooling estimator.  Thus, shared subspace estimation can be an
+appropriate choice when the amount of similarity between groups is not
+known a priori.
+
+The shared subspace model combined with the rank
+estimation procedure suggested by \citet{Gavish2014} leads to an
+estimator which adapts to the amount of similarity between groups.
+Importantly, the shared subspace model can be viewed as a
+generalization of the ``no pooling'' spiked covariance model.  The two
+are equivalent when we let the shared subspace span the full space,
+e.g. $V \in \mathcal{O}(p)$.  Figure \ref{fig:sdimension} depicts the
+loss for data generated under the no pooling model and fit with shared
+subspace estimators with different dimension.  As the shared subspace
+dimension increases to the dimension of the feature space,
+$s \rightarrow p$, the loss for the shared subspace estimator quickly
+coverges to the loss for the individual spiked covariance estimator.
+Note that span($[U_1, ..., U_K]) \leq rK$, thus even when there is
+little similarity between the eigenvectors from each group, the shared
 subspace estimator with $s = rK$ will perform comparably to the ``no
@@
 
-In many real world examples, we are more interested in finding a
-useful space to interpret and compare differences between covariance
-matrices than we are in finding the lowest-loss estimator for the full
-covariance matrices.  In this context, the relevant quantity is often
-an estimate of the projected covariance, $V^T\Sigma V$.  In Figure
-\ref{fig:lossBoxplot} we depict boxplots for the distribution of
-losses,
+For any given subspace we can identify whether the space is
+inappropriate for some groups, either because the dimension is too
+small or the space itself does not represent every group equally.
+Specifically, we estimate which groups are well represented on the
+subspace and which groups are not using the proportion of variance
+estimator from Section \ref{sec:em} (Equation \label{eqn:ratio}).
+Figure \ref{fig:evalRatios} depicts the estimated proportion of
+variance on the subspace for each group in two different simulated
+datasets.  When the shared subspace model is appropriate, the subspace
+spanned by $\hat{V}$ explains most of the variance in $\Sigma_k$, for
+all groups.  When the estimated subspace is too small, the proportion
+of variance will be small for at least some of the groups.
+
+Ultimately, in many real world examples, we are more interested in
+finding a useful space to interpret and compare differences between
+covariance matrices than we are in finding the lowest-loss estimator
+for the full covariance matrices.  In this context, the relevant
+quantity is often an estimate of the projected covariance,
+$V^T\Sigma V$.  In Figure \ref{fig:lossBoxplot} we depict boxplots for
+the distribution of losses,
 $L_S(\hat{V}^T\Sigma_k\hat{V}, \hat{V}^T\hat{\Sigma}_k\hat{V})$, on
@@
 come from a 10-dimensional shared subspace model.  For both boxplots,
-estimates
-$\hat{V}^T\hat{\Sigma}_k\hat{V}$ are derived from assumed 2-dimensional
-model. This figure demonstrates that the loss for estimating the
-projected covariance matrix is not significantly larger when we
-underestimate the dimension of the true shared subspace.
+estimates $\hat{V}^T\hat{\Sigma}_k\hat{V}$ are derived from assumed
+2-dimensional model. This figure demonstrates that the loss for
+estimating the projected covariance matrix is not significantly larger
+when we underestimate the dimension of the true shared subspace.
 
@@
     \multirow{3}{*}{\rotatebox[origin=c]{90}{\textbf{Data Model}}} & 
-   Shared Subspace & 0.8 & 2.1 & 3.0 \\ \cline{2-5}
-   & Complete Pooling & 0.8 & 0.7 & 3.0 \\ \cline{2-5}
-   & No pooling & 8.3 & 136.3 & 3.0 \\ \cline{2-5}
+   Shared Subspace & 0.8 (0.7, 0.9) & 2.1 (1.7, 2.6) & 3.0 (2.9, 3.2) \\ \cline{2-5}
+   & Complete Pooling & 0.8 (0.7, 0.9) & 0.7 (0.6, 0.8) & 3.0 (2.9, 3.2)\\ \cline{2-5}
+   & No pooling & 7.1 (6.2, 8.0) & 138.2 (119, 153) & 3.0 (2.9, 3.2) \\ \cline{2-5}
   \end{tabular}
-  \caption[Table caption text]{Stein's loss for different inferential
-    and data generating models.  For each of the $K=10$ groups, we
-    simulate a $p=200$ dimensional spiked covariance matrix for rank
-    $r=2$.  For data from a shared subspace model the eigenvectors
-    of the covariance matrix share a subspace dimension $s=2$.  All results are
-    based on $n_k = 50$ observations per group.  }
+  \caption[Table caption text]{Stein's loss (and 95\% intervals) for
+    different inferential and data generating models.  For each of
+    $K=10$ groups, we simulate a $p=200$ dimensional spiked covariance
+    matrix for rank $r=2$.  For data from a shared subspace model the
+    eigenvectors of the covariance matrix share a subspace dimension
+    $s=2$.  All results are based on $n_k = 50$ observations per
+    group.  }
 \label{table:groupLoss}
@@
       converges to the loss from the independent spiked covariance
-      estimates. Right) Distribution of Stein's loss
+      estimates (dashed line). Right) Distribution of Stein's loss
       $L_S(\hat{V}^T\Sigma_k\hat{V}, \hat{V}^T\hat{\Sigma_k}\hat{V})$
@@
 \end{figure}
-
-\subsection{Model Checking and Uncertainty Quantification}
-
-If the shared subspace model is reasonable and we succeed in
-identifying an appropriate subspace, $\hat{V}$, then for each $k$,
-most of the variation in $\Sigma_k$ should lie on the subspace spanned
-by $\hat{V}$.  To quantify the extent to which this is true for
-different groups, we propose a simple estimator for the proportion of
-``signal'' variance that lies on any given subspace.  Specifically, we
-apply the following estimator for ratio of the sum of the first $s$ eigenvalues
-of $\hat{V}^T\bl \Sigma_k \hat{V}$ to the sum of the first $s$
-eigenvalues of $\Sigma_k$:
-
-\begin{equation}
- \frac{||Y_k^T\hat{V}||_F}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{sup}
-  ||Y_k^T\widetilde{V}||_F - \hat{\sigma}_k^2ps}
-\label{eqn:ratio}
-\end{equation}
-
-If the inferred subspace is the true subspace, e.g. $\hat{V}\hat{V}^T = VV^T$,
-then this estimator will asymptotically approach one.  To see this, note that
-$\underset{V \in \mathcal{V}_{p, s}}{sup}\frac{||Y_k^TV||_F}{n_k}$ is
-equivalent to the sum of the first $s$ eigenvalues of the sample
-covariance matrix $\frac{\bl S_k}{n_k}$.  With $\hat{\lambda}^{(k)}_i$
-the $i$-th eigenvalue of $\frac{\bl S_k}{n_k}$, \citet{Baik2006}
-demonstrated that asymptotically as $p, n_k \rightarrow \infty$ with
-$\frac{p}{n_k} = \gamma_k$ fixed
-
-\begin{eqnarray}
-\hat{\lambda}^{(k)}_i &\rightarrow& \lambda^{(k)}_i\left(1 +
-                                    \frac{\sigma_k^2\gamma}{\lambda^{(k)}_i
-                                    - 1}\right)\\
-& \approx& \lambda^{(k)} + \sigma^2_k\frac{p}{n_k}
-\end{eqnarray}
-
-Thus,
-$\sum_i^s \hat{\lambda}^{(k)}_i \approx \sum_i^s \lambda^{(k)}_i +
-\sigma_k^2ps$.
-If we assume that $s$ is fixed in this asymptotic regime, then
-$||Y_k^TV||_F/n_k$ is a consistent estimator for
-$\sum_i^s \lambda^{(k)}_i$, and thus the above estimator will be close
-to one when $\hat{V}\hat{V}^T = VV^T$ and smaller if not.  Figure
-\ref{fig:evalRatios} depicts this estimate for each group in two
-different simulated datasets.  When the shared subspace model is
-appropriate, the subspace spanned by $\hat{V}$ explains a large
-proportion of variance in $\Sigma_k$ in all groups.  When the
-estimated subspace is too small, the proportion of variance will be
-small for at least some of the groups.  The metric thus provides a
-useful indicator for which groups can be reasonably compared on a
-given subspace and which groups cannot.
 
@@
 
-\paragraph{Uncertainty Quantification}
-
-The EM algorithm presented in Section \ref{sec:em} provides a fast way
-to estimate the shared subspace $VV^T$, the rank $r$ covariance,
-$U_k\Lambda_kU_k^T$, and the homoscedastic noise $\sigma_k^2$.
-Although the EM algorithm is useful for deriving point estimates, it
-does not provide associated uncertainty bounds.  Uncertainty
-quantification is particularly important in the $n << p$ setting where
-seemingly large differences in the low rank covariance matrices across groups may not
-actually be statistically significant.
-
-\begin{figure}[!ht]
-  \centering
-    \includegraphics[width=0.5\textwidth]{Figs/posteriorRegions}
-    \caption{95\% posterior regions for the ratio of the eigenvalues,
-      $\frac{\lambda_1}{\lambda_2}$ and the angle on the shared
-      subspace determined spanned by $V$.  The horizontal dashed line
-      represents the true ratio of eigenvalues
-      $\frac{\lambda_1}{\lambda_2} = 10$.  Colored tick marks at the
-      bottom of the figure represent ``true'' orientation of the
-      principle axis for each group relative to axes determined by
-      $\hat{V}R$.  Since $V$ is only identifiable up to rtation, we
-      fix R using the Procrustes rotation that maximizes the
-      similarity of $\hat{V}R$ to the true data generating $V$. }
-\label{fig:posteriorRegions}
-\end{figure}
-
-In order to quantify the uncertainty associated with the relevant
-parameter estimates, we use the Bayesian inference algorithm outlined
-in Section \ref{sec:bayes}. In Figure \ref{fig:posteriorRegions} we
-depict the 95\% posterior regions for statistics based on the
-eigendecomposition of $\hat{V}^T\hat{\Sigma}\hat{V}$.  Specifically,
-for the two dimensional shared subspace model, we plot the eigenvalue
-ratio $\frac{\lambda_1}{\lambda_2}$ and the angle of the first
-eigenvector on the $\text{arctan}(\frac{O_1}{O_2})$, where
-$\lambda_1 > \lambda_2$.  Non-overlapping 95\% posterior regions are
-indicative of significant differences in the covariance structure
-between groups.
-
-In this example, the posterior regions for all groups cover the true
-eigenvalue ratio, $\frac{\lambda_1}{\lambda_2} = 10$, even though the
-posterior mean for this quantity varies by a factor of 2 across groups.  We also
-estimate how well we recover the true orientation of the covariance
-matrix.  Since $V$ is only identifiable up to rotations, we estimate
-the Procustes transformation, R, which maximizes the similarity
-between $\hat{V}R$ and $V$.  The posterior regions for each group also
-cover the correct orientation.
-
 \section{Analysis of Biological Data}
-
+\label{sec:bio}
 As in many natural systems, high dimensional biological data is often
@@
 compare covariance matrices in a metabolomic analysis of fly data.  In
-this example, after controlling for mean effects of fly age and
-gender, we find little evidence of differences between metabolite
-covariances across groups.
+this example, after controlling for mean effects due to fly age, gender and
+genotype, we find little evidence of differences in metabolite
+covariances between groups.
+
+In these analyses we employ a variation of the shared
+subspace model.  Specifically, we define the $s$ by $s$ matrix $\psi_k$ as
+
+\begin{equation}
+\psi_k =\left( \begin{array}{cc}
+O_k\Lambda_kO_k^T & 0  \\
+0 & D  \end{array} \right)
+\end{equation}
+
+For the following analyses $O_k\Lambda_kO_k^T$ is a rank $r=2$ matrix
+and $D$ is an $(s-r)$-dimensional diagonal matrix.  We write
+$V = [V_1\, V_2]$, with $V_1 \in \mathcal{V}_{p,r}$ as the basis for a
+r-dimensional shared subspace that explains the differences between
+groups and $V_2 \in \mathcal{V}_{p,(s-r)}$ corresponding to the remaining $(s-r)$
+eigenvectors common to all groups.  Importantly, we make no assumptions about the
+magntidue of the eigenvalues $\Lambda_k$ relative to the eigenvalues
+$D$, so that the largest eigenvectors of $\psi_k$ may be in $V_2$.
+
+We find this formulation useful becasue in real world analyses,
+differences between groups may not manifest themselves in the first
+principle components.  For instance, in genetic analyses of humans,
+there may be large person to person variability common to all groups
+which is unrelated to the grouping of those subjects based on the
+phenotype of interest.  Differences between groups may manifest themselves
+as changes in the smaller principle components. This formulation
+allows, complete pooling of most eigenvectors, yet a space to identify
+and compare relevant differences.  Moreover, visualizing and
+interpreting differences between groups on higher dimensional spaces
+is a challenging problem.  Using this framework we can easily understand the
+largest differences across groups in a two dimensional space while
+explaining the remaining variability with a anisoptropic shared
+structure.
 
@@
 Further intuition about the differences in covariances between groups
-can be understood in the biplot in Figure
-\ref{fig:leukemiaBiplot}.  Here, we plot the contours of the
-covariance matrices for three of subtypes of Leukemia.  We also plot
-points for the 1\% of genes with the largest loadings on the first two columns
-of $V$.  The genes in each cluster are
-listed in the corresponding table.  Even though we remove all
-mean-level differences in gene expressions between groups, we are
-still able to identify genes with known connections to cancer and
-leukemia.  As one example, MYC (top right) is an oncogene with well
-established association to many cancers \citep{Dang2012}.
+can be understood in the biplot in Figure \ref{fig:leukemiaBiplot}.
+Here, we plot the contours of the covariance matrices for three of
+subtypes of Leukemia.  We also plot different shapes for the 1\% of genes with
+the largest loadings on the first two columns of $V$ and the
+remaining values as light grey dots.  The genes with the largest
+loadings are clustered by quadrant and listed in the corresponding
+table.  Even though we remove all mean-level differences in gene
+expressions between groups, we are still able to identify genes with
+known connections to cancer and leukemia.  As one example, MYC (top
+right) is an oncogene with well established association to many
+cancers \citep{Dang2012}.
 
@@
 
-In this figure, genes that fall in the upper right quadrant
-($\blacksquare$) have positive loadings in both columns of $V$. On the
-other hand, genes in the upper left corner have positive loadings on
-the first column of $V$ but negative loadings on the second column of
-$V$.  When the principle axis has an orientation close to
-$\frac{\pi}{4}$, genes that fall in the upper right or lower left
-quadrant will exhibit larger variability, whereas genes in the upper
-left or lower right exhibit little variability since their loadings
-effectively cancel out.  On the other hand, when the principle axis is
-closer to $-\frac{\pi}{4}$, the genes in the upper left
-($\blacktriangle$) and lower right exhibit more variability.  In this case,
-the genes in the upper left portion of the plot show large correlated
-variability for the Hyperdip50 group but not for E2A-PBX1 group.
-Similarly, genes in the upper right tend to vary together more for
-E2A-PBX1 but not for the Hyperdip50 group.  Both of these sets of
-genes have large variability for the BCR-ABL group because the
-principle axis of variation aligns predominantly with $V2$.
+In this figure, genes that fall in the upper right quadrant have
+positive loadings in both columns of $V$. Genes in the upper left
+corner have positive loadings on the first column of $V$ but negative
+loadings on the second column of $V$.  The principle axis for a group
+aligns in the direction of genes that exhibit the largest variability
+in that group.  Genes which lie in a direction orthogonal to the
+principle axis exhibit reduced variability, relative to the other
+groups. For instance, genes in the upper left (triangles) exhibit
+large, positively correlated variability in the Hyperdip50 group.  In
+this same group, there is reduced varaibility, relative to the other
+groups, among the genes in the upper right of the plot (squares),
+since this cluster of genes lie in a direction orthogonal to the
+principle axis.  In contrast, the BCR-ABL group is aligned primarily
+with the second column of $V$, which means that the genes indicated by
+squares and circles vary significantly in this group.  Genes in the
+square group are anti-correlated with those in the circle group, since
+the their loadings have opposite signs on $V2$.
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{Figs/leukemia-biplot-05-06}
+    \includegraphics[width=0.5\textwidth]{Figs/leukemia-biplot-05-13}
     \qquad
-\raisebox{\height}{
+\raisebox{1.25\height}{
 \footnotesize
@@
 \caption{Left) Variant of a biplot with contours for three leukemia
-  subtypes and the loadings of the f Right) Gene's with }
+  subtypes and the loadings for each gene on the first two columns of
+  $\hat{V}$.  All gene's loadings are displayed in light gray, and the
+  top 1\% of genes with with the largest magnitude loadings are
+  displayed as either a triangle, square or circle depending on which
+  quadrant they lie in.  Right) List of the gene's with the largest
+  loadings, grouped by quadrant. }
 \label{fig:leukemiaBiplot}
@@
 
-Next, we apply shared subspace group covariance estimation to
-a metabolomic data analysis on fly aging \citep{Hoffman2014}.  We bin the data to include
-groups of flies less than 10 days (young), between 10 and 51 days (middle) and greater
-than 51 days (old), further split by gender.   We analyze metabolomic
-data corresponding to metabolites with 3714 mass-charge ratios.  After removing mean
-effects due to age and gender we fit the shared subspace estimator and
-compare covariances between groups.  Interestingly, we identify a
-subspace that explains almost all of the variability in the first few
-components of $\Sigma_k$ for all groups (Figure
-\ref{fig:dmelanRatio}). However, we see little evidence for
-differences in metabolite covariances on this subspace (Figure
-\ref{dmelanPosterior}). This is indicative of large amount of
-variation which is common across all groups. In this example, each
-group contains flies from multiple genotypes.  As such, much of the
-variability is probably due differences in genotype rather than age or
-gender.
+Next, we apply shared subspace group covariance estimation to a
+metabolomic data analysis on fly aging \citep{Hoffman2014}.  We bin
+the data to include groups of flies less than 10 days (young), between
+10 and 51 days (middle) and greater than 51 days (old), and further
+split by gender. The sample sizes range from 43 to 59 flies per group.
+We analyze metabolomic data corresponding to metabolites with 3714
+mass-charge ratios.  After removing mean effects due to age and gender
+we fit the shared subspace estimator and compare covariances between
+groups.  Interestingly, we identify a subspace that explains almost
+all of the variability in the first few components of $\Sigma_k$ for
+all groups (Figure \ref{fig:dmelanRatio}). However, we see little
+evidence for differences in metabolite covariances on this subspace,
+as evidenced by the large overlap in posterior distributions for each group
+(Figure \ref{fig:dmelanPosterior}). This is indicative of large
+amount of variation which is common across all groups for the first
+$s$ principle components. Differences in covariation across may
+manifest themselves for the smaller principle components, but are too
+small to detect given the sample sizes.
 

### 2016-05-28 paper/mscov.tex
First complete draft of paper

@@
 % between other pairs may be consistently negative. In such cases much
-% of the similarity across covari- ance matrices can be described by
+% of the similarity across covariance matrices can be described by
 % similarities in their principal axes, which are the axes that are
@@
 
-  We develop model-based methods for evaluating similarities and
+  We develop a model-based method for evaluating similarities and
   differences among several $p\times p$ covariance matrices in the
@@
   eigenvectors and eigenvalues.  Finally, we demonstrate the utility
-  of our method in analyses exploring covariability in biological data.
+  of our method in an analysis of biological data.
 
@@
 
-% \section{NOTES}
-
-% \begin{itemize}
-% \item For MSCov to work in no-pooeld case need $R < S$ (doens't work if $S=P$).
-% \item Explore the effect on estimation when R=2 two eigenvalues converge/diverge.
-% \item TODO: Experiments: 
-% \begin{itemize}
-% \item S=20, R=2, data generated under no-pooled model.
-% \item Fit with R=2 to R=20, Fix S=20
-% \item Fit with R=2, S=5 to 20
-% \item Fix $1/(1-Y^2)$ bug for Y=1
-% \item debeta_un = 0 for large sample sizes (underflows)
-% \item Check that CP models yield same results using fitBayesianSpike and fit-subspace.
-% \end{itemize}
-% \end{itemize}
-
 \section{Introduction}
 
-Introduction to covariance estimation and spiked covariance model.
+It has long been known that the sample covariance matrix is a poor
+estimator of the population covariance matrix, particularly when the
+sample size is comparable to or smaller than the number of features
+($n \leq p$) \citep{Dempster1969, Stein1975}.  A host of estimators
+have been devised to address this fact. These primarily include of
+shrinkage based approaches \citep{Schafer2005, Ledoit2011} and low
+rank factor models \citep{Mardia1980, Ullman2003, Fan2008}.  In many
+cases, the covariance matrices are highly structured and thus can be
+sparsely parameterized even if the matrix itself is not low rank
+\citep{Burg1982, Williams1993} .  Along these lines, there has been a
+recent interest in sparse Gaussian graphical model, in which the there
+are assumed to be many zeros in the inverse covariance matrix
+\citep{Meinshausen2006, Friedman2008}.
+
+Most of these approaches assume that the goal is to estimate a
+covariance matrix from a single dataset.  In many applications,
+however, the data can be split into groups, each of which represent
+samples from populations with distinct (but possibly related)
+covariance matrices.  In this case, a comparative analysis of
+covariance patterns across multiple groups of observations is often
+warranted.  Evaluating similarities and differences between covariance
+matrices can be an important complement to classical analyses which
+center primarily on estimating differences in means across groups.
+For instance, in case control-studies, mean level effects may be small
+relative to subject variability; distributional differences between
+groups may still be evident as differences in the covariances between
+features.  Even when mean level differences are detectable, better
+estimates of the covariability of features across groups may lead to
+an improved understanding of the mechanisms underlying these apparent
+mean level differences.  Further, accurate covariance estimation is an
+essential part of many prediction tasks (e.g. quadratic discriminant
+analysis).
+
+By exploiting similarity across related groups of high-dimensional
+data, we can increase the efficiency of existing covariance
+estimators.  Some of the earliest approaches for modeling multiple
+covariance matrices focused on estimation in terms of the spectral
+decompositions for these matrices.  \cite{Flury1987} developed
+estimation and testing procedures for the ``common principal
+components'' model, in which a set of covariance matrices were assumed
+to share the same eigenvectors.  \citet{Schott1991, Schott1999}
+considered cases in which only certain eigenvectors are shared across
+populations, and \citet{Boik2002} described an even more general model
+in which eigenvectors can be shared between all or some of the groups.
+More recently, \citet{Hoff2009}, noting that eigenvectors are unlikely
+to be shared exactly between groups, introduced a hierarchical model
+for eigenvector shrinkage based on the matrix Bingham distribution.
+There has also been a significant interest in estimating covariance
+matrices using Gaussian graphical models. For Gaussian graphical
+models, zeros in the precision matrix correspond to conditional
+independences between pairs of features given the remaining features
+\citep{Meinshausen2006}.  \citet{Witten2014} extended existing work in
+this area to the multi-group setting, by pooling information about the
+pattern of zeros in across precision matrices.
+
+% When the
+% data can be grouped according to a set of continuous covariates
+% continuously parameterized covariance models can be appropriate
+% \citep{Chiu1996, Yin2010, Hoff2011}.
+
+Another popular method for modeling relationships between
+high-dimensional multivariate data related that is closely related to
+low rank covariance estimation, is known partial least squares
+regression (PLS). This approach, which is a special case of a bilinear
+factor model, involves projecting the data into a lower dimension
+space which maximizes the similarity of the two groups; this technique
+does not require the data from each group to share the same feature
+set.  An oft used variant for prediction, partial least squares discriminant analysis
+(PLS-DA) is especially common in the chemometrics and bioinformatics
+ \citep{Barker2003} .  Although closely related to
+the approaches we will consider here, the primarily goal of of PLS based models
+is to create regression or discrimination models, not to explicitly
+infer covariance matrices from multiple groups of data.  Nevertheless,
+the basic idea that data can often be well represented on a low
+dimensional space is an appealing one that we leverage as well.
+
+In many natural systems, high dimensional data is often very
+structured and thus can be well understood on a relatively low
+dimensional subspace. For example, with gene expression data, the
+effective dimensionality is thought to scale with the number of gene
+regulatory modules, not the number of genes themselves
+\citep{Heimberg2016}.  As such, differences in gene expression across
+groups should be expressed in terms of differences between these
+regulatory modules rather than strict differences between expression
+levels.  Such differences can be examined on a subspace that reflects
+the correlations resulting from these modules.
+
+Motivated by this notion, in this paper we propose a multi-group
+covariance estimation model by sharing information about the subspace
+spanned by group-level eigenvectors.  The shared subspace assumption
+can be used to improve estimates and facilitate the interpretation of
+differences between covariance matrices across groups.  For each
+group, we assume ``the spiked covariance model'', a well studied
+variant of the factor model \citep{Johnstone2001}.  In Section
+\ref{sec:shared} we briefly review the behavior of spiked covariance
+models for estimating a single covariance matrix before introducing
+our extension to the multi-group setting.  In Section
+\ref{sec:inference} we describe an optimization algorithm for
+inferring the shared subspace and then outline a MCMC algorithm for
+estimating the posterior distribution of the projected covariance
+matrices.  Finally, in Section \ref{sec:simulation} we investigate the
+behavior of these models in simulation before demonstrating the
+utility of this model with real biological data in Section
+\ref{sec:app}.
+
+\section{The Shared Subspace Spiked Covariance Model}
+\label{sec:shared}
+
+\paragraph{The Spiked Covariance Model:}
 
 Suppose $\bl S$ has a possibly degenerate Wishart$(\Sigma,n$)
-distribution with density 
-given by 
+distribution with density given by
 \begin{equation} 
@@
 \end{equation}
-where $\Sigma \in \mathcal S_p^+$ and $n$ may be less than $p$.
-
-% However, when the population size is large and comparable with the
-% sample size, as is in many contemporary data, it is known that the
-% sample covariance matrix is no longer a good approximation to the
-% covari- ance matrix. Baik
-
-Such a likelihood results from $\bl S$ being, for example, a residual 
-sum of squares matrix from a multivariate regression analysis. In  this case,
-$n$ is the number of independent observations minus the 
-rank of the design matrix. 
-The spiked principle 
-components model (spiked PCA) introduced by \citet{Johnstone2001} assumes that 
+%
+\noindent where $\Sigma \in \mathcal S_p^+$ and $n$ may be less than $p$.  Such
+a likelihood results from $\bl S$ being, for example, a residual sum
+of squares matrix from a multivariate regression analysis. In this
+case, $n$ is the number of independent observations minus the rank of
+the design matrix.  The spiked principle components model (spiked PCA)
+introduced by \citet{Johnstone2001} assumes that
 \begin{equation} 
@@
 \end{equation}
-
-where $\Lambda$ is an $r\times r$ diagonal matrix and 
-$\bl U \in \mathcal V_{r,p}$, with 
-%$r< ( n \wedge p)$. 
-$r<p$. Thus, the spiked covariance model is a low rank factor model
-with homoscedastic error \citep{}.  Such a formulation is appealing because it explicitly
-partitions the covariance into a tractable low rank ``signal'' and homscedastic ``noise''.
-
-For such a covariance matrix, we have
-$\Sigma^{-1} = (\bl U \Lambda \bl U^T + \bl I )^{-1} = \bl I - \bl U
-\Omega \bl U^T $
-where $\Omega = \text{diag}(\omega_1,\ldots, \omega_r)$ with
-$\omega_s = \lambda_j/(\lambda_s+1)$.  The likelihood (\ref{eqn:lik})
-in terms of $(\sigma^2,\bl U, \Omega)$ is then
-\begin{equation}
-L(\sigma^2,\bl U , \Omega : \bl Y)  = 
-(\sigma^2)^{-np/2} \etr(-\bl S/[2\sigma^2]) \times 
-   \etr( \Omega \bl U^T [\bl S/(2\sigma^2)] \bl U ) \times  
- \left\{ \prod_{s=1}^r (1-\omega_s)^{n/2} \right \} .  
-\label{eqn:rplik}
-\end{equation}
-
-Classical results for parametric models 
-(e.g., \citet{schwartz_1965})  
-imply that 
-asymptotically in $n$ for fixed $p$, a Bayes estimator
-will be consistent for a spiked  population covariance 
-as long as the assumed number of spikes is greater than or equal to 
-the true number. 
-However, the situation is more difficult 
-when both $n$ and $p$ are large.  
-%Since a Bayes estimator for an invariant loss function 
-%is orthogonally equivariant under this prior, 
-%it can be expressed as 
-%$\hat \Sigma(\bl S) = \bl E \bl D(\bl L) \bl E^T$, 
-%where $\bl E$ and $\bl L$ are the matrices of eigenvectors and 
-%eigenvalues of $\bl S$, and $\bl D$ denotes a diagonal matrix 
-%that is a function of $\bl L$. 
-%As a result 
-Under the spiked covariance model, 
-results of 
-\citet{Baik2006, Paul2007} indicate that if 
-$p/n \rightarrow \gamma >0$ as $n\rightarrow \infty$,   
-the top $r$ eigenvalues of $\bl S/(n\sigma^2)$ will converge 
-to upwardly  biased versions of $(\lambda_1+1,\ldots, \lambda_r+1)$, 
-if each $\lambda_s $ is greater than $\sqrt{\gamma}$.   
-This has led several authors to suggest estimating 
-$\Sigma$ via shrinkage of the eigenvalues of the sample covariance 
-matrix. In particular, 
-in the setting that $\sigma^2$ is known, 
-\citet{Donoho2013} propose  
-setting all sample eigenvalues that fall below 
-the value $(1+\sqrt{\gamma})^2$ to be one, and shrinking the larger 
-eigenvalues in a way that depends on the particular 
-loss function being used. 
-These shrinkage functions are shown to be asymptotically 
-optimal in the $p/n\rightarrow \gamma$ setting.
-
-As demonstrated by Gavish et al \citep{Gavish2014}, even when the
-true rank is large, if blah it is still asymptotically preferable to fit a lower rank
-blah .  it is always better than hard thresholding at any other value, and is always better than ideal truncated singular value decomposition (TSVD)
-
-Bayes estimators under the prior distribution described above 
-can also be viewed as eigenvalue shrinkage estimators: 
-Since a Bayes estimator for an invariant loss function 
-is orthogonally equivariant under this prior, 
-it can be expressed as 
-$\hat \Sigma(\bl S) = \bl E  \hat \Sigma (\bl L) \bl E^T$, 
-where $\bl E$ and $\bl L$ are the matrices of eigenvectors and 
-eigenvalues of $\bl S$. Furthermore, it is straightforward to 
-show that $\hat \Sigma (\bl L)$ must be a diagonal matrix, 
-and so the Bayes estimator is obtained by applying the 
-shrinkage function $\hat\Sigma$ to the sample eigenvalues $\bl L$. 
-Understanding of this shrinkage function could in principle 
-be pursued by investigation of the posterior distribution of the 
-eigenvalue parameters $\Lambda$, or equivalently their  
-transformed values $\Omega$.
-
-However, in many modern applications, it is not enough to
-independently estimate an individual covariance matrix.  Rather, a
-comparative analysis of covariance patterns across multiple groups of
-observations would be most informative.  For example, in case
-control-studies, mean level effects may be small relative to subject
-variability.  Differences in the covariance
-matrices may lead to better insights about the differences between groups.
-Alternatively, when mean level differences are detectable, better
-estimates of the covariability of features across groups may lead to improved
-understanding of the mechanisms underlying these apparent mean level
-differences.
-
-With this situation in mind, Flury (1984) developed estimation and
-testing pro-cedures for the ‘common principal components’ model, in
-which a set of covariance matrices
-\citep{Flury1987} and \citep{Schott1991, Schott1999} considered
-cases in which only certain columns or subspaces of U are shared across populations, and \citep{Boik2002} described a very general model in which eigenspaces can be shared between all or some of the populations.
-
-Continuously parameterized covariance matrices. \citep{Hoff2011}, \citep{Chiu1996}, \citep{Yin2010}
-
-In such models, it is no longer reasonable to assume orthogonal equivariance .
-To address these issues, \citet{Hoff2009} introduced a model for
-eigenvector shrinkage based on the matrix Bingham distribution.
-
-Partial least squares discriminant analysis refs?
-\citep{Westerhuis2008, Szymanska2012, Eriksson2001, Barker2003, Saebo2008}
-
-Recently there has been significant interest gaussian graphical
-models, .   In gaussian graphical models, zeros in the
-precision matrix correspond to conditional independences between pairs
-of features given the remaining features \citep{}.  \citet{Witten2014} developed a
-method for pooled estimation inverse covariance matrices across
-multiple groups.  Identifying a shared pattern of zero's in the inverse covariance
-matrices for multiple classes of observations.
-
-In the following section, we
-propose a model for estimating and exploring differences between
-covariance matrices from multiple groups of related data.
-
-Below we describe our approach for comparaing covariance matrices on
-a common low dimensional subspace.
-
-\section{The Shared Subspace Spiked Covariance Model}
-\label{sec:shared}
-
-Assume we have $k=1, ..., K$ groups of matrices $\bl S_k$ each of
-which follow a (degenerate) Wishart distribution, as in Equation
-\ref{eqn:lik}.  Each group consists of of $n_k$ measurements of $p$
-features, typically with $n_k << p$.  We pool information about the
-covariance matrices by assuming that most of the relevant variation
-for each group shares a low rank subspace spanned by
-the columns of a matrix $V$.  That is,
+%
+\noindent where $\Lambda$ is an $r\times r$ diagonal matrix and
+$\bl U \in \mathcal V_{r,p}$, with $r \ll p$. Here $\mathcal V_{r,p}$,
+known as the Stiefel manifold, comprises the set of $p \times r$
+orthonormal matrices in $\mathcal{R}^p$.  The spiked covariance model
+is thus a low rank factor model with homoscedastic error.  Such a
+formulation is appealing because it explicitly partitions the
+covariance matrix into a tractable low rank ``signal'' and isotropic
+``noise''.
+
+%  For such a covariance matrix, we have
+% $\Sigma^{-1} = (\bl U \Lambda \bl U^T + \bl I )^{-1} = \bl I - \bl U
+% \Omega \bl U^T $
+% where $\Omega = \text{diag}(\omega_1,\ldots, \omega_r)$ with
+% $\omega_s = \lambda_j/(\lambda_s+1)$.  The likelihood (\ref{eqn:lik})
+% in terms of $(\sigma^2,\bl U, \Omega)$ is then
+% \begin{equation}
+% L(\sigma^2,\bl U , \Omega : \bl Y)  = 
+% (\sigma^2)^{-np/2} \etr(-\bl S/[2\sigma^2]) \times 
+%    \etr( \Omega \bl U^T [\bl S/(2\sigma^2)] \bl U ) \times  
+%  \left\{ \prod_{s=1}^r (1-\omega_s)^{n/2} \right \} .  
+% \label{eqn:rplik}
+% \end{equation}
+
+Classical results for parametric models (e.g., \citet{Schwartz1965})
+imply that asymptotically in $n$ for fixed $p$, a Bayes estimator will
+be consistent for a spiked population covariance as long as the
+assumed number of spikes (non-unit eigenvalues) is greater than or
+equal to the true number.  However, when $p$ is large relative to $n$,
+as is the case for the examples considered here, things are more
+difficult.  Under the spiked covariance model, it has been shown that
+$p/n \rightarrow \gamma >0$ as $n\rightarrow \infty$, the top $r$
+eigenvalues of $\bl S/(n\sigma^2)$ will converge to upwardly biased
+versions of $(\lambda_1+1,\ldots, \lambda_r+1)$, if each $\lambda_s $
+is greater than $\sqrt{\gamma}$ \citep{Baik2006, Paul2007} .  This has
+led several authors to suggest estimating $\Sigma$ via shrinkage of
+the eigenvalues of the sample covariance matrix. In particular, in the
+setting that $\sigma^2$ is known, \citet{Donoho2013} propose setting
+all sample eigenvalues that fall below the value
+$\sigma^2(1+\sqrt{\gamma})^2$ to be $\sigma^2$, and shrinking the
+larger eigenvalues in a way that depends on the particular loss
+function being used.  These shrinkage functions are shown to be
+asymptotically optimal in the $p/n\rightarrow \gamma$ setting.
+
+% As demonstrated by Gavish et al \citep{Gavish2014}, even when the
+% true rank is large, if blah it is still asymptotically preferable to fit a lower rank
+% blah .  it is always better than hard thresholding at any other value, and is always better than ideal truncated singular value decomposition (TSVD)
+
+Many eigenvalue shrinkage estimators can be viewed in the Bayesian
+context as derivative of a Bayes estimator under a particular prior.
+Note that in the absence of real prior information, invariance
+considerations suggest the use of priors on $\sigma^2$ and $\bl U$
+that lead to equivariant estimators.  For instance, an (improper)
+prior having density $p(\sigma^2,\bl U) = 1/\sigma^2 $, with respect
+to the product of Lebesgue measure on $\mathbb R^+$ and the uniform
+probability measure on $\mathcal V_{r,p}$, leads to Bayes estimators
+$\hat \Sigma(\bl S)$ that are equivariant with respect to rotations
+and scale changes.  For multi-group covariance estimation, however, we
+often have reason to believe that a priori the eigenvectors are
+related across groups.  In this case, group-level equivariance to
+rotations is an unreasonable assumption.  Here, both eigenvalue
+\emph{and} eigenvector shrinkage can play an important role in
+improving covariance estimates.  Below, we introduce our extension of
+the spiked covariance model to the group setting by including an
+assumption about the space shared by the span of the group-level eigenvectors.
+
+\paragraph{The Shared Subspace Model:}
+
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}[b]{0.25\textwidth}
+        \includegraphics[width=\textwidth]{Figs/3dplot}
+        \caption{Projection in $\mathcal{R}^3$}
+        \label{fig:dmelanRatio}
+    \end{subfigure}
+\quad
+ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.25\textwidth}
+        \includegraphics[width=\textwidth]{Figs/2d-scatter}
+        \caption{$Y_kV$}
+        \label{fig:dmelanPosterior}
+    \end{subfigure}
+\quad 
+    \begin{subfigure}[b]{0.25\textwidth}
+        \includegraphics[width=\textwidth]{Figs/2d-scatter-orth}
+        \caption{$Y_kV_{\perp}$}
+        \label{fig:dmelanPosterior}
+    \end{subfigure}
+    \caption{ Two groups of four-dimensional data (red and blue)
+      projected into different subspaces.  a) To visualize $Y_k$ we
+      can project the data into $R^3$.  In this illustration, the
+      distributional differences between the groups are confined to a
+      two-dimensional shared subspace ($VV^T$, grey plane).  b) The
+      data projected onto the two-dimensional shared subspace,
+      $Y_kV$, have covariances $\psi_k$ that differ between
+      groups. c) The orthogonal projection, $Y_kV_{\perp}$
+      has isotropic covariance, $\sigma_k^2I$, for all groups.  }
+\label{fig:shared}
+\end{figure}
+
+The spiked covariance model assumes that the goal is to estimate a
+single covariance matrix.  The situation is different when we are
+interested in estimating multiple covariance matrices from distinct,
+but related groups.  For multi-group estimation, we can often improve
+upon the efficiency of covariance estimators by pooling information
+across the groups of data.  Here, we assume we have $k=1, ..., K$
+groups of mean-zero normal data, $Y_k$, where $Y_k$ consists of $n_k$
+independent rows of measurements on $p$ features, typically with
+$n_k \ll p$.  Then, $Y_k^TY_k = \bl S_k$ follow a (degenerate) Wishart
+distribution as in Equation \ref{eqn:lik}.  To improve estimation, we
+seek estimators for the population covariance from one group that depend on the
+data from all $K$ groups, e.g. of the form
+$\hat \Sigma_k(\bl S_1, ... \bl S_K)$.  To this end, we extend the
+spiked covariance model to grouped data, by assuming that the
+anisotropic variability from each group can be expressed on a common
+low dimensional subspace.  Specifically, we posit that the covariance
+matrix for each group can be written as
 
@@
 \end{equation}
-
-Here, $V \in \mathcal{V}_{s,p}$ and $p \geq s$.  $V$ is only
-identifiable up to right rotations.  Consequently,
-$VV^T \in \mathcal{G}_{s, p}$, which defines the plane of variation
-shared by all groups, is identifiable.  $\psi_k$ is a rank-$r$
-covariance matrix on the $s$-dimensional subspace.  Such a model is
-reasonable for many applications involving groups of highly structure
-data on the same set of $p$ features.  For instance, in biological
-applications, only certain subsets of molecules can physically
-interact which constrains the space of possible correlation patterns.
-Our model exploits this notion, by estimating differences between
-groups on a common lower dimensional subpace.
-
-\begin{figure}[t]
-    \centering
-    \begin{subfigure}[b]{0.25\textwidth}
-        \includegraphics[width=\textwidth]{Figs/3dplot}
-        \caption{Ratio}
-        \label{fig:dmelanRatio}
-    \end{subfigure}
-    \qquad     \qquad     \qquad 
- %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-      %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.25\textwidth}r
-        \includegraphics[width=\textwidth]{Figs/2d-scatter}
-        \caption{Posterior}
-        \label{fig:dmelanPosterior}
-    \end{subfigure}
-    \caption{ }
-\end{figure}
+%
+\noindent Here, the columns of $V \in \mathcal{V}_{p, s}$ determine a subspace
+shared by all groups $k$ with $p \geq s$.  Throughout this paper we
+will denote to the shared subspace as $VV^T \in \mathcal G_{r,p}$, where
+$\mathcal G_{r,p}$, known as the Grassmann manifold, consists of all
+$s$-dimensional linear subspaces of $\mathcal{R}^p$
+\citep{Chikuse2012}.  Although $V$ is only identifiable up to right
+rotations, $VV^T$, which defines the plane of variation shared by all
+groups, is identifiable.  Later, to emphasize the connection to the
+spiked PCA model (\ref{eqn:spiked}), we will write $\psi_k$ in terms
+of its eigendecomposition, $\psi_k = O_k\Lambda_kO_k$, where $O_k$ are
+eigenvectors and $\Lambda_k$ are the eigenvalues of $\psi_k$ (see
+Section \ref{sec:bayes}).
+
+Note that for the shared subspace model,
+$V^T\Sigma_kV = \psi_k + \sigma_k^2I$ is an anisotropic
+$s$-dimensional covariance matrix for the projected data, $Y_kV$.
+In contrast, the data projected onto the orthogonal space,
+$Y_kV_{\perp}$, is isotropic, and typically small, for all
+groups.  In Figure \ref{fig:shared} we provide a simple illustration
+of this fact using simulated $4$-dimensional data from two groups.  The
+differences in distribution between the groups of data can be
+expressed on a two dimensional subspace spanned by the columns of
+$V \in \mathcal{V}_{4, 2}$.  Differences in the correlations between
+the two groups manifest themselves on this shared subspace, whereas
+only the magnitude of the isotropic variability can differ between groups
+on the orthogonal space.  Thus, the the shared subspace model can be
+viewed as a covariance partition model, where one partition includes
+the anisotropic variability from all groups and the other partition is
+constrained to the isotropic variability from each group.  This
+isotropic variability is often taken to be ``measurement noise''.
+
+%Such a modelis reasonable for many applications involving groups of
+%highly structure data on the same set of $p$ features.  % For instance,
+% in biological applications, only certain subsets of molecules can
+% physically interact which constrains the space of possible correlation
+% patterns.  Our model exploits this notion, by estimating differences
+% between groups on a common lower dimensional subpace.
 
 \section{Inference}
-
+\label{sec:inference}
 In this section we outline Bayesian inference algorithms including
@@
 can quantify the uncertainty associated with each estimate.
-Uncertainty quantification is particularly important in the $n << p$
+Uncertainty quantification is particularly important in the $n \ll p$
 setting where seemingly large differences in the point estimates of
 the low rank covariance matrices across groups may not actually
-reflect significant differences.  Although, uncertainty quantification
+reflect significant differences.  Although uncertainty quantification
 for all parameters is desirable, for the size of the problems we are
 generally interested in, we found MCMC inference for the subspace
-$VV^T$ to be intractable.  Our optimziation based approached is
-signifcantly faster and works well for large large dimensional
-datasets.  As such, we start, in Section \ref{sec:em}, by outlining an
-expectation-maximization algorithm for estimating the shared subspace
-and derving point estimates of the projected covariance $\psi_k$. In
-Section \ref{sec:bayes} we describe a Gibbs sampler for posterior
-inference of $\psi_k$ given the common subspace.  These two algorithms
-can be used in conjuction to first identify and evaluate a shared
-subspace and then to quantify the uncertainty associated with the
-projected covariance matrices on this space.
-
-\subsection{An Expectation-Maximization Algorithm}
+$VV^T$ to be intractable.  Our optimization based approached is
+significantly faster and works well for high-dimensional datasets.  As
+such, we start, in Section \ref{sec:em}, by outlining an
+expectation-maximization algorithm for estimating the shared
+subspace. In Section \ref{sec:bayes} we describe a Gibbs sampler for
+posterior inference of $\psi_k$ given the common subspace.  These two
+algorithms can be used in conjunction to first identify and evaluate a
+shared subspace and then to quantify the uncertainty associated with
+the projected covariance matrices.
+
+\subsection{Estimating the Shared Subspace}
 \label{sec:em}
@@
 In this section we describe a computationally efficient algorithm for
-estimating the shared subspace, $VV^T$, projected covariance matrix
+estimating the shared subspace, $VV^T$, the projected covariance matrix
 $\psi_k$ and noise parameter $\sigma_k^2$.  First, note that the full
@@
 \end{align}
-
-Where we define $M_k^{-1} = (\psi_k + I) ^{-1}/\sigma_k^2$.  The log-likelihood in
+%
+\noindent where we define $M_k^{-1} = (\psi_k + I) ^{-1}/\sigma_k^2$.  The log-likelihood in
 $V$ (up to additive constant) is
@@
 
-We propose an expectation-maximization (EM)
-algorithm, to estimate the relevant quantities.  In particular we
-maximize the marginal posterior distribution of $V$, treating 
-$M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as the ``missing'' parameters to
-be imputed.  We assume independent Jeffreys' prior distributions for
-both $\frac{1}{\sigma_k^2}$ and $M_k^{-1}$ and a uniform prior for $V
-\in \mathcal{V}_{s, p}$.  In particular we let
+We propose an expectation-maximization (EM) algorithm, to estimate the
+relevant quantities.  In particular we maximize the marginal posterior
+distribution of $V$, treating $M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as
+the ``missing'' parameters to be imputed.  We assume independent
+Jeffreys' prior distributions for both $\frac{1}{\sigma_k^2}$ and
+$M_k^{-1}$ and a uniform prior for $V \in \mathcal{V}_{s, p}$.
+Jeffreys' prior for these quantities corresponds to
 $p(\frac{1}{\sigma_k^2}) \propto \sigma_k^2$ and
@@
 
-\paragraph{Evaluating a Shared Subspace}
+\paragraph{Evaluating a Shared Subspace:}
 
@@
 in practice there are no guarantees that the inferred subspace is
-equally appropriate for all groups.  Moreover, the shared subspace can
-be chosen in other ways, including more ad hoc methods
-involving the singular value dcomposition or simply by incorporating
-subject matter considerations.  Regardless of how a particular
-subspace is identified, it is usually important to quantify how useful
-this subspace is for explaining the variability in each group.  In
-particular, if $\hat{V}$ is a basis for a subspace shared across
-groups, then for each $k$, most of the variation in $\Sigma_k$ should
-lie on the subspace spanned by $\hat{V}$.  To quantify the extent to
-which this is true for different groups, we propose a simple estimator
-for the proportion of ``signal'' variance that lies on any given
-subspace.  Specifically, we apply the following estimator for ratio of
+appropriate for all groups.  Moreover, the shared subspace can be
+chosen in other ways, including more ad hoc methods involving the
+singular value decomposition or simply by incorporating subject matter
+considerations.  Regardless of how a particular subspace is specified,
+it is important to quantify how useful this subspace is for explaining
+the variability in each group.  In particular, if $V$ is a basis for
+the shared subspace, then for each $k$, most of the variation in the
+data should be preserved when projecting the data to this subspace.
+To quantify the extent to which this is true for different groups, we
+propose a simple estimator for the proportion of ``signal'' variance
+that lies on any given subspace.  Specifically, for any inferred
+subspace $\hat{V}$, we apply the following estimator for ratio of
   the sum of the first $s$ eigenvalues of
@@
   eigenvalues of $\Sigma_k$:
-
+%
 \begin{equation}
- \frac{||Y_k^T\hat{V}||_F}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{sup}
-  ||Y_k^T\widetilde{V}||_F - \hat{\sigma}_k^2ps}
+ \frac{||Y_k\hat{V}||_F}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{sup}
+  ||Y_k\widetilde{V}||_F - \hat{\sigma}_k^2ps}
 \label{eqn:ratio}
 \end{equation}
-
-where $||.||_F$ is the Frobenius norm. If the inferred subspace is the
-true subspace, e.g. $\hat{V}\hat{V}^T = VV^T$, then this estimator
-will asymptotically approach one.  To see this, note that
-$\underset{V \in \mathcal{V}_{p, s}}{sup}\frac{||Y_k^TV||_F}{n_k}$ is
+%
+\noindent where $||.||_F$ is the Frobenius norm. If the inferred
+subspace is in fact the true subspace, i.e. $\hat{V}\hat{V}^T = VV^T$,
+then this estimator will asymptotically approach one.  To see this,
+note that
+$\underset{V \in \mathcal{V}_{p, s}}{sup}\frac{||Y_kV||_F}{n_k}$ is
 equivalent to the sum of the first $s$ eigenvalues of the sample
@@
 If we assume that $s$ is fixed in this asymptotic regime, then
-$||Y_k^TV||_F/n_k$ is a consistent estimator for
-$\sum_i^s \lambda^{(k)}_i$, and thus the above estimator will be close
+$||Y_kV||_F/n_k$ is a consistent estimator for
+$\sum_i^s \lambda^{(k)}_i$.  As a consequence, the proposed estimator will be close
 to one when $\hat{V}\hat{V}^T = VV^T$ and smaller if not.  The metric provides a
@@
 given subspace and which groups cannot.  In Section \ref{sec:simulation} we plot
-these estimates for different simulated datasets.  In Section \ref{sec:bio}
-we plot the estimates for inferred subspaces for real biological data.
+these estimates for different simulated datasets.  In Section \ref{sec:app}
+we plot the estimates for subspaces  inferred with real biological data.
 
@@
 
-Although the EM algorithm presented in the previous section yields a
-fast way to derive point estimates for $VV^T$ and $\psi_k$, it does
-not lead to natural uncertainty bounds for these estimates. Full
-uncertainty quantification fo $VV^T$ is desirable but we found this
+Although the EM algorithm presented in the previous section yields
+point estimates for $VV^T$ and $\psi_k$, it does not lead to natural
+uncertainty bounds for these estimates. Full uncertainty
+quantification for $VV^T$ may be desirable but we found this
 challenging for the dimensions of the problems we would like to
 consider.  As such, in this section, we assume that the subspace
-$VV^T$ is fixed and known and focus on uncertainty quantificatoin for
+$VV^T$ is fixed and known and focus on uncertainty quantification for
 $\psi_k$ only.  Further, by focusing on group differences on a fixed
 subspace, it is simpler to interpret differences across covariance
-matrices.  The shared subspace can be an estimated using the EM algorithm
-presented in Section \ref{sec:em} or chosen based on subejct matter
-considerations.  Note that when the subspace is known, the posterior
-distribution of $\Sigma_k$ is conditionally independent from the other
-groups, i.e.
+matrices.  %The shared subspace can be an estimated using the EM
+% algorithm presented in Section \ref{sec:em} or chosen based on
+% subject matter considerations.  
+Note that when the subspace is known, the posterior distribution of
+$\Sigma_k$ is conditionally independent from the other groups, i.e.
 $P(\Sigma_{k} | V, \bl S_k, \Sigma_{-k}) = P(\Sigma_{k} | V, \bl
@@
 support of $\psi_k$ is not constrained to the cone of positive
-semi-definite matrices.
-
-Direct prior specifications for $\psi_k$ are more appealing, but many
-obvious choices, like the inverse-Wishart prior distribution for
-$\psi_k$ do not lead to simple posterior distributions.  As such,
-building on recent interest in spiked covariance models
-\citep{Hoff2009, Donoho2013, Paul2007} we focus on tractable models
-based on the eigendecomposition of $\psi_k$.  Such models are
-especially appealing because it is simple to summarize and compare
-$\psi_k$ across groups in terms their eigenstructure (See Section
-\ref{sec:simulation} and \ref{sec:bio}).
+semi-definite matrices.  Direct prior specifications for $\psi_k$ are
+more appealing, but many obvious choices, like the inverse-Wishart
+prior distribution for $\psi_k$ do not lead to simple posterior
+distributions.  As such, building on recent interest in spiked
+covariance models \citep{Donoho2013, Paul2007} we focus on
+tractable models based on the eigendecomposition of $\psi_k$.  Such
+models are especially appealing because it is simple to summarize and
+compare $\psi_k$ across groups in terms their eigenstructure (See
+Section \ref{sec:simulation} and \ref{sec:app}).
 
 By modeling the eigenstructure, we can now view each covariance
-$\Sigma_k$ as a spiked covariance matrix.  Equation \ref{eqn:sspsi}, written as a function $V$, becomes
+$\Sigma_k$ in terms of the spiked principle components model.
+Equation \ref{eqn:sspsi}, written as a function $V$, becomes
 
@@
 \end{equation}
-
-Here, we make one further generalization to allow
+%
+\noindent Here, we make one further generalization to allow
 $\psi_k = O_k\Lambda_KO_k^T$ be a rank $r \leq s$ dimensional covariance
@@
 $\psi_k$.  For any individual group, this corresponds to the original spiked
-covariance model (Equation \ref{eqn:spiked}) with $U_k = VO_k \in
+PCA model (Equation \ref{eqn:spiked}) with $U_k = VO_k \in
 \mathcal{V}_{p, r}$.
 
-The likelihood for $\Sigma_k$ given the sufficient statistic $\mathbf{S}_k = Y_kY_k^T$ is simply
-
-\begin{equation}
-p(S | \Sigma_k,n_k) \propto |\Sigma_k|^{-n/2}\etr(-\Sigma_k^{-1}\mathbf{S}_k/2)
-\end{equation}
-
-First note that by the Woodbury matrix identity we can show that 
+The likelihood for $\Sigma_k$ given the sufficient statistic
+$\mathbf{S}_k = Y_kY_k^T$ is given in Equation \ref{eqn:lik}.  For the
+spiked PCA formulation, we must rewrite this likelihood in terms of $V$, $O_k$,
+$\Lambda_k$ and $\sigma_k^2$.  First note that by the Woodbury matrix
+identity
+
 \begin{align}
@@
 \end{align}
-
-Where the diagonal matrix $\Omega = \Lambda(I+\Lambda)^{-1}$, e.g. $\omega_i = \frac{\lambda_i}{\lambda_{i}+1}$.  Further,
+%
+\noindent where the diagonal matrix $\Omega = \Lambda(I+\Lambda)^{-1}$, e.g. $\omega_i = \frac{\lambda_i}{\lambda_{i}+1}$.  Further,
 
@@
 \end{align}
-
-where the second line is due to Sylvester's determinant theorem.  The
-likelihood of $V$, $O_k$, $\Lambda_k$ and $\sigma_k^2$ is available
-from Equation\ref{eqn:rplik} by substituting in the appropriate
-quantities in for $\Sigma^{-1}_k$ and $|\Sigma_k|$ and replacing $U_k$
-with $VO_k$:
+%
+\noindent where the second line is due to Sylvester's determinant
+theorem.  Now, the likelihood of $V$, $O_k$, $\Lambda_k$ and
+$\sigma_k^2$ is available from Equation \ref{eqn:lik} by substituting
+the appropriate quantities for $\Sigma^{-1}_k$ and $|\Sigma_k|$ and
+replacing $U_k$ with $VO_k$:
 
@@
 \end{equation}
-
-We use conjugate and semiconjugate priors for the parameters $O_k$,
-$\sigma^2_k$ and $\Omega_k$ to faciliate inference via a Gibbs
+%
+\noindent We use conjugate and semi-conjugate priors for the parameters $O_k$,
+$\sigma^2_k$ and $\Omega_k$ to facilitate inference via a Gibbs
 sampling algorithm.  In the absence of real prior information,
@@
 class of prior distributions is conjugate for$\sigma_k^2$.  We chose a
-default prior distriubtion for $\sigma^2$ that is equivariant with
+default prior distribution for $\sigma^2$ that is equivariant with
 respect to scale changes.  Specifically, we use Jeffreys' prior, an
@@
 inverse-gamma$( n_k p/2 , \tr(\bl S_k(\bl I -\bl U_k\Omega_k\bl
-U_k)/2)$, where $U_k = \hat{V}O_k$
+U_k)/2)$, where $U_k = VO_k$
 
@@
 \label{lik_vo}
- p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto \etr(\Omega_kO^T_k\hat{V}^T[S_k/(2\sigma^2_k)]\hat{V}O_k)
+ p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto \etr(\Omega_kO^T_kV^T[S_k/(2\sigma^2_k)]VO_k)
 \end{align}
-
+%
 \noindent This is a Bingham$(\Omega, \hat{V}^T \bl S_k \hat{V}/(2\sigma^2)$
@@
 \begin{align}
-p(\omega_{ki}|V, O_k, \bl S_k) \propto_{\omega_{ki}}
+p(\omega_{ki}|V, O_k, \bl S_k) &\propto_{\omega_{ki}}
   \left(\prod_{i=1}^r(1-\omega_{ki})^{n_k/2}  \right)
   \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) \\
-&  \propto & (1-\omega_{ki})^{n/2} e^{c \omega_s  n/2},    
+&  \propto  (1-\omega_{ki})^{n/2} e^{c \omega_s  n/2},    
 \label{eqn:wpost}
 \end{align}
-
-where $c = \bl u_{ki}^T S_k \bl u_{ki}/(n \sigma^2)$ and $\bl u_{ki}$ is
+%
+\noindent where $c = \bl u_{ki}^T S_k \bl u_{ki}/(n \sigma^2)$ and $\bl u_{ki}$ is
 column $i$ of $U_k = VO_k$.  While not proportional to a density
@@
 
-\subsection{Model Works}
-First, we do a single simulation that shows that we can recover the
-true parameters using our inferential algorithm when the assumed model
-is correct.
+\subsection{Validation of Inference}
+
+In this section we start by validating, in simulation, that our
+inferential algorithm recovers the true parameters of the data
+generating mechanism.  We simulate five groups of $p=200$ dimensional
+data each with sample size $n_k=50$.  We choose the eigenvectors of
+the anisotropic component of the covariance matrix, $\psi_k$, to share
+a common two dimensional subspace.  To fit the model to this simulated
+data, we first estimate the shared subspace using the EM algorithm outlined
+in Section \ref{sec:em} and then use the MCMC algorithm described in
+Section \ref{sec:bayes} to estimate the posterior distribution of the
+projected covariance matrices $\psi_k$ and noise variance $\sigma_k^2$.
 
@@
       %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.45\textwidth}
+    \begin{subfigure}[b]{0.4\textwidth}
         \includegraphics[width=\textwidth]{Figs/simRatio-05-19}
@@
     \end{subfigure}
-    \caption{a) 95\% posterior regions for the ratio of the eigenvalues,
-      $\frac{\lambda_1}{\lambda_2}$, of $\psi_k$ and the orientation
-      of the principle axis on space spanned by $\hat{V}$ cover the
-      truth.  True parameter values on $\hat{V}^T\Sigma_k\hat{V}$ are
-      indicated by the crosses.  Since $V$ is only identifiable up to
-      rotation we find the Procrustes rotation that maximizes the
-      similarity of $\hat{V}$ to the true data generating basis. True
-      eigenvalue ratios were 10 (red and black), 3 (green and blue)
-      and 1 (light blue).  True orientations were $\pi/4$ (black), $-\pi/4$
-      (red and green) and 0 (blue and cyan).  b)  Estimated proportion
-    of variance explained by the subspace (Equation \ref{}).  As expected, for each
-    group, all of the variance can be explained on the inferred subspace.  }
+    \caption{a) 95\% posterior regions for the ratio of the
+      eigenvalues, $\frac{\lambda_1}{\lambda_2}$, of $\psi_k$ and the
+      orientation of the principle axis on space spanned by $\hat{V}$
+      cover the truth.  True parameter values on
+      $\hat{V}^T\Sigma_k\hat{V}$ are indicated by the crosses.  Since
+      $V$ is only identifiable up to rotation we find the Procrustes
+      rotation that maximizes the similarity of $\hat{V}$ to the true
+      data generating basis. True eigenvalue ratios were 10 (red and
+      black), 3 (green and blue) and 1 (light blue).  True
+      orientations were $\pi/4$ (black), $-\pi/4$ (red and green) and
+      0 (blue and cyan).  b) Estimated proportion of variance
+      explained by the subspace (Equation \ref{eqn:ratio}).  As
+      expected, for each group, the inferred subspace explains
+      essentially all of the variance from the first two eigenvectors of
+      $\Sigma_k$.  }
 \end{figure}
 
-In order to quantify the uncertainty associated with the relevant
-parameters conditional on a shared subspace, we use the Bayesian
-inference algorithm outlined in Section \ref{sec:bayes}. In Figure
-\ref{fig:posteriorRegions} we depict the 95\% posterior regions for
-statistics based on the eigendecomposition of
-$\psi_k$.  Specifically, for the two dimensional
-shared subspace model, we plot the eigenvalue ratio
-$\frac{\lambda_1}{\lambda_2}$ and the angle of the first eigenvector
-on the $\text{arctan}(\frac{O_2}{O_1})$, where
-$\lambda_1 > \lambda_2$.  Non-overlapping 95\% posterior regions are
-indicative of significant differences in the covariance structure
-between groups.
-
-In this example, the posterior regions for all groups cover the true
-eigenvalue ratio and orienntation of the primary axis for the matrix
-$\hat{V}^T\Sigma_k\hat{V}$.  Note that for smaller eigenvalue ratios,
-there is larger uncertainty about the orientation of the primary axis.
-When the ratio is one (light blue) there is no information about
-orientation of the primary axis since the contours are spherical.
-
-\subsection{Model Misspecification}
-
-We present results of a simulation which demonstrates the
-relative performance of shared subspace group covariance estimation
-under different data generating models.  In this simulation, we consider three different
-spiked covariance data models: 1) ``complete pooling'', in which the
-spiked covariance matrices from all groups are identical, e.g.
+For this two dimensional shared subspace model we summarize the
+estimates of $\psi_k$ in terms of its eigendecomposition.
+Specifically, we compute the eigenvalue ratio
+$\frac{\lambda_1}{\lambda_2}$, with $\lambda_1 > \lambda_2$, and the
+angle of the first eigenvector on this subspace,
+$\text{arctan}(\frac{O_{12}}{O_{21}})$, relative to the first column
+of of $V$.  Figure \ref{fig:simPosterior} depicts the 95\% posterior
+regions for these quantities.  We select the displayed posterior
+samples in each group by iteratively remove the points corresponding
+to the vertices of the convex hull until only 95\% of the original
+samples remain.  Non-overlapping posterior regions are indicative of
+significant differences in the covariance structure between groups.
+
+In this example, the ratio of the eigenvalues of the true covariance
+matrices are $10$ (red and black groups), $3$ (green and blue
+groups) and $1$ (cyan groups).  Larger values of this ratio
+correspond to more correlated contours and a value of $1$ implies
+isotropic covariance.  The posterior regions for all groups cover the
+true eigenvalue ratio and primary axis orientation of $\hat{V}^T\Sigma_k\hat{V}$.
+Note that for smaller eigenvalue ratios, there is larger uncertainty
+about the orientation of the primary axis.  When the ratio is one, as
+is the case for the cyan colored group, there is no information
+about orientation of the primary axis since the contours are
+spherical.
+
+We also verify that the data projected onto the inferred subspace
+preserves most of the variability in the data.  To do this, we compute
+estimates for the proportion of explained variance, as given in
+Equation \ref{eqn:ratio}.  Figure \ref{fig:simRatio} depicts a barplot
+of this estimate for each group.  For each group, the estimate is
+nearly one, which reflects the fact that we were able to infer a two
+dimensional shared subspace that explains most of the variance from
+the first two eigenvectors of $\Sigma_k$.  In this example, by
+construction we ensured that such a subspace exists. However, this is
+not always the case.  In the next section we investigate examples for
+which the inferred subspace may not capture the relevant variation for
+all groups.
+
+\subsection{Rank Selection and Model Misspecification}
+
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{Figs/LossVsDimension}
+        \caption{Stein's Loss vs s}
+        \label{fig:sdimension}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{Figs/simRatio-s5}
+        \caption{s = 5}
+        \label{fig:ratio-s5}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{Figs/simRatio-s20}
+        \caption{s = 20}
+        \label{fig:ratio-s20}
+    \end{subfigure}
+    \caption{a) Average Stein's loss as a function of the dimension of
+      the shared subspace dimension.  Data from ten groups, $U_k$
+      uniform on the Stiefel manifold $\mathcal{V}_{200, 2}$.  As the
+      shared subspace dimension increases, $s \rightarrow p$, the loss
+      converges to the loss from the independent spiked covariance
+      estimates (dashed line). b) For a simulated dataset, estimated
+      fraction of variance explained when the assumed shared subspace
+      is dimension 5.  c).  For a single simulated dataset, estimated
+      fraction of variance explained when the assumed shared subspace
+      is dimension 20.  We can capture nearly all of the variability
+      in each of the 10 groups using an $s=rK=20$ dimensional shared
+      subspace. }
+\label{fig:dimensionPlots}
+\end{figure}
+
+Naturally, shared subspace inference works well when the model is
+correctly specified.  What happens when the model is not well
+specified?  We explore this question in silico by simulating data from
+different data generating models and evaluating the efficiency of
+various covariance estimators.  In all of the following simulations we
+evaluate covariance estimates using Stein's loss,
+$L_S( \Sigma_k , \hat\Sigma_k) = \text{tr}( \Sigma_k^{-1} \hat
+\Sigma_k ) - \log |\Sigma_k^{-1} \Sigma_k | - p$.
+Since we compute estimates for multiple groups of data, we report the
+average Stein's loss
+$L(\Sigma_1, ..., \Sigma_K; \hat\Sigma_1, ..., \hat\Sigma_K ) =
+\frac{1}{K} \sum_k L_S( \Sigma_k , \hat\Sigma_k)$.
+Note that under Stein's loss, the Bayes estimator is the inverse of
+the posterior mean of the precision matrix,
+$\hat \Sigma_{k} = \Exp{ \Sigma_k^{-1} | \bl S_k}^{-1}$ which we
+estimate using MCMC samples.
+
+We start by investigating the behavior of our model when we
+underestimate the true dimension of the shared subspace.  In this
+simulation, we generate data from five groups of mean-zero normally
+distributed data in a $p=200$ dimensional feature space.  For each
+group we use the spiked covariance model of Equation \ref{eqn:spiked},
+where we generate eigenvectors uniformly from the Stiefel manifold,
+$U_k \in \mathcal{V}_{200, 2}$ with eigenvalues
+($\lambda_1, \lambda_2) = (250, 25)$ .  In this case, although most of
+the variation in each group can be captured by a two dimensional
+subspace, these are not the same subspaces across groups.  We make no
+assumption, a priori, that the subspaces are similar across groups.
+Below, we evaluate how well the shared subspace estimator performs with such
+data for inferred subspaces of increasing dimension.
+
+In Figure \ref{fig:sdimension} we plot the average Stein's loss, for
+estimates derived from a shared subspace model for different values
+of the shared subspace dimension, $s$.  For each value of $s$, we
+report the loss averaged over ten independent simulations.  In this
+figure, the dashed blue line corresponds to the average Stein's loss
+for covariance matrices estimated independently, e.g. without the
+shared subspace assumption.  As the shared subspace dimension increases to the
+dimension of the feature space, $s \rightarrow p$, the loss for the
+shared subspace estimator quickly converges to the loss for the
+individual spiked covariance estimator.  Note that
+it is always true that span($[U_1, ..., U_K]) \leq rK$.  Even when there is little
+similarity between the eigenvectors from each group, the shared
+subspace estimator with $s = rK$ will perform comparably to the
+estimator, provided that we can identify a subspace, $\hat{V}\hat{V}^T$, where
+span($\hat{V})$ is close to $\text{span}([U_1, ..., U_K])$.  This is
+immediately apparent in Figure \ref{fig:sdimension} from the phase transition
+in Stein's loss at $s = 20$.
+
+From the above simulation, it is clear that correctly specifying the
+dimension of the shared subspace is important for efficient covariance
+estimation.  When the dimension of the shared subspace is too small,
+there is larger loss for estimate relative to the truth.  We can also
+use the available data, independent of the true covariance matrices,
+to evaluate the appropriateness of the inferred subspace.  To this
+end, we calculate the estimate of the ``proportion of explained
+variance'' estimator from Section \ref{sec:em} (Equation
+\ref{eqn:ratio}).
+
+Figure \ref{fig:dimensionPlots}b and c depict the estimated proportion of
+variance explained for a single dataset for each of five group, using
+inferred subspaces of two different dimensions.  For the results
+displayed in Figure \ref{fig:ratio-s5}, we fixed the dimension of the
+subspace to $s=5$ (clearly too small) and for the results displayed in
+Figure \ref{fig:ratio-s5} we fixed the dimension of the shared subspace
+to $s=20$.  When the shared subspace model is appropriate, the
+subspace spanned by $\hat{V}$ explains most of the variance in
+$\Sigma_k$, for all groups.  When the estimated subspace is too small,
+the proportion of variance will be smaller than one for at least some of the
+groups.  For a given fixed subspace, these plots can be used to decide
+which groups can be fairly be compared on this space and which groups
+cannot and whether we would benefit from fitting a model with
+increased dimension $s$.
+
+\paragraph{Model Comparison and Rank Estimation:}
+
+In the previous section, it was apparent that correct specification
+for the rank of the shared subspace is important for efficient
+inference.  So far, we have assumed that the group rank, $r$, and
+shared subspace dimension, $s$, are fixed and known, but in practice
+this is not the case.  Prior to fitting a model we should estimate
+these quantities.  \citet{Gavish2014} provide an asymptotically
+optimal (in mean squared error) singular value threshold for low rank
+matrix recovery with noisy data.  We use their rank estimator, which
+is a function of the median singular value of the data matrix and the
+ratio $\gamma_k =\frac{p}{n_k}$, to estimate the rank, $r$, of a
+spiked covariance matrix.  Although their estimator was derived for
+individual covariance estimation, we found as a useful heuristic,
+Gavish and Donoho's estimator can also be applied to the pooled data
+to estimate the shared subspace dimension, $s$.  Specifically, we concatenate
+the data from all groups to create an $(\sum_k n_k) \times p$
+dimensional matrix and apply their rank estimator to this matrix to
+choose $s$.
+
+Using these rank estimators, we conduct a simulation which demonstrates
+the relative performance of shared subspace group covariance
+estimation under different data generating models.  In this
+simulation, we consider three different spiked covariance data models:
+1) ``complete pooling'', in which the spiked covariance matrices from
+all groups are identical, e.g.
 $\Sigma_1 = \Sigma_2 = ... \Sigma_K = U\Lambda U^T + \sigma^2I$; 2)
@@
 uniform sample from the Stiefel manifold $\mathcal{V}_{r, p}$.  For
-each of these data generating models, we simulate datasets
-with $p=200$ features, rank $r=2$ and eigenvalues
+each of these data generating models, we simulate datasets with
+$p=200$ features, rank $r=2$ and eigenvalues
 $(\lambda_1, \lambda_2) = (250, 25)$.  For data generated under the
-shared subspace model, we additionally assume that the first two
-eigenvectors span a common two dimensional subspace, e.g. we fix
-$s=2$.
-
-We evaluate the performance of the three corresponding inferential
-models on thirty simulated datasets for each data generating
-mechanism.  In practice, the true rank of the covariance matrix is not
-known, and thus, prior to fitting each of these inferential models we
-must first estimate the rank.  \citet{Gavish2014} provide an
-asymptotically optimal (in mean squared error) singular value
-threshold for low rank matrix recovery with noisy data.  We use their
-rank estimator, which is a function of the median singular value of
-the data matrix and $\frac{p}{n_k}$, to independently estimate the
-rank, $r$, of the spiked covariance matrices.  For data generated
-under the shared subspace model we must also estimate the dimension of
-the shared subspace.  As a useful heuristic, we use Gavish and
-Donoho's estimator on the pooled data to estimate the shared subspace
-dimension, $s$.
-
-Given estimates of $s$ and $r$, we then proceed with Bayesian
-inference of the eigenvectors, eigenvalues, and noise variance using
-each of the three inferential models.  When applying the shared
-subspace estimator, we use the EM algorithm presented in Section
-\ref{sec:em} first, to estimate $VV^T$ and then use Bayesian inference
-to estimate the posterior distribution of all remaining quantities
-conditional on $V$.  For all covariance estimates, we compute Stein's
-loss
-$L_S( \Sigma_k , \hat\Sigma_k) = \text{tr}( \Sigma_k^{-1} \hat
-\Sigma_k ) - \log |\Sigma_k^{-1} \Sigma_k | - p$
-for each group.  Under Stein's loss, the Bayes estimator is the
-inverse of the posterior mean of the precision matrix,
-$\hat \Sigma_{k} = \Exp{ \Sigma_k^{-1} | \bl S_k}^{-1}$ which we
-estimate using MCMC samples.  In Table \ref{table:groupLoss} we report
-the average group loss
-$L(\Sigma_1, ..., \Sigma_K; \hat\Sigma_1, ..., \hat\Sigma_K ) =
-\frac{1}{K} \sum_k L_S( \Sigma_k , \hat\Sigma_k)$
-and 95\% intervals for the estimates derived using each inferential
-model.
-
-Naturally, the estimates with the lowest loss are derived from the
-inferential model that matches the data generating model. However, the
-shared subspace estimator has small loss under model misspecification
-relative to the alternative.  For example, when the data is generated
-using the complete pooling model, the shared subspace estimator has
-almost four times smaller loss than the no pooling estimator.  When
-the data is generated under the no pooling model, the shared subspace
-estimator is over an order of magnitude better than the complete
-pooling estimator.  Thus, shared subspace estimation can be an
-appropriate choice when the amount of similarity between groups is not
-known a priori.
-
-The shared subspace model combined with the rank
-estimation procedure suggested by \citet{Gavish2014} leads to an
-estimator which adapts to the amount of similarity between groups.
-Importantly, the shared subspace model can be viewed as a
-generalization of the ``no pooling'' spiked covariance model.  The two
-are equivalent when we let the shared subspace span the full space,
-e.g. $V \in \mathcal{O}(p)$.  Figure \ref{fig:sdimension} depicts the
-loss for data generated under the no pooling model and fit with shared
-subspace estimators with different dimension.  As the shared subspace
-dimension increases to the dimension of the feature space,
-$s \rightarrow p$, the loss for the shared subspace estimator quickly
-coverges to the loss for the individual spiked covariance estimator.
-Note that span($[U_1, ..., U_K]) \leq rK$, thus even when there is
-little similarity between the eigenvectors from each group, the shared
-subspace estimator with $s = rK$ will perform comparably to the ``no
-pooling'' estimator, provided that we can identify a subspace where
-span($\hat{V})$ is close to $\text{span}([U_1, ..., U_K])$.
-
-For any given subspace we can identify whether the space is
-inappropriate for some groups, either because the dimension is too
-small or the space itself does not represent every group equally.
-Specifically, we estimate which groups are well represented on the
-subspace and which groups are not using the proportion of variance
-estimator from Section \ref{sec:em} (Equation \label{eqn:ratio}).
-Figure \ref{fig:evalRatios} depicts the estimated proportion of
-variance on the subspace for each group in two different simulated
-datasets.  When the shared subspace model is appropriate, the subspace
-spanned by $\hat{V}$ explains most of the variance in $\Sigma_k$, for
-all groups.  When the estimated subspace is too small, the proportion
-of variance will be small for at least some of the groups.
-
-Ultimately, in many real world examples, we are more interested in
-finding a useful space to interpret and compare differences between
-covariance matrices than we are in finding the lowest-loss estimator
-for the full covariance matrices.  In this context, the relevant
-quantity is often an estimate of the projected covariance,
-$V^T\Sigma V$.  In Figure \ref{fig:lossBoxplot} we depict boxplots for
-the distribution of losses,
-$L_S(\hat{V}^T\Sigma_k\hat{V}, \hat{V}^T\hat{\Sigma}_k\hat{V})$, on
-simulated data.  In the left boxplot the data come from a
-2-dimensional shared subspace model and on the right plot the data
-come from a 10-dimensional shared subspace model.  For both boxplots,
-estimates $\hat{V}^T\hat{\Sigma}_k\hat{V}$ are derived from assumed
-2-dimensional model. This figure demonstrates that the loss for
-estimating the projected covariance matrix is not significantly larger
-when we underestimate the dimension of the true shared subspace.
+shared subspace model, we additionally assume that the two
+eigenvectors of $\psi_k$ span a common two dimensional subspace, e.g. we fix
+$s=2$.  We simulate 100 independent datasets for each data generating
+mechanism.
+
+For each data generating mechanism we fit each of three corresponding
+inferential models (``complete pooling'', ``shared subspace'' and ``no
+pooling''). For the complete pooling and no pooling models, which
+involve independent estimation of individual covariance matrices, we
+estimate the rank $r$, and then proceed with Bayesian inference of the
+eigenvectors, eigenvalues, and noise variance.  For the shared
+subspace estimator, we first estimate the shared subspace dimension
+$s$ and then infer the optimal $s$-dimensional shared subspace,
+$\hat{V}\hat{V}^T$, using the EM algorithm presented in Section
+\ref{sec:em}.  We then use Bayesian inference to estimate the
+posterior distribution of the projected covariance matrices
+conditional on $\hat{V}\hat{V}^T$.  In Table \ref{table:groupLoss} we report the
+average Stein's loss, with corresponding 95\% intervals, for the
+estimates derived from each inferential model.
 
@@
 
-\begin{figure}[t]
-    \centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/LossVsDimension}
-        \caption{Ratio}
-        \label{fig:sdimension}
-    \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-      %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/lossBoxplot}
-        \caption{Posterior}
-        \label{fig:lossBoxplot}
-    \end{subfigure}
-    \caption{Left) Stein's loss for data generated under the ``no
-      pooling'' model and fit using the shared subspace model.  As the
-      shared subspace dimension increases, $s \rightarrow p$, the loss
-      converges to the loss from the independent spiked covariance
-      estimates (dashed line). Right) Distribution of Stein's loss
-      $L_S(\hat{V}^T\Sigma_k\hat{V}, \hat{V}^T\hat{\Sigma_k}\hat{V})$
-      for data generated from a two dimensional shared subspace (left
-      box) and data generated with a ten dimensional shared subspace
-      (right).  In both cases the estimates were computed assuming a
-      two-dimensional subspace, $s$ and matrix rank $r=2$.  Even when
-      we underestimate the dimension of the shared subspace, the
-      average loss of the estimator for the projected covariance matrices is
-      similar to the loss under the correctly specified model. }
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-    \includegraphics[width=0.75\textwidth]{Figs/evalRatios}
-    \caption{Proportion of variance explained by the estimated
-      subspace, $\hat{V}\hat{V}^T$, for each of 10 groups.  Left) Data
-      generated and fit under the shared subspace with $s=r=2$.  The
-      estimated two dimensional subspace explains nearly all of the
-      variance in each of the ten groups.  Right) Data generated from
-      the shared subspace model with $s=10$ and $r=2$ but estimated
-      assuming $s=r=2$.  In this case, a two dimensional subspace does
-      not capture the proportion of variance for all groups.  Here,
-      the inferred subspace explains over 50\% of the variance for
-      groups 2, 5, 7, 9 and 10 but almost none of the variance for
-      groups 6 and 8. }
-\label{fig:evalRatios}
-\end{figure}
+As expected, the estimates with the lowest loss are derived from the
+inferential model that matches the data generating model. However, the
+shared subspace estimator has small loss under model misspecification
+relative to the alternative.  For example, when the data is generated
+using the complete pooling model, the shared subspace estimator has
+almost four times smaller loss than the no pooling estimator.  When
+the data is generated under the no pooling model, the shared subspace
+estimator is over an order of magnitude better than the complete
+pooling estimator.  Thus, shared subspace estimation can be an
+appropriate choice when the amount of similarity between groups is not
+known a priori.
+
+The shared subspace model combined with the rank estimation procedure
+suggested by \citet{Gavish2014} can be widely applied to group
+covariance estimation because the estimator adapts to the amount of
+similarity between groups.  Importantly, the shared subspace model can
+be viewed as a generalization of the standard ``no pooling'' spiked covariance
+model. The two are equivalent when we let the shared subspace span the
+full space, e.g. $V \in \mathcal{O}(p)$.  As shown in Figure
+\ref{fig:sdimension}, the shared subspace model performs nearly as
+well as the no pooling estimator when the subspace dimension is equal to $s = rK$.
+In the simulations used to create Table \ref{table:groupLoss}, we
+typically estimated $s$ to be slightly smaller than $rK$ for the ``no
+pooling'' data, which explains the slightly higher loss for the shared
+subspace model. With improved estimators for $s$, the shared subspace
+model should be even more competitive.
+
+% Ultimately, in many real world examples, we are more interested in
+% finding a useful space to interpret and compare differences between
+% covariance matrices than we are in finding the lowest-loss estimator
+% for the full covariance matrices.  In this context, the relevant
+% quantity is often an estimate of the projected covariance, $V^T\Sigma
+% V$.
+
+% In Figure \ref{fig:lossBoxplot} we depict boxplots for
+% the distribution of losses,
+% $L_S(\hat{V}^T\Sigma_k\hat{V}, \hat{V}^T\hat{\Sigma}_k\hat{V})$, on
+% simulated data.  In the left boxplot the data come from a
+% 2-dimensional shared subspace model and on the right plot the data
+% come from a 10-dimensional shared subspace model.  For both boxplots,
+% estimates $\hat{V}^T\hat{\Sigma}_k\hat{V}$ are derived from assumed
+% 2-dimensional model. This figure demonstrates that the loss for
+% estimating the projected covariance matrix is not significantly larger
+% when we underestimate the dimension of the true shared subspace.
 
 \section{Analysis of Biological Data}
-\label{sec:bio}
-As in many natural systems, high dimensional biological data is often
-very structured and thus can be well understood on a relatively low
-dimensional subspace. For example, with gene expression
-data, the effective dimensionality is thought to scale with the number
-of gene regulatory modules, not the number of genes themselves
-\citep{Heimberg2016}.  As such, differences in gene expression across
-groups should be expressed in terms of differences between these
-regulatory modules rather than strict differences between expression
-levels.  Such differences can be examined on a subspace that reflects
-the correlations resulting from these modules.
-
-With this in mind, we apply the shared subspace spiked covariance model to data in two
+\label{sec:app}
+
+In this Section, we apply the shared subspace spiked covariance model to data in two
 different biological examples.  In the first, we compare gene
@@
 this example, after controlling for mean effects due to fly age, gender and
-genotype, we find little evidence of differences in metabolite
+genotype, we find little compelling evidence of differences in metabolite
 covariances between groups.
@@
 $V = [V_1\, V_2]$, with $V_1 \in \mathcal{V}_{p,r}$ as the basis for a
-r-dimensional shared subspace that explains the differences between
+$r$-dimensional shared subspace that explains the differences between
 groups and $V_2 \in \mathcal{V}_{p,(s-r)}$ corresponding to the remaining $(s-r)$
 eigenvectors common to all groups.  Importantly, we make no assumptions about the
-magntidue of the eigenvalues $\Lambda_k$ relative to the eigenvalues
-$D$, so that the largest eigenvectors of $\psi_k$ may be in $V_2$.
-
-We find this formulation useful becasue in real world analyses,
+magnitude of the eigenvalues $\Lambda_k$ relative to the eigenvalues
+$D$, so that the largest eigenvectors of $\psi_k$ may correspond to
+the eigenvectors $V_2$.
+
+We find this formulation useful because in real world analyses,
 differences between groups may not manifest themselves in the first
@@
 there may be large person to person variability common to all groups
-which is unrelated to the grouping of those subjects based on the
-phenotype of interest.  Differences between groups may manifest themselves
-as changes in the smaller principle components. This formulation
-allows, complete pooling of most eigenvectors, yet a space to identify
-and compare relevant differences.  Moreover, visualizing and
-interpreting differences between groups on higher dimensional spaces
-is a challenging problem.  Using this framework we can easily understand the
-largest differences across groups in a two dimensional space while
-explaining the remaining variability with a anisoptropic shared
-structure.
+which is unrelated to the grouping of those subjects. Differences
+between groups may still manifest themselves in the smaller principle
+components. This formulation allows complete pooling of most
+eigenvectors, yet a space to identify and compare relevant
+differences.  Moreover, visualizing and interpreting differences
+between groups on higher dimensional spaces is a challenging problem.
+Using this framework we can easily understand the largest differences
+across groups in a two dimensional space while completely pooling
+remaining anisotropic variability among the groups.
 
@@
 
-Here, we demonstrate the utility of the shared subspace
-covariance estimator for exploring differences in the covariability of
-gene expression levels in young adults with different subtypes of
-pediatric acute lymphoblastic leukemia \citep{Yeoh2002}.  The raw data
-consists of gene expression levels for over 12,000 genes in seven
-different subtypes of leukemia: BCR-ABL, E2A-PBX1, Hyperdip50, MLL,
-T-ALL, TEL-AML1 and a seventh group for unidentified subtypes.  The
-groups have corresponding sample sizes of
-$n = (15, 27, 64, 20, 43, 79, 79)$.  Although there are over 12,000
-genes, the vast majority of expression levels are missing.
-Thus, we restrict our attention to the genes for which less than half
-of the values are missing and use Amelia, a software package for
-missing value imputation, to fill in the remaining missing values
-\citep{Amelia}.  After restricting the data in this way, $p=3124$
-genes remain.  Prior to analysis, we demean both the rows and columns of
-the gene expression levels in each group.
+Here, we demonstrate the utility of the shared subspace covariance
+estimator for exploring differences in the covariability of gene
+expression levels in young adults with different subtypes of pediatric
+acute lymphoblastic leukemia \citep{Yeoh2002}.  The raw data consists
+of gene expression levels for over 12,000 genes in seven different
+subtypes of leukemia: BCR-ABL, E2A-PBX1, Hyperdip50, MLL, T-ALL,
+TEL-AML1 and a seventh group for unidentified subtypes.  The groups
+have corresponding sample sizes of $n = (15, 27, 64, 20, 43, 79, 79)$.
+Although there are over 12,000 genes, the vast majority of expression
+levels are missing.  Thus, we restrict our attention to the genes for
+which less than half of the values are missing and use Amelia, a
+software package for missing value imputation, to fill in the
+remaining missing values \citep{Amelia}.  After restricting the data
+in this way, $p=3124$ genes remain.  Prior to analysis, we demean both
+the rows and columns of the gene expression levels in each group.
 
@@
 positive loadings in both columns of $V$. Genes in the upper left
-corner have positive loadings on the first column of $V$ but negative
-loadings on the second column of $V$.  The principle axis for a group
-aligns in the direction of genes that exhibit the largest variability
-in that group.  Genes which lie in a direction orthogonal to the
-principle axis exhibit reduced variability, relative to the other
-groups. For instance, genes in the upper left (triangles) exhibit
-large, positively correlated variability in the Hyperdip50 group.  In
-this same group, there is reduced varaibility, relative to the other
-groups, among the genes in the upper right of the plot (squares),
-since this cluster of genes lie in a direction orthogonal to the
-principle axis.  In contrast, the BCR-ABL group is aligned primarily
-with the second column of $V$, which means that the genes indicated by
-squares and circles vary significantly in this group.  Genes in the
-square group are anti-correlated with those in the circle group, since
-the their loadings have opposite signs on $V2$.
+corner of the figure have positive loadings on the first column of $V$
+but negative loadings on the second column of $V$.  The principle axis
+for a group aligns in the direction of genes that exhibit the largest
+variability in that group.  Genes which lie in a direction orthogonal
+to the principle axis exhibit reduced variability, relative to the
+other groups. For instance, genes in the upper left corner of this
+plot (triangles) exhibit large, positively correlated variability in
+the Hyperdip50 group.  In this same group, there is reduced
+variability, relative to the other groups, among the genes in the
+upper right corner of the plot (squares), since this cluster of genes lie in
+a direction orthogonal to the principle axis.  In contrast, the
+BCR-ABL group is aligned primarily with the second column of $V$,
+which means that the genes indicated by squares and circles vary
+significantly in this group.  Genes in the square group are
+anti-correlated with those in the circle group, since the their
+loadings have opposite signs on $V2$.
 
@@
 
-\paragraph{Metabolomic Analysis}
+\paragraph{Analysis of Metabolomic Data}
 
@@
 
-In this paper, we present a method for comparing differences between
-covariance matrices on a low dimensional subspace.  We provide an EM
-algorithm for estimating the subspace which explains the largest
-proportion of variance .
-
-Variation that differs across groups versus variation which is the
-same across groups.
-
-Importantly, our approach can be combined with existing hierarchical
-models for multi-group covariance estimation.  For instance, we can
-incorporate additional shrinkage as in \citet{Hoff2009} by utilizing
-non-uniform Bingham prior distributions for the subspace eigenvectors.
-Alternatively, we can summarize the estimated group covariance
-matrices by thresholding entries of the precision matrix to visualize
-differences between groups using a graphical model \citep{}.
-
-Full Bayesian approach using Riemmanian manifold sampler?
+In this paper, we proposed a hierarchical model for estimating and
+comparing differences in covariance matrices across multiple groups on
+a common low dimensional subspace.  We described an EM algorithm for
+estimating this common subspace and a MCMC algorithm for inferring the
+projected covariance matrices and their associated uncertainty.
+Estimates of both the shared subspace and the projected covariance
+matrices can be useful summaries of the data.  For example, with the
+biological data, the shared subspace highlights the full set of genes
+or molecules that might be correlated in any group.  Differences between
+group covariance matrices can be understood in terms of differences in
+these sets of correlated and anti-correlated genes.  In this applied
+analysis, we demonstrated how we can use these notions to visualize
+and contrast the posterior distributions of covariance matrices
+projected onto a particular subspace.
+
+In simulation, we showed that the shared subspace model can
+still be a reasonable choice for modeling multi-group covariance
+matrices even when we have little knowledge, a priori, of how similar
+the groups are.  Even when there is little to no similarity, the shared
+subspace model is appropriate, as long as the dimension of the shared
+subspace is large enough.  However, selecting the rank of the shared
+subspace remains a significant practical challenge.  Although we
+propose a useful heuristic for choosing the dimension of the shared
+subspace based on the rank selection estimators of \citet{Gavish2014},
+a more principled approach is warranted.
+
+It was also a challenge to estimate the subspace $VV^T$ once we had
+chosen a rank.  In our approach, we estimate the shared subspace
+$VV^T$ separately from the MCMC inference of $\psi_k$ because we found
+estimation of $VV^T$ using MCMC inference to be
+particularly difficult.  Although fixing the shared subspace
+simplifies interpretation of the results, in many cases it may be
+desirable to develop a fully Bayesian approach for jointly estimating
+$VV^T$ and $\psi_k$.  Recent Markov chain Monte Carlo algorithms, like
+Riemmanian manifold Hamilton Monte Carlo, which can exploit the
+geometry of the Grassmann manifold, may be useful here \citep{Byrne2013,
+  Girolami2011}.  It may also be possible, though computationally
+intensive, to jointly estimate the rank of the shared subspace, using
+for instance, a reversible-jump MCMC algorithm.
+
+Fundamentally, our approach is quite general and can be integrated
+with other existing approaches for hierarchical multi-group covariance
+estimation.  In particular, we can incorporate additional shrinkage on
+the projected covariance matrices $\psi_k$.  As in \citet{Hoff2009} we
+can employ non-uniform Bingham prior distributions for the
+eigenvectors of $\psi_k$ or model $\psi_k$ as a function of continuous
+covariates as in \citet{Yin2010, Hoff2011}.  Alternatively, we can
+summarize the estimated covariance matrices by thresholding entries of
+the precision matrix, $\psi_k^{-1}$ to visualize differences between
+groups using a graphical model \citet{Meinshausen2006}.  The specifics
+of the problem at hand should dictate which shrinkage models are
+appropriate, but the shared subspace assumption should be useful in a
+wide range of analyses, especially when the number of features is very
+large.
 
@@
 
-\section{Appendix}
+%\section{Appendix}
 

### 2016-06-19 paper/mscov.tex
Intro edits for Peter

@@
-\documentclass{article}
-
+\documentclass[12pt]{article}
+
+\usepackage{amsmath,amssymb,amsthm}
 \usepackage{graphicx}
@@
 \usepackage{multirow, booktabs}
-\usepackage{amsfonts,amsmath,amssymb}
 \usepackage{natbib}
@@
   differences among several $p\times p$ covariance matrices in the
-  $p > n$ setting.  This is done by assuming a spiked covariance model
-  for each group, and sharing information about the subspace spanned
-  by the group-level eigenvectors.  We propose an EM algorithm for
-  estimating the shared subspace and the covariance matrix of the data
-  projected onto this subspace.  We also provide a Gibbs sampling
-  algorithm for estimating the posterior uncertainty of group-level
-  eigenvectors and eigenvalues.  Finally, we demonstrate the utility
-  of our method in an analysis of biological data.
+  $p \gg n$ setting.  This is done by assuming a spiked covariance
+  model for each group and sharing information about the space spanned
+  by the group-level eigenvectors.  We use empirical Bayes to identify
+  a low-dimensional subspace which explains as a large amount of
+  variation across all groups, and, given this subspace, show how a
+  Gibbs sampling algorithm can be used to estimate the posterior
+  uncertainty of group-level eigenvectors and eigenvalues.  The
+  implementation and utility of our model is illustrated with analyses
+  of high-dimensional multivariate biological data.
 
@@
 \section{Introduction}
+
+In many natural systems, high dimensional data is often very
+structured and thus can be well understood on a relatively low
+dimensional subspace. For example, with gene expression data, the
+effective dimensionality is thought to scale with the number of gene
+regulatory modules, not the number of genes themselves
+\citep{Heimberg2016}.  As such, differences in gene expression across
+groups should be expressed in terms of differences between these
+regulatory modules rather than strict differences between expression
+levels.  Such differences can be examined on a subspace that reflects
+the correlations resulting from these modules.
 
@@
 ($n \leq p$) \citep{Dempster1969, Stein1975}.  A host of estimators
-have been devised to address this fact. These primarily include of
-shrinkage based approaches \citep{Schafer2005, Ledoit2011} and low
+have been devised to address this fact. These primarily include
+shrinkage approaches \citep{Schafer2005, Ledoit2011} and low
 rank factor models \citep{Mardia1980, Ullman2003, Fan2008}.  In many
-cases, the covariance matrices are highly structured and thus can be
+cases, a covariance matrix is highly structured and thus can be
 sparsely parameterized even if the matrix itself is not low rank
@@
 
-Most of these approaches assume that the goal is to estimate a
-covariance matrix from a single dataset.  In many applications,
+Most of these approaches assume that the goal is to estimate the
+covariance matrix of a single population.  In many applications,
 however, the data can be split into groups, each of which represent
@@
 
-By exploiting similarity across related groups of high-dimensional
-data, we can increase the efficiency of existing covariance
-estimators.  Some of the earliest approaches for modeling multiple
+Some of the earliest approaches for modeling multiple
 covariance matrices focused on estimation in terms of the spectral
@@
 this area to the multi-group setting, by pooling information about the
-pattern of zeros in across precision matrices.
+pattern of zeros across precision matrices.
 
@@
 Another popular method for modeling relationships between
-high-dimensional multivariate data related that is closely related to
-low rank covariance estimation, is known partial least squares
-regression (PLS). This approach, which is a special case of a bilinear
-factor model, involves projecting the data into a lower dimension
-space which maximizes the similarity of the two groups; this technique
-does not require the data from each group to share the same feature
-set.  An oft used variant for prediction, partial least squares discriminant analysis
-(PLS-DA) is especially common in the chemometrics and bioinformatics
- \citep{Barker2003} .  Although closely related to
-the approaches we will consider here, the primarily goal of of PLS based models
-is to create regression or discrimination models, not to explicitly
-infer covariance matrices from multiple groups of data.  Nevertheless,
-the basic idea that data can often be well represented on a low
-dimensional space is an appealing one that we leverage as well.
-
-In many natural systems, high dimensional data is often very
-structured and thus can be well understood on a relatively low
-dimensional subspace. For example, with gene expression data, the
-effective dimensionality is thought to scale with the number of gene
-regulatory modules, not the number of genes themselves
-\citep{Heimberg2016}.  As such, differences in gene expression across
-groups should be expressed in terms of differences between these
-regulatory modules rather than strict differences between expression
-levels.  Such differences can be examined on a subspace that reflects
-the correlations resulting from these modules.
-
-Motivated by this notion, in this paper we propose a multi-group
-covariance estimation model by sharing information about the subspace
-spanned by group-level eigenvectors.  The shared subspace assumption
-can be used to improve estimates and facilitate the interpretation of
-differences between covariance matrices across groups.  For each
-group, we assume ``the spiked covariance model'', a well studied
-variant of the factor model \citep{Johnstone2001}.  In Section
-\ref{sec:shared} we briefly review the behavior of spiked covariance
-models for estimating a single covariance matrix before introducing
-our extension to the multi-group setting.  In Section
-\ref{sec:inference} we describe an optimization algorithm for
-inferring the shared subspace and then outline a MCMC algorithm for
-estimating the posterior distribution of the projected covariance
-matrices.  Finally, in Section \ref{sec:simulation} we investigate the
-behavior of these models in simulation before demonstrating the
-utility of this model with real biological data in Section
-\ref{sec:app}.
+high-dimensional multivariate data is partial least squares regression
+(PLS). This approach, which is a special case of a bilinear factor
+model, involves projecting the data into a lower dimension space which
+maximizes the similarity of the two groups.  This technique does not
+require the data from each group to share the same feature set.  An
+common variant for prediction, partial least squares discriminant
+analysis (PLS-DA) is especially common in chemometrics and
+bioinformatics \citep{Barker2003} .  Although closely related to the
+approaches we will consider here, the primarily goal of of PLS-based
+models is to create regression or discrimination models, not to
+explicitly infer covariance matrices from multiple groups of data.
+Nevertheless, the basic idea that data can often be well represented
+on a low dimensional space is an appealing one that we leverage as
+well.
+
+In this paper we propose a multi-group covariance estimation model by
+sharing information about the subspace spanned by group-level
+eigenvectors.  The shared subspace assumption can be used to improve
+estimates and facilitate the interpretation of differences between
+covariance matrices across groups.  For each group, we assume ``the
+spiked covariance model'', a well studied variant of the factor model
+\citep{Johnstone2001}.  In Section \ref{sec:shared} we briefly review
+the behavior of spiked covariance models for estimating a single
+covariance matrix and then introduce our extension to the multi-group
+setting.  In Section \ref{sec:inference} we describe an optimization
+algorithm for inferring the shared subspace and then outline a MCMC
+algorithm for estimating the posterior distribution of the projected
+covariance matrices.  In Section \ref{sec:simulation} we
+investigate the behavior of these models in simulation and then
+demonstrate the utility of the model with real biological data in
+Section \ref{sec:app}.
 
@@
 
-\paragraph{The Spiked Covariance Model:}
+%%\paragraph{The Spiked Covariance Model:}
 
@@
 \begin{equation} 
-p(\bl S | \Sigma, n) \propto_{\Sigma} l(\Sigma:\bl S) =  |\Sigma|^{-n/2} \text{etr}( - \Sigma^{-1} \bl S/2 ) ,  
+p(\bl S | \Sigma, n) \propto l(\Sigma:\bl S) =  |\Sigma|^{-n/2} \text{etr}( - \Sigma^{-1} \bl S/2 ) ,  
 \label{eqn:lik}
@@
 the design matrix.  The spiked principle components model (spiked PCA)
-introduced by \citet{Johnstone2001} assumes that
+studied by \citet{Johnstone2001} and others assumes that
 \begin{equation} 
@@
 %
-\noindent where $\Lambda$ is an $r\times r$ diagonal matrix and
-$\bl U \in \mathcal V_{r,p}$, with $r \ll p$. Here $\mathcal V_{r,p}$,
-known as the Stiefel manifold, comprises the set of $p \times r$
-orthonormal matrices in $\mathcal{R}^p$.  The spiked covariance model
-is thus a low rank factor model with homoscedastic error.  Such a
-formulation is appealing because it explicitly partitions the
-covariance matrix into a tractable low rank ``signal'' and isotropic
-``noise''.
+\noindent where for $r \ll p$, $\Lambda$ is an $r\times r$ diagonal
+matrix and $\bl U \in \mathcal V_{r,p}$.  $\mathcal V_{r,p}$ is the
+Stiefel manifold, which consists of the set of $p \times r$
+orthonormal matrices in $\mathbb{R}^p$.  % The spiked covariance model
+% is thus a low rank factor model with homoscedastic error. 
+The spiked covariance formulation is appealing because it explicitly
+partitions the covariance matrix into a tractable low rank ``signal''
+and isotropic ``noise''.
 
@@
 Classical results for parametric models (e.g., \citet{Schwartz1965})
-imply that asymptotically in $n$ for fixed $p$, a Bayes estimator will
+imply that asymptotically in $n$ for fixed $p$, an estimator will
 be consistent for a spiked population covariance as long as the
-assumed number of spikes (non-unit eigenvalues) is greater than or
+assumed number of spikes (eigenvalues larger than $\sigma^2$) is greater than or
 equal to the true number.  However, when $p$ is large relative to $n$,
@@
 difficult.  Under the spiked covariance model, it has been shown that
-$p/n \rightarrow \gamma >0$ as $n\rightarrow \infty$, the top $r$
-eigenvalues of $\bl S/(n\sigma^2)$ will converge to upwardly biased
-versions of $(\lambda_1+1,\ldots, \lambda_r+1)$, if each $\lambda_s $
+if $p/n \rightarrow \gamma >0$ as $n\rightarrow \infty$, the $k$th largest
+eigenvalue of $\bl S/(n\sigma^2)$ will converge to an upwardly biased
+version of $\lambda_{k}+1$, if $\lambda_k$
 is greater than $\sqrt{\gamma}$ \citep{Baik2006, Paul2007} .  This has
@@
 the eigenvalues of the sample covariance matrix. In particular, in the
-setting that $\sigma^2$ is known, \citet{Donoho2013} propose setting
-all sample eigenvalues that fall below the value
-$\sigma^2(1+\sqrt{\gamma})^2$ to be $\sigma^2$, and shrinking the
+setting that $\sigma^2$ is known, \citet{Donoho2013} propose estimating
+all eigenvalues whose sample estimates are smaller than 
+$\sigma^2(1+\sqrt{\gamma})^2$ by $\sigma^2$, and shrinking the
 larger eigenvalues in a way that depends on the particular loss
@@
 
-Many eigenvalue shrinkage estimators can be viewed in the Bayesian
-context as derivative of a Bayes estimator under a particular prior.
-Note that in the absence of real prior information, invariance
-considerations suggest the use of priors on $\sigma^2$ and $\bl U$
-that lead to equivariant estimators.  For instance, an (improper)
-prior having density $p(\sigma^2,\bl U) = 1/\sigma^2 $, with respect
-to the product of Lebesgue measure on $\mathbb R^+$ and the uniform
-probability measure on $\mathcal V_{r,p}$, leads to Bayes estimators
-$\hat \Sigma(\bl S)$ that are equivariant with respect to rotations
-and scale changes.  For multi-group covariance estimation, however, we
-often have reason to believe that a priori the eigenvectors are
-related across groups.  In this case, group-level equivariance to
-rotations is an unreasonable assumption.  Here, both eigenvalue
-\emph{and} eigenvector shrinkage can play an important role in
-improving covariance estimates.  Below, we introduce our extension of
-the spiked covariance model to the group setting by including an
-assumption about the space shared by the span of the group-level eigenvectors.
-
-\paragraph{The Shared Subspace Model:}
+% Many eigenvalue shrinkage estimators can be viewed in the Bayesian
+% context as derivative of a Bayes estimator under a particular prior.
+
+%\paragraph{The Shared Subspace Model:}
 
@@
       data projected onto the two-dimensional shared subspace,
-      $Y_kV$, have covariances $\psi_k$ that differ between
+      $Y_kV$, have covariances $\Psi_k$ that differ between
       groups. c) The orthogonal projection, $Y_kV_{\perp}$
@@
 
-The spiked covariance model assumes that the goal is to estimate a
-single covariance matrix.  The situation is different when we are
+When estimating a single covariance matrix, the absence of prior
+knowledge suggests estimators that are equivariant with respect to
+rotations and scale changes.  The situation is different when we are
 interested in estimating multiple covariance matrices from distinct,
-but related groups.  For multi-group estimation, we can often improve
-upon the efficiency of covariance estimators by pooling information
-across the groups of data.  Here, we assume we have $k=1, ..., K$
-groups of mean-zero normal data, $Y_k$, where $Y_k$ consists of $n_k$
-independent rows of measurements on $p$ features, typically with
-$n_k \ll p$.  Then, $Y_k^TY_k = \bl S_k$ follow a (degenerate) Wishart
-distribution as in Equation \ref{eqn:lik}.  To improve estimation, we
-seek estimators for the population covariance from one group that depend on the
-data from all $K$ groups, e.g. of the form
-$\hat \Sigma_k(\bl S_1, ... \bl S_K)$.  To this end, we extend the
-spiked covariance model to grouped data, by assuming that the
-anisotropic variability from each group can be expressed on a common
-low dimensional subspace.  Specifically, we posit that the covariance
+but related groups.  Here, group-level equivariance to rotations is an
+unreasonable assumption; both eigenvalue \emph{and} eigenvector
+shrinkage can play an important role in improving covariance
+estimates.
+
+For multi-group covariance estimation, we assume that we have $K$
+matrices, $Y_1, ... Y_K$, where $Y_k$ conists of where $Y_k$ consists
+of $n_k$ independent rows of mean-zero normal data of dimension $p$,
+typically with $n_k \ll p$.  Then, $Y_k^TY_k = \bl S_k$ has a
+(degenerate) Wishart distribution as in Equation \ref{eqn:lik}.  To
+improve estimation, we seek estimators for the population covariance
+from one group that depend on the data from all $K$ groups, e.g. of
+the form $\hat \Sigma_k(\bl S_1, ... \bl S_K)$.  To this end, we
+extend the spiked covariance model to grouped data, by assuming that
+the anisotropic variability from each group occurs on a common low
+dimensional subspace.  Specifically, we posit that the covariance
 matrix for each group can be written as
@@
 \begin{equation}
-\Sigma_k = V\psi_kV^T + \sigma^2_kI
+\Sigma_k = V\Psi_kV^T + \sigma^2_kI.
 \label{eqn:sspsi}
@@
 \citep{Chikuse2012}.  Although $V$ is only identifiable up to right
-rotations, $VV^T$, which defines the plane of variation shared by all
+rotations, the matrix $VV^T$, which defines the plane of variation shared by all
 groups, is identifiable.  Later, to emphasize the connection to the
-spiked PCA model (\ref{eqn:spiked}), we will write $\psi_k$ in terms
-of its eigendecomposition, $\psi_k = O_k\Lambda_kO_k$, where $O_k$ are
-eigenvectors and $\Lambda_k$ are the eigenvalues of $\psi_k$ (see
+spiked PCA model (\ref{eqn:spiked}), we will write $\Psi_k$ in terms
+of its eigendecomposition, $\Psi_k = O_k\Lambda_kO_k$, where $O_k$ are
+eigenvectors and $\Lambda_k$ are the eigenvalues of $\Psi_k$ (see
 Section \ref{sec:bayes}).
@@
 Note that for the shared subspace model,
-$V^T\Sigma_kV = \psi_k + \sigma_k^2I$ is an anisotropic
+$V^T\Sigma_kV = \Psi_k + \sigma_k^2I$ is an anisotropic
 $s$-dimensional covariance matrix for the projected data, $Y_kV$.
 In contrast, the data projected onto the orthogonal space,
-$Y_kV_{\perp}$, is isotropic, and typically small, for all
+$Y_kV_{\perp}$, is isotropic, and assumed to be small, for all
 groups.  In Figure \ref{fig:shared} we provide a simple illustration
-of this fact using simulated $4$-dimensional data from two groups.  The
+using simulated $4$-dimensional data from two groups.  In this example, the
 differences in distribution between the groups of data can be
@@
 constrained to the isotropic variability from each group.  This
-isotropic variability is often taken to be ``measurement noise''.
+isotropic variability is often characterized as measurement noise.
 
@@
 
-\section{Inference}
+\section{Empirical Bayes Inference}
 \label{sec:inference}
-In this section we outline Bayesian inference algorithms including
-both an optimization algorithm and an MCMC algorithm for estimating
-covariance parameters and their associated uncertainty.  MCMC
-inference is particularly appealing because of the ease with which we
-can quantify the uncertainty associated with each estimate.
+
+In this section we outline our empirical Bayesian approach for estimating
+covariance parameters and their associated uncertainty.  
 Uncertainty quantification is particularly important in the $n \ll p$
@@
 subspace. In Section \ref{sec:bayes} we describe a Gibbs sampler for
-posterior inference of $\psi_k$ given the common subspace.  These two
+posterior inference of $\Psi_k$ given the common subspace.  These two
 algorithms can be used in conjunction to first identify and evaluate a
@@
 
+TODO: emphasize 1) justification 2) objective of \emph{identifying} this
+subspace when p much greater than n.
+
 \subsection{Estimating the  Shared Subspace}
@@
 
-In this section we describe a computationally efficient algorithm for
-estimating the shared subspace, $VV^T$, the projected covariance matrix
-$\psi_k$ and noise parameter $\sigma_k^2$.  First, note that the full
-likelihood for the shared subspace model can be written as
+In this section we describe a maximum marginal likelihood procedure for
+estimating the shared subspace, $VV^T$ based on the
+expectation-maximization algorithm.  The full likelihood
+for the shared subspace model can be written as
 
 \begin{align}
-p(S_1, ... S_k | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}\etr(-\Sigma_k^{-1}\mathbf{S}_k/2)  \\
-&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-(\sigma_k^2(V\psi_kV^T +
+\nonumber p(S_1, ... S_k | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}\etr(-\Sigma_k^{-1}\mathbf{S}_k/2)  \\
+\nonumber &\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-(\sigma_k^2(V\Psi_kV^T +
   I))^{-1}\mathbf{S}_k/2) \\
-&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-\left[V(\psi_k +
+\nonumber &\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-\left[V(\Psi_k +
   I)^{-1}/\sigma_k^2 V^T + (I-VV^T)/\sigma^2_k\right]\mathbf{S}_k/2)
   \\
-&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-s)/2}|M_k|^{-n_k/2}\etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]\mathbf{S}_k/2) 
+&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-s)/2}|M_k|^{-n_k/2}\etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]\mathbf{S}_k/2) ,
 \end{align}
 %
-\noindent where we define $M_k^{-1} = (\psi_k + I) ^{-1}/\sigma_k^2$.  The log-likelihood in
-$V$ (up to additive constant) is
+\noindent where we define $M_k^{-1} = (\Psi_k + I) ^{-1}/\sigma_k^2$.  The log-likelihood in
+$V$ (up to an additive constant) is
 
 \begin{align}
-l(V) &= \sum_k tr\left(-VM_k^{-1}V^T +
+\nonumber l(V) &= \sum_k \tr\left(-VM_k^{-1}V^T +
        VV^T/\sigma^2_k\right)\mathbf{S}_k/2)\\
-&= \sum_k \frac{1}{2}tr\left(-M_k^{-1}V^T \mathbf{S}_kV\right) + \frac{1}{2\sigma_k^2}tr\left(VV^T \mathbf{S}_k\right)
+&= \sum_k \frac{1}{2}\tr\left(-M_k^{-1}V^T \mathbf{S}_kV\right) + \frac{1}{2\sigma_k^2}tr\left(VV^T \mathbf{S}_k\right).
 \end{align}
@@
 
-\noindent which is a Wishart($V^T\mathbf{S}_kV$, $n_k$).  The
+\noindent which is a Wishart($V^T\mathbf{S}_kV$, $n_k$) distribution.  The
 conditional posterior distribution of $\frac{1}{\sigma_k^2}$ is simply
@@
 
-\noindent which is a Gamma($n_k(p-r)/2$,
-$\frac{1}{2}tr((I-VV^T)\mathbf{S}_k)$).  Thus, for each iteration,
-$t$, of the EM algorithm:
+\noindent which is a gamma($n_k(p-r)/2$,
+$\frac{1}{2}tr((I-VV^T)\mathbf{S}_k)$) distribution.  Based on these
+results the following EM algorithm can be used.
 
@@
 \end{enumerate}
-\item Compute $V_{t}$ = $\underset{V}{argmax}  \sum_k tr\left(-VE[M^{-1}|V_{(t-1)}]^T +
+\item Compute $V_{t}$ = $\underset{V}{\text{arg } \text{max}}  \sum_k tr\left(-VE[M^{-1}|V_{(t-1)}]^T +
        E[\frac{1}{2\sigma_k^2}|V_{(t-1)}]VV^T\right)\mathbf{S}_k/2)$ 
@@
 
-\paragraph{Evaluating a Shared Subspace:}
-
-The above EM algorithm provides an estimate of a shared subspace, but
-in practice there are no guarantees that the inferred subspace is
-appropriate for all groups.  Moreover, the shared subspace can be
-chosen in other ways, including more ad hoc methods involving the
-singular value decomposition or simply by incorporating subject matter
-considerations.  Regardless of how a particular subspace is specified,
-it is important to quantify how useful this subspace is for explaining
-the variability in each group.  In particular, if $V$ is a basis for
-the shared subspace, then for each $k$, most of the variation in the
-data should be preserved when projecting the data to this subspace.
-To quantify the extent to which this is true for different groups, we
+\paragraph{Evaluating Goodness of Fit:}
+
+It is important to quantify how much of the variability in each group can be
+explained by an inferred subspace.  If $V$ is a basis for
+the shared subspace, then for each group $k$, most of the variation in
+$Y_k$ should be preserved when projecting the data to this subspace.
+To characterize the extent to which this is true for different groups, we
 propose a simple estimator for the proportion of ``signal'' variance
-that lies on any given subspace.  Specifically, for any inferred
-subspace $\hat{V}$, we apply the following estimator for ratio of
-  the sum of the first $s$ eigenvalues of
-  $\hat{V}^T\bl \Sigma_k \hat{V}$ to the sum of the first $s$
-  eigenvalues of $\Sigma_k$:
+that lies on any given subspace.  Specifically, for an estimated
+subspace $\hat{V}$, we apply the following statistic for the ratio of the
+sum of the first $s$ eigenvalues of $\hat{V}^T\bl \Sigma_k \hat{V}$ to
+the sum of the first $s$ eigenvalues of $\Sigma_k$:
 %
 \begin{equation}
- \frac{||Y_k\hat{V}||_F}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{sup}
+r(Y_k: \hat{V}, \hat{\sigma}_k^2) = \frac{||Y_k\hat{V}||_F}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
   ||Y_k\widetilde{V}||_F - \hat{\sigma}_k^2ps}
@@
 %
-\noindent where $||.||_F$ is the Frobenius norm. If the inferred
-subspace is in fact the true subspace, i.e. $\hat{V}\hat{V}^T = VV^T$,
-then this estimator will asymptotically approach one.  To see this,
-note that
-$\underset{V \in \mathcal{V}_{p, s}}{sup}\frac{||Y_kV||_F}{n_k}$ is
-equivalent to the sum of the first $s$ eigenvalues of the sample
-covariance matrix $\frac{\bl S_k}{n_k}$.  With $\hat{\lambda}^{(k)}_i$
+\noindent where $||.||_F$ is the Frobenius norm. Our motivation for
+this statisic is the fact that if
+$\Sigma_k = \hat{V}\Psi_k\hat{V}^T + \sigma_k^2I$ then
+$r(Y_k: \hat{V}, \hat{\sigma}_k^2) \rightarrow 1$.  To see this, note
+that $\underset{V \in \mathcal{V}_{p, s}}{max}\frac{||Y_kV||_F}{n_k}$
+is equivalent to the sum of the first $s$ eigenvalues of the sample
+covariance matrix $\bl S_k/n_k$.  With $\hat{\lambda}^{(k)}_i$
 the $i$-th eigenvalue of $\frac{\bl S_k}{n_k}$, \citet{Baik2006}
@@
 $||Y_kV||_F/n_k$ is a consistent estimator for
-$\sum_i^s \lambda^{(k)}_i$.  As a consequence, the proposed estimator will be close
-to one when $\hat{V}\hat{V}^T = VV^T$ and smaller if not.  The metric provides a
-useful indicator for which groups can be reasonably compared on a
-given subspace and which groups cannot.  In Section \ref{sec:simulation} we plot
-these estimates for different simulated datasets.  In Section \ref{sec:app}
-we plot the estimates for subspaces  inferred with real biological data.
-
-\subsection{Uncertainty Quantification}
+$\sum_i^s \lambda^{(k)}_i$.  As a consequence, the proposed estimator
+will be close to one for all groups $k$ when $\hat{V}\hat{V}^T = VV^T$
+and smaller if not.  The metric can provides a useful indicator for
+which groups can be reasonably compared on a given subspace and which
+groups cannot.  When $r(Y_k: \hat{V}, \hat{\sigma})k^2)$ is small for
+some groups, it may suggest that the rank, $s$, of the inferred
+subspace needs to be larger to capture the variation in all
+groups. The use of this statistic is illustrated in Section
+\ref{sec:simulation}, where we plot these estimates for different
+simulated datasets, and in Section \ref{sec:app}, where we plot the
+estimates for subspaces inferred with real biological data.
+
+\subsection{Inference for Projected Covariance Matrices}
 \label{sec:bayes}
@@
 Although the EM algorithm presented in the previous section yields
-point estimates for $VV^T$ and $\psi_k$, it does not lead to natural
-uncertainty bounds for these estimates. Full uncertainty
-quantification for $VV^T$ may be desirable but we found this
-challenging for the dimensions of the problems we would like to
-consider.  As such, in this section, we assume that the subspace
-$VV^T$ is fixed and known and focus on uncertainty quantification for
-$\psi_k$ only.  Further, by focusing on group differences on a fixed
-subspace, it is simpler to interpret differences across covariance
-matrices.  %The shared subspace can be an estimated using the EM
-% algorithm presented in Section \ref{sec:em} or chosen based on
-% subject matter considerations.  
-Note that when the subspace is known, the posterior distribution of
-$\Sigma_k$ is conditionally independent from the other groups, i.e.
-$P(\Sigma_{k} | V, \bl S_k, \Sigma_{-k}) = P(\Sigma_{k} | V, \bl
-S_k)$,
-so that we can independently estimate the conditional posterior
-distribution for each group.
-
-There are many different prior specifications for $\psi_k$ and
-$\sigma_k^2$ that we could consider for estimating posterior
-uncertainty.  One simple approach would be to use the same prior
-specification used in Section \ref{sec:em}.  That is, we can use
-Jeffreys' prior for $\frac{1}{\sigma_k^2}$ and
-$M_k^{-1} = (\psi_k + I) ^{-1}/\sigma_k^2$.  In fact, it is easy to
-show that the Gamma distribution for $\frac{1}{\sigma_k^2}$ and the
-Wishart distribution for $M_k^{-1}$ are conjugate priors for the
-respective quantities.  From posterior samples of $M_k$ and
-$\sigma_k^2$ we can easily recover samples of $\psi_k$.
-Unfortunately, if $M_k^{-1}$ has a Wishart distribution, the implied
-support of $\psi_k$ is not constrained to the cone of positive
-semi-definite matrices.  Direct prior specifications for $\psi_k$ are
-more appealing, but many obvious choices, like the inverse-Wishart
-prior distribution for $\psi_k$ do not lead to simple posterior
-distributions.  As such, building on recent interest in spiked
-covariance models \citep{Donoho2013, Paul2007} we focus on
-tractable models based on the eigendecomposition of $\psi_k$.  Such
-models are especially appealing because it is simple to summarize and
-compare $\psi_k$ across groups in terms their eigenstructure (See
-Section \ref{sec:simulation} and \ref{sec:app}).
+point estimates for $VV^T$ and $\Psi_k$, it does not lead to natural
+uncertainty for these estimates. As such, in this section, we assume
+that the subspace $VV^T$ is fixed and known and focus on uncertainty
+quantification for $\Psi_k$ only.  Note that when the subspace is
+known, the posterior distribution of $\Sigma_k$ is conditionally
+independent from the other groups, so that we can independently
+estimate the conditional posterior distribution for each group.
+
+There are many different prior specifications for $\Psi_k$ and
+$\sigma^2_k$ that we could consider for estimating the posterior
+uncertaity.  In this paper, we build on recent interest in the spiked
+covariance model \citep{Donoho2013, Paul2007} and develop tractable
+models for the eigendecomposition of $\Psi_k$.  Such models are
+especially appealing because it is simple to summarize and compare
+$\Psi_k$ across groups in terms their eigenstructure (See Section
+\ref{sec:simulation} and \ref{sec:app}).
 
 By modeling the eigenstructure, we can now view each covariance
-$\Sigma_k$ in terms of the spiked principle components model.
+$\Sigma_k$ in terms of the spiked principle components model (\ref{eqn:spiked}).
 Equation \ref{eqn:sspsi}, written as a function $V$, becomes
 
-\begin{equation}
-\Sigma_k = VO_k\Lambda_KO_k^TV^T + \sigma^2_kI
+\begin{align}
+\nonumber \Psi_k &= O_k\Lambda_KO_k^T\\
+\Sigma_k &= VO_k\Lambda_KO_k^TV^T + \sigma^2_kI
 \label{eqn:ss}
-\end{equation}
+\end{align}
 %
 \noindent Here, we make one further generalization to allow
-$\psi_k = O_k\Lambda_KO_k^T$ be a rank $r \leq s$ dimensional covariance
+$\Psi_k$ to be a rank $r \leq s$ dimensional covariance
 matrix on the $s$-dimensional subspace.  Thus, $\Lambda_k$ is an $r \times
@@
 \mathcal{V}_{s,r}$ is the matrix of eigenvectors of
-$\psi_k$.  For any individual group, this corresponds to the original spiked
+$\Psi_k$.  For any individual group, this corresponds to the original spiked
 PCA model (Equation \ref{eqn:spiked}) with $U_k = VO_k \in
-\mathcal{V}_{p, r}$.
+\mathcal{V}_{p, r}$.    This ``double spiked'' covariance model TODO
 
@@
 \begin{align}
-\Sigma^{-1}_k &=  (\sigma_k^2(U_k\Lambda_kU_k^T+I))^{-1}\\
-&= \frac{1}{\sigma_k^2}(U_k\Lambda_kU_k^T+I)^{-1}\\
-&= \frac{1}{\sigma_k^2}(I-U_k\Omega_kU_k^T)\\
+\nonumber \Sigma^{-1}_k &=  (\sigma_k^2(U_k\Lambda_kU_k^T+I))^{-1}\\
+\nonumber &= \frac{1}{\sigma_k^2}(U_k\Lambda_kU_k^T+I)^{-1}\\
+&= \frac{1}{\sigma_k^2}(I-U_k\Omega_kU_k^T),
 \end{align}
@@
 \begin{align}
-|\Sigma_k| &= (\sigma_k^2)^{p}|U_k\Lambda_kU_k^T+I|\\
-&= (\sigma_k^2)^{p}|\Lambda_k+I| \\
-&= (\sigma_k^2)^{p}\prod_{i=1}^r(\lambda_i+1)\\
-&= (\sigma_k^2)^{p}\prod_{i=1}^r(1-\omega_i)
+\nonumber |\Sigma_k| &= (\sigma_k^2)^{p}|U_k\Lambda_kU_k^T+I|\\
+\nonumber &= (\sigma_k^2)^{p}|\Lambda_k+I| \\
+\nonumber &= (\sigma_k^2)^{p}\prod_{i=1}^r(\lambda_i+1)\\
+&= (\sigma_k^2)^{p}\prod_{i=1}^r(1-\omega_i),
 \end{align}
@@
    \right) ^{n_k/2}
-   \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k)
+   \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k).
 \label{eqn:sslik}
@@
 $\sigma^2_k$ and $\Omega_k$ to facilitate inference via a Gibbs
-sampling algorithm.  In the absence of real prior information,
+sampling algorithm.  In the absence of specific prior information,
 invariance considerations suggest the use of priors that lead to
-equivariant estimators.  Below we describe our choices for the prior
+invariant estimators.  Below we describe our choices for the prior
 distributions of each parameter and the resultant conditional posterior
@@
 data each with sample size $n_k=50$.  We choose the eigenvectors of
-the anisotropic component of the covariance matrix, $\psi_k$, to share
+the anisotropic component of the covariance matrix, $\Psi_k$, to share
 a common two dimensional subspace.  To fit the model to this simulated
@@
 Section \ref{sec:bayes} to estimate the posterior distribution of the
-projected covariance matrices $\psi_k$ and noise variance $\sigma_k^2$.
+projected covariance matrices $\Psi_k$ and noise variance $\sigma_k^2$.
 
@@
     \caption{a) 95\% posterior regions for the ratio of the
-      eigenvalues, $\frac{\lambda_1}{\lambda_2}$, of $\psi_k$ and the
+      eigenvalues, $\frac{\lambda_1}{\lambda_2}$, of $\Psi_k$ and the
       orientation of the principle axis on space spanned by $\hat{V}$
@@
       orientations were $\pi/4$ (black), $-\pi/4$ (red and green) and
-      0 (blue and cyan).  b) Estimated proportion of variance
-      explained by the subspace (Equation \ref{eqn:ratio}).  As
-      expected, for each group, the inferred subspace explains
-      essentially all of the variance from the first two eigenvectors of
-      $\Sigma_k$.  }
+      0 (blue and cyan).  b) Goodness of fit,
+      $r(Y_k: \hat{V}, \hat{\sigma}^2_k$ (Equation \ref{eqn:ratio}).
+      As expected, for each group, the inferred subspace explains
+      practically all of the variance from the first two eigenvectors
+      of $\Sigma_k$.  }
 \end{figure}
@@
 For this two dimensional shared subspace model we summarize the
-estimates of $\psi_k$ in terms of its eigendecomposition.
+estimates of $\Psi_k$ in terms of its eigendecomposition.
 Specifically, we compute the eigenvalue ratio
@@
         \includegraphics[width=\textwidth]{Figs/LossVsDimension}
-        \caption{Stein's Loss vs s}
+        \caption{Stein's risk vs s}
         \label{fig:sdimension}
@@
     \end{subfigure}
-    \caption{a) Average Stein's loss as a function of the dimension of
-      the shared subspace dimension.  Data from ten groups, $U_k$
-      uniform on the Stiefel manifold $\mathcal{V}_{200, 2}$.  As the
-      shared subspace dimension increases, $s \rightarrow p$, the loss
+    \caption{a) Stein's risk as a function of the dimension of the
+      shared subspace dimension.  Data from ten groups, $U_k$ uniform
+      on the Stiefel manifold $\mathcal{V}_{200, 2}$.  As the shared
+      subspace dimension increases, $s \rightarrow p$, the loss
       converges to the loss from the independent spiked covariance
-      estimates (dashed line). b) For a simulated dataset, estimated
-      fraction of variance explained when the assumed shared subspace
-      is dimension 5.  c).  For a single simulated dataset, estimated
-      fraction of variance explained when the assumed shared subspace
-      is dimension 20.  We can capture nearly all of the variability
-      in each of the 10 groups using an $s=rK=20$ dimensional shared
-      subspace. }
+      estimates (dashed line). b) For a simulated dataset, the
+      goodness of fit statistic, $r(Y_k: \hat{V}, \hat{\sigma_k}^2$,
+      when the assumed shared subspace is dimension 5.  c).  For a
+      single simulated dataset, goodness of fit when the assumed
+      shared subspace is dimension 20.  We can capture nearly all of
+      the variability in each of the 10 groups using an $s=rK=20$
+      dimensional shared subspace. }
 \label{fig:dimensionPlots}
@@
 shared subspace model, we additionally assume that the two
-eigenvectors of $\psi_k$ span a common two dimensional subspace, e.g. we fix
+eigenvectors of $\Psi_k$ span a common two dimensional subspace, e.g. we fix
 $s=2$.  We simulate 100 independent datasets for each data generating
@@
 posterior distribution of the projected covariance matrices
-conditional on $\hat{V}\hat{V}^T$.  In Table \ref{table:groupLoss} we report the
-average Stein's loss, with corresponding 95\% intervals, for the
-estimates derived from each inferential model.
+conditional on $\hat{V}\hat{V}^T$.  In Table \ref{table:groupLoss} we
+report the average Stein's risk, with corresponding 95\% loss
+intervals, for the estimates derived from each inferential model.
 
@@
   \end{tabular}
-  \caption[Table caption text]{Stein's loss (and 95\% intervals) for
+  \caption[Table caption text]{Stein's risk (and 95\% loss intervals) for
     different inferential and data generating models.  For each of
@@
 
-As expected, the estimates with the lowest loss are derived from the
+As expected, the estimates with the lowest risk are derived from the
 inferential model that matches the data generating model. However, the
-shared subspace estimator has small loss under model misspecification
+shared subspace estimator has small risk under model misspecification
 relative to the alternative.  For example, when the data is generated
 using the complete pooling model, the shared subspace estimator has
-almost four times smaller loss than the no pooling estimator.  When
+almost four times smaller risk than the no pooling estimator.  When
 the data is generated under the no pooling model, the shared subspace
@@
 typically estimated $s$ to be slightly smaller than $rK$ for the ``no
-pooling'' data, which explains the slightly higher loss for the shared
+pooling'' data, which explains the slightly higher risk for the shared
 subspace model. With improved estimators for $s$, the shared subspace
@@
 In these analyses we employ a variation of the shared
-subspace model.  Specifically, we define the $s$ by $s$ matrix $\psi_k$ as
+subspace model.  Specifically, we define the $s$ by $s$ matrix $\Psi_k$ as
 
 \begin{equation}
-\psi_k =\left( \begin{array}{cc}
+\Psi_k =\left( \begin{array}{cc}
 O_k\Lambda_kO_k^T & 0  \\
@@
 magnitude of the eigenvalues $\Lambda_k$ relative to the eigenvalues
-$D$, so that the largest eigenvectors of $\psi_k$ may correspond to
+$D$, so that the largest eigenvectors of $\Psi_k$ may correspond to
 the eigenvectors $V_2$.
@@
 eigenvectors between groups.  The $x$-axis corresponds to the
-orientation of the principle axis of $\psi_k$ on $V$ and the $y$-axis
+orientation of the principle axis of $\Psi_k$ on $V$ and the $y$-axis
 corresponds to the ratio of the larger eigenvalue over the smaller
@@
 It was also a challenge to estimate the subspace $VV^T$ once we had
-chosen a rank.  In our approach, we estimate the shared subspace
-$VV^T$ separately from the MCMC inference of $\psi_k$ because we found
+chosen a rank.  We used an empirical bayes approach, we estimate the shared subspace
+$VV^T$ separately from the MCMC inference of $\Psi_k$ because we found
 estimation of $VV^T$ using MCMC inference to be
-particularly difficult.  Although fixing the shared subspace
+particularly difficult.
+
+Full uncertainty
+quantification for $VV^T$ may be desirable but we found this
+challenging for the dimensions of the problems we would like to
+consider.  Further, by focusing on group differences on a fixed
+subspace, it is simpler to interpret differences across covariance
+matrices.
+
+ Although fixing the shared subspace
 simplifies interpretation of the results, in many cases it may be
 desirable to develop a fully Bayesian approach for jointly estimating
-$VV^T$ and $\psi_k$.  Recent Markov chain Monte Carlo algorithms, like
+$VV^T$ and $\Psi_k$.  Recent Markov chain Monte Carlo algorithms, like
 Riemmanian manifold Hamilton Monte Carlo, which can exploit the
@@
 
+% There are many different prior specifications for $\Psi_k$ and
+% $\sigma_k^2$ that we could consider for estimating posterior
+% uncertainty.  One simple approach would be to use the same prior
+% specification used in Section \ref{sec:em}.  That is, we can use
+% Jeffreys' prior for $\frac{1}{\sigma_k^2}$ and
+% $M_k^{-1} = (\Psi_k + I) ^{-1}/\sigma_k^2$.  In fact, it is easy to
+% show that the Gamma distribution for $\frac{1}{\sigma_k^2}$ and the
+% Wishart distribution for $M_k^{-1}$ are conjugate priors for the
+% respective quantities.  From posterior samples of $M_k$ and
+% $\sigma_k^2$ we can easily recover samples of $\Psi_k$.
+% Unfortunately, if $M_k^{-1}$ has a Wishart distribution, the implied
+% support of $\Psi_k$ is not constrained to the cone of positive
+% semi-definite matrices.  Direct prior specifications for $\Psi_k$ are
+% more appealing, but many obvious choices, like the inverse-Wishart
+% prior distribution for $\Psi_k$ do not lead to simple posterior
+% distributions.
+
 Fundamentally, our approach is quite general and can be integrated
@@
 estimation.  In particular, we can incorporate additional shrinkage on
-the projected covariance matrices $\psi_k$.  As in \citet{Hoff2009} we
+the projected covariance matrices $\Psi_k$.  As in \citet{Hoff2009} we
 can employ non-uniform Bingham prior distributions for the
-eigenvectors of $\psi_k$ or model $\psi_k$ as a function of continuous
+eigenvectors of $\Psi_k$ or model $\Psi_k$ as a function of continuous
 covariates as in \citet{Yin2010, Hoff2011}.  Alternatively, we can
 summarize the estimated covariance matrices by thresholding entries of
-the precision matrix, $\psi_k^{-1}$ to visualize differences between
+the precision matrix, $\Psi_k^{-1}$ to visualize differences between
 groups using a graphical model \citet{Meinshausen2006}.  The specifics

### 2016-06-20 bayes-coverage.R
compute coverage

@@ -1,0 +1,133 @@
+rm(list=ls())
+
+## Need these for Rscript
+library("methods")
+library("utils")
+library(rstiefel)
+library(sp)
+
+source("fit-subspace.R")
+source("generateData.R")
+source("helper.R")
+
+########################################################
+############# Data Generation #########################
+########################################################
+
+n <- 50
+S <- 2
+R <- 2
+P <- 200
+ngroups <- 5
+
+niters <- 1000
+nwarmup <- niters/2
+
+LambdaList <- Olist <- list()
+
+LambdaList[[1]] <- c(100, 10)
+LambdaList[[2]] <- c(100, 10)
+LambdaList[[3]] <- c(100, 100/3)
+LambdaList[[4]] <- c(100, 100/3)
+LambdaList[[5]] <- c(100, 100)
+
+Olist[[1]] <- matrix(c(sqrt(2)/2, sqrt(2)/2, -sqrt(2)/2, sqrt(2)/2), nrow=2)
+Olist[[2]] <- matrix(c(-sqrt(2)/2, sqrt(2)/2, sqrt(2)/2, sqrt(2)/2), nrow=2)
+Olist[[3]] <- matrix(c(-sqrt(2)/2, sqrt(2)/2, sqrt(2)/2, sqrt(2)/2), nrow=2)
+Olist[[4]] <- diag(2)
+Olist[[5]] <- diag(2)
+
+nsim <- 100
+
+for(i in 1:nsim) {
+
+    dat <- generateData(P=P, S=S, R=2, ngroups=ngroups,
+                        LambdaList=LambdaList, Olist=Olist,
+                        nvec=rep(n, ngroups))
+
+    dat$genType <- "Shared subspace"
+
+    Ypooled <- c()
+    for(i in 1:length(dat$Ylist)){
+        Ypooled <- rbind(Ypooled, dat$Ylist[[i]])
+    }
+    dat$Ypooled <- Ypooled
+
+
+########################################################
+############# Data Fit #########################
+########################################################
+
+####### FIT DATA USING SHARED SUBSPACE ###############
+
+    ## Initialize sampler
+    ngroups <- dat$ngroups
+    nvec <- dat$nvec
+    Slist <- dat$Slist
+
+    Vinit <- svd(do.call(cbind, lapply(1:ngroups, function(k) svd(t(dat$Ylist[[k]]))$u[, 1:R])))$u[, 1:S]
[…]

### 2016-06-20 helper.R
Updates to plotPosterior function

@@ -1,7 +1,7 @@
-library(rstan)
 library(mvtnorm)
 library(Matrix)
 library(magrittr)
+library(scales)
 
 ## stanstart <- stan(file="vectorBMF.stan", data=list(M=9,lam=rep(1,10), gamma=rep(1, 10)), chains=1, iter=1)
 
@@ -69,13 +69,21 @@ getAgeStat <- function(normsMat) {
 
 }
 
-posteriorPlot <- function(Osamps, OmegaSamps, s2samps, nsamps, groups, probRegion=0.95, hline=NULL, col=NULL, pch=NULL) {
+posteriorPlot <- function(Osamps, OmegaSamps, s2samps, nsamps, groups, probRegion=0.95, hline=NULL, col=NULL, pch=NULL, ymax=30, plotPoints=TRUE) {
 
-
+    ngroups <- length(groups)
+    
+    if(is.null(col)){
+        col <- 1:ngroups
+    }
+    if(is.null(pch)){
+        pch=rep(19, ngroups)
+    }
+    par(mar=c(5.1, 5.1, 4.1, 2.1))
     plot(0, 0, xlim=c(-pi/2, pi/2),
-         ylim=c(0, 60), cex=0, xlab="Angle", ylab=expression(lambda[1]/lambda[2]))
-    if(!is.null(hline))
-        abline(h=hline, lty=2)
+         ylim=c(0, ymax), cex=0, xlab="Angle", ylab=expression(lambda[1]/lambda[2]), xaxt="n", cex.axis=1.5, cex.lab=1.5)
+    axis(1, at=seq(-pi/2, pi/2, by=pi/4), labels=expression(-pi/2, -pi/4, 0, pi/4, pi/2), cex.axis=1.5, cex.lab=1.5)
+
     
     
     for(g in groups) {
@@ -86,46 +94,59 @@ posteriorPlot <- function(Osamps, OmegaSamps, s2samps, nsamps, groups, probRegio
         pmValues <- eigPsi$values
         pmVectors <- eigPsi$vectors
         maxIndex <- which.max(pmValues)
+
         pmPoint <- pmValues[maxIndex]*pmVectors[, maxIndex]
         if(pmPoint[1] < 0)
             pmPoint <- pmPoint * c(-1, -1)
 
+        hp <- getHullPoints(nsamps, OmegaSamps[, g, ], Osamps[, , g, ])
+        pts <- hp$pts
+        hullPoints <- hp$hullPoints
[…]

### 2016-06-20 fit-subspace.R
Fixed bug for R=S

@@ -4,6 +4,7 @@
 fitSubspace <- function(P, S, R, Slist, nvec, ngroups=length(Slist),
                         niters=100, nskip=1, init=NULL, binaryO=FALSE,
                         verbose=TRUE, sigmaTruthList=NULL, draw=c(),
+                        saveU=FALSE,
                         printLoss=FALSE,
                         Vmode="hmc") {
     
@@ -12,7 +13,10 @@ fitSubspace <- function(P, S, R, Slist, nvec, ngroups=length(Slist),
     nskip <- nskip
     nsamps <- floor(niters/nskip)
 
-    Usamps <- array(dim=c(P, S, ngroups, nsamps))
+    if(saveU)
+        Usamps <- array(dim=c(P, S, ngroups, nsamps))
+    else
+        Usamps <- NULL
     Osamps <- array(dim=c(S, S, ngroups, nsamps))
     Vsamps <- array(dim=c(P, S, nsamps))
     omegaSamps <- array(dim=c(S, ngroups, ncol=nsamps))
@@ -92,13 +96,18 @@ fitSubspace <- function(P, S, R, Slist, nvec, ngroups=length(Slist),
 
         ## Sample common omega
         Uk <- Ulist[[1]]
+
+        if(R < S) {
+            Omega2 <- sampleOmega(Ssum, Uk[, (R+1):S],
+                                  1, sum(nvec[k]))
         
-        Omega2 <- sampleOmega(Ssum, Uk[, (R+1):S],
-                              1, sum(nvec[k]))
-        
-        ## Sample common O
-        O2 <- sampleO(Ssum, Uk[, (R+1):S], 1, OmegaList[[k]][(R+1):S],
-                      V[, (R+1):S])
+            ## Sample common O
+            O2 <- sampleO(Ssum, Uk[, (R+1):S], 1, OmegaList[[k]][(R+1):S],
+                          V[, (R+1):S])
+        } else {
+            Omega2 <- c()
+            O2 <- matrix(nrow=0, ncol=0)
+        }
         
         for ( k in 1:ngroups ) {
             
@@ -134,8 +143,9 @@ fitSubspace <- function(P, S, R, Slist, nvec, ngroups=length(Slist),
             
             ## save samples        
             if(i %% nskip==0) {
+                if(saveU)
+                    Usamps[, , k, i/nskip] <- Ulist[[k]]
 
[…]

### 2016-06-20 generateData.R
Q = S - R

@@ -11,11 +11,14 @@ library(mvtnorm)
 library(rstiefel)
 source("helper.R")
 
-generateData <- function(P=200, S=10, R=S, ngroups=10, nvec=rep(100, ngroups),
+generateData <- function(P=200, S=10, R=S, Q=S-R, ngroups=10,
+                         nvec=rep(100, ngroups),
                          V=rbind(diag(S), matrix(0, nrow=P-S, ncol=S)),
-                         s2vec=rep(1, ngroups), LambdaList=NULL, Olist=NULL) {
+                         s2vec=rep(1, ngroups), LambdaList=NULL,
+                         Olist=NULL) {
 
-
+    if(Q + R != S)
+        stop("Q + R must be equal S")
 
     if( is.null(LambdaList) ) {
         LambdaList <- vector("list", ngroups)
@@ -31,21 +34,31 @@ generateData <- function(P=200, S=10, R=S, ngroups=10, nvec=rep(100, ngroups),
     
     Slist <- Ylist <- Ulist <- OmegaList <- SigmaList <- list()
 
+    if(Q > 0) {
+        Oshared <- rbind(matrix(0, nrow=(S-Q), ncol=Q), rustiefel(Q, Q))
+    } else {
+        Oshared <- matrix(ncol=0, nrow=S)
+    }
+
+    LambdaShared <- sort(rexp(Q, 1/10), decreasing=TRUE)
     for ( k in 1:ngroups ) {
 
         if( is.null(Olist[[k]]) ) {
             print(k)
             ## Generate Ok
-            Ok <- rustiefel(S, R)
+            Odiff <- rbind(rustiefel(R, R), matrix(0, nrow=S-R, ncol=R))
+            Ok <- cbind(Odiff, Oshared)
             Olist[[k]] <- Ok
         }
 
+
         if( is.null(LambdaList[[k]]) ){
-            LambdaList[[k]] <- sort(rexp(R, 1/10), decreasing=TRUE)
+            LambdaDiff <- sort(rexp(R, 1/10), decreasing=TRUE)
+            LambdaList[[k]] <- c(LambdaDiff, LambdaShared)
         }
 
         Ok <- Olist[[k]]
-        Uk <- V%*%Ok
+        Uk <- V %*% Ok
         Ulist[[k]] <- Uk
         Lamk <- LambdaList[[k]]
[…]

### 2016-06-28 helper.R
Convex hull plotting

@@ -69,7 +69,7 @@ getAgeStat <- function(normsMat) {
 
 }
 
-posteriorPlot <- function(Osamps, OmegaSamps, s2samps, nsamps, groups, probRegion=0.95, hline=NULL, col=NULL, pch=NULL, ymax=30, plotPoints=TRUE) {
+posteriorPlot <- function(Osamps, OmegaSamps, s2samps, nsamps, groups, probRegion=0.95, hline=NULL, col=NULL, pch=NULL, lty=NULL, ymax=30, plotPoints=TRUE) {
 
     ngroups <- length(groups)
     
@@ -79,17 +79,20 @@ posteriorPlot <- function(Osamps, OmegaSamps, s2samps, nsamps, groups, probRegio
     if(is.null(pch)){
         pch=rep(19, ngroups)
     }
+    if(is.null(lty)){
+        lty=rep(1, ngroups)
+    }
     par(mar=c(5.1, 5.1, 4.1, 2.1))
     plot(0, 0, xlim=c(-pi/2, pi/2),
          ylim=c(0, ymax), cex=0, xlab="Angle", ylab=expression(lambda[1]/lambda[2]), xaxt="n", cex.axis=1.5, cex.lab=1.5)
     axis(1, at=seq(-pi/2, pi/2, by=pi/4), labels=expression(-pi/2, -pi/4, 0, pi/4, pi/2), cex.axis=1.5, cex.lab=1.5)
 
     
-    
     for(g in groups) {
 
         pmPsi <- getPostMeanPsi(Osamps[, , g , ], OmegaSamps[, g, ],
                                 s2samps[g, ], nsamps)
+        
         eigPsi <- eigen(pmPsi)
         pmValues <- eigPsi$values
         pmVectors <- eigPsi$vectors
@@ -106,7 +109,7 @@ posteriorPlot <- function(Osamps, OmegaSamps, s2samps, nsamps, groups, probRegio
         if(plotPoints) {
             points(pts[1, ], pts[2, ], col=alpha(col[g], 1/2), pch=pch[g], cex=0.5)
         } else {
-            polygon(pts[1, hullPoints], pts[2, hullPoints], lwd=3, border=col[g], col=alpha(col[g], 1/4))
+            polygon(pts[1, hullPoints], pts[2, hullPoints], lwd=3, border=col[g], col=alpha(col[g], 1/4), lty=lty[g])
         }
     }
 
[…]

### 2016-06-28 simulationFigures.R
Simulation figures for paper

@@ -1,0 +1,91 @@
+load("SdimLoss.Rdata")
+pdf("paper/Figs/LossVsDimension.pdf")
+par(mar=c(5.1, 5.1, 4.1, 2.1))
+plot(5:50, lossLst, type="l", lwd=3, ylim=c(0, 100), xlim=c(0, 50), ylab="Stein's Risk", xlab=expression(hat(s)), cex.axis=2, cex.lab=2)
+grid()
+lines(5:50, lossLst, type="l", lwd=3, ylim=c(0, 120))
+abline(h=3, lty=2, col="blue", lwd=3)
+abline(h=0.89, lty=2, col="red", lwd=3)
+
+legend("topright", lty=2, lwd=3, col=c("blue", "red"),
+       legend=c(expression(VV^T==I[p]),
+                expression(paste(VV^T, "= span(", U[1], ",..., ", U[K], ")"))),
+                bty='n', cex=2)
+dev.off()
+
+pdf(sprintf("paper/Figs/simRatio-s5.pdf", format(Sys.Date(), "%m-%d")))
+load("sdimtest-5-1.RData")
+evalRatios <- with(resList[[1]],
+    sapply(1:ngroups, function(k) {
+    numer <- sum(eigen(t(V[, 1:S]) %*% dat$Slist[[k]] %*% V[, 1:S])$values)/dat$nvec[k]
+    denom <- sum(svd(dat$Ylist[[k]])$d[1:S]^2)/dat$nvec[k]
+    correction <-  (denom - 1/mean(s2samps[k, ]) * S * P / dat$nvec[k])  / denom
+    (numer/denom) / correction
+    }))
+
+par(mar=c(7.1, 4.5, 4.1, 2.1))
+barplot(evalRatios, ylim=c(0, 1), xlab="Group", ylab="Goodness of Fit", main="",
+        space=0.2, cex.axis=1.5, col="#646464", col.axis="#646464",
+        col.main="#646464", border=NA,
+        names.arg=1:10, cex.names=2, cex.main=2, cex.lab=2, col.lab="#646464")
+dev.off()
+
+pdf(sprintf("paper/Figs/simRatio-s20.pdf", format(Sys.Date(), "%m-%d")))
+load("sdimtest-20-1.RData")
+evalRatios <- with(resList[[1]],
+    sapply(1:ngroups, function(k) {
[…]

### 2016-06-30 paper/mscov.tex
Latest iteration.

@@
 
-\title{Shared Subspace Model}
-
-\date{\today}
-
-\title{Shared Subspace Covariance Model}
+\title{A Shared Subspace Model for Multi-Group Covariance Estimation}
 \author{Alexander Franks and Peter Hoff}
@@
   by the group-level eigenvectors.  We use empirical Bayes to identify
-  a low-dimensional subspace which explains as a large amount of
-  variation across all groups, and, given this subspace, show how a
-  Gibbs sampling algorithm can be used to estimate the posterior
-  uncertainty of group-level eigenvectors and eigenvalues.  The
+  a low-dimensional subspace which explains variation across all
+  groups and use an MCMC algorithm to estimate the posterior
+  uncertainty of eigenvectors and eigenvalues on this subspace.  The
   implementation and utility of our model is illustrated with analyses
-  of high-dimensional multivariate biological data.
+  of high-dimensional multivariate gene-expression data.
 
@@
 
-In many natural systems, high dimensional data is often very
-structured and thus can be well understood on a relatively low
-dimensional subspace. For example, with gene expression data, the
-effective dimensionality is thought to scale with the number of gene
-regulatory modules, not the number of genes themselves
-\citep{Heimberg2016}.  As such, differences in gene expression across
-groups should be expressed in terms of differences between these
-regulatory modules rather than strict differences between expression
-levels.  Such differences can be examined on a subspace that reflects
-the correlations resulting from these modules.
-
-It has long been known that the sample covariance matrix is a poor
-estimator of the population covariance matrix, particularly when the
-sample size is comparable to or smaller than the number of features
-($n \leq p$) \citep{Dempster1969, Stein1975}.  A host of estimators
-have been devised to address this fact. These primarily include
-shrinkage approaches \citep{Schafer2005, Ledoit2011} and low
-rank factor models \citep{Mardia1980, Ullman2003, Fan2008}.  In many
-cases, a covariance matrix is highly structured and thus can be
-sparsely parameterized even if the matrix itself is not low rank
-\citep{Burg1982, Williams1993} .  Along these lines, there has been a
-recent interest in sparse Gaussian graphical model, in which the there
-are assumed to be many zeros in the inverse covariance matrix
-\citep{Meinshausen2006, Friedman2008}.
-
-Most of these approaches assume that the goal is to estimate the
-covariance matrix of a single population.  In many applications,
-however, the data can be split into groups, each of which represent
-samples from populations with distinct (but possibly related)
-covariance matrices.  In this case, a comparative analysis of
-covariance patterns across multiple groups of observations is often
-warranted.  Evaluating similarities and differences between covariance
-matrices can be an important complement to classical analyses which
-center primarily on estimating differences in means across groups.
-For instance, in case control-studies, mean level effects may be small
-relative to subject variability; distributional differences between
-groups may still be evident as differences in the covariances between
-features.  Even when mean level differences are detectable, better
-estimates of the covariability of features across groups may lead to
-an improved understanding of the mechanisms underlying these apparent
-mean level differences.  Further, accurate covariance estimation is an
-essential part of many prediction tasks (e.g. quadratic discriminant
-analysis).
-
-Some of the earliest approaches for modeling multiple
-covariance matrices focused on estimation in terms of the spectral
-decompositions for these matrices.  \cite{Flury1987} developed
-estimation and testing procedures for the ``common principal
-components'' model, in which a set of covariance matrices were assumed
-to share the same eigenvectors.  \citet{Schott1991, Schott1999}
-considered cases in which only certain eigenvectors are shared across
-populations, and \citet{Boik2002} described an even more general model
-in which eigenvectors can be shared between all or some of the groups.
-More recently, \citet{Hoff2009}, noting that eigenvectors are unlikely
-to be shared exactly between groups, introduced a hierarchical model
-for eigenvector shrinkage based on the matrix Bingham distribution.
-There has also been a significant interest in estimating covariance
-matrices using Gaussian graphical models. For Gaussian graphical
-models, zeros in the precision matrix correspond to conditional
-independences between pairs of features given the remaining features
+% For example, it has long been known that the sample covariance matrix
+% is a poor estimator of the population covariance matrix, particularly
+% when the sample size is comparable to or smaller than the number of
+% features ($n \leq p$) \citep{Dempster1969, Stein1975}.  A host of
+% estimators have been devised to address this fact. These primarily
+% include shrinkage approaches \citep{Schafer2005, Ledoit2011} and low
+% rank factor models \citep{Mardia1980, Ullman2003, Fan2008}.  In many
+% cases, a covariance matrix is highly structured and thus can be
+% sparsely parameterized even if the matrix itself is not low rank
+% \citep{Burg1982, Williams1993} .  Along these lines, there has been a
+% recent interest in sparse Gaussian graphical model, in which the there
+% are assumed to be many zeros in the inverse covariance matrix
+% \citep{Meinshausen2006, Friedman2008}.
+
+In modern analyses, multivariate data is often partitioned into
+groups, each of which represent samples from populations with
+distinct, but possibly related, distributions.  Although historically the
+primary focus has been on identifying mean-level differences between
+populations, there has been a growing need to identify differences in
+population covariances as well.  For instance, in case
+control-studies, mean level effects may be small relative to subject
+variability; distributional differences between groups may still be
+evident as differences in the covariances between features.  Even when
+mean level differences are detectable, better estimates of the
+covariability of features across groups may lead to an improved
+understanding of the mechanisms underlying these apparent mean level
+differences.  Further, accurate covariance estimation is an essential
+part of many prediction tasks (e.g. quadratic discriminant analysis).
+Thus, evaluating similarities and differences between covariance
+matrices can be an important complement to more traditional analyses
+for estimating differences in means across
+groups.
+
+To address this needs, we develop a novel method for
+group-covariance estimation.  In our method, we exploit the fact that
+in many natural systems, high dimensional data is often very
+structured and thus can be best understood on a lower dimensional
+subspace. For example, with gene expression data, the effective
+dimensionality is thought to scale with the number of gene regulatory
+modules, not the number of genes themselves \citep{Heimberg2016}.  As
+such, differences in gene expression across groups should be expressed
+in terms of differences between these regulatory modules rather than
+strict differences between expression levels.  Such differences can be
+examined on a subspace that reflects the correlations resulting from
+these modules.  In contrast to most existing approaches for group
+covariance estimation, our approach is to directly infer such
+subspaces from groups of related data.
+
+% Existing approaches incorporate a variety of
+% different structural assumptions about the relationship between
+% multiple groups of covariance matrices.
+
+Some of the earliest approaches for multi-group covariance estimation
+focus on estimation in terms of spectral decompositions.
+\cite{Flury1987} developed estimation and testing procedures for the
+``common principal components'' model, in which a set of covariance
+matrices were assumed to share the same eigenvectors.
+\citet{Schott1991, Schott1999} considered cases in which only certain
+eigenvectors are shared across populations, and \citet{Boik2002}
+described an even more general model in which eigenvectors can be
+shared between some or all of the groups.  More recently,
+\citet{Hoff2009}, noting that eigenvectors are unlikely to be shared
+exactly between groups, introduced a hierarchical model for
+eigenvector shrinkage based on the matrix Bingham distribution.  There
+has also been a significant interest in estimating covariance matrices
+using Gaussian graphical models. For Gaussian graphical models, zeros
+in the precision matrix correspond to conditional independences
+between pairs of features given the remaining features
 \citep{Meinshausen2006}.  \citet{Witten2014} extended existing work in
@@
 maximizes the similarity of the two groups.  This technique does not
-require the data from each group to share the same feature set.  An
+require the data from each group to share the same feature set.  A
 common variant for prediction, partial least squares discriminant
@@
 Nevertheless, the basic idea that data can often be well represented
-on a low dimensional space is an appealing one that we leverage as
-well.
+on a low dimensional space is an appealing one that we leverage.
 
@@
 covariance matrix and then introduce our extension to the multi-group
-setting.  In Section \ref{sec:inference} we describe an optimization
-algorithm for inferring the shared subspace and then outline a MCMC
-algorithm for estimating the posterior distribution of the projected
-covariance matrices.  In Section \ref{sec:simulation} we
-investigate the behavior of these models in simulation and then
-demonstrate the utility of the model with real biological data in
-Section \ref{sec:app}.
+setting.  In Section \ref{sec:inference} we describe an empirical
+Bayes algorithm for inferring the shared subspace and estimating the
+posterior distribution of the covariance matrices of the data
+projected onto this subspace.  In Section \ref{sec:simulation} we
+investigate the behavior of these models in simulation and demonstrate
+how the shared subspace is widely applicable, even when there is
+little similarity in the covariance matrices across groups.  We then
+demonstrate the utility of the model in an analysis of gene expression
+data from juvenile leukemia patients in Section \ref{sec:app}.  Despite the
+large feature size ($p > 3000$) and small sample size ($n < 100$ per
+group), we identify interpretable similarities and differences in gene
+covariances on a low dimensional subspace.
 
@@
 Classical results for parametric models (e.g., \citet{Schwartz1965})
-imply that asymptotically in $n$ for fixed $p$, an estimator will
-be consistent for a spiked population covariance as long as the
-assumed number of spikes (eigenvalues larger than $\sigma^2$) is greater than or
-equal to the true number.  However, when $p$ is large relative to $n$,
-as is the case for the examples considered here, things are more
+imply that asymptotically in $n$ for fixed $p$, an estimator will be
+consistent for a spiked population covariance as long as the assumed
+number of spikes (eigenvalues larger than $\sigma^2$) is greater than
+or equal to the true number.  However, when $p$ is large relative to
+$n$, as is the case for the examples considered here, things are more
 difficult.  Under the spiked covariance model, it has been shown that
-if $p/n \rightarrow \gamma >0$ as $n\rightarrow \infty$, the $k$th largest
-eigenvalue of $\bl S/(n\sigma^2)$ will converge to an upwardly biased
-version of $\lambda_{k}+1$, if $\lambda_k$
-is greater than $\sqrt{\gamma}$ \citep{Baik2006, Paul2007} .  This has
-led several authors to suggest estimating $\Sigma$ via shrinkage of
-the eigenvalues of the sample covariance matrix. In particular, in the
-setting that $\sigma^2$ is known, \citet{Donoho2013} propose estimating
-all eigenvalues whose sample estimates are smaller than 
-$\sigma^2(1+\sqrt{\gamma})^2$ by $\sigma^2$, and shrinking the
-larger eigenvalues in a way that depends on the particular loss
-function being used.  These shrinkage functions are shown to be
-asymptotically optimal in the $p/n\rightarrow \gamma$ setting.
+if $p/n \rightarrow \alpha >0$ as $n\rightarrow \infty$, the $k$th
+largest eigenvalue of $\bl S/(n\sigma^2)$ will converge to an upwardly
+biased version of $\lambda_{k}+1$, if $\lambda_k$ is greater than
+$\sqrt{\alpha}$ \citep{Baik2006, Paul2007} .  This has led several
+authors to suggest estimating $\Sigma$ via shrinkage of the
+eigenvalues of the sample covariance matrix. In particular, in the
+setting that $\sigma^2$ is known, \citet{Donoho2013} propose
+estimating all eigenvalues whose sample estimates are smaller than
+$\sigma^2(1+\sqrt{\alpha})^2$ by $\sigma^2$, and shrinking the larger
+eigenvalues in a way that depends on the particular loss function
+being used.  These shrinkage functions are shown to be asymptotically
+optimal in the $p/n\rightarrow \alpha$ setting.  Note that covariance
+estimators of this form are are equivariant with respect to rotations
+and scale changes.  The situation should be different, however, when
+we are interested in estimating multiple covariance matrices from
+distinct, but related groups.  Here, group-level equivariance to
+rotations is an unreasonable assumption; both eigenvalue \emph{and}
+eigenvector shrinkage can play an important role in improving
+covariance estimates.
 
@@
 
-When estimating a single covariance matrix, the absence of prior
-knowledge suggests estimators that are equivariant with respect to
-rotations and scale changes.  The situation is different when we are
-interested in estimating multiple covariance matrices from distinct,
-but related groups.  Here, group-level equivariance to rotations is an
-unreasonable assumption; both eigenvalue \emph{and} eigenvector
-shrinkage can play an important role in improving covariance
-estimates.
-
 For multi-group covariance estimation, we assume that we have $K$
-matrices, $Y_1, ... Y_K$, where $Y_k$ conists of where $Y_k$ consists
+matrices, $Y_1, ... Y_K$, where $Y_k$ consists
 of $n_k$ independent rows of mean-zero normal data of dimension $p$,
@@
 \noindent Here, the columns of $V \in \mathcal{V}_{p, s}$ determine a subspace
-shared by all groups $k$ with $p \geq s$.  Throughout this paper we
-will denote to the shared subspace as $VV^T \in \mathcal G_{r,p}$, where
-$\mathcal G_{r,p}$, known as the Grassmann manifold, consists of all
+shared by all groups $k$ with $s \geq p$.  Throughout this paper we
+will denote to the shared subspace as $VV^T \in \mathcal G_{p,s}$, where
+$\mathcal G_{p,s}$, known as the Grassmann manifold, consists of all
 $s$-dimensional linear subspaces of $\mathcal{R}^p$
@@
 
-In this section we outline our empirical Bayesian approach for estimating
-covariance parameters and their associated uncertainty.  
-Uncertainty quantification is particularly important in the $n \ll p$
-setting where seemingly large differences in the point estimates of
-the low rank covariance matrices across groups may not actually
-reflect significant differences.  Although uncertainty quantification
-for all parameters is desirable, for the size of the problems we are
-generally interested in, we found MCMC inference for the subspace
-$VV^T$ to be intractable.  Our optimization based approached is
-significantly faster and works well for high-dimensional datasets.  As
-such, we start, in Section \ref{sec:em}, by outlining an
-expectation-maximization algorithm for estimating the shared
-subspace. In Section \ref{sec:bayes} we describe a Gibbs sampler for
-posterior inference of $\Psi_k$ given the common subspace.  These two
-algorithms can be used in conjunction to first identify and evaluate a
-shared subspace and then to quantify the uncertainty associated with
-the projected covariance matrices.
-
-TODO: emphasize 1) justification 2) objective of \emph{identifying} this
-subspace when p much greater than n.
+In this section we outline an empirical Bayesian approach for
+estimating a low-dimensional shared subspace and the covariance
+matrices of the data projected onto this space. As we highlight in
+Section \ref{sec:simulation}, if the spiked covariance model holds for
+each group individually, then the shared subspace model also
+holds because we can take the shared subspace to be the full
+$p$-dimensional space. In practice, we can often identify a shared
+subspace of dimension $s \ll p$ that preserves most of the variation
+in the data.  Our primary objective is to identify the ``best'' shared
+subspace of fixed dimension $s < p$.  In Section \ref{sec:em} we
+describe an expectation maximization algorithm for estimating the
+maximum marginal likelihood of the shared subspace, $VV^T$.  This
+approach is computationally tractable for high-dimensional datasets.
+Given an inferred subspace, we then seek estimators for the covariance
+matrices of the data projected onto this space.  Because seemingly
+large differences in the point estimates of covariance matrices across
+groups may not actually reflect statistically significant differences,
+in Section \ref{sec:bayes} we describe a Gibbs sampler that can be
+used to generate estimates of the projected covariance matrices,
+$\Psi_k$ and and their associated uncertainty.
+
+% However, in the $n \ll p$ setting, seemingly large
+% differences in the point estimates of the covariance matrices across
+% groups may not actually reflect statistically significant differences.
+% Thus, in Section \ref{sec:bayes} we describe a Gibbs sampler that can
+% be used to infer both point estimates and the associated uncertainty
+% of the projected covariance matrices, $\Psi_k$.
+
+%TODO: emphasize 1) justification 2) objective of \emph{identifying} this
+%subspace when p much greater than n.
 
@@
 We propose an expectation-maximization (EM) algorithm, to estimate the
-relevant quantities.  In particular we maximize the marginal posterior
-distribution of $V$, treating $M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as
+relevant quantities.  In particular we maximize the marginal
+likelihood of $V$, treating $M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as
 the ``missing'' parameters to be imputed.  We assume independent
 Jeffreys' prior distributions for both $\frac{1}{\sigma_k^2}$ and
-$M_k^{-1}$ and a uniform prior for $V \in \mathcal{V}_{s, p}$.
-Jeffreys' prior for these quantities corresponds to
+$M_k^{-1}$.  Jeffreys' prior for these quantities corresponds to
 $p(\frac{1}{\sigma_k^2}) \propto \sigma_k^2$ and
@@
 $\frac{1}{2}tr((I-VV^T)\mathbf{S}_k)$) distribution.  Based on these
-results the following EM algorithm can be used.
+results the following EM algorithm can be used:
 
@@
 
-For the ``M-step'', we use a numerical optimization algorithm based on
+For the second step (``M-step''), we use a numerical optimization algorithm based on
 the Cayley transform to preserve the orthogonality constraints in $V$
@@
 
-It is important to quantify how much of the variability in each group can be
-explained by an inferred subspace.  If $V$ is a basis for
-the shared subspace, then for each group $k$, most of the variation in
-$Y_k$ should be preserved when projecting the data to this subspace.
-To characterize the extent to which this is true for different groups, we
-propose a simple estimator for the proportion of ``signal'' variance
-that lies on any given subspace.  Specifically, for an estimated
-subspace $\hat{V}$, we apply the following statistic for the ratio of the
-sum of the first $s$ eigenvalues of $\hat{V}^T\bl \Sigma_k \hat{V}$ to
-the sum of the first $s$ eigenvalues of $\Sigma_k$:
+If $V$ is a basis for a shared subspace, then for each group $k$, most
+of the variation in $Y_k$ should be preserved when projecting the data
+to this space.  To characterize the extent to which this is true for
+different groups, we propose a simple estimator for the proportion of
+``signal'' variance that lies on any given subspace.  Specifically,
+for an estimated subspace $\hat{V}$, we use the following statistic
+for the ratio of the sum of the first $s$ eigenvalues of
+$\hat{V}^T\bl \Sigma_k \hat{V}$ to the sum of the first $s$
+eigenvalues of $\Sigma_k$:
 %
 \begin{equation}
-r(Y_k: \hat{V}, \hat{\sigma}_k^2) = \frac{||Y_k\hat{V}||_F}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
-  ||Y_k\widetilde{V}||_F - \hat{\sigma}_k^2ps}
+\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2) = \frac{||Y_k\hat{V}||_F}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
+  ||Y_k\widetilde{V}||_F - \hat{\sigma}_k^2ps/n_k}
 \label{eqn:ratio}
@@
 \noindent where $||.||_F$ is the Frobenius norm. Our motivation for
-this statisic is the fact that if
+this statistic is the fact that if
 $\Sigma_k = \hat{V}\Psi_k\hat{V}^T + \sigma_k^2I$ then
-$r(Y_k: \hat{V}, \hat{\sigma}_k^2) \rightarrow 1$.  To see this, note
-that $\underset{V \in \mathcal{V}_{p, s}}{max}\frac{||Y_kV||_F}{n_k}$
+$\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2) \approx 1$.  To see this, note
+that $\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}\frac{||Y_k\widetilde{V}||_F}{n_k}$
 is equivalent to the sum of the first $s$ eigenvalues of the sample
 covariance matrix $\bl S_k/n_k$.  With $\hat{\lambda}^{(k)}_i$
-the $i$-th eigenvalue of $\frac{\bl S_k}{n_k}$, \citet{Baik2006}
+the $i$-th eigenvalue of $\bl S_k/n_k$, \citet{Baik2006}
 demonstrated that asymptotically as $p, n_k \rightarrow \infty$ with
-$\frac{p}{n_k} = \gamma_k$ fixed
+$p/n_k = \alpha_k$ fixed
 
 \begin{eqnarray}
-\hat{\lambda}^{(k)}_i &\rightarrow& \lambda^{(k)}_i\left(1 +
-                                    \frac{\sigma_k^2\gamma}{\lambda^{(k)}_i
+\nonumber \hat{\lambda}^{(k)}_i &\rightarrow& \lambda^{(k)}_i\left(1 +
+                                    \frac{\sigma_k^2\alpha}{\lambda^{(k)}_i
                                     - 1}\right)\\
@@
 $\sum_i^s \hat{\lambda}^{(k)}_i \approx \sum_i^s \lambda^{(k)}_i +
-\sigma_k^2ps$.
+\sigma_k^2ps/n_k$.
 If we assume that $s$ is fixed in this asymptotic regime, then
@@
 which groups can be reasonably compared on a given subspace and which
-groups cannot.  When $r(Y_k: \hat{V}, \hat{\sigma})k^2)$ is small for
-some groups, it may suggest that the rank, $s$, of the inferred
-subspace needs to be larger to capture the variation in all
-groups. The use of this statistic is illustrated in Section
-\ref{sec:simulation}, where we plot these estimates for different
-simulated datasets, and in Section \ref{sec:app}, where we plot the
-estimates for subspaces inferred with real biological data.
+groups cannot.  When $\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2)$ is small
+for some groups, it may suggest that the rank, $s$, of the inferred
+subspace needs to be larger to capture the variation in all groups. To
+demonstrate this, in Section \ref{sec:simulation}, we compute these
+statistics for inferred subspaces of different dimensions on a single
+dataset. In Section \ref{sec:app}, we plot the estimates for subspaces
+inferred with real biological data.
+
+%% Problem here! Can't compare different sample sizes??
 
@@
 
-Although the EM algorithm presented in the previous section yields
-point estimates for $VV^T$ and $\Psi_k$, it does not lead to natural
-uncertainty for these estimates. As such, in this section, we assume
-that the subspace $VV^T$ is fixed and known and focus on uncertainty
-quantification for $\Psi_k$ only.  Note that when the subspace is
-known, the posterior distribution of $\Sigma_k$ is conditionally
-independent from the other groups, so that we can independently
-estimate the conditional posterior distribution for each group.
-
-There are many different prior specifications for $\Psi_k$ and
-$\sigma^2_k$ that we could consider for estimating the posterior
-uncertaity.  In this paper, we build on recent interest in the spiked
-covariance model \citep{Donoho2013, Paul2007} and develop tractable
-models for the eigendecomposition of $\Psi_k$.  Such models are
-especially appealing because it is simple to summarize and compare
-$\Psi_k$ across groups in terms their eigenstructure (See Section
-\ref{sec:simulation} and \ref{sec:app}).
-
-By modeling the eigenstructure, we can now view each covariance
-$\Sigma_k$ in terms of the spiked principle components model (\ref{eqn:spiked}).
-Equation \ref{eqn:sspsi}, written as a function $V$, becomes
+The EM algorithm presented in the previous section yields point
+estimates for $VV^T$ and $\Psi_k$, but does not lead to natural
+uncertainty quantification for these estimates.  In this section, we
+assume that the subspace $VV^T$ is fixed and known and demonstrate how
+we can estimate the posterior distribution for $\Psi_k$.   Note that when
+the subspace is known, the posterior distribution of $\Sigma_k$ is
+conditionally independent from the other groups, so that we can
+independently estimate the conditional posterior distributions for each
+group.
+
+% There are many different prior specifications for $\Psi_k$ and
+% $\sigma^2_k$ that we could consider for estimating the posterior
+% uncertainty.  In this paper, we build on recent interest in the spiked
+% covariance model \citep{Donoho2013, Paul2007} and develop tractable
+% models for the eigendecomposition of $\Psi_k$.  Such models are
+% especially appealing because it is simple to summarize and compare
+% $\Psi_k$ across groups in terms their eigenstructure (See Section
+% \ref{sec:simulation} and \ref{sec:app}).
+
+Building on recent interest in the spiked covariance model
+\citep{Donoho2013, Paul2007} we propose a tractable MCMC algorithm by
+specifying priors on the eigenvalues and eigenvectors of $\Psi_k$.  By
+modeling the eigenstructure, we can now view each covariance
+$\Sigma_k$ in terms of the original spiked principle components model
+(\ref{eqn:spiked}).  Equation \ref{eqn:sspsi}, written as a function
+$V$, becomes
 
@@
 \nonumber \Psi_k &= O_k\Lambda_KO_k^T\\
-\Sigma_k &= VO_k\Lambda_KO_k^TV^T + \sigma^2_kI
+\Sigma_k &= V\Psi_kV^T + \sigma^2_kI
 \label{eqn:ss}
@@
 %
-\noindent Here, we make one further generalization to allow
-$\Psi_k$ to be a rank $r \leq s$ dimensional covariance
-matrix on the $s$-dimensional subspace.  Thus, $\Lambda_k$ is an $r \times
-r$ diagonal matrix of eigenvalues, and $O_k \in
-\mathcal{V}_{s,r}$ is the matrix of eigenvectors of
-$\Psi_k$.  For any individual group, this corresponds to the original spiked
-PCA model (Equation \ref{eqn:spiked}) with $U_k = VO_k \in
-\mathcal{V}_{p, r}$.    This ``double spiked'' covariance model TODO
+\noindent Here, we make one further generalization to allow $\Psi_k$
+to be a rank $r \leq s$ dimensional covariance matrix on the
+$s$-dimensional subspace.  Thus, $\Lambda_k$ is an $r \times r$
+diagonal matrix of eigenvalues, and $O_k \in \mathcal{V}_{s,r}$ is the
+matrix of eigenvectors of $\Psi_k$.  For any individual group, this
+corresponds to the original spiked PCA model (Equation
+\ref{eqn:spiked}) with $U_k = VO_k \in \mathcal{V}_{p, r}$.  This
+generalization is helpful because it enables us to independently
+specify a subspace common to all groups and the possibly lower rank
+features on this space that are specific to individual groups.
+Although our model is most useful when the covariance matrices are
+related across groups, we can use this formulation to specify models
+for multiple unrelated spiked covariance models by letting the shared
+subspace be full rank, e.g. $s=p$.  We explore this in detail in
+Section \ref{sec:simulation}.
 
@@
 
-\paragraph{Conditional distribution for $\Omega$:} Here we specify the conditional
-distribution of the diagonal matrix $\Omega_k =
-\Lambda_k(I+\Lambda_k)^{-1} = \text{diag}(\omega_{k1}, ... \omega_{kr})$.  We first consider a uniform(0,1) prior distribution for each element of $\Omega$, or
-equivalently, an $F_{2,2}$ prior distribution for the elements of
-$\Lambda$.  The full conditional distribution of an
+\paragraph{Conditional distribution for $\Omega$:} Here we specify the
+conditional distribution of the diagonal matrix
+$\Omega_k = \Lambda_k(I+\Lambda_k)^{-1} = \text{diag}(\omega_{k1},
+... \omega_{kr})$.
+We consider a uniform(0,1) prior distribution for each element of
+$\Omega$, or equivalently, an $F_{2,2}$ prior distribution for the
+elements of $\Lambda$.  The full conditional distribution of an
 element $\omega_i$ of $\Omega$ is proportional to the likelihood
@@
 
-\subsection{Validation of Inference}
-
-In this section we start by validating, in simulation, that our
-inferential algorithm recovers the true parameters of the data
-generating mechanism.  We simulate five groups of $p=200$ dimensional
-data each with sample size $n_k=50$.  We choose the eigenvectors of
-the anisotropic component of the covariance matrix, $\Psi_k$, to share
-a common two dimensional subspace.  To fit the model to this simulated
-data, we first estimate the shared subspace using the EM algorithm outlined
-in Section \ref{sec:em} and then use the MCMC algorithm described in
-Section \ref{sec:bayes} to estimate the posterior distribution of the
-projected covariance matrices $\Psi_k$ and noise variance $\sigma_k^2$.
+%\subsection{Validation of Inference}
+
+We start with an example demonstrating how the shared subspace
+model can be used to identify statistically significant differences between
+covariance matrices on a low dimensional subspace. Here, we simulate
+$K=5$ groups of data from the shared subspace spiked covariance
+model with $p=200$, $s=r=2$, $\sigma_k^2=1$, and
+$n_k=50$.  We fix the first eigenvalue of $\Psi_k$ from each group to
+$\lambda_1=100$ and vary $\lambda_2$ so that $\lambda_1/\lambda_2$ is
+either $10$, $3$ or $1$.  For this two dimensional shared subspace
+model we summarize $\Psi_k$ in terms of its eigendecomposition by
+computing posterior distributions for the eigenvalue ratio
+$\frac{\lambda_1}{\lambda_2}$, with $\lambda_1 > \lambda_2$, and the
+angle of the first eigenvector on this subspace,
+$\text{arctan}(\frac{O_{12}}{O_{21}})$, relative to the first column
+of $V$.
+
+Figure \ref{fig:simPosterior} depicts the 95\% posterior regions for
+these quantities from a single simulation.  Dots correspond to the
+true ratios and orientations of $\hat{V}^T\Sigma_k\hat{V}$. To compute
+these regions, we iteratively remove posterior samples corresponding to the
+vertices of the convex hull until only 95\% of the original samples
+remain.  Non-overlapping posterior regions imply that differences in
+the covariances are statistically significant between groups.  In this
+example, the ratio of the eigenvalues of the true covariance matrices
+were $10$ (black and red groups), $3$ (green and blue groups) and $1$
+(cyan group).  Larger values of the eigenvalue ratio correspond to more
+correlated contours and a value of $1$ implies isotropic covariance.
+Note that for the smaller eigenvalue ratio of $3$, there is larger
+uncertainty about the orientation of the primary axis.  When the ratio
+is one, as is the case for the cyan colored group, there is no
+information about the orientation of the primary axis since the contours
+are spherical.
 
@@
     \centering
-    \begin{subfigure}[b]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{Figs/posteriorRegions}
-        \caption{Posterior}
+%    \begin{subfigure}[b]{0.45\textwidth}
+    \includegraphics[width=0.5\textwidth]{Figs/posteriorRegions-chull}
+        \caption{95\% posterior regions}
         \label{fig:simPosterior}
-    \end{subfigure}
+%    \end{subfigure}
     ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
       %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=\textwidth]{Figs/simRatio-05-19}
-        \caption{Ratio}
-        \label{fig:simRatio}
-    \end{subfigure}
+    % \begin{subfigure}[b]{0.4\textwidth}
+    %     \includegraphics[width=\textwidth]{Figs/simRatio-05-19}
+    %     \caption{Ratio}
+    %     \label{fig:simRatio}
+    % \end{subfigure}
     \caption{a) 95\% posterior regions for the ratio of the
@@
       orientation of the principle axis on space spanned by $\hat{V}$
-      cover the truth.  True parameter values on
-      $\hat{V}^T\Sigma_k\hat{V}$ are indicated by the crosses.  Since
-      $V$ is only identifiable up to rotation we find the Procrustes
+      cover the truth.  Dots correspond to true data generating
+      parameter values on $\hat{V}^T\Sigma_k\hat{V}$ .  Since $V$
+      is only identifiable up to rotation we find the Procrustes
       rotation that maximizes the similarity of $\hat{V}$ to the true
@@
       orientations were $\pi/4$ (black), $-\pi/4$ (red and green) and
-      0 (blue and cyan).  b) Goodness of fit,
-      $r(Y_k: \hat{V}, \hat{\sigma}^2_k$ (Equation \ref{eqn:ratio}).
-      As expected, for each group, the inferred subspace explains
-      practically all of the variance from the first two eigenvectors
-      of $\Sigma_k$.  }
+      0 (blue and cyan). % b) Goodness of fit,
+      % $\gamma(Y_k: \hat{V}, \hat{\sigma}^2_k$ (Equation \ref{eqn:ratio}).
+      % As expected, for each group, the inferred subspace explains
+      % practically all of the variance from the first two eigenvectors
+      % of $\Sigma_k$. 
+    }
 \end{figure}
 
-For this two dimensional shared subspace model we summarize the
-estimates of $\Psi_k$ in terms of its eigendecomposition.
-Specifically, we compute the eigenvalue ratio
-$\frac{\lambda_1}{\lambda_2}$, with $\lambda_1 > \lambda_2$, and the
-angle of the first eigenvector on this subspace,
-$\text{arctan}(\frac{O_{12}}{O_{21}})$, relative to the first column
-of of $V$.  Figure \ref{fig:simPosterior} depicts the 95\% posterior
-regions for these quantities.  We select the displayed posterior
-samples in each group by iteratively remove the points corresponding
-to the vertices of the convex hull until only 95\% of the original
-samples remain.  Non-overlapping posterior regions are indicative of
-significant differences in the covariance structure between groups.
-
-In this example, the ratio of the eigenvalues of the true covariance
-matrices are $10$ (red and black groups), $3$ (green and blue
-groups) and $1$ (cyan groups).  Larger values of this ratio
-correspond to more correlated contours and a value of $1$ implies
-isotropic covariance.  The posterior regions for all groups cover the
-true eigenvalue ratio and primary axis orientation of $\hat{V}^T\Sigma_k\hat{V}$.
-Note that for smaller eigenvalue ratios, there is larger uncertainty
-about the orientation of the primary axis.  When the ratio is one, as
-is the case for the cyan colored group, there is no information
-about orientation of the primary axis since the contours are
-spherical.
-
-We also verify that the data projected onto the inferred subspace
-preserves most of the variability in the data.  To do this, we compute
-estimates for the proportion of explained variance, as given in
-Equation \ref{eqn:ratio}.  Figure \ref{fig:simRatio} depicts a barplot
-of this estimate for each group.  For each group, the estimate is
-nearly one, which reflects the fact that we were able to infer a two
-dimensional shared subspace that explains most of the variance from
-the first two eigenvectors of $\Sigma_k$.  In this example, by
-construction we ensured that such a subspace exists. However, this is
-not always the case.  In the next section we investigate examples for
-which the inferred subspace may not capture the relevant variation for
-all groups.
+To demonstrate the overall validity of the shared subspace approach,
+we compute the frequentist coverage of these 95\% bayesian credible
+regions for the eigenvalue ratio and primary axis orientation using
+one thousand simulations.  For the two groups with eigenvalue ratio
+$\lambda_1/\lambda_2 = 3$ the frequentist coverage was close to
+nominal at approximately 0.94.  For the groups with
+$\lambda_1/\lambda_2 = 10$ the coverage was approximately 0.92.  We
+did not evaluate the coverage for the group with
+$\lambda_1/\lambda_2 = 1$ since this value is on the edge of the
+parameter space and would not be covered by the 95\% posterior regions
+as constructed.  The slight under coverage is likely due to the fact
+that we infer $VV^T$ using maximum marginal likelihood, and thus
+ignore the extra variability due to the uncertainty about the shared
+subspace estimate.
 
@@
         \includegraphics[width=\textwidth]{Figs/LossVsDimension}
-        \caption{Stein's risk vs s}
+        \caption{Stein's risk vs $\hat{s}$}
         \label{fig:sdimension}
@@
         \includegraphics[width=\textwidth]{Figs/simRatio-s5}
-        \caption{s = 5}
+        \caption{$\hat{s}$ = 5}
         \label{fig:ratio-s5}
@@
         \includegraphics[width=\textwidth]{Figs/simRatio-s20}
-        \caption{s = 20}
+        \caption{$\hat{s}$ = 20}
         \label{fig:ratio-s20}
@@
     \caption{a) Stein's risk as a function of the dimension of the
-      shared subspace dimension.  Data from ten groups, $U_k$ uniform
-      on the Stiefel manifold $\mathcal{V}_{200, 2}$.  As the shared
-      subspace dimension increases, $s \rightarrow p$, the loss
-      converges to the loss from the independent spiked covariance
-      estimates (dashed line). b) For a simulated dataset, the
-      goodness of fit statistic, $r(Y_k: \hat{V}, \hat{\sigma_k}^2$,
-      when the assumed shared subspace is dimension 5.  c).  For a
-      single simulated dataset, goodness of fit when the assumed
-      shared subspace is dimension 20.  We can capture nearly all of
-      the variability in each of the 10 groups using an $s=rK=20$
+      shared subspace dimension.  Data from ten groups, with $U_k$
+      generated uniformly on the Stiefel manifold
+      $\mathcal{V}_{200, 2}$.  As the shared subspace dimension
+      increases, $\hat{s} \rightarrow p$, the risk converges to the
+      risk from independently estimated spiked covariance matrices
+      (dashed blue line).  The data also fit a shared subspace model
+      with $s=rK$.  If $VV^T = \text{span}(U_1, ..., U_k)$ were known
+      exactly, shared subspace estimation yields lower risk than
+      independent covariance estimation (dashed red line).  b) For a
+      single simulated dataset, the goodness of fit statistic,
+      $\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$, when the assumed
+      shared subspace is dimension $\hat{s} = 5$.  c).  For the same
+      dataset, goodness of fit when the assumed shared subspace is
+      dimension $\hat{s} = 20$.  We can capture nearly all of the
+      variability in each of the 10 groups using an $\hat{s}=rK=20$
       dimensional shared subspace. }
@@
 \Sigma_k ) - \log |\Sigma_k^{-1} \Sigma_k | - p$.
-Since we compute estimates for multiple groups of data, we report the
+Since we compute multi-group estimates, we report the
 average Stein's loss
@@
 \frac{1}{K} \sum_k L_S( \Sigma_k , \hat\Sigma_k)$.
-Note that under Stein's loss, the Bayes estimator is the inverse of
+Under Stein's loss, the Bayes estimator is the inverse of
 the posterior mean of the precision matrix,
@@
 underestimate the true dimension of the shared subspace.  In this
-simulation, we generate data from five groups of mean-zero normally
-distributed data in a $p=200$ dimensional feature space.  For each
-group we use the spiked covariance model of Equation \ref{eqn:spiked},
-where we generate eigenvectors uniformly from the Stiefel manifold,
-$U_k \in \mathcal{V}_{200, 2}$ with eigenvalues
-($\lambda_1, \lambda_2) = (250, 25)$ .  In this case, although most of
-the variation in each group can be captured by a two dimensional
-subspace, these are not the same subspaces across groups.  We make no
-assumption, a priori, that the subspaces are similar across groups.
-Below, we evaluate how well the shared subspace estimator performs with such
-data for inferred subspaces of increasing dimension.
-
-In Figure \ref{fig:sdimension} we plot the average Stein's loss, for
-estimates derived from a shared subspace model for different values
-of the shared subspace dimension, $s$.  For each value of $s$, we
-report the loss averaged over ten independent simulations.  In this
-figure, the dashed blue line corresponds to the average Stein's loss
-for covariance matrices estimated independently, e.g. without the
-shared subspace assumption.  As the shared subspace dimension increases to the
-dimension of the feature space, $s \rightarrow p$, the loss for the
-shared subspace estimator quickly converges to the loss for the
-individual spiked covariance estimator.  Note that
-it is always true that span($[U_1, ..., U_K]) \leq rK$.  Even when there is little
-similarity between the eigenvectors from each group, the shared
-subspace estimator with $s = rK$ will perform comparably to the
-estimator, provided that we can identify a subspace, $\hat{V}\hat{V}^T$, where
-span($\hat{V})$ is close to $\text{span}([U_1, ..., U_K])$.  This is
-immediately apparent in Figure \ref{fig:sdimension} from the phase transition
-in Stein's loss at $s = 20$.
-
-From the above simulation, it is clear that correctly specifying the
+simulation, we generate data $K=10$ groups of mean-zero normally
+distributed data with $p=200$, $r=2$, $s=p$ and $\sigma_k^2=1$.  We
+fix the eigenvalues for covariance matrix to
+($\lambda_1, \lambda_2) = (250, 25)$.  Although the signal variance
+from each group individually is preserved on a two dimensional
+subspace, these subspaces are not similar across groups-- the
+eigenvectors from each group are generated uniformly from the Stiefel
+manifold, $U_k \in \mathcal{V}_{p, r}$.
+
+We use this data to evaluate how well the shared subspace estimator
+performs when we fit the data using a shared subspace model of
+dimension $\hat{s} < s$.  In Figure \ref{fig:sdimension} we plot
+Stein's risk as a function of $\hat{s}$, estimating the risk
+empirically using ten independent simulations per value of $\hat{s}$.
+The dashed blue line corresponds to the multi-group Stein's risk for
+covariance matrices estimated independently.  Independent covariance
+estimation is equivalent to shared subspace inference with
+$\hat{s} = p$ because this implies $VV^T = I_p$.  Although the risk is
+large for small values of $\hat{s}$, as the shared subspace dimension
+increases to the dimension of the feature space,
+$\hat{s} \rightarrow p$, the risk for the shared subspace estimator
+quickly decreases.  Importantly, it is always true that
+rank($[U_1, ..., U_K]) \leq rK$ so it can equivalently be assumed that
+the data were generated from a shared subspace model with dimension
+$s = rK < p$.  As such, even when there is little similarity between the
+eigenvectors from each group, the shared subspace estimator with
+$\hat{s} = rK$ will perform well, provided that we can identify a
+subspace, $\hat{V}\hat{V}^T$, that is close to
+$\text{span}([U_1, ..., U_K])$. When
+$\hat{V}\hat{V}^T = \text{span}([U_1, ..., U_K])$ exactly,
+(\ref{fig:sdimension}, dashed red line),
+shared subspace estimation outperforms independent covariance
+estimation.
+
+From this simulation, it is clear that correctly specifying the
 dimension of the shared subspace is important for efficient covariance
 estimation.  When the dimension of the shared subspace is too small,
-there is larger loss for estimate relative to the truth.  We can also
-use the available data, independent of the true covariance matrices,
-to evaluate the appropriateness of the inferred subspace.  To this
-end, we calculate the estimate of the ``proportion of explained
-variance'' estimator from Section \ref{sec:em} (Equation
-\ref{eqn:ratio}).
-
-Figure \ref{fig:dimensionPlots}b and c depict the estimated proportion of
-variance explained for a single dataset for each of five group, using
-inferred subspaces of two different dimensions.  For the results
-displayed in Figure \ref{fig:ratio-s5}, we fixed the dimension of the
-subspace to $s=5$ (clearly too small) and for the results displayed in
-Figure \ref{fig:ratio-s5} we fixed the dimension of the shared subspace
-to $s=20$.  When the shared subspace model is appropriate, the
-subspace spanned by $\hat{V}$ explains most of the variance in
-$\Sigma_k$, for all groups.  When the estimated subspace is too small,
-the proportion of variance will be smaller than one for at least some of the
-groups.  For a given fixed subspace, these plots can be used to decide
-which groups can be fairly be compared on this space and which groups
-cannot and whether we would benefit from fitting a model with
-increased dimension $s$.
+we accrue higher risk.  The goodness of fit statistic,
+$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$, can be used to identify when
+a larger shared subspace is warranted.  When $\hat{s}$ is too small
+(e.g. Figure \ref{fig:ratio-s5}),
+$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$ will be significantly smaller
+than one for at least some of the groups, regardless of $\hat{V}$.
+When $\hat{s}$ is large enough, we are able to use maximum marginal
+likelihood to identify a shared subspace which preserves most of the
+variation in the data for all groups (Figure \ref{fig:ratio-s20}).
+Thus, for a given $\hat{V}\hat{V}^T$, the goodness of fit statistic
+can be used to identify the groups that can be fairly compared on this
+subspace and whether we would benefit from fitting a model with a
+larger value of $\hat{s}$.
 
@@
 
-In the previous section, it was apparent that correct specification
-for the rank of the shared subspace is important for efficient
-inference.  So far, we have assumed that the group rank, $r$, and
-shared subspace dimension, $s$, are fixed and known, but in practice
-this is not the case.  Prior to fitting a model we should estimate
-these quantities.  \citet{Gavish2014} provide an asymptotically
-optimal (in mean squared error) singular value threshold for low rank
-matrix recovery with noisy data.  We use their rank estimator, which
-is a function of the median singular value of the data matrix and the
-ratio $\gamma_k =\frac{p}{n_k}$, to estimate the rank, $r$, of a
-spiked covariance matrix.  Although their estimator was derived for
-individual covariance estimation, we found as a useful heuristic,
-Gavish and Donoho's estimator can also be applied to the pooled data
-to estimate the shared subspace dimension, $s$.  Specifically, we concatenate
-the data from all groups to create an $(\sum_k n_k) \times p$
-dimensional matrix and apply their rank estimator to this matrix to
-choose $s$.
-
-Using these rank estimators, we conduct a simulation which demonstrates
-the relative performance of shared subspace group covariance
-estimation under different data generating models.  In this
-simulation, we consider three different spiked covariance data models:
-1) ``complete pooling'', in which the spiked covariance matrices from
-all groups are identical, e.g.
-$\Sigma_1 = \Sigma_2 = ... \Sigma_K = U\Lambda U^T + \sigma^2I$; 2)
-the ``shared subspace'' model, outlined in Section \ref{sec:shared};
-and 3) ``no pooling'', e.g. the data from each group is generated from
-a spiked covariance matrix with each $U_k$ drawn as an independent
-uniform sample from the Stiefel manifold $\mathcal{V}_{r, p}$.  For
-each of these data generating models, we simulate datasets with
-$p=200$ features, rank $r=2$ and eigenvalues
-$(\lambda_1, \lambda_2) = (250, 25)$.  For data generated under the
-shared subspace model, we additionally assume that the two
-eigenvectors of $\Psi_k$ span a common two dimensional subspace, e.g. we fix
-$s=2$.  We simulate 100 independent datasets for each data generating
-mechanism.
-
-For each data generating mechanism we fit each of three corresponding
-inferential models (``complete pooling'', ``shared subspace'' and ``no
-pooling''). For the complete pooling and no pooling models, which
-involve independent estimation of individual covariance matrices, we
-estimate the rank $r$, and then proceed with Bayesian inference of the
-eigenvectors, eigenvalues, and noise variance.  For the shared
-subspace estimator, we first estimate the shared subspace dimension
-$s$ and then infer the optimal $s$-dimensional shared subspace,
-$\hat{V}\hat{V}^T$, using the EM algorithm presented in Section
-\ref{sec:em}.  We then use Bayesian inference to estimate the
-posterior distribution of the projected covariance matrices
-conditional on $\hat{V}\hat{V}^T$.  In Table \ref{table:groupLoss} we
-report the average Stein's risk, with corresponding 95\% loss
-intervals, for the estimates derived from each inferential model.
+It is apparent that correct specification for the rank of the shared
+subspace is important for efficient inference.  So far in this
+section, we have assumed that the group rank, $r$, and shared subspace
+dimension, $s$, are fixed and known.  In practice this is not the
+case.  Prior to fitting a model we should estimate these quantities.
+\citet{Gavish2014} provide an asymptotically optimal (in mean squared
+error) singular value threshold for low rank matrix recovery with
+noisy data.  We use their rank estimator, which is a function of the
+median singular value of the data matrix and the ratio
+$\alpha_k =\frac{p}{n_k}$, to estimate the rank, $r$, of a spiked
+covariance matrix.  Although their estimator was derived for
+individual covariance estimation, we found that, Gavish and Donoho's
+estimator can also be applied to the pooled data to estimate the
+shared subspace dimension, $s$.  Specifically, we concatenate the data
+from all groups to create an $(\sum_k n_k) \times p$ dimensional
+matrix and apply their rank estimator to this matrix to choose $s$.
+
+Using these rank estimators, we conduct a simulation which
+demonstrates the relative performance of shared subspace group
+covariance estimation under different data generating models.  In
+these simulations we assume the data have $p=200$ features, $r=2$
+spikes, $\sigma^2_k=1$, and $n_k = 50$.  We fix the eigenvalues of
+$\Psi_k$, $(\lambda_1, \lambda_2) = (250, 25)$.  We consider three
+different shared subspace data models: 1) a low dimensional shared
+subspace model with $s=r=2$; 2) a model in which the spiked covariance
+matrices from all groups are identical, e.g.
+$\Sigma_k = \Sigma = U\Lambda U^T + \sigma^2I$; and 3) a full rank
+shared subspace model with $s=p=200$.  We simulate 100 independent
+datasets for each of these data generating mechanisms.
+
+We fit these datasets using three different versions of the shared
+subspace model.  For each of these fits we estimate the number of
+spikes, $r$, from the data.  First, we estimate a single spiked
+covariance matrix from the pooled data and let
+$\hat{\Sigma}_k = \hat{\Sigma}$.  Second, we fit the full rank shared
+subspace model in which we independently estimate each spiked
+covariance matrix (since $s=p$ implies $VV^T = I_p$).  Finally, we use
+an ``adaptive'' shared subspace estimator, in which we estimate both
+$s$ and $r$ and apply empirical Bayes inference for $VV^T$.  In Table
+\ref{table:groupLoss} we report the average Stein's risk and
+corresponding 95\% loss intervals for the estimates derived from each
+inferential model.
 
@@
     \multicolumn{2}{c}{} & \multicolumn{3}{c}{\textbf{Inferential Model}} \\
-  \multicolumn{2}{c|}{}  & Shared Subspace & Complete Pooling & No pooling \\  \cline{2-5}
-    \multirow{3}{*}{\rotatebox[origin=c]{90}{\textbf{Data Model}}} & 
-   Shared Subspace & 0.8 (0.7, 0.9) & 2.1 (1.7, 2.6) & 3.0 (2.9, 3.2) \\ \cline{2-5}
-   & Complete Pooling & 0.8 (0.7, 0.9) & 0.7 (0.6, 0.8) & 3.0 (2.9, 3.2)\\ \cline{2-5}
-   & No pooling & 7.1 (6.2, 8.0) & 138.2 (119, 153) & 3.0 (2.9, 3.2) \\ \cline{2-5}
+  \multicolumn{2}{c|}{}  & Adaptive & $\hat{\Sigma}_k=\hat{\Sigma}$
+                                                           & $\hat{s} = p$ \\  \cline{2-5}
+    \multirow{3}{*}{\rotatebox[origin=c]{90}{\textbf{Data Model}}} 
+& $s=r=2$ & 0.8 (0.7, 0.9) & 2.1 (1.7, 2.6) & 3.0 (2.9, 3.2) \\ \cline{2-5}
+   &   $s=r=2$, $\Sigma_k = \Sigma$ & 0.8 (0.7, 0.9) & 0.7 (0.6, 0.8) & 3.0 (2.9, 3.2)\\ \cline{2-5}
+   &  $s=p=200$ & 7.1 (6.2, 8.0) & 138.2 (119, 153) & 3.0 (2.9, 3.2) \\ \cline{2-5}
   \end{tabular}
-  \caption[Table caption text]{Stein's risk (and 95\% loss intervals) for
-    different inferential and data generating models.  For each of
-    $K=10$ groups, we simulate a $p=200$ dimensional spiked covariance
-    matrix for rank $r=2$.  For data from a shared subspace model the
-    eigenvectors of the covariance matrix share a subspace dimension
-    $s=2$.  All results are based on $n_k = 50$ observations per
-    group.  }
+  \caption[Table caption text]{Stein's risk (and 95\% loss intervals)
+    for different inferential models and data generating models with
+    varying degrees of between-group covariance similarity.  For each
+    of $K=10$ groups, we simulate from spiked covariance models with
+    $p=200$, $r=2$, $\sigma_k^2=1$ and $n_k=50$.  We fit the data to
+    a shared subspace model in which $s$, $r$ and $VV^T$ are all
+    estimated from the data (``adaptive''), a spiked covariance model
+    in which the covariance matrices from each group are assumed to be
+    identical ($\hat{\Sigma}_k=\hat{\Sigma}$) and a model in which we
+    assume the data do \emph{not} share a lower dimensional subspace
+    across groups ($\hat{s} = p$). The estimators which most closely
+    match the data generating model have the lowest risk (diagonal)
+    but the adaptive estimator performs well relative to the
+    alternative misspecified model.}
 \label{table:groupLoss}
@@
 As expected, the estimates with the lowest risk are derived from the
-inferential model that matches the data generating model. However, the
-shared subspace estimator has small risk under model misspecification
-relative to the alternative.  For example, when the data is generated
-using the complete pooling model, the shared subspace estimator has
-almost four times smaller risk than the no pooling estimator.  When
-the data is generated under the no pooling model, the shared subspace
-estimator is over an order of magnitude better than the complete
-pooling estimator.  Thus, shared subspace estimation can be an
-appropriate choice when the amount of similarity between groups is not
-known a priori.
-
-The shared subspace model combined with the rank estimation procedure
-suggested by \citet{Gavish2014} can be widely applied to group
-covariance estimation because the estimator adapts to the amount of
-similarity between groups.  Importantly, the shared subspace model can
-be viewed as a generalization of the standard ``no pooling'' spiked covariance
-model. The two are equivalent when we let the shared subspace span the
-full space, e.g. $V \in \mathcal{O}(p)$.  As shown in Figure
-\ref{fig:sdimension}, the shared subspace model performs nearly as
-well as the no pooling estimator when the subspace dimension is equal to $s = rK$.
-In the simulations used to create Table \ref{table:groupLoss}, we
-typically estimated $s$ to be slightly smaller than $rK$ for the ``no
-pooling'' data, which explains the slightly higher risk for the shared
-subspace model. With improved estimators for $s$, the shared subspace
-model should be even more competitive.
+inferential model that most closely match the data generating
+specifications. However, the adaptive estimator has small risk under
+model misspecification relative to the alternatives.  For example,
+when $\Sigma_k = \Sigma$, the adaptive shared subspace estimator has
+almost four times smaller risk than the full rank estimator, in which
+each covariance matrix is estimated independently.  When we generated
+data from a model with $s=p$, that is, the eigenvectors of $\Psi_k$
+are generated uniformly from $\mathcal{V}_{p,r}$, the adaptive
+estimator is over an order of magnitude better than the estimator
+which assumes no differences between groups.  These results suggest
+that empirical Bayes inference for $VV^T$ combined with the rank
+estimation procedure suggested by \citet{Gavish2014} can be widely
+applied to group covariance estimation because the estimator adapts to
+the amount of similarity across groups.  Thus, shared subspace estimation
+can be especially appropriate choice when the similarity between
+groups is not known a priori.  % In the simulations
+% used to create Table \ref{table:groupLoss}, we typically estimated $s$
+% to be smaller than $rK$ for the data generated under ``full rank''
+% shared subspace model data.  This explains the slightly higher risk
+% for the shared subspace model. With improved estimators for $s$, the
+% shared subspace model should be even more competitive.
 
@@
 and $D$ is an $(s-r)$-dimensional diagonal matrix.  We write
-$V = [V_1\, V_2]$, with $V_1 \in \mathcal{V}_{p,r}$ as the basis for a
+$V = [V_1, V_2]$, with $V_1 \in \mathcal{V}_{p,r}$ as the basis for a
 $r$-dimensional shared subspace that explains the differences between
@@
 $D$, so that the largest eigenvectors of $\Psi_k$ may correspond to
-the eigenvectors $V_2$.
+the eigenvalues from columns of $V_2$.
 
@@
 differences between groups may not manifest themselves in the first
-principle components.  For instance, in genetic analyses of humans,
-there may be large person to person variability common to all groups
-which is unrelated to the grouping of those subjects. Differences
-between groups may still manifest themselves in the smaller principle
+principle components.  For instance, in genetic analyses, there may be
+large between subject variability common to all groups which is
+unrelated to how those subjects are grouped. Differences between
+groups may manifest themselves in the smaller principle
 components. This formulation allows complete pooling of most
 eigenvectors, yet a space to identify and compare relevant
-differences.  Moreover, visualizing and interpreting differences
-between groups on higher dimensional spaces is a challenging problem.
-Using this framework we can easily understand the largest differences
-across groups in a two dimensional space while completely pooling
-remaining anisotropic variability among the groups.
-
-\paragraph{Analysis of Gene Expression Data}
-
-Here, we demonstrate the utility of the shared subspace covariance
+differences.  Using this framework we can easily visualize and compare the
+largest differences between groups in a two dimensional space while
+completely pooling remaining anisotropic variability that is not
+specific to the grouping.
+
+\paragraph{Analysis of Gene Expression Data:}
+
+We demonstrate the utility of the shared subspace covariance
 estimator for exploring differences in the covariability of gene
@@
 acute lymphoblastic leukemia \citep{Yeoh2002}.  The raw data consists
-of gene expression levels for over 12,000 genes in seven different
+of gene expression levels for thousands of genes in seven different
 subtypes of leukemia: BCR-ABL, E2A-PBX1, Hyperdip50, MLL, T-ALL,
@@
 have corresponding sample sizes of $n = (15, 27, 64, 20, 43, 79, 79)$.
-Although there are over 12,000 genes, the vast majority of expression
+Although there are over 12,000 genes in the dataset, the vast majority of expression
 levels are missing.  Thus, we restrict our attention to the genes for
@@
 Panel \ref{fig:leukemiaRatio} shows that over $40\%$ of the variance
-can be explained by the estimated two-dimensional subspace for all groups, with 4 of
-the 6 groups over 60\%.  Panel \ref{fig:leukemiaPosterior} reflects some significant
-differences in the posterior distribution of eigenvalues and
-eigenvectors between groups.  The $x$-axis corresponds to the
-orientation of the principle axis of $\Psi_k$ on $V$ and the $y$-axis
-corresponds to the ratio of the larger eigenvalue over the smaller
-eigenvalue.  Several groups show significantly different orientations
-for the first principle component of the subspace $\hat{V}$.  Further,
-the subtype E2A-PBX1 has a larger eigenvalue ratio which reflects a
-more correlated distribution on the subspace spanned by $V$.  The
-TEL-AML1 and Hyperdip50 appear to cluster together which suggests
-there is little detectable difference between their covariance
-structure.  Note that when the eigenvalue ratio is close to one (as is
-the case for the ``other'' subtype), the distribution of expression
-values on the subspace is nearly spherical and thus the orientation of
-the primary eigenvector is at best weakly identifiable.  This is
-reflected in the wide posterior range of orientations of the first
-principle component.
+can be explained by the estimated two-dimensional subspace for all
+groups, with 4 of the 6 groups over 60\%.  Panel
+\ref{fig:leukemiaPosterior} reflects some significant differences in
+the posterior distribution of eigenvalues and eigenvectors between
+groups.  The $x$-axis corresponds to the orientation of the principle
+axis of $\Psi_k$ on $V$ and the $y$-axis corresponds to the ratio of
+the first eigenvalue to the second.  Several groups have significantly
+different orientations for the first principle component of the
+projected covariance matrix, as well as differences in the ratio of
+eigenvalues.  For instance, the subtype E2A-PBX1 has a larger
+eigenvalue ratio which reflects a more correlated distribution on this
+subspace.  The posterior regions for the TEL-AML1 and Hyperdip50
+groups are almost entirely overlapping, which suggests there is little
+detectable difference between their covariance structures.  Note that
+when the eigenvalue ratio is close to one (as is the case for the
+``other'' subtype), the distribution of expression values on the
+subspace is nearly spherical and thus the orientation of the primary
+eigenvector is at best weakly identifiable.  This is reflected in the
+wide posterior range of orientations of the first principle component.
 
@@
 can be understood in the biplot in Figure \ref{fig:leukemiaBiplot}.
-Here, we plot the contours of the covariance matrices for three of
-subtypes of Leukemia.  We also plot different shapes for the 1\% of genes with
-the largest loadings on the first two columns of $V$ and the
-remaining values as light grey dots.  The genes with the largest
+Here, we plot the contours of the covariance matrices for three
+leukemia subtypes. The 1\% of genes with the largest loadings on the
+first two columns of $\hat{V}$ are indicated with black shapes and the
+remaining loadings with light grey dots.  The genes with the largest
 loadings are clustered by quadrant and listed in the corresponding
@@
         \includegraphics[width=\textwidth]{Figs/leukemiaRatio-04-27}
-        \caption{Proportion of Variance on $\hat{V}$}
+        \caption{$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$}
         \label{fig:leukemiaRatio}
@@
     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/leukemiaPosterior-04-27}
+        \includegraphics[width=\textwidth]{Figs/leukemiaPosterior-06-23}
         \caption{95\% Posterior Regions }
@@
     \end{subfigure}
-\caption{}
+    \caption{a) Goodness of shared subspace fit for each of the seven
+      Leukemia groups.  Over 40\% of the variance is explained on this
+      subspace in each of the seven groups, with as much as 85\%
+      explained in some groups.  b).  95\% posterior regions for the
+      eigenvalue ratio and primary eigenvector orientation.  Regions
+      for some pairs of groups are disjoint, suggesting significant
+      differences in the projected covariance matrices.  For other
+      groups (e.g. Hyperdip50 and TEL-AML1) overlap in the posterior
+      regions indicate that differences are not detectable on this subspace. }
 \label{fig:leukemia}
@@
 In this figure, genes that fall in the upper right quadrant have
-positive loadings in both columns of $V$. Genes in the upper left
-corner of the figure have positive loadings on the first column of $V$
-but negative loadings on the second column of $V$.  The principle axis
+positive loadings in both columns of $\hat{V}$. Genes in the upper left
+corner of the figure have positive loadings on the first column of $\hat{V}$
+but negative loadings on the second column of $\hat{V}$.  The principle axis
 for a group aligns in the direction of genes that exhibit the largest
@@
 to the principle axis exhibit reduced variability, relative to the
-other groups. For instance, genes in the upper left corner of this
+other groups. As an example, genes in the upper left corner of this
 plot (triangles) exhibit large, positively correlated variability in
@@
 a direction orthogonal to the principle axis.  In contrast, the
-BCR-ABL group is aligned primarily with the second column of $V$,
+BCR-ABL group is aligned primarily with the second column of $\hat{V}$,
 which means that the genes indicated by squares and circles vary
 significantly in this group.  Genes in the square group are
-anti-correlated with those in the circle group, since the their
-loadings have opposite signs on $V2$.
+anti-correlated with those in the circle group, since their
+loadings have opposite signs on the second column of $\hat{V}$.
 
@@
 
-\paragraph{Analysis of Metabolomic Data}
-
-Next, we apply shared subspace group covariance estimation to a
-metabolomic data analysis on fly aging \citep{Hoffman2014}.  We bin
-the data to include groups of flies less than 10 days (young), between
-10 and 51 days (middle) and greater than 51 days (old), and further
-split by gender. The sample sizes range from 43 to 59 flies per group.
-We analyze metabolomic data corresponding to metabolites with 3714
-mass-charge ratios.  After removing mean effects due to age and gender
-we fit the shared subspace estimator and compare covariances between
-groups.  Interestingly, we identify a subspace that explains almost
-all of the variability in the first few components of $\Sigma_k$ for
-all groups (Figure \ref{fig:dmelanRatio}). However, we see little
-evidence for differences in metabolite covariances on this subspace,
-as evidenced by the large overlap in posterior distributions for each group
-(Figure \ref{fig:dmelanPosterior}). This is indicative of large
-amount of variation which is common across all groups for the first
-$s$ principle components. Differences in covariation across may
-manifest themselves for the smaller principle components, but are too
+\paragraph{Analysis of Metabolomic Data:}
+
+Next, we briefly discuss an example in which we could not identify
+significant covariance differences between groups.  Here, we apply
+shared subspace group covariance estimation to a metabolomic data
+analysis on fly aging \citep{Hoffman2014}.  We bin the data to include
+groups of flies less than 10 days (young), between 10 and 51 days
+(middle) and greater than 51 days (old), and further split by
+gender. The sample sizes range from 43 to 59 flies per group.  We
+analyze metabolomic data corresponding to metabolites with 3714
+mass-charge ratios.  After removing mean effects due to age and
+gender, we fit the shared subspace estimator and compare covariances
+between groups.  We identify a subspace that explains almost all of
+the variability in the first few components of $\Sigma_k$ for all
+groups (Figure \ref{fig:dmelanRatio}) but see little evidence for
+differences in metabolite covariances on this subspace, as evidenced
+by the large overlap in posterior distributions for each group (Figure
+\ref{fig:dmelanPosterior}). This could be indicative of a large amount
+of variation that is common across all groups for the first $\hat{s}$
+principle components. Differences in covariation across groups likely
+manifest themselves in the smaller principle components but are too
 small to detect given the sample sizes.
@@
         \includegraphics[width=\textwidth]{Figs/dmelanRatio-04-28}
-        \caption{Ratio}
+        \caption{$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$}
         \label{fig:dmelanRatio}
@@
     \begin{subfigure}[b]{0.45\textwidth}r
-        \includegraphics[width=\textwidth]{Figs/dmelanPosterior-04-28}
-        \caption{Posterior}
+        \includegraphics[width=\textwidth]{Figs/dmelanPosterior-06-23}
+        \caption{95\% Posterior Regions}
         \label{fig:dmelanPosterior}
     \end{subfigure}
-    \caption{a) Proportion of variance on $\hat{V}$ is close to one
-      for all groups suggesting that blah. However, there is
-      significant overlap in the posterior distribution of $\Lambda_k$
-      and $U_k$. This suggests that there is little difference in the
-      covariance matrices between groups, on this particular subspace.  }
+    \caption{a) Goodness of fit is close to one for all groups
+      suggesting that the inferred subspace preserves most of the
+      variability across groups. b) There is significant overlap in
+      the posterior distribution of the eigenvalue ratio and primary
+      eigenvector orientation across groups. Thus, there is little
+      evidence of significant differences in the covariance matrices
+      between groups on this inferred subspace.  }
 \end{figure}
@@
 comparing differences in covariance matrices across multiple groups on
-a common low dimensional subspace.  We described an EM algorithm for
-estimating this common subspace and a MCMC algorithm for inferring the
-projected covariance matrices and their associated uncertainty.
-Estimates of both the shared subspace and the projected covariance
-matrices can be useful summaries of the data.  For example, with the
-biological data, the shared subspace highlights the full set of genes
-or molecules that might be correlated in any group.  Differences between
-group covariance matrices can be understood in terms of differences in
-these sets of correlated and anti-correlated genes.  In this applied
-analysis, we demonstrated how we can use these notions to visualize
-and contrast the posterior distributions of covariance matrices
-projected onto a particular subspace.
-
-In simulation, we showed that the shared subspace model can
-still be a reasonable choice for modeling multi-group covariance
-matrices even when we have little knowledge, a priori, of how similar
-the groups are.  Even when there is little to no similarity, the shared
-subspace model is appropriate, as long as the dimension of the shared
-subspace is large enough.  However, selecting the rank of the shared
-subspace remains a significant practical challenge.  Although we
-propose a useful heuristic for choosing the dimension of the shared
-subspace based on the rank selection estimators of \citet{Gavish2014},
-a more principled approach is warranted.
-
-It was also a challenge to estimate the subspace $VV^T$ once we had
-chosen a rank.  We used an empirical bayes approach, we estimate the shared subspace
-$VV^T$ separately from the MCMC inference of $\Psi_k$ because we found
-estimation of $VV^T$ using MCMC inference to be
-particularly difficult.
-
-Full uncertainty
-quantification for $VV^T$ may be desirable but we found this
-challenging for the dimensions of the problems we would like to
-consider.  Further, by focusing on group differences on a fixed
-subspace, it is simpler to interpret differences across covariance
-matrices.
-
- Although fixing the shared subspace
-simplifies interpretation of the results, in many cases it may be
-desirable to develop a fully Bayesian approach for jointly estimating
-$VV^T$ and $\Psi_k$.  Recent Markov chain Monte Carlo algorithms, like
-Riemmanian manifold Hamilton Monte Carlo, which can exploit the
-geometry of the Grassmann manifold, may be useful here \citep{Byrne2013,
-  Girolami2011}.  It may also be possible, though computationally
-intensive, to jointly estimate the rank of the shared subspace, using
-for instance, a reversible-jump MCMC algorithm.
+a common low dimensional subspace.  We described an empirical Bayes
+algorithm for estimating this common subspace and a Gibbs sampler for
+inferring the projected covariance matrices and their associated
+uncertainty.  Estimates of both the shared subspace and the projected
+covariance matrices can both be useful summaries of the data.  For
+example, with the biological data, the shared subspace highlights the
+full set of genes or metabolites that are correlated across group.
+Differences between group covariance matrices can be understood in
+terms of differences in these sets of correlated and
+molecules.  In the applied analysis, we demonstrated how
+we can use these notions to visualize and contrast the posterior
+distributions of covariance matrices projected onto a particular
+subspace.
+
+In simulation, we showed that the shared subspace model can still be a
+reasonable choice for modeling multi-group covariance matrices even
+when the groups may be largely dissimilar.  When there is little
+similarity between groups, the shared subspace model can still be appropriate as
+long as the dimension of the shared subspace is large enough.
+However, selecting the rank of the shared subspace remains a practical
+challenge.  Although we propose a useful heuristic for choosing the
+dimension of the shared subspace based on the rank selection
+estimators of \citet{Gavish2014}, a more principled approach is
+warranted.  Improved rank estimators would further improve the
+performance of the adaptive shared subspace estimator discussed in
+Section \ref{sec:simulation}.
+
+It is also a challenging problem to estimate the ``best'' subspace,
+$VV^T$, once the rank of the space is specified.  We used maximum
+marginal likelihood to estimate the shared subspace $VV^T$ and then
+use MCMC to infer $\Psi_k$.  By focusing on
+group differences for $\Psi_k$ on a \emph{fixed} subspace, it is much
+simpler to interpret similarities and differences.  Nevertheless, full
+uncertainty quantification for $VV^T$ can be desirable.  We found
+MCMC inference for $VV^T$ to be challenging for the problems considered
+in this paper and leave it for future work to develop an efficient
+fully Bayesian approach for jointly estimating $VV^T$ and $\Psi_k$.
+Recently developed Markov chain Monte Carlo algorithms, like Riemmanian manifold
+Hamilton Monte Carlo, which can exploit the geometry of the Grassmann
+manifold, may be useful here \citep{Byrne2013, Girolami2011}.  It may
+also be possible, though computationally intensive, to jointly
+estimate the $s$ and $VV^T$ using for instance, a
+reversible-jump MCMC algorithm.
 
@@
 Fundamentally, our approach is quite general and can be integrated
-with other existing approaches for hierarchical multi-group covariance
+with existing approaches for hierarchical multi-group covariance
 estimation.  In particular, we can incorporate additional shrinkage on
@@
 eigenvectors of $\Psi_k$ or model $\Psi_k$ as a function of continuous
-covariates as in \citet{Yin2010, Hoff2011}.  Alternatively, we can
+covariates as in \citet{Yin2010} and \citet{Hoff2011}.  Alternatively, we can
 summarize the estimated covariance matrices by thresholding entries of
 the precision matrix, $\Psi_k^{-1}$ to visualize differences between
-groups using a graphical model \citet{Meinshausen2006}.  The specifics
+groups using a graphical model \citep{Meinshausen2006}.  The specifics
 of the problem at hand should dictate which shrinkage models are
-appropriate, but the shared subspace assumption should be useful in a
+appropriate, but the shared subspace assumption can be useful in a
 wide range of analyses, especially when the number of features is very

### 2016-07-02 paper/mscov.tex
Revised goodness of fit and priors (3.1)

@@
 
-\title{A Shared Subspace Model for Multi-Group Covariance Estimation}
-\author{Alexander Franks and Peter Hoff}
-\date{\today}
+\title{Shared Subspace Models for Multi-Group Covariance Estimation
+%
+  \protect\thanks{Alexander M. Franks is a Moore/Sloan Data Science
+    and WRF Innovation in Data Science Postdoctoral Fellow and a
+    graduate of the Department of Statistics at Harvard University
+    (\href{mailto:amfranks@uw.edu}{amfranks@uw.edu}).  Peter Hoff is a
+    Professor of Statistics at The University of Washington
+    (\href{mailto:pdhoff@uw.edu}{pdfhoff@uw.edu}).
+%
+This work was partially supported 
+ by the Washington Research Foundation Fund for Innovation in Data-Intensive Discovery and the Moore/Sloan Data Science Environments Project at the University of Washington.
+%
+ The authors are grateful to Dr.\ Daniel Promislow (Department of
+ Pathology, University of Washington) and Dr.\ Jessica Hoffman
+ (University of Alabama at Birmingham) for sharing data and ideas
+ which contributed to the framing of the ideas this paper.}}
+\author{Alexander Franks and Peter Hoff} \date{\today}
 \maketitle
@@
 
-  We develop a model-based method for evaluating similarities and
-  differences among several $p\times p$ covariance matrices in the
-  $p \gg n$ setting.  This is done by assuming a spiked covariance
-  model for each group and sharing information about the space spanned
-  by the group-level eigenvectors.  We use empirical Bayes to identify
-  a low-dimensional subspace which explains variation across all
-  groups and use an MCMC algorithm to estimate the posterior
-  uncertainty of eigenvectors and eigenvalues on this subspace.  The
-  implementation and utility of our model is illustrated with analyses
-  of high-dimensional multivariate gene-expression data.
-
+  We develop a model-based method for evaluating heterogeneity among
+  several $p\times p$ covariance matrices in the big p, small n setting.
+  This is done by assuming a spiked covariance model for each group
+  and sharing information about the space spanned by the group-level
+  eigenvectors.  We use an empirical Bayes method to identify a low-dimensional
+  subspace which explains variation across all groups and use an MCMC
+  algorithm to estimate the posterior uncertainty of eigenvectors and
+  eigenvalues on this subspace.  The implementation and utility of our
+  model is illustrated with analyses of high-dimensional multivariate
+  gene expression and metabolomics data.
+\vfill
+\noindent\textbf{Keywords:} covariance estimation; spiked covariance
+model; Stiefel manifold; Grassmann manifold; big p, small n;
+high-dimensional data; empirical Bayes.
 \end{abstract}
@@
 
-In modern analyses, multivariate data is often partitioned into
+Multivariate data is often partitioned into
 groups, each of which represent samples from populations with
@@
 populations, there has been a growing need to identify differences in
-population covariances as well.  For instance, in case
-control-studies, mean level effects may be small relative to subject
+population covariances as well.  For instance, in case-control studies, mean level effects may be small relative to subject
 variability; distributional differences between groups may still be
@@
 part of many prediction tasks (e.g. quadratic discriminant analysis).
-Thus, evaluating similarities and differences between covariance
+Thus, evaluating heterogeneity between covariance
 matrices can be an important complement to more traditional analyses
@@
 
-To address this needs, we develop a novel method for
+To address this need, we develop a novel method for
 group-covariance estimation.  In our method, we exploit the fact that
@@
 using Gaussian graphical models. For Gaussian graphical models, zeros
-in the precision matrix correspond to conditional independences
+in the precision matrix correspond to conditional independencies
 between pairs of features given the remaining features
@@
 (PLS). This approach, which is a special case of a bilinear factor
-model, involves projecting the data into a lower dimension space which
+model, involves projecting the data onto a lower dimensional space which
 maximizes the similarity of the two groups.  This technique does not
@@
 covariance matrices across groups.  For each group, we assume ``the
-spiked covariance model'', a well studied variant of the factor model
-\citep{Johnstone2001}.  In Section \ref{sec:shared} we briefly review
-the behavior of spiked covariance models for estimating a single
-covariance matrix and then introduce our extension to the multi-group
-setting.  In Section \ref{sec:inference} we describe an empirical
-Bayes algorithm for inferring the shared subspace and estimating the
-posterior distribution of the covariance matrices of the data
-projected onto this subspace.  In Section \ref{sec:simulation} we
-investigate the behavior of these models in simulation and demonstrate
-how the shared subspace is widely applicable, even when there is
-little similarity in the covariance matrices across groups.  We then
-demonstrate the utility of the model in an analysis of gene expression
-data from juvenile leukemia patients in Section \ref{sec:app}.  Despite the
-large feature size ($p > 3000$) and small sample size ($n < 100$ per
-group), we identify interpretable similarities and differences in gene
-covariances on a low dimensional subspace.
-
-\section{The Shared Subspace Spiked Covariance Model}
+spiked covariance model'' (also known as the ``partial isotropy
+model''), a well studied variant of the factor model
+\citep{Mardia1980, Johnstone2001}.  In Section \ref{sec:shared} we
+briefly review the behavior of spiked covariance models for estimating
+a single covariance matrix and then introduce our extension to the
+multi-group setting.  In Section \ref{sec:inference} we describe an
+empirical Bayes algorithm for inferring the shared subspace and
+estimating the posterior distribution of the covariance matrices of
+the data projected onto this subspace.  In Section
+\ref{sec:simulation} we investigate the behavior of this class of
+models in simulation and demonstrate how the shared subspace
+assumption is widely applicable, even when there is little similarity
+in the covariance matrices across groups.  In particular, shared
+subspace models include those in which there is no information shared
+across groups because in the worst case we can assume the subspace
+spanned by the group-level eigenvectors is simply the full
+$p$-dimensional space.  We then demonstrate the utility of the model
+in an analysis of gene expression data from juvenile leukemia patients
+in Section \ref{sec:app}.  Despite the large feature size ($p > 3000$)
+and small sample size ($n < 100$ per group), we identify interpretable
+similarities and differences in gene covariances on a low dimensional
+subspace.  Finally, we conclude with a brief exploration of
+heterogeneity in metabolomic data from a study on aging in the fly
+species \textit{Drosophila melanogaster}.
+
+\section{A Shared Subspace Spiked Covariance Model}
 \label{sec:shared}
@@
 
-Suppose $\bl S$ has a possibly degenerate Wishart$(\Sigma,n$)
+Suppose $S$ has a possibly degenerate Wishart$(\Sigma,n$)
 distribution with density given by
 \begin{equation} 
-p(\bl S | \Sigma, n) \propto l(\Sigma:\bl S) =  |\Sigma|^{-n/2} \text{etr}( - \Sigma^{-1} \bl S/2 ) ,  
+p(S | \Sigma, n) \propto l(\Sigma: S) =  |\Sigma|^{-n/2} \text{etr}( - \Sigma^{-1}  S/2 ) ,  
 \label{eqn:lik}
@@
 \noindent where $\Sigma \in \mathcal S_p^+$ and $n$ may be less than $p$.  Such
-a likelihood results from $\bl S$ being, for example, a residual sum
+a likelihood results from $S$ being, for example, a residual sum
 of squares matrix from a multivariate regression analysis. In this
@@
 \begin{equation} 
-\Sigma = \sigma^2 (  \bl U  \Lambda  \bl U^T  + \bl I )
+\Sigma = \sigma^2 ( U  \Lambda  U^T  + I )
 \label{eqn:spiked}
@@
 \noindent where for $r \ll p$, $\Lambda$ is an $r\times r$ diagonal
-matrix and $\bl U \in \mathcal V_{r,p}$.  $\mathcal V_{r,p}$ is the
+matrix and $U \in \mathcal V_{r,p}$.  $\mathcal V_{r,p}$ is the
 Stiefel manifold, which consists of the set of $p \times r$
-orthonormal matrices in $\mathbb{R}^p$.  % The spiked covariance model
-% is thus a low rank factor model with homoscedastic error. 
-The spiked covariance formulation is appealing because it explicitly
+orthonormal matrices in $\mathbb{R}^p$, so that $U^TU =
+I$.  The spiked covariance formulation is appealing because it explicitly
 partitions the covariance matrix into a tractable low rank ``signal''
@@
 if $p/n \rightarrow \alpha >0$ as $n\rightarrow \infty$, the $k$th
-largest eigenvalue of $\bl S/(n\sigma^2)$ will converge to an upwardly
+largest eigenvalue of $S/(n\sigma^2)$ will converge to an upwardly
 biased version of $\lambda_{k}+1$, if $\lambda_k$ is greater than
@@
 being used.  These shrinkage functions are shown to be asymptotically
-optimal in the $p/n\rightarrow \alpha$ setting.  Note that covariance
-estimators of this form are are equivariant with respect to rotations
-and scale changes.  The situation should be different, however, when
-we are interested in estimating multiple covariance matrices from
-distinct, but related groups.  Here, group-level equivariance to
-rotations is an unreasonable assumption; both eigenvalue \emph{and}
-eigenvector shrinkage can play an important role in improving
-covariance estimates.
+optimal in the $p/n\rightarrow \alpha$ setting.
+
+Note that covariance estimators of this form are are equivariant with
+respect to rotations and scale changes.  The situation should be
+different, however, when we are interested in estimating multiple
+covariance matrices from distinct, but related groups.  Here,
+group-level equivariance to rotations is an unreasonable assumption;
+both eigenvalue \emph{and} eigenvector shrinkage can play an important
+role in improving covariance estimates.  Consider multi-group
+covariance estimation based on $K$ matrices, $Y_1, ..., Y_K$, where
+$Y_k$ is assumed to be an $n_k$ by $p$ matrix of mean-zero normal
+data, typically with $n_k \ll p$.  Then, $Y_k^TY_k = S_k$ has a
+(degenerate) Wishart distribution as in Equation \ref{eqn:lik}.  To
+improve estimation, we seek estimators that share information across
+groups, that is, where the estimator $\hat{\Sigma}_k$ may depend on
+data from all groups.  To this end, we extend the spiked covariance
+model to grouped data, by assuming that the anisotropic variability
+from each group occurs on a common low dimensional subspace.
+Specifically, we posit that the covariance matrix for each group can
+be written as
+\begin{equation}
+\Sigma_k = V\Psi_kV^T + \sigma^2_kI,
+\label{eqn:sspsi}
+\end{equation}
 
@@
         \caption{Projection in $\mathcal{R}^3$}
-        \label{fig:dmelanRatio}
+        \label{fig:3dplot}
     \end{subfigure}
@@
         \caption{$Y_kV$}
-        \label{fig:dmelanPosterior}
+        \label{fig:2dscatter}
     \end{subfigure}
@@
         \caption{$Y_kV_{\perp}$}
-        \label{fig:dmelanPosterior}
+        \label{fig:2dscatterorth}
     \end{subfigure}
@@
 \end{figure}
-
-For multi-group covariance estimation, we assume that we have $K$
-matrices, $Y_1, ... Y_K$, where $Y_k$ consists
-of $n_k$ independent rows of mean-zero normal data of dimension $p$,
-typically with $n_k \ll p$.  Then, $Y_k^TY_k = \bl S_k$ has a
-(degenerate) Wishart distribution as in Equation \ref{eqn:lik}.  To
-improve estimation, we seek estimators for the population covariance
-from one group that depend on the data from all $K$ groups, e.g. of
-the form $\hat \Sigma_k(\bl S_1, ... \bl S_K)$.  To this end, we
-extend the spiked covariance model to grouped data, by assuming that
-the anisotropic variability from each group occurs on a common low
-dimensional subspace.  Specifically, we posit that the covariance
-matrix for each group can be written as
-
-\begin{equation}
-\Sigma_k = V\Psi_kV^T + \sigma^2_kI.
-\label{eqn:sspsi}
-\end{equation}
-%
-\noindent Here, the columns of $V \in \mathcal{V}_{p, s}$ determine a subspace
-shared by all groups $k$ with $s \geq p$.  Throughout this paper we
-will denote to the shared subspace as $VV^T \in \mathcal G_{p,s}$, where
-$\mathcal G_{p,s}$, known as the Grassmann manifold, consists of all
-$s$-dimensional linear subspaces of $\mathcal{R}^p$
-\citep{Chikuse2012}.  Although $V$ is only identifiable up to right
-rotations, the matrix $VV^T$, which defines the plane of variation shared by all
-groups, is identifiable.  Later, to emphasize the connection to the
-spiked PCA model (\ref{eqn:spiked}), we will write $\Psi_k$ in terms
-of its eigendecomposition, $\Psi_k = O_k\Lambda_kO_k$, where $O_k$ are
+% For multi-group covariance estimation, we assume that we have $K$
+% matrices, $Y_1, ... Y_K$, where $Y_k$ consists of $n_k$ independent
+% rows of mean-zero normal data of dimension $p$, typically with
+% $n_k \ll p$.
+
+%
+\noindent where, the columns of $V \in \mathcal{V}_{p, s}$, with
+$s \ll p$, determine a subspace shared by all groups.  Throughout this
+paper we will denote to the shared subspace as
+$VV^T \in \mathcal G_{p,s}$, where $\mathcal G_{p,s}$, known as the
+Grassmann manifold, consists of all $s$-dimensional linear subspaces
+of $\mathbb{R}^p$ \citep{Chikuse2012}.  Although $V$ is only
+identifiable up to right rotations, the matrix $VV^T$, which defines
+the plane of variation shared by all groups, is identifiable.  Later,
+to emphasize the connection to the spiked PCA model
+(\ref{eqn:spiked}), we will write $\Psi_k$ in terms of its
+eigendecomposition, $\Psi_k = O_k\Lambda_kO_k$, where $O_k$ are
 eigenvectors and $\Lambda_k$ are the eigenvalues of $\Psi_k$ (see
@@
 
-Note that for the shared subspace model,
+Note that for a shared subspace model,
 $V^T\Sigma_kV = \Psi_k + \sigma_k^2I$ is an anisotropic
-$s$-dimensional covariance matrix for the projected data, $Y_kV$.
-In contrast, the data projected onto the orthogonal space,
-$Y_kV_{\perp}$, is isotropic, and assumed to be small, for all
-groups.  In Figure \ref{fig:shared} we provide a simple illustration
-using simulated $4$-dimensional data from two groups.  In this example, the
+$s$-dimensional covariance matrix for the projected data, $Y_kV$.  In
+contrast, the data projected onto the orthogonal space,
+$Y_kV_{\perp}$, is isotropic for all groups.  In Figure
+\ref{fig:shared} we provide a simple illustration using simulated
+$4$-dimensional data from two groups.  In this example, the
 differences in distribution between the groups of data can be
@@
 the two groups manifest themselves on this shared subspace, whereas
-only the magnitude of the isotropic variability can differ between groups
-on the orthogonal space.  Thus, the the shared subspace model can be
-viewed as a covariance partition model, where one partition includes
-the anisotropic variability from all groups and the other partition is
-constrained to the isotropic variability from each group.  This
-isotropic variability is often characterized as measurement noise.
+only the magnitude of the isotropic variability can differ between
+groups on the orthogonal space.  Thus, a shared subspace model
+can be viewed as a covariance partition model, where one partition
+includes the anisotropic variability from all groups and the other
+partition is constrained to the isotropic variability from each group.
+This isotropic variability is often characterized as measurement
+noise.
 
@@
 Section \ref{sec:simulation}, if the spiked covariance model holds for
-each group individually, then the shared subspace model also
-holds because we can take the shared subspace to be the full
-$p$-dimensional space. In practice, we can often identify a shared
+each group individually, then the shared subspace assumption also
+holds because at worst we can take the shared subspace ot be the full
+$p$-dimensional space. In practice, we can usually identify a shared
 subspace of dimension $s \ll p$ that preserves most of the variation
@@
 groups may not actually reflect statistically significant differences,
-in Section \ref{sec:bayes} we describe a Gibbs sampler that can be
-used to generate estimates of the projected covariance matrices,
-$\Psi_k$ and and their associated uncertainty.
+in Section \ref{sec:bayes} we also describe a Gibbs sampler that can
+be used to generate estimates of the projected covariance matrices,
+$\Psi_k$, and and their associated uncertainty.  Later, we discuss
+strategies for infering an appropriate value for $s$ and explore how
+shared subspace models can be used for exploratory data analysis by
+visualizing covariance heterogeneity on two or three dimensional
+subspaces.
 
@@
 \begin{align}
-\nonumber p(S_1, ... S_k | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}\etr(-\Sigma_k^{-1}\mathbf{S}_k/2)  \\
+\nonumber p(S_1, ... S_k | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}\etr(-\Sigma_k^{-1}S_k/2)  \\
 \nonumber &\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-(\sigma_k^2(V\Psi_kV^T +
-  I))^{-1}\mathbf{S}_k/2) \\
+  I))^{-1}S_k/2) \\
 \nonumber &\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}\etr(-\left[V(\Psi_k +
-  I)^{-1}/\sigma_k^2 V^T + (I-VV^T)/\sigma^2_k\right]\mathbf{S}_k/2)
+  I)^{-1}/\sigma_k^2 V^T + (I-VV^T)/\sigma^2_k\right]S_k/2)
   \\
-&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-s)/2}|M_k|^{-n_k/2}\etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]\mathbf{S}_k/2) ,
+&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-s)/2}|M_k|^{-n_k/2}\etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]S_k/2) ,
 \end{align}
 %
-\noindent where we define $M_k^{-1} = (\Psi_k + I) ^{-1}/\sigma_k^2$.  The log-likelihood in
+\noindent where we define $M_k = \sigma_k^2(\Psi_k + I)$.  The log-likelihood in
 $V$ (up to an additive constant) is
-
+%
 \begin{align}
 \nonumber l(V) &= \sum_k \tr\left(-VM_k^{-1}V^T +
-       VV^T/\sigma^2_k\right)\mathbf{S}_k/2)\\
-&= \sum_k \frac{1}{2}\tr\left(-M_k^{-1}V^T \mathbf{S}_kV\right) + \frac{1}{2\sigma_k^2}tr\left(VV^T \mathbf{S}_k\right).
+       VV^T/\sigma^2_k\right)S_k/2)\\
+&= \sum_k \frac{1}{2}\tr\left(-M_k^{-1}V^T S_kV\right) + \frac{1}{2\sigma_k^2}tr\left(VV^T S_k\right).
 \end{align}
@@
 
-We propose an expectation-maximization (EM) algorithm, to estimate the
+We propose an expectation-maximization (EM) algorithm to estimate the
 relevant quantities.  In particular we maximize the marginal
@@
 the ``missing'' parameters to be imputed.  We assume independent
-Jeffreys' prior distributions for both $\frac{1}{\sigma_k^2}$ and
-$M_k^{-1}$.  Jeffreys' prior for these quantities corresponds to
-$p(\frac{1}{\sigma_k^2}) \propto \sigma_k^2$ and
-$p(M_k^{-1}) \propto |M_k^{-1}|^{-(s+1)/2}$.  From the likelihood it
-can easily be shown that the conditional posterior for $M_k^{-1}$ is
-
-$$p(M_k^{-1} | V) \propto |M_k^{-1}|^{(n_k - s -1)/2}\etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) $$
-
-\noindent which is a Wishart($V^T\mathbf{S}_kV$, $n_k$) distribution.  The
-conditional posterior distribution of $\frac{1}{\sigma_k^2}$ is simply
-
-$$p\left(\frac{1}{\sigma^2} | V\right) \propto \left(\frac{1}{\sigma_k^2}\right)^{n_k(p-s)/2-1}\etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)\mathbf{S}_k\right)  $$
-
-\noindent which is a gamma($n_k(p-r)/2$,
-$\frac{1}{2}tr((I-VV^T)\mathbf{S}_k)$) distribution.  Based on these
+Jeffreys' prior distributions for both $\sigma_k^2$ and
+$M_k$.  Jeffreys' prior for these quantities corresponds to
+$p(\sigma_k^2) \propto 1/\sigma_k^2$ and
+$p(M_k) \propto |M_k|^{-(s+1)/2}$.  From the likelihood it
+can easily be shown that the conditional posterior for $M_k$ is
+%
+$$p(M_k | V) \propto |M_k|^{(n_k + s + 1)/2}\etr(-1/2(M_k^{-1}V^TS_kV)) $$
+%
+\noindent which is an inverse-Wishart($V^TS_kV$, $n_k$) distribution.  The
+conditional posterior distribution of $\sigma_k^2$ is simply
+%
+$$p\left(\sigma^2 | V\right) \propto (\sigma_k^2)^{-n_k(p-s)/2-1}\etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)S_k\right)  $$
+%
+\noindent which is an inverse-gamma($n_k(p-s)/2$,
+$\frac{1}{2}\text{tr}((I-VV^T)S_k)$) distribution.  Based on these
 results the following EM algorithm can be used:
@@
 \begin{enumerate}
-\item $E[M_k^{-1} | V_{(t-1)}] = n_k(V_{(t-1)}^T \mathbf{S}_kV_{(t-1)})^{-1}$
-\item $E[\frac{1}{2\sigma_k^2}|V_{(t-1)}] = \frac{n_k(p-r)}{tr((I-V_{(t-1)}V_{(t-1)}^T)S_k)}$
+\item $E[M_k^{-1} | V_{(t-1)}] = n_k(V_{(t-1)}^T S_kV_{(t-1)})^{-1}$
+\item $E[\frac{1}{\sigma_k^2}|V_{(t-1)}] = \frac{n_k(p-s)}{tr((I-V_{(t-1)}V_{(t-1)}^T)S_k)}$
 \end{enumerate}
 \item Compute $V_{t}$ = $\underset{V}{\text{arg } \text{max}}  \sum_k tr\left(-VE[M^{-1}|V_{(t-1)}]^T +
-       E[\frac{1}{2\sigma_k^2}|V_{(t-1)}]VV^T\right)\mathbf{S}_k/2)$ 
+       E[\frac{1}{2\sigma_k^2}|V_{(t-1)}]VV^T\right)S_k/2)$ 
 \end{enumerate}
@@
 If $V$ is a basis for a shared subspace, then for each group $k$, most
-of the variation in $Y_k$ should be preserved when projecting the data
+of the non-isotropic variation in $Y_k$ should be preserved when projecting the data
 to this space.  To characterize the extent to which this is true for
@@
 for the ratio of the sum of the first $s$ eigenvalues of
-$\hat{V}^T\bl \Sigma_k \hat{V}$ to the sum of the first $s$
+$\hat{V}^T \Sigma_k \hat{V}$ to the sum of the first $s$
 eigenvalues of $\Sigma_k$:
@@
 \begin{equation}
-\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2) = \frac{||Y_k\hat{V}||_F}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
-  ||Y_k\widetilde{V}||_F - \hat{\sigma}_k^2ps/n_k}
+\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2) = \frac{||Y_k\hat{V}||_F/n_k}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
+  ||Y_k\widetilde{V}||_F/n_k - \hat{\sigma}_k^2ps/n_k}
 \label{eqn:ratio}
@@
 %
-\noindent where $||.||_F$ is the Frobenius norm. Our motivation for
+\noindent where $||\cdot||_F$ is the Frobenius norm. Our motivation for
 this statistic is the fact that if
 $\Sigma_k = \hat{V}\Psi_k\hat{V}^T + \sigma_k^2I$ then
-$\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2) \approx 1$.  To see this, note
-that $\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}\frac{||Y_k\widetilde{V}||_F}{n_k}$
+$\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2) \approx 1$.  We demonstrate
+this in an asymptotic regime where $s$ is fixed, $p$,
+$n_k \rightarrow \infty$ and $p/n_k = \alpha_k$ constant.
+
+First, since $s$ is fixed and $n_k$ is growing, the numerator,
+$||Y_k\hat{V}||_F/n_k$, is a consistent estimator for the sum of the
+eigenvalues of $\hat{V}^T\Sigma_k\hat{V}$.  The denominator,
+$\underset{\widetilde{V} \in \mathcal{V}_{p,
+    s}}{\text{max}}||Y_k\widetilde{V}||_F/n_k$
 is equivalent to the sum of the first $s$ eigenvalues of the sample
-covariance matrix $\bl S_k/n_k$.  With $\hat{\lambda}^{(k)}_i$
-the $i$-th eigenvalue of $\bl S_k/n_k$, \citet{Baik2006}
-demonstrated that asymptotically as $p, n_k \rightarrow \infty$ with
-$p/n_k = \alpha_k$ fixed
-
+covariance matrix $S_k/n_k$.  With $\hat{\lambda}^{(k)}_i$ the
+$i$-th eigenvalue of $S_k/n_k$, \citet{Baik2006} demonstrated that
+asymptotically as $p, n_k \rightarrow \infty$ with $p/n_k = \alpha_k$
+fixed
+%
 \begin{eqnarray}
 \nonumber \hat{\lambda}^{(k)}_i &\rightarrow& \lambda^{(k)}_i\left(1 +
-                                    \frac{\sigma_k^2\alpha}{\lambda^{(k)}_i
+                                    \frac{\sigma_k^2\alpha_k}{\lambda^{(k)}_i
                                     - 1}\right)\\
@@
 \end{eqnarray}
-
-Thus,
+%
+\noindent where the approximation in the second line is due to the
+fact that $\lambda_k/(\lambda_k - 1) \approx 1$ for $\lambda_k$
+large. Thus,
 $\sum_i^s \hat{\lambda}^{(k)}_i \approx \sum_i^s \lambda^{(k)}_i +
-\sigma_k^2ps/n_k$.
-If we assume that $s$ is fixed in this asymptotic regime, then
-$||Y_kV||_F/n_k$ is a consistent estimator for
-$\sum_i^s \lambda^{(k)}_i$.  As a consequence, the proposed estimator
-will be close to one for all groups $k$ when $\hat{V}\hat{V}^T = VV^T$
-and smaller if not.  The metric can provides a useful indicator for
-which groups can be reasonably compared on a given subspace and which
-groups cannot.  When $\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2)$ is small
-for some groups, it may suggest that the rank, $s$, of the inferred
+\sigma_k^2ps/n_k$
+and $\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
+  ||Y_k\widetilde{V}||_F/n_k - \hat{\sigma}_k^2ps/n_k$ is a reasonable approximation to the sum of the
+first $s$ eigenvalues of $\Sigma_k$.
+
+As a consequence, the proposed statistic will be
+close to one for all groups when $\hat{V}\hat{V}^T = VV^T$ and smaller
+if not.  The metric can provides a useful indicator for which groups
+can be reasonably compared on a given subspace and which groups
+cannot.  When $\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2)$ is small for
+some groups, it may suggest that the rank, $s$, of the inferred
 subspace needs to be larger to capture the variation in all groups. To
@@
 
-% There are many different prior specifications for $\Psi_k$ and
-% $\sigma^2_k$ that we could consider for estimating the posterior
-% uncertainty.  In this paper, we build on recent interest in the spiked
+%  In this paper, we build on recent interest in the spiked
 % covariance model \citep{Donoho2013, Paul2007} and develop tractable
@@
 
-Building on recent interest in the spiked covariance model
+There are many different ways in which we could choose to parameterize
+$\Psi_k$.  Building on recent interest in the spiked covariance model
 \citep{Donoho2013, Paul2007} we propose a tractable MCMC algorithm by
@@
 $\Sigma_k$ in terms of the original spiked principle components model
-(\ref{eqn:spiked}).  Equation \ref{eqn:sspsi}, written as a function
+(\ref{eqn:spiked}).  Equation \ref{eqn:sspsi}, written as a function of
 $V$, becomes
-
+%
 \begin{align}
 \nonumber \Psi_k &= O_k\Lambda_KO_k^T\\
-\Sigma_k &= V\Psi_kV^T + \sigma^2_kI
+\Sigma_k &= V\Psi_kV^T + \sigma^2_kI.
 \label{eqn:ss}
@@
 %
-\noindent Here, we make one further generalization to allow $\Psi_k$
-to be a rank $r \leq s$ dimensional covariance matrix on the
-$s$-dimensional subspace.  Thus, $\Lambda_k$ is an $r \times r$
-diagonal matrix of eigenvalues, and $O_k \in \mathcal{V}_{s,r}$ is the
-matrix of eigenvectors of $\Psi_k$.  For any individual group, this
-corresponds to the original spiked PCA model (Equation
-\ref{eqn:spiked}) with $U_k = VO_k \in \mathcal{V}_{p, r}$.  This
-generalization is helpful because it enables us to independently
-specify a subspace common to all groups and the possibly lower rank
-features on this space that are specific to individual groups.
-Although our model is most useful when the covariance matrices are
-related across groups, we can use this formulation to specify models
-for multiple unrelated spiked covariance models by letting the shared
-subspace be full rank, e.g. $s=p$.  We explore this in detail in
-Section \ref{sec:simulation}.
+\noindent Here, we allow $\Psi_k$ to be of rank $r \leq s$ dimensional
+covariance matrix on the $s$-dimensional subspace.  Thus, $\Lambda_k$
+is an $r \times r$ diagonal matrix of eigenvalues, and
+$O_k \in \mathcal{V}_{s,r}$ is the matrix of eigenvectors of $\Psi_k$.
+For any individual group, this corresponds to the original spiked PCA
+model (Equation \ref{eqn:spiked}) with
+$U_k = VO_k \in \mathcal{V}_{p, r}$.  This generalization is helpful
+because it enables us to independently specify a subspace common to
+all groups and the possibly lower rank features on this space that are
+specific to individual groups.  Although our model is most useful when
+the covariance matrices are related across groups, we can also use
+this formulation to specify models for multiple unrelated spiked
+covariance models.  We explore this in detail in Section
+\ref{sec:simulation}.  In Section \ref{sec:app} we introduce a shared
+subspace model with additional structure on the eigenvectors and
+eigenvalues of $\Psi_k$ to facilitate interpretation of covariance
+heterogeneity on a two-dimensional subspace.
 
 The likelihood for $\Sigma_k$ given the sufficient statistic
-$\mathbf{S}_k = Y_kY_k^T$ is given in Equation \ref{eqn:lik}.  For the
+$S_k = Y_kY_k^T$ is given in Equation \ref{eqn:lik}.  For the
 spiked PCA formulation, we must rewrite this likelihood in terms of $V$, $O_k$,
@@
 identity
-
+ %
 \begin{align}
@@
 \noindent where the diagonal matrix $\Omega = \Lambda(I+\Lambda)^{-1}$, e.g. $\omega_i = \frac{\lambda_i}{\lambda_{i}+1}$.  Further, 
-
+%
 \begin{align}
@@
 replacing $U_k$ with $VO_k$:
-
+%
 \begin{equation}
- L(\sigma_k^2,\bl V , O_k \Omega_k : \bl Y_k) \propto
-    (\sigma_k^2)^{-n_kp/2}\etr(-\frac{1}{2\sigma_k^2}\mathbf{S}_k)\left(\prod_{i=1}^r(1-\omega_{ki})
+ L(\sigma_k^2, V , O_k \Omega_k : Y_k) \propto
+    (\sigma_k^2)^{-n_kp/2}\etr(-\frac{1}{2\sigma_k^2}S_k)\left(\prod_{i=1}^r(1-\omega_{ki})
    \right) ^{n_k/2}
-   \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k).
+   \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)S_k).
 \label{eqn:sslik}
@@
 invariance considerations suggest the use of priors that lead to
-invariant estimators.  Below we describe our choices for the prior
+equivariant estimators.  Below we describe our choices for the prior
 distributions of each parameter and the resultant conditional posterior
@@
 From Equation \ref{eqn:sslik} it is clear that the the inverse-gamma
-class of prior distributions is conjugate for$\sigma_k^2$.  We chose a
+class of prior distributions is conjugate for $\sigma_k^2$.  We chose a
 default prior distribution for $\sigma^2$ that is equivariant with
@@
 conditional distribution of $\sigma_k^2$ is
-inverse-gamma$( n_k p/2 , \tr(\bl S_k(\bl I -\bl U_k\Omega_k\bl
-U_k)/2)$, where $U_k = VO_k$
+inverse-gamma$( n_k p/2 , \tr[S_k( I - U_k\Omega_k
+U_k)/2])$, where $U_k = VO_k$.
 
@@
 from Equation \ref{eqn:sslik}, it is easy to show that the class of
-von Mises-Fisher-Bingham distributions are conjugate for $O_k$
-\citep{Hoff2009, Hoff2012}.  Again, invariance considerations
-lead us to use a rotationally invariant uniform probability measure on
+Bingham distributions are conjugate for $O_k$ \citep{Hoff2009,
+  Hoff2012}.  Again, invariance considerations lead us to use a
+rotationally invariant uniform probability measure on
 $\mathcal V_{s,p}$.  Under this uniform prior, the full conditional
-distribution of $\bl O_k$ has a density proportional to the
+distribution of $O_k$ has a density proportional to the
 likelihood, so
-
+%
 \begin{align}
 \label{lik_vo}
- p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto \etr(\Omega_kO^T_kV^T[S_k/(2\sigma^2_k)]VO_k)
+ p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto \etr(\Omega_kO^T_kV^T[S_k/(2\sigma^2_k)]VO_k).
 \end{align}
 %
-\noindent This is a Bingham$(\Omega, \hat{V}^T \bl S_k \hat{V}/(2\sigma^2)$
+\noindent This is a Bingham$(\Omega, \hat{V}^T S_k \hat{V}/(2\sigma^2))$
 distribution on $\mathcal V_{s, r}$ \citep{Khatri1977}. A
@@
 Together, the prior for $\sigma_k^2$ and $O_k$ leads to conditional
-(on $V$) Bayes estimators $\hat \Sigma(V^T \bl S V)$ that are
+(on $V$) Bayes estimators $\hat \Sigma(V^T S V)$ that are
 equivariant with respect to scale changes and rotations on the
 subspace spanned by $V$, so
-that $\hat \Sigma(a \bl W V^T \bl S V \bl W^T) = a \bl W \hat\Sigma(V^T
-\bl S V) \bl W$
-for all $a>0$ and $\bl W\in \mathcal O_{s}$ (assuming an invariant
+that $\hat \Sigma(a W V^T S V W^T) = a W \hat\Sigma(V^T
+S V)  W$
+for all $a>0$ and $ W\in \mathcal O_{s}$ (assuming an invariant
 loss function). Interestingly, if $\Omega$ were known (which it is
@@
 \begin{align}
-p(\omega_{ki}|V, O_k, \bl S_k) &\propto_{\omega_{ki}}
+p(\omega_{ki}|V, O_k, S_k) &\propto_{\omega_{ki}}
   \left(\prod_{i=1}^r(1-\omega_{ki})^{n_k/2}  \right)
-  \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) \\
+  \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)S_k) \\
 &  \propto  (1-\omega_{ki})^{n/2} e^{c \omega_s  n/2},    
@@
 %
-\noindent where $c = \bl u_{ki}^T S_k \bl u_{ki}/(n \sigma^2)$ and $\bl u_{ki}$ is
+\noindent where $c = u_{ki}^T S_k u_{ki}/(n \sigma^2)$ and $ u_{ki}$ is
 column $i$ of $U_k = VO_k$.  While not proportional to a density
@@
 this makes sense, note that the likelihood is maximized when the
-columns of $\bl U_k$ are equal to the eigenvectors of $\bl S_k$
+columns of $ U_k$ are equal to the eigenvectors of $S_k$
 corresponding to its top $r$ eigenvalues
-\citep{Tipping1999}. At this value of $\bl U_k$, $c$ will then
-equal one of the top $r$ eigenvalues of $\bl S_k/(n\sigma_k^2)$.  In the
+\citep{Tipping1999}. At this value of $U_k$, $c$ will then
+equal one of the top $r$ eigenvalues of $ S_k/(n\sigma_k^2)$.  In the
 case that $n\gg p$, we expect
-$\bl S_k/(n_k\sigma_k^2)\approx \Sigma_k/\sigma_k^2$, the true (scaled)
+$ S_k/(n_k\sigma_k^2)\approx \Sigma_k/\sigma_k^2$, the true (scaled)
 population covariance, and so we expect $c$ to be near one of the top
@@
 
-We start with an example demonstrating how the shared subspace
+We start with an example demonstrating how a shared subspace
 model can be used to identify statistically significant differences between
@@
 model we summarize $\Psi_k$ in terms of its eigendecomposition by
-computing posterior distributions for the eigenvalue ratio
-$\frac{\lambda_1}{\lambda_2}$, with $\lambda_1 > \lambda_2$, and the
+computing posterior distributions for the log eigenvalue ratio
+$\text{log}(\frac{\lambda_1}{\lambda_2})$, with $\lambda_1 > \lambda_2$, and the
 angle of the first eigenvector on this subspace,
@@
 these quantities from a single simulation.  Dots correspond to the
-true ratios and orientations of $\hat{V}^T\Sigma_k\hat{V}$. To compute
+true log ratios and orientations of $\hat{V}^T\Sigma_k\hat{V}$. To compute
 these regions, we iteratively remove posterior samples corresponding to the
@@
     % \end{subfigure}
-    \caption{a) 95\% posterior regions for the ratio of the
-      eigenvalues, $\frac{\lambda_1}{\lambda_2}$, of $\Psi_k$ and the
+    \caption{a) 95\% posterior regions for the log of the ratio of
+      eigenvalues, $\text{log}(\frac{\lambda_1}{\lambda_2})$, of $\Psi_k$ and the
       orientation of the principle axis on space spanned by $\hat{V}$
@@
       parameter values on $\hat{V}^T\Sigma_k\hat{V}$ .  Since $V$
-      is only identifiable up to rotation we find the Procrustes
+      is only identifiable up to rotation, for this plot we find the Procrustes
       rotation that maximizes the similarity of $\hat{V}$ to the true
@@
 the posterior mean of the precision matrix,
-$\hat \Sigma_{k} = \Exp{ \Sigma_k^{-1} | \bl S_k}^{-1}$ which we
+$\hat \Sigma_{k} = \Exp{ \Sigma_k^{-1} | S_k}^{-1}$ which we
 estimate using MCMC samples.
@@
 
-We fit these datasets using three different versions of the shared
+We fit these datasets using three different variants of the shared
 subspace model.  For each of these fits we estimate the number of
@@
 
-In this Section, we apply the shared subspace spiked covariance model to data in two
+In this Section we apply a shared subspace spiked covariance model to data in two
 different biological examples.  In the first, we compare gene
@@
 in covariance matrices between groups.  In the second example, we
-compare covariance matrices in a metabolomic analysis of fly data.  In
+compare covariance matrices in a metabolomic analysis of \textit{Drosophila} data.  In
 this example, after controlling for mean effects due to fly age, gender and
@@
 
-In these analyses we employ a variation of the shared
-subspace model.  Specifically, we define the $s$ by $s$ matrix $\Psi_k$ as
-
+In these analyses we employ a shared subspace model in
+which we define the $s$ by $s$ matrix $\Psi_k$ as
+%
 \begin{equation}
@@
 \end{equation}
-
+%
 For the following analyses $O_k\Lambda_kO_k^T$ is a rank $r=2$ matrix
@@
 groups may manifest themselves in the smaller principle
-components. This formulation allows complete pooling of most
+components. This formulation allows a complete pooling of several
 eigenvectors, yet a space to identify and compare relevant
@@
 In Figure \ref{fig:leukemia} we plot the results of our analysis.
-Panel \ref{fig:leukemiaRatio} shows that over $40\%$ of the variance
-can be explained by the estimated two-dimensional subspace for all
-groups, with 4 of the 6 groups over 60\%.  Panel
+Panel \ref{fig:leukemiaRatio} shows that over $60\%$ of the variance
+can be explained by the  shared subspace for all
+groups, with 4 of the 6 groups over 80\%.  Panel
 \ref{fig:leukemiaPosterior} reflects some significant differences in
@@
 groups.  The $x$-axis corresponds to the orientation of the principle
-axis of $\Psi_k$ on $V$ and the $y$-axis corresponds to the ratio of
-the first eigenvalue to the second.  Several groups have significantly
+axis of $\Psi_k$ on $V$ and the $y$-axis corresponds to the log ratio of
+the first two eigenvalues.  Several groups have significantly
 different orientations for the first principle component of the
@@
     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/leukemiaRatio-04-27}
+        \includegraphics[width=\textwidth]{Figs/leukemiaRatio-07-01}
         \caption{$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$}
@@
     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/leukemiaPosterior-06-23}
+        \includegraphics[width=\textwidth]{Figs/leukemiaPosterior-07-01}
         \caption{95\% Posterior Regions }
@@
       \end{subfigure}
-    \caption{a) Goodness of shared subspace fit for each of the seven
-      Leukemia groups.  Over 40\% of the variance is explained on this
-      subspace in each of the seven groups, with as much as 85\%
-      explained in some groups.  b).  95\% posterior regions for the
-      eigenvalue ratio and primary eigenvector orientation.  Regions
-      for some pairs of groups are disjoint, suggesting significant
-      differences in the projected covariance matrices.  For other
-      groups (e.g. Hyperdip50 and TEL-AML1) overlap in the posterior
-      regions indicate that differences are not detectable on this subspace. }
+      \caption{a) Goodness of shared subspace fit for each of the
+        seven Leukemia groups.  Over 60\% of the variance is explained
+        on the shared subspace in each of the seven groups, with
+        more than 80\% explained in most groups.  b) 95\% posterior
+        regions for the eigenvalue ratio and primary eigenvector
+        orientation.  Regions for some pairs of groups are disjoint,
+        suggesting significant differences in the projected covariance
+        matrices.  For other groups (e.g. Hyperdip50 and TEL-AML1)
+        overlap in the posterior regions indicate that differences are
+        not detectable on this subspace. }
 \label{fig:leukemia}
@@
 (middle) and greater than 51 days (old), and further split by
-gender. The sample sizes range from 43 to 59 flies per group.  We
+sex. The sample sizes range from 43 to 59 flies per group.  We
 analyze metabolomic data corresponding to metabolites with 3714
 mass-charge ratios.  After removing mean effects due to age and
-gender, we fit the shared subspace estimator and compare covariances
+sex, we fit the shared subspace estimator and compare covariances
 between groups.  We identify a subspace that explains almost all of
@@
 \ref{fig:dmelanPosterior}). This could be indicative of a large amount
-of variation that is common across all groups for the first $\hat{s}$
+of variation that is common across all groups for the first $s$
 principle components. Differences in covariation across groups likely
@@
     \begin{subfigure}[b]{0.45\textwidth}r
-        \includegraphics[width=\textwidth]{Figs/dmelanPosterior-06-23}
+        \includegraphics[width=\textwidth]{Figs/dmelanPosterior-07-01}
         \caption{95\% Posterior Regions}
@@
       variability across groups. b) There is significant overlap in
-      the posterior distribution of the eigenvalue ratio and primary
+      the posterior distribution of the log eigenvalue ratio and primary
       eigenvector orientation across groups. Thus, there is little
@@
 
-In this paper, we proposed a hierarchical model for estimating and
+In this paper, we proposed a class of models for estimating and
 comparing differences in covariance matrices across multiple groups on
@@
 in this paper and leave it for future work to develop an efficient
-fully Bayesian approach for jointly estimating $VV^T$ and $\Psi_k$.
+fully Bayesian approach for estimating the joint posterior of $VV^T$ and $\Psi_k$.
 Recently developed Markov chain Monte Carlo algorithms, like Riemmanian manifold
@@
 Fundamentally, our approach is quite general and can be integrated
-with existing approaches for hierarchical multi-group covariance
+with existing approaches for multi-group covariance
 estimation.  In particular, we can incorporate additional shrinkage on

### 2016-07-06 paper/mscov.tex
Minor edits from proofread

@@
     graduate of the Department of Statistics at Harvard University
-    (\href{mailto:amfranks@uw.edu}{amfranks@uw.edu}).  Peter Hoff is a
-    Professor of Statistics at The University of Washington
+    (\href{mailto:amfranks@uw.edu}{amfranks@uw.edu}).  Peter D. Hoff
+    is a Professor in the Departments of Statistics, Biostatistics and
+    the Center for Statistics and the Social Sciences at the
+    University of Washington
     (\href{mailto:pdhoff@uw.edu}{pdfhoff@uw.edu}).
@@
  (University of Alabama at Birmingham) for sharing data and ideas
- which contributed to the framing of the ideas this paper.}}
+ which contributed to framing of this paper.}}
 \author{Alexander Franks and Peter Hoff} \date{\today}
@@
   We develop a model-based method for evaluating heterogeneity among
-  several $p\times p$ covariance matrices in the big p, small n setting.
+  several $p\times p$ covariance matrices in the big $p$, small $n$ setting.
   This is done by assuming a spiked covariance model for each group
@@
 high-dimensional multivariate data is partial least squares regression
-(PLS). This approach, which is a special case of a bilinear factor
+(PLS) \citep{Wold2001}. This approach, which is a special case of a bilinear factor
 model, involves projecting the data onto a lower dimensional space which
@@
 estimating the posterior distribution of the covariance matrices of
-the data projected onto this subspace.  In Section
-\ref{sec:simulation} we investigate the behavior of this class of
-models in simulation and demonstrate how the shared subspace
+the data projected onto this subspace.
+
+In Section \ref{sec:simulation} we investigate the behavior of this
+class of models in simulation and demonstrate how the shared subspace
 assumption is widely applicable, even when there is little similarity
@@
 spanned by the group-level eigenvectors is simply the full
-$p$-dimensional space.  We then demonstrate the utility of the model
-in an analysis of gene expression data from juvenile leukemia patients
-in Section \ref{sec:app}.  Despite the large feature size ($p > 3000$)
-and small sample size ($n < 100$ per group), we identify interpretable
-similarities and differences in gene covariances on a low dimensional
-subspace.  Finally, we conclude with a brief exploration of
-heterogeneity in metabolomic data from a study on aging in the fly
-species \textit{Drosophila melanogaster}.
+$p$-dimensional space.  In Section \ref{sec:app} we demonstrate the
+utility of a shared subspace model in an analysis of gene expression
+data from juvenile leukemia patients .  Despite the large feature size
+($p > 3000$) and small sample size ($n < 100$ per group), we identify
+interpretable similarities and differences in gene covariances on a
+low dimensional subspace.  Finally, we conclude with a brief
+exploration of heterogeneity in metabolomic data from a study on aging
+in the fly species \textit{Drosophila melanogaster}.
 
@@
 \noindent where for $r \ll p$, $\Lambda$ is an $r\times r$ diagonal
-matrix and $U \in \mathcal V_{r,p}$.  $\mathcal V_{r,p}$ is the
+matrix and $U \in \mathcal V_{p,r}$.  $\mathcal V_{p,r}$ is the
 Stiefel manifold, which consists of the set of $p \times r$
@@
 (degenerate) Wishart distribution as in Equation \ref{eqn:lik}.  To
-improve estimation, we seek estimators that share information across
-groups, that is, where the estimator $\hat{\Sigma}_k$ may depend on
-data from all groups.  To this end, we extend the spiked covariance
-model to grouped data, by assuming that the anisotropic variability
-from each group occurs on a common low dimensional subspace.
-Specifically, we posit that the covariance matrix for each group can
-be written as
+improve estimation, we seek estimators of each covariance matrix,
+$\hat{\Sigma}_k$, that may depend on data from all groups.  To this
+end, we extend the spiked covariance model to grouped data, by
+assuming that the anisotropic variability from each group occurs on a
+common low dimensional subspace.  Specifically, we posit that the
+covariance matrix for each group can be written as
 \begin{equation}
@@
 each group individually, then the shared subspace assumption also
-holds because at worst we can take the shared subspace ot be the full
+holds because at worst we can take the shared subspace to be the full
 $p$-dimensional space. In practice, we can usually identify a shared
@@
 In this section we describe a maximum marginal likelihood procedure for
-estimating the shared subspace, $VV^T$ based on the
-expectation-maximization algorithm.  The full likelihood
+estimating the shared subspace, $VV^T$, based on the
+expectation-maximization (EM) algorithm.  The full likelihood
 for the shared subspace model can be written as
@@
 
-We propose an expectation-maximization (EM) algorithm to estimate the
-relevant quantities.  In particular we maximize the marginal
-likelihood of $V$, treating $M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as
-the ``missing'' parameters to be imputed.  We assume independent
-Jeffreys' prior distributions for both $\sigma_k^2$ and
-$M_k$.  Jeffreys' prior for these quantities corresponds to
+We maximize the marginal likelihood of $V$ with an EM algorithm, where
+$M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as the considered the
+``missing'' parameters to be imputed.  We assume independent Jeffreys'
+prior distributions for both $\sigma_k^2$ and $M_k$.  Jeffreys' prior
+for these quantities corresponds to
 $p(\sigma_k^2) \propto 1/\sigma_k^2$ and
-$p(M_k) \propto |M_k|^{-(s+1)/2}$.  From the likelihood it
-can easily be shown that the conditional posterior for $M_k$ is
+$p(M_k) \propto |M_k|^{-(s+1)/2}$.  From the likelihood it can easily
+be shown that the conditional posterior for $M_k$ is
 %
@@
 \noindent which is an inverse-gamma($n_k(p-s)/2$,
-$\frac{1}{2}\text{tr}((I-VV^T)S_k)$) distribution.  Based on these
-results the following EM algorithm can be used:
+$\frac{1}{2}\text{tr}[(I-VV^T)S_k]$) distribution.  Based on these
+results, we complete the following steps for each iteration, $t$, of
+the EM algorithm until convergence:
 
@@
 \item $E[M_k^{-1} | V_{(t-1)}] = n_k(V_{(t-1)}^T S_kV_{(t-1)})^{-1}$
-\item $E[\frac{1}{\sigma_k^2}|V_{(t-1)}] = \frac{n_k(p-s)}{tr((I-V_{(t-1)}V_{(t-1)}^T)S_k)}$
+\item $E[\frac{1}{\sigma_k^2}|V_{(t-1)}] = \frac{n_k(p-s)}{\text{tr}[(I-V_{(t-1)}V_{(t-1)}^T)S_k]}$
 \end{enumerate}
-\item Compute $V_{t}$ = $\underset{V}{\text{arg } \text{max}}  \sum_k tr\left(-VE[M^{-1}|V_{(t-1)}]^T +
+\item Compute $V_{t}$ = $\underset{V}{\text{arg } \text{max}}  \sum_k \text{tr}\left(-VE[M^{-1}|V_{(t-1)}]^T +
        E[\frac{1}{2\sigma_k^2}|V_{(t-1)}]VV^T\right)S_k/2)$ 
@@
 \citep{Wen2013}.  Importantly, the complexity of this algorithm is
-dominated by dimension of the shared subspace, $s$, not the number of
-features $p$.  Thus, our approach is computationally efficient for
+dominated by the dimension of the shared subspace not the number of
+features.  Thus, our approach is computationally efficient for
 relatively small values of $s$, even when $p$ is large.
@@
 
-If $V$ is a basis for a shared subspace, then for each group $k$, most
-of the non-isotropic variation in $Y_k$ should be preserved when projecting the data
-to this space.  To characterize the extent to which this is true for
-different groups, we propose a simple estimator for the proportion of
-``signal'' variance that lies on any given subspace.  Specifically,
-for an estimated subspace $\hat{V}$, we use the following statistic
-for the ratio of the sum of the first $s$ eigenvalues of
-$\hat{V}^T \Sigma_k \hat{V}$ to the sum of the first $s$
-eigenvalues of $\Sigma_k$:
+If $V$ is a basis for a shared subspace, then for each group, most
+of the non-isotropic variation in $Y_k$ should be preserved when
+projecting the data to this space.  To characterize the extent to
+which this is true for different groups, we propose a simple estimator
+for the proportion of ``signal'' variance that lies on a given
+subspace.  Specifically, we use the following statistic for the ratio
+of the sum of the first $s$ eigenvalues of
+$V^T \Sigma_k V$ to the sum of the first $s$ eigenvalues
+of $\Sigma_k$:
 %
 \begin{equation}
-\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2) = \frac{||Y_k\hat{V}||_F/n_k}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
-  ||Y_k\widetilde{V}||_F/n_k - \hat{\sigma}_k^2ps/n_k}
+\gamma(Y_k: V, \sigma_k^2) = \frac{||Y_kV||_F/n_k}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
+  ||Y_k\widetilde{V}||_F/n_k - \sigma_k^2ps/n_k}
 \label{eqn:ratio}
@@
 this statistic is the fact that if
-$\Sigma_k = \hat{V}\Psi_k\hat{V}^T + \sigma_k^2I$ then
-$\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2) \approx 1$.  We demonstrate
+$\Sigma_k = V\Psi_kV^T + \sigma_k^2I$ then
+$\gamma(Y_k: V, \sigma_k^2) \approx 1$.  We establish
 this in an asymptotic regime where $s$ is fixed, $p$,
@@
 First, since $s$ is fixed and $n_k$ is growing, the numerator,
-$||Y_k\hat{V}||_F/n_k$, is a consistent estimator for the sum of the
-eigenvalues of $\hat{V}^T\Sigma_k\hat{V}$.  The denominator,
+$||Y_kV||_F/n_k$, is a consistent estimator for the sum of the
+eigenvalues of $V^T\Sigma_kV$.  The denominator,
 $\underset{\widetilde{V} \in \mathcal{V}_{p,
@@
 & \approx& \lambda^{(k)} + \sigma^2_k\frac{p}{n_k}
+\label{eqn:lamAsymp}
 \end{eqnarray}
@@
 \noindent where the approximation in the second line is due to the
-fact that $\lambda_k/(\lambda_k - 1) \approx 1$ for $\lambda_k$
-large. Thus,
+fact that $\lambda_k/(\lambda_k - 1) \approx 1$ for $\lambda_k$ large.
+Thus,
 $\sum_i^s \hat{\lambda}^{(k)}_i \approx \sum_i^s \lambda^{(k)}_i +
 \sigma_k^2ps/n_k$
-and $\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
-  ||Y_k\widetilde{V}||_F/n_k - \hat{\sigma}_k^2ps/n_k$ is a reasonable approximation to the sum of the
-first $s$ eigenvalues of $\Sigma_k$.
-
-As a consequence, the proposed statistic will be
-close to one for all groups when $\hat{V}\hat{V}^T = VV^T$ and smaller
-if not.  The metric can provides a useful indicator for which groups
-can be reasonably compared on a given subspace and which groups
-cannot.  When $\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2)$ is small for
-some groups, it may suggest that the rank, $s$, of the inferred
-subspace needs to be larger to capture the variation in all groups. To
-demonstrate this, in Section \ref{sec:simulation}, we compute these
-statistics for inferred subspaces of different dimensions on a single
-dataset. In Section \ref{sec:app}, we plot the estimates for subspaces
-inferred with real biological data.
+and
+$\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
+||Y_k\widetilde{V}||_F/n_k - \hat{\sigma}_k^2ps/n_k$
+is a reasonable approximation to the sum of the first $s$ eigenvalues
+of $\Sigma_k$.  Alternatively, when $\lambda_k$ is small, a better
+approximation may be obtained by solving the quadratic equation
+implied by the first line of Equation \ref{eqn:lamAsymp}.
+
+As a consequence, the proposed statistic will be close to one for all
+groups when $VV^T$ is a shared subspace for the data and smaller if
+not.  The metric provides a useful indicator for which groups can be
+reasonably compared on a given subspace and which groups cannot.  In
+practice, we estimate a shared subspace, $\hat{V}$, and the isotropic
+variances, $\hat{\sigma}_k^2$, using EM and compute the plug-in
+estimate $\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2)$.  When this
+statistic is small for some groups, it may suggest that the rank, $s$,
+of the inferred subspace needs to be larger to capture the variation
+in all groups. We investigate this in Section \ref{sec:simulation}, by
+computing the goodness of fit statistic for inferred subspaces of
+different dimensions on a single dataset. In Section \ref{sec:app}, we
+compute the estimates for subspaces inferred with real biological
+data.
 
@@
 modeling the eigenstructure, we can now view each covariance
-$\Sigma_k$ in terms of the original spiked principle components model
-(\ref{eqn:spiked}).  Equation \ref{eqn:sspsi}, written as a function of
+$\Sigma_k$ in terms of the original spiked principle components model.  Equation \ref{eqn:sspsi}, written as a function of
 $V$, becomes
@@
 model (Equation \ref{eqn:spiked}) with
-$U_k = VO_k \in \mathcal{V}_{p, r}$.  This generalization is helpful
-because it enables us to independently specify a subspace common to
-all groups and the possibly lower rank features on this space that are
-specific to individual groups.  Although our model is most useful when
-the covariance matrices are related across groups, we can also use
-this formulation to specify models for multiple unrelated spiked
-covariance models.  We explore this in detail in Section
-\ref{sec:simulation}.  In Section \ref{sec:app} we introduce a shared
-subspace model with additional structure on the eigenvectors and
-eigenvalues of $\Psi_k$ to facilitate interpretation of covariance
+$U_k = VO_k \in \mathcal{V}_{p, r}$.  Differentiating the ranks $r$
+and $s$ is helpful because it enables us to independently specify a
+subspace common to all groups and the possibly lower rank features on
+this space that are specific to individual groups.  Although our model
+is most useful when the covariance matrices are related across groups,
+we can also use this formulation to specify models for multiple
+unrelated spiked covariance models.  We explore this in detail in
+Section \ref{sec:simulation}.  In Section \ref{sec:app} we introduce a
+shared subspace model with additional structure on the eigenvectors
+and eigenvalues of $\Psi_k$ to facilitate interpretation of covariance
 heterogeneity on a two-dimensional subspace.
@@
 class of prior distributions is conjugate for $\sigma_k^2$.  We chose a
-default prior distribution for $\sigma^2$ that is equivariant with
+default prior distribution for $\sigma^2_k$ that is equivariant with
 respect to scale changes.  Specifically, we use Jeffreys' prior, an
-improper prior with density $p(\sigma^2) \propto 1/\sigma^2 $.  Under
+improper prior with density $p(\sigma^2_k) \propto 1/\sigma^2_k $.  Under
 this prior, straightforward calculations show that the full
@@
 likelihood, so
-%
 \begin{align}
@@
 %
-\noindent This is a Bingham$(\Omega, \hat{V}^T S_k \hat{V}/(2\sigma^2))$
+\noindent This is a Bingham$(\Omega, V^T S_k V/(2\sigma^2))$
 distribution on $\mathcal V_{s, r}$ \citep{Khatri1977}. A
@@
 Together, the prior for $\sigma_k^2$ and $O_k$ leads to conditional
-(on $V$) Bayes estimators $\hat \Sigma(V^T S V)$ that are
+(on $V$) Bayes estimators $\hat \Sigma(V^T S_k V)$ that are
 equivariant with respect to scale changes and rotations on the
 subspace spanned by $V$, so
-that $\hat \Sigma(a W V^T S V W^T) = a W \hat\Sigma(V^T
-S V)  W$
+that $\hat \Sigma(a W V^T S_k V W^T) = a W \hat\Sigma(V^T
+S_k V)  W$
 for all $a>0$ and $ W\in \mathcal O_{s}$ (assuming an invariant
-loss function). Interestingly, if $\Omega$ were known (which it is
+loss function). Interestingly, if $\Omega_k$ were known (which it is
 not), then for a given invariant loss function the Bayes estimator
@@
 
-\paragraph{Conditional distribution for $\Omega$:} Here we specify the
+\paragraph{Conditional distribution for $\Omega_k$:} Here we specify the
 conditional distribution of the diagonal matrix
@@
   \etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)S_k) \\
-&  \propto  (1-\omega_{ki})^{n/2} e^{c \omega_s  n/2},    
+&  \propto  (1-\omega_{ki})^{n/2} e^{c_{ki} \omega_{ki}  n/2},    
 \label{eqn:wpost}
@@
 %
-\noindent where $c = u_{ki}^T S_k u_{ki}/(n \sigma^2)$ and $ u_{ki}$ is
+\noindent where $c_{ki} = u_{ki}^T S_k u_{ki}/(n_k \sigma^2_k)$ and $ u_{ki}$ is
 column $i$ of $U_k = VO_k$.  While not proportional to a density
@@
 corresponding univariate distribution numerically.  The behavior of
-this distribution is straightforward to understand: if $c\leq 1$, then
-the the function has a maximum at $\omega_i =0$, and decays
-monotonically to zero as $\omega \rightarrow 1$.  If $c>1$ then the
-function is uniquely maximized at $(c-1)/c \in (0,1)$.  To see why
+this distribution is straightforward to understand: if $c_{ki}\leq 1$, then
+the the function has a maximum at $\omega_{ki} =0$, and decays
+monotonically to zero as $\omega_{ki} \rightarrow 1$.  If $c_{ki}>1$ then the
+function is uniquely maximized at $(c_{ki}-1)/c_{ki} \in (0,1)$.  To see why
 this makes sense, note that the likelihood is maximized when the
@@
 corresponding to its top $r$ eigenvalues
-\citep{Tipping1999}. At this value of $U_k$, $c$ will then
-equal one of the top $r$ eigenvalues of $ S_k/(n\sigma_k^2)$.  In the
-case that $n\gg p$, we expect
+\citep{Tipping1999}. At this value of $U_k$, $c_{ki}$ will then
+equal one of the top $r$ eigenvalues of $ S_k/(n_k\sigma_k^2)$.  In the
+case that $n_k\gg p$, we expect
 $ S_k/(n_k\sigma_k^2)\approx \Sigma_k/\sigma_k^2$, the true (scaled)
-population covariance, and so we expect $c$ to be near one of the top
-$r$ eigenvalues of $\Sigma_0/\sigma^2$, say $\lambda_0+1$.  If indeed
-$\Sigma_k$ has $r$ spikes, then $\lambda_k>0$,
-$c \approx \lambda_0 +1 > 1$, and so the conditional mode of $w_i$ is
-approximately $(c-1)/c = \lambda_i/(\lambda_i+1)$, the correct value.
+population covariance, and so we expect $c_{ki}$ to be near one of the top
+$r$ eigenvalues of $\Sigma_k/\sigma^2_k$, say $\lambda_{ki}+1$.  If indeed
+$\Sigma_k$ has $r$ spikes, then $\lambda_{ki}>0$,
+$c_{ki} \approx \lambda_{ki} +1 > 1$, and so the conditional mode of $w_{ki}$ is
+approximately $(c_{ki}-1)/c_{ki} = \lambda_{ki}/(\lambda_{ki}+1)$, the correct value.
 On the other hand, if we have assumed the existence of a spike when
-there is none, then $\lambda_0=0$, $c\approx 1$ and the Bayes estimate
-of $w_i$ will be shrunk towards zero, as it should be.
+there is none, then $\lambda_{ki}=0$, $c_{ki}\approx 1$ and the Bayes estimate
+of $w_{ki}$ will be shrunk towards zero, as it should be.
 
@@
 $n_k=50$.  We fix the first eigenvalue of $\Psi_k$ from each group to
-$\lambda_1=100$ and vary $\lambda_2$ so that $\lambda_1/\lambda_2$ is
-either $10$, $3$ or $1$.  For this two dimensional shared subspace
+$\lambda_1=100$ and vary $\lambda_2$.  For this two dimensional shared subspace
 model we summarize $\Psi_k$ in terms of its eigendecomposition by
-computing posterior distributions for the log eigenvalue ratio
+computing posterior distributions for the log eigenvalue ratio,
 $\text{log}(\frac{\lambda_1}{\lambda_2})$, with $\lambda_1 > \lambda_2$, and the
@@
 these quantities from a single simulation.  Dots correspond to the
-true log ratios and orientations of $\hat{V}^T\Sigma_k\hat{V}$. To compute
-these regions, we iteratively remove posterior samples corresponding to the
+true log ratios and orientations of $\hat{V}^T\Sigma_k\hat{V}$, where
+$\hat{V}$ is the maximum marginal likelihood for $V$. To compute
+the posterior regions, we iteratively remove posterior samples corresponding to the
 vertices of the convex hull until only 95\% of the original samples
@@
 were $10$ (black and red groups), $3$ (green and blue groups) and $1$
-(cyan group).  Larger values of the eigenvalue ratio correspond to more
+(cyan group).  Larger eigenvalue ratios correspond to more
 correlated contours and a value of $1$ implies isotropic covariance.
@@
       eigenvalues, $\text{log}(\frac{\lambda_1}{\lambda_2})$, of $\Psi_k$ and the
-      orientation of the principle axis on space spanned by $\hat{V}$
-      cover the truth.  Dots correspond to true data generating
+      orientation of the principle axis on the space spanned by $\hat{V}$
+      cover the truth in this simulation.  Dots correspond to true data generating
       parameter values on $\hat{V}^T\Sigma_k\hat{V}$ .  Since $V$
-      is only identifiable up to rotation, for this plot we find the Procrustes
+      is only identifiable up to rotation, for this figure we find the Procrustes
       rotation that maximizes the similarity of $\hat{V}$ to the true
@@
 did not evaluate the coverage for the group with
-$\lambda_1/\lambda_2 = 1$ since this value is on the edge of the
-parameter space and would not be covered by the 95\% posterior regions
-as constructed.  The slight under coverage is likely due to the fact
+$\lambda_1/\lambda_2 = 1$ (cyan) since this value is on the edge of the
+parameter space and is not be covered by the 95\% posterior regions
+as constructed.  The slight under coverage for the other groups is likely due to the fact
 that we infer $VV^T$ using maximum marginal likelihood, and thus
@@
     \end{subfigure}
-    \caption{a) Stein's risk as a function of the dimension of the
+    \caption{a) Stein's risk as a function of the
       shared subspace dimension.  Data from ten groups, with $U_k$
       generated uniformly on the Stiefel manifold
-      $\mathcal{V}_{200, 2}$.  As the shared subspace dimension
-      increases, $\hat{s} \rightarrow p$, the risk converges to the
+      $\mathcal{V}_{200, 2}$.  As $\hat{s} \rightarrow p$, the risk converges to the
       risk from independently estimated spiked covariance matrices
@@
 underestimate the true dimension of the shared subspace.  In this
-simulation, we generate data $K=10$ groups of mean-zero normally
+simulation, we generate $K=10$ groups of mean-zero normally
 distributed data with $p=200$, $r=2$, $s=p$ and $\sigma_k^2=1$.  We
-fix the eigenvalues for covariance matrix to
+fix the eigenvalues of $\Psi_k$ to
 ($\lambda_1, \lambda_2) = (250, 25)$.  Although the signal variance
 from each group individually is preserved on a two dimensional
-subspace, these subspaces are not similar across groups-- the
+subspace, these subspaces are not similar across groups since the
 eigenvectors from each group are generated uniformly from the Stiefel
@@
 empirically using ten independent simulations per value of $\hat{s}$.
-The dashed blue line corresponds to the multi-group Stein's risk for
-covariance matrices estimated independently.  Independent covariance
-estimation is equivalent to shared subspace inference with
-$\hat{s} = p$ because this implies $VV^T = I_p$.  Although the risk is
-large for small values of $\hat{s}$, as the shared subspace dimension
-increases to the dimension of the feature space,
-$\hat{s} \rightarrow p$, the risk for the shared subspace estimator
-quickly decreases.  Importantly, it is always true that
-rank($[U_1, ..., U_K]) \leq rK$ so it can equivalently be assumed that
-the data were generated from a shared subspace model with dimension
-$s = rK < p$.  As such, even when there is little similarity between the
-eigenvectors from each group, the shared subspace estimator with
-$\hat{s} = rK$ will perform well, provided that we can identify a
-subspace, $\hat{V}\hat{V}^T$, that is close to
-$\text{span}([U_1, ..., U_K])$. When
-$\hat{V}\hat{V}^T = \text{span}([U_1, ..., U_K])$ exactly,
-(\ref{fig:sdimension}, dashed red line),
-shared subspace estimation outperforms independent covariance
-estimation.
+The dashed blue line corresponds to Stein's risk for covariance
+matrices estimated independently.  Independent covariance estimation
+is equivalent to shared subspace inference with $\hat{s} = p$ because
+this implies $VV^T = I_p$.  Although the risk is large for small
+values of $\hat{s}$, as the shared subspace dimension increases to the
+dimension of the feature space, that is $\hat{s} \rightarrow p$, the
+risk for the shared subspace estimator quickly decreases.
+Importantly, it is always true that rank($[U_1, ..., U_K]) \leq rK$ so
+it can equivalently be assumed that the data were generated from a
+shared subspace model with dimension $s = rK < p$.  As such, even when
+there is little similarity between the eigenvectors from each group,
+the shared subspace estimator with $\hat{s} = rK$ will perform well,
+provided that we can identify a subspace, $\hat{V}\hat{V}^T$, that is
+close to $\text{span}([U_1, ..., U_K])$. When
+$\hat{V}\hat{V}^T = \text{span}([U_1, ..., U_K])$ exactly, shared
+subspace estimation outperforms independent covariance estimation
+(\ref{fig:sdimension}, dashed red line).
 
@@
 $\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$, can be used to identify when
-a larger shared subspace is warranted.  When $\hat{s}$ is too small
-(e.g. Figure \ref{fig:ratio-s5}),
+a larger shared subspace is warranted.  When $\hat{s}$ is too small,
 $\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$ will be significantly smaller
-than one for at least some of the groups, regardless of $\hat{V}$.
-When $\hat{s}$ is large enough, we are able to use maximum marginal
-likelihood to identify a shared subspace which preserves most of the
-variation in the data for all groups (Figure \ref{fig:ratio-s20}).
-Thus, for a given $\hat{V}\hat{V}^T$, the goodness of fit statistic
-can be used to identify the groups that can be fairly compared on this
-subspace and whether we would benefit from fitting a model with a
-larger value of $\hat{s}$.
+than one for at least some of the groups, regardless of $\hat{V}$
+(e.g. Figure \ref{fig:ratio-s5}).  When $\hat{s}$ is large enough, we
+are able to use maximum marginal likelihood to identify a shared
+subspace which preserves most of the variation in the data for all
+groups (Figure \ref{fig:ratio-s20}).  Thus, for any estimated
+subspace, the goodness of fit statistic can be used to identify the
+groups that can be fairly compared on this subspace and whether we
+would benefit from fitting a model with a larger value of $\hat{s}$.
 
@@
 
-It is apparent that correct specification for the rank of the shared
-subspace is important for efficient inference.  So far in this
-section, we have assumed that the group rank, $r$, and shared subspace
-dimension, $s$, are fixed and known.  In practice this is not the
-case.  Prior to fitting a model we should estimate these quantities.
+Clearly, correct specification for the rank of the shared subspace is
+important for efficient inference.  So far in this section, we have
+assumed that the group rank, $r$, and shared subspace dimension, $s$,
+are fixed and known.  In practice this is not the case.  Prior to
+fitting a model we should estimate these quantities.
 \citet{Gavish2014} provide an asymptotically optimal (in mean squared
@@
 median singular value of the data matrix and the ratio
-$\alpha_k =\frac{p}{n_k}$, to estimate the rank, $r$, of a spiked
-covariance matrix.  Although their estimator was derived for
-individual covariance estimation, we found that, Gavish and Donoho's
-estimator can also be applied to the pooled data to estimate the
-shared subspace dimension, $s$.  Specifically, we concatenate the data
-from all groups to create an $(\sum_k n_k) \times p$ dimensional
-matrix and apply their rank estimator to this matrix to choose $s$.
+$\alpha_k =\frac{p}{n_k}$, to estimate $r$.  Although their estimator
+was derived for individual covariance estimation, we found that,
+Gavish and Donoho's estimator can also be applied to the pooled data
+to estimate $s$.  Specifically, we concatenate the data from all
+groups to create an $(\sum_k n_k) \times p$ dimensional matrix and
+apply their rank estimator to this matrix to choose $s$.
 
@@
 these simulations we assume the data have $p=200$ features, $r=2$
-spikes, $\sigma^2_k=1$, and $n_k = 50$.  We fix the eigenvalues of
-$\Psi_k$, $(\lambda_1, \lambda_2) = (250, 25)$.  We consider three
+spikes, $\sigma^2_k=1$, and $n_k = 50$.  We fix the non-zero eigenvalues of
+$\Psi_k$ to $(\lambda_1, \lambda_2) = (250, 25)$.  We consider three
 different shared subspace data models: 1) a low dimensional shared
@@
 We fit these datasets using three different variants of the shared
-subspace model.  For each of these fits we estimate the number of
-spikes, $r$, from the data.  First, we estimate a single spiked
-covariance matrix from the pooled data and let
-$\hat{\Sigma}_k = \hat{\Sigma}$.  Second, we fit the full rank shared
-subspace model in which we independently estimate each spiked
-covariance matrix (since $s=p$ implies $VV^T = I_p$).  Finally, we use
-an ``adaptive'' shared subspace estimator, in which we estimate both
-$s$ and $r$ and apply empirical Bayes inference for $VV^T$.  In Table
+subspace model.  For each of these fits we estimate $r$.  First, we
+estimate a single spiked covariance matrix from the pooled data and
+let $\hat{\Sigma}_k = \hat{\Sigma}$.  Second, we fit the full rank
+shared subspace model.  This corresponds to a procedure in which we
+estimate each spiked covariance matrix independently, since $s=p$
+implies $VV^T = I_p$.  Finally, we use an ``adaptive'' shared subspace
+estimator, in which we estimate both $s$, $r$ and $VV^T$.  In Table
 \ref{table:groupLoss} we report the average Stein's risk and
 corresponding 95\% loss intervals for the estimates derived from each
-inferential model.
+of these inferential models.
 
@@
     varying degrees of between-group covariance similarity.  For each
-    of $K=10$ groups, we simulate from spiked covariance models with
-    $p=200$, $r=2$, $\sigma_k^2=1$ and $n_k=50$.  We fit the data to
-    a shared subspace model in which $s$, $r$ and $VV^T$ are all
-    estimated from the data (``adaptive''), a spiked covariance model
-    in which the covariance matrices from each group are assumed to be
-    identical ($\hat{\Sigma}_k=\hat{\Sigma}$) and a model in which we
-    assume the data do \emph{not} share a lower dimensional subspace
-    across groups ($\hat{s} = p$). The estimators which most closely
-    match the data generating model have the lowest risk (diagonal)
-    but the adaptive estimator performs well relative to the
-    alternative misspecified model.}
+    of $K=10$ groups, we simulate data from three different types of
+    shared subspace models.  For each of these models, $p=200$, $r=2$,
+    $\sigma_k^2=1$ and $n_k=50$.  We also fit the data using three
+    different shared subspace models: a model in which $s$, $r$ and
+    $VV^T$ are all estimated from the data (``adaptive''), a spiked
+    covariance model in which the covariance matrices from each group
+    are assumed to be identical ($\hat{\Sigma}_k=\hat{\Sigma}$) and a
+    model in which we assume the data do \emph{not} share a lower
+    dimensional subspace across groups (i.e. $\hat{s} = p$). The estimators
+    which most closely match the data generating model have the lowest
+    risk (diagonal) but the adaptive estimator performs well relative
+    to the alternative misspecified model.}
 \label{table:groupLoss}
@@
 almost four times smaller risk than the full rank estimator, in which
-each covariance matrix is estimated independently.  When we generated
-data from a model with $s=p$, that is, the eigenvectors of $\Psi_k$
+each covariance matrix is estimated independently.  When the
+data come from a model in which $s=p$, that is, the eigenvectors of $\Psi_k$
 are generated uniformly from $\mathcal{V}_{p,r}$, the adaptive
@@
 the amount of similarity across groups.  Thus, shared subspace estimation
-can be especially appropriate choice when the similarity between
+can be an especially appropriate choice when the similarity between
 groups is not known a priori.  % In the simulations
@@
 
-In this Section we apply a shared subspace spiked covariance model to data in two
+In this Section we apply shared subspace spiked covariance models to data in two
 different biological examples.  In the first, we compare gene
 expression data from juveniles with different subtypes of leukemia.  Even
-after removing mean effects, the data indicate significant differences
+after removing mean effects, our results indicate significant differences
 in covariance matrices between groups.  In the second example, we
@@
 subtypes of leukemia: BCR-ABL, E2A-PBX1, Hyperdip50, MLL, T-ALL,
-TEL-AML1 and a seventh group for unidentified subtypes.  The groups
+TEL-AML1 and a seventh group for unidentified subtypes (``Others'').  The groups
 have corresponding sample sizes of $n = (15, 27, 64, 20, 43, 79, 79)$.
@@
 can be explained by the  shared subspace for all
-groups, with 4 of the 6 groups over 80\%.  Panel
+groups, with over 80\% of the variance explained in  4 of the 6 groups.  Panel
 \ref{fig:leukemiaPosterior} reflects some significant differences in
@@
 groups.  The $x$-axis corresponds to the orientation of the principle
-axis of $\Psi_k$ on $V$ and the $y$-axis corresponds to the log ratio of
+axis of $\Psi_k$ on $\hat{V}_1$ and the $y$-axis corresponds to the log ratio of
 the first two eigenvalues.  Several groups have significantly
@@
 leukemia subtypes. The 1\% of genes with the largest loadings on the
-first two columns of $\hat{V}$ are indicated with black shapes and the
+first two columns of $\hat{V}_1$ are indicated with black shapes and the
 remaining loadings with light grey dots.  The genes with the largest
@@
 In this figure, genes that fall in the upper right quadrant have
-positive loadings in both columns of $\hat{V}$. Genes in the upper left
-corner of the figure have positive loadings on the first column of $\hat{V}$
-but negative loadings on the second column of $\hat{V}$.  The principle axis
+positive loadings in both columns of $\hat{V}_1$. Genes in the upper left
+corner of the figure have positive loadings on the first column of $\hat{V}_1$
+but negative loadings on the second column of $\hat{V}_1$.  The principle axis
 for a group aligns in the direction of genes that exhibit the largest
@@
 upper right corner of the plot (squares), since this cluster of genes lie in
-a direction orthogonal to the principle axis.  In contrast, the
-BCR-ABL group is aligned primarily with the second column of $\hat{V}$,
+a direction nearly orthogonal to the principle axis.  In contrast, the
+BCR-ABL group is aligned primarily with the second column of $\hat{V}_1$,
 which means that the genes indicated by squares and circles vary
@@
 anti-correlated with those in the circle group, since their
-loadings have opposite signs on the second column of $\hat{V}$.
+loadings have opposite signs on the second column of $\hat{V}_1$.
 
@@
 
-Next, we briefly discuss an example in which we could not identify
+Next, we briefly discuss an example in which we could not readily identify
 significant covariance differences between groups.  Here, we apply
@@
 Differences between group covariance matrices can be understood in
-terms of differences in these sets of correlated and
-molecules.  In the applied analysis, we demonstrated how
+terms of differences in these sets of correlated molecules.  In the applied analysis, we demonstrated how
 we can use these notions to visualize and contrast the posterior
@@
 
-It is also a challenging problem to estimate the ``best'' subspace,
-$VV^T$, once the rank of the space is specified.  We used maximum
-marginal likelihood to estimate the shared subspace $VV^T$ and then
-use MCMC to infer $\Psi_k$.  By focusing on
-group differences for $\Psi_k$ on a \emph{fixed} subspace, it is much
-simpler to interpret similarities and differences.  Nevertheless, full
-uncertainty quantification for $VV^T$ can be desirable.  We found
-MCMC inference for $VV^T$ to be challenging for the problems considered
-in this paper and leave it for future work to develop an efficient
-fully Bayesian approach for estimating the joint posterior of $VV^T$ and $\Psi_k$.
-Recently developed Markov chain Monte Carlo algorithms, like Riemmanian manifold
-Hamilton Monte Carlo, which can exploit the geometry of the Grassmann
-manifold, may be useful here \citep{Byrne2013, Girolami2011}.  It may
-also be possible, though computationally intensive, to jointly
-estimate the $s$ and $VV^T$ using for instance, a
-reversible-jump MCMC algorithm.
+It is also a challenging problem to estimate the ``best'' subspace
+once the rank of the space is specified.  We used maximum marginal
+likelihood to estimate $VV^T$ and then use MCMC to
+infer $\Psi_k$.  By focusing on group differences for $\Psi_k$ on a
+\emph{fixed} subspace, it is much simpler to interpret similarities
+and differences.  Nevertheless, full uncertainty quantification for
+$VV^T$ can be desirable.  We found MCMC inference for $VV^T$ to be
+challenging for the problems considered in this paper and leave it for
+future work to develop an efficient fully Bayesian approach for
+estimating the joint posterior of $VV^T$ and $\Psi_k$.  Recently
+developed Markov chain Monte Carlo algorithms, like Riemmanian
+manifold Hamilton Monte Carlo, which can exploit the geometry of the
+Grassmann manifold, may be useful here \citep{Byrne2013,
+  Girolami2011}.  It may also be possible, though computationally
+intensive, to jointly estimate the $s$ and $VV^T$ using for instance,
+a reversible-jump MCMC algorithm.
 
@@
 can employ non-uniform Bingham prior distributions for the
-eigenvectors of $\Psi_k$ or model $\Psi_k$ as a function of continuous
+eigenvectors of $\Psi_k$ or we can model $\Psi_k$ as a function of continuous
 covariates as in \citet{Yin2010} and \citet{Hoff2011}.  Alternatively, we can

### 2016-07-09 paper/mscov.tex
Prettifyed figures.  Fixed typos.

@@
   \protect\thanks{Alexander M. Franks is a Moore/Sloan Data Science
-    and WRF Innovation in Data Science Postdoctoral Fellow and a
-    graduate of the Department of Statistics at Harvard University
+    and WRF Innovation in Data Science Postdoctoral Fellow
     (\href{mailto:amfranks@uw.edu}{amfranks@uw.edu}).  Peter D. Hoff
-    is a Professor in the Departments of Statistics, Biostatistics and
-    the Center for Statistics and the Social Sciences at the
-    University of Washington
+    is a Professor at the Departments of Statistics and Biostatistics
+    at the University of Washington
     (\href{mailto:pdhoff@uw.edu}{pdfhoff@uw.edu}).
@@
 This work was partially supported 
- by the Washington Research Foundation Fund for Innovation in Data-Intensive Discovery and the Moore/Sloan Data Science Environments Project at the University of Washington.
+ by the Washington Research Foundation Fund for Innovation in
+ Data-Intensive Discovery, the Moore/Sloan Data Science Environments
+ Project at the University of Washington and NSF grant DMS-1505136.
 %
  The authors are grateful to Dr.\ Daniel Promislow (Department of
- Pathology, University of Washington) and Dr.\ Jessica Hoffman
- (University of Alabama at Birmingham) for sharing data and ideas
- which contributed to framing of this paper.}}
-\author{Alexander Franks and Peter Hoff} \date{\today}
+ Pathology, University of Washington), Dr.\ Jessica Hoffman
+ (University of Alabama at Birmingham) and Dr. Julie Jossie (INRIA)
+ for sharing data and ideas which contributed to framing of this
+ paper.}}  \author{Alexander Franks and Peter Hoff} \date{\today}
 \maketitle 
-
 \begin{abstract}
@@
   gene expression and metabolomics data.
+\\
+\\
+\noindent\textbf{Keywords:} covariance estimation; spiked covariance
+model; Stiefel manifold; Grassmann manifold; big $p$, small $n$;
+high-dimensional data; empirical Bayes.
 \vfill
-\noindent\textbf{Keywords:} covariance estimation; spiked covariance
-model; Stiefel manifold; Grassmann manifold; big p, small n;
-high-dimensional data; empirical Bayes.
 \end{abstract}
@@
 groups, each of which represent samples from populations with
-distinct, but possibly related, distributions.  Although historically the
+distinct but possibly related distributions.  Although historically the
 primary focus has been on identifying mean-level differences between
 populations, there has been a growing need to identify differences in
-population covariances as well.  For instance, in case-control studies, mean level effects may be small relative to subject
+population covariances as well.  For instance, in case-control studies, mean-level effects may be small relative to subject
 variability; distributional differences between groups may still be
 evident as differences in the covariances between features.  Even when
-mean level differences are detectable, better estimates of the
+mean-level differences are detectable, better estimates of the
 covariability of features across groups may lead to an improved
-understanding of the mechanisms underlying these apparent mean level
+understanding of the mechanisms underlying these apparent mean-level
 differences.  Further, accurate covariance estimation is an essential
@@
 To address this need, we develop a novel method for
-group-covariance estimation.  In our method, we exploit the fact that
+multi-group covariance estimation.  Our method exploits the fact that
 in many natural systems, high dimensional data is often very
@@
 analysis (PLS-DA) is especially common in chemometrics and
-bioinformatics \citep{Barker2003} .  Although closely related to the
-approaches we will consider here, the primarily goal of of PLS-based
+bioinformatics \citep{Barker2003}.  Although closely related to the
+approaches we will consider here, the primarily goal of PLS-based
 models is to create regression or discrimination models, not to
@@
 assumption is widely applicable, even when there is little similarity
-in the covariance matrices across groups.  In particular, shared
-subspace models include those in which there is no information shared
-across groups because in the worst case we can assume the subspace
-spanned by the group-level eigenvectors is simply the full
-$p$-dimensional space.  In Section \ref{sec:app} we demonstrate the
-utility of a shared subspace model in an analysis of gene expression
-data from juvenile leukemia patients .  Despite the large feature size
-($p > 3000$) and small sample size ($n < 100$ per group), we identify
-interpretable similarities and differences in gene covariances on a
-low dimensional subspace.  Finally, we conclude with a brief
-exploration of heterogeneity in metabolomic data from a study on aging
-in the fly species \textit{Drosophila melanogaster}.
+in the covariance matrices across groups.  In particular, independent
+covariance estimation is equivalent to shared subspace
+estimation with a sufficiently large shared subspace.  In Section
+\ref{sec:app} we demonstrate the utility of a shared subspace model in
+an analysis of gene expression data from juvenile leukemia patients .
+Despite the large feature size ($p > 3000$) and small sample size
+($n < 100$ per group), we identify interpretable similarities and
+differences in gene covariances on a low dimensional subspace.
+Finally, we conclude with a brief exploration of heterogeneity in
+metabolomic data from a study on aging in the fly species
+\textit{Drosophila melanogaster}.
 
@@
 
-Suppose $S$ has a possibly degenerate Wishart$(\Sigma,n$)
+Suppose a random matrix $S$ has a possibly degenerate Wishart$(\Sigma,n$)
 distribution with density given by
@@
 case, $n$ is the number of independent observations minus the rank of
-the design matrix.  The spiked principle components model (spiked PCA)
+the design matrix.  The spiked principal components model (spiked PCA)
 studied by \citet{Johnstone2001} and others assumes that
@@
 \noindent where for $r \ll p$, $\Lambda$ is an $r\times r$ diagonal
-matrix and $U \in \mathcal V_{p,r}$.  $\mathcal V_{p,r}$ is the
-Stiefel manifold, which consists of the set of $p \times r$
-orthonormal matrices in $\mathbb{R}^p$, so that $U^TU =
-I$.  The spiked covariance formulation is appealing because it explicitly
-partitions the covariance matrix into a tractable low rank ``signal''
-and isotropic ``noise''.
+matrix and $U \in \mathcal V_{p,r}$, where $\mathcal V_{p,r}$ is the
+Stiefel manifold consisting of all $p \times r$ orthonormal matrices in $\mathbb{R}^p$, so that
+$U^TU = I_r$.  The spiked covariance formulation is appealing because it
+explicitly partitions the covariance matrix into a tractable low rank
+``signal'' and isotropic ``noise''.
 
@@
 largest eigenvalue of $S/(n\sigma^2)$ will converge to an upwardly
-biased version of $\lambda_{k}+1$, if $\lambda_k$ is greater than
-$\sqrt{\alpha}$ \citep{Baik2006, Paul2007} .  This has led several
+biased version of $\lambda_{k}+1$ if $\lambda_k$ is greater than
+$\sqrt{\alpha}$ \citep{Baik2006, Paul2007}.  This has led several
 authors to suggest estimating $\Sigma$ via shrinkage of the
@@
 different, however, when we are interested in estimating multiple
-covariance matrices from distinct, but related groups.  Here,
+covariance matrices from distinct but related groups.  Here,
 group-level equivariance to rotations is an unreasonable assumption;
@@
         \includegraphics[width=\textwidth]{Figs/3dplot}
-        \caption{Projection in $\mathcal{R}^3$}
+        \caption{Projection in $\mathbb{R}^3$}
         \label{fig:3dplot}
@@
       projected into different subspaces.  a) To visualize $Y_k$ we
-      can project the data into $R^3$.  In this illustration, the
+      can project the data into $\mathbb{R}^3$.  In this illustration, the
       distributional differences between the groups are confined to a
@@
 %
-\noindent where, the columns of $V \in \mathcal{V}_{p, s}$, with
+\noindent where the columns of $V \in \mathcal{V}_{p, s}$, with
 $s \ll p$, determine a subspace shared by all groups.  Throughout this
 paper we will denote to the shared subspace as
-$VV^T \in \mathcal G_{p,s}$, where $\mathcal G_{p,s}$, known as the
-Grassmann manifold, consists of all $s$-dimensional linear subspaces
+$VV^T \in \mathcal G_{p,s}$, where $\mathcal G_{p,s}$ is the
+Grassmann manifold consisting of all $s$-dimensional linear subspaces
 of $\mathbb{R}^p$ \citep{Chikuse2012}.  Although $V$ is only
@@
 
-In this section we outline an empirical Bayesian approach for
-estimating a low-dimensional shared subspace and the covariance
-matrices of the data projected onto this space. As we highlight in
-Section \ref{sec:simulation}, if the spiked covariance model holds for
-each group individually, then the shared subspace assumption also
-holds because at worst we can take the shared subspace to be the full
-$p$-dimensional space. In practice, we can usually identify a shared
-subspace of dimension $s \ll p$ that preserves most of the variation
-in the data.  Our primary objective is to identify the ``best'' shared
-subspace of fixed dimension $s < p$.  In Section \ref{sec:em} we
-describe an expectation maximization algorithm for estimating the
-maximum marginal likelihood of the shared subspace, $VV^T$.  This
-approach is computationally tractable for high-dimensional datasets.
-Given an inferred subspace, we then seek estimators for the covariance
-matrices of the data projected onto this space.  Because seemingly
-large differences in the point estimates of covariance matrices across
-groups may not actually reflect statistically significant differences,
-in Section \ref{sec:bayes} we also describe a Gibbs sampler that can
-be used to generate estimates of the projected covariance matrices,
-$\Psi_k$, and and their associated uncertainty.  Later, we discuss
-strategies for infering an appropriate value for $s$ and explore how
-shared subspace models can be used for exploratory data analysis by
-visualizing covariance heterogeneity on two or three dimensional
-subspaces.
+In this section we outline an empirical Bayes approach for estimating
+a low-dimensional shared subspace and the covariance matrices of the
+data projected onto this space. As we discuss in Section
+\ref{sec:simulation}, if the spiked covariance model holds for each
+group individually, then the shared subspace assumption also holds,
+where the shared subspace is simply the span of the group-specific
+eigenvectors, $U_1, ..., U_K$. In practice, we can usually identify a
+shared subspace of dimension $s \ll p$ that preserves most of the
+variation in the data.  Our primary objective is to identify the
+``best'' shared subspace of fixed dimension $s < p$.  In Section
+\ref{sec:em} we describe an expectation-maximization algorithm for
+estimating the maximum marginal likelihood of the shared subspace,
+$VV^T$.  This approach is computationally tractable for
+high-dimensional datasets.  Given an inferred subspace, we then seek
+estimators for the covariance matrices of the data projected onto this
+space.  Because seemingly large differences in the point estimates of
+covariance matrices across groups may not actually reflect
+statistically significant differences, in Section \ref{sec:bayes} we
+also describe a Gibbs sampler that can be used to generate estimates
+of the projected covariance matrices, $\Psi_k$, and and their
+associated uncertainty.  Later, in Section \ref{sec:simulation} we
+discuss strategies for inferring an appropriate value for $s$ and
+explore how shared subspace models can be used for exploratory data
+analysis by visualizing covariance heterogeneity on two or three
+dimensional subspaces.
 
@@
        VV^T/\sigma^2_k\right)S_k/2)\\
-&= \sum_k \frac{1}{2}\tr\left(-M_k^{-1}V^T S_kV\right) + \frac{1}{2\sigma_k^2}tr\left(VV^T S_k\right).
+&= \frac{1}{2}\sum_k \tr\left(-M_k^{-1}V^T S_kV\right) + \text{tr}\left(VV^T S_k\right)/\sigma_k^2.
 \end{align}
@@
 We maximize the marginal likelihood of $V$ with an EM algorithm, where
-$M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as the considered the
-``missing'' parameters to be imputed.  We assume independent Jeffreys'
+$M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ are considered the
+``missing'' parameters.  We assume independent Jeffreys'
 prior distributions for both $\sigma_k^2$ and $M_k$.  Jeffreys' prior
@@
 %
-$$p(M_k | V) \propto |M_k|^{(n_k + s + 1)/2}\etr(-1/2(M_k^{-1}V^TS_kV)) $$
+$$p(M_k | V) \propto |M_k|^{(n_k + s + 1)/2}\etr(-(M_k^{-1}V^TS_kV)/2) $$
 %
@@
 %
-$$p\left(\sigma^2 | V\right) \propto (\sigma_k^2)^{-n_k(p-s)/2-1}\etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)S_k\right)  $$
+$$p\left(\sigma^2 | V\right) \propto (\sigma_k^2)^{-n_k(p-s)/2-1}\etr\left(- (I-VV^T)S_k/[2\sigma_lk^2]\right)  $$
 %
 \noindent which is an inverse-gamma($n_k(p-s)/2$,
-$\frac{1}{2}\text{tr}[(I-VV^T)S_k]$) distribution.  Based on these
-results, we complete the following steps for each iteration, $t$, of
+$\text{tr}[(I-VV^T)S_k]/2$) distribution.  Based on these
+results, we complete the following steps for each iteration $t$ of
 the EM algorithm until convergence:
@@
 \citep{Wen2013}.  Importantly, the complexity of this algorithm is
-dominated by the dimension of the shared subspace not the number of
+dominated by the dimension of the shared subspace, not the number of
 features.  Thus, our approach is computationally efficient for
@@
 of the non-isotropic variation in $Y_k$ should be preserved when
-projecting the data to this space.  To characterize the extent to
+projecting the data onto this space.  To characterize the extent to
 which this is true for different groups, we propose a simple estimator
@@
 groups when $VV^T$ is a shared subspace for the data and smaller if
-not.  The metric provides a useful indicator for which groups can be
+not.  The metric provides a useful indicator of which groups can be
 reasonably compared on a given subspace and which groups cannot.  In
-practice, we estimate a shared subspace, $\hat{V}$, and the isotropic
-variances, $\hat{\sigma}_k^2$, using EM and compute the plug-in
+practice, we estimate a shared subspace $\hat{V}$ and the isotropic
+variances $\hat{\sigma}_k^2$ using EM and compute the plug-in
 estimate $\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2)$.  When this
-statistic is small for some groups, it may suggest that the rank, $s$,
+statistic is small for some groups, it may suggest that the rank $s$
 of the inferred subspace needs to be larger to capture the variation
@@
 The EM algorithm presented in the previous section yields point
-estimates for $VV^T$ and $\Psi_k$, but does not lead to natural
+estimates for $VV^T$, $\Psi_k$, and $\sigma_k^2$ but does not lead to natural
 uncertainty quantification for these estimates.  In this section, we
@@
 modeling the eigenstructure, we can now view each covariance
-$\Sigma_k$ in terms of the original spiked principle components model.  Equation \ref{eqn:sspsi}, written as a function of
+$\Sigma_k$ in terms of the original spiked principal components model.  Equation \ref{eqn:sspsi}, written as a function of
 $V$, becomes
@@
 \begin{align}
-\nonumber \Psi_k &= O_k\Lambda_KO_k^T\\
+\nonumber \Psi_k &= O_k\Lambda_kO_k^T\\
 \Sigma_k &= V\Psi_kV^T + \sigma^2_kI.
@@
 distribution of $O_k$ has a density proportional to the
-likelihood, so
+likelihood
 \begin{align}
@@
 element $\omega_i$ of $\Omega$ is proportional to the likelihood
-function, so
-
+function
 \begin{align}
@@
 true log ratios and orientations of $\hat{V}^T\Sigma_k\hat{V}$, where
-$\hat{V}$ is the maximum marginal likelihood for $V$. To compute
-the posterior regions, we iteratively remove posterior samples corresponding to the
-vertices of the convex hull until only 95\% of the original samples
-remain.  Non-overlapping posterior regions imply that differences in
-the covariances are statistically significant between groups.  In this
-example, the ratio of the eigenvalues of the true covariance matrices
-were $10$ (black and red groups), $3$ (green and blue groups) and $1$
-(cyan group).  Larger eigenvalue ratios correspond to more
-correlated contours and a value of $1$ implies isotropic covariance.
-Note that for the smaller eigenvalue ratio of $3$, there is larger
-uncertainty about the orientation of the primary axis.  When the ratio
-is one, as is the case for the cyan colored group, there is no
-information about the orientation of the primary axis since the contours
-are spherical.
+$\hat{V}$ is the maximum marginal likelihood for $V$. To compute the
+posterior regions, we iteratively remove posterior samples
+corresponding to the vertices of the convex hull until only 95\% of
+the original samples remain.  Non-overlapping posterior regions
+provide evidence that differences in the covariances are
+``statistically significant'' between groups.  In this example, the
+ratio of the eigenvalues of the true covariance matrices were $10$
+(black and red groups), $3$ (green and blue groups) and $1$ (cyan
+group).  Larger eigenvalue ratios correspond to more correlated
+contours and a value of $1$ implies isotropic covariance.  Note that
+for the smaller eigenvalue ratio of $3$, there is larger uncertainty
+about the orientation of the primary axis.  When the ratio is one, as
+is the case for the cyan colored group, there is no information about
+the orientation of the primary axis since the contours are spherical.
 
@@
       eigenvalues, $\text{log}(\frac{\lambda_1}{\lambda_2})$, of $\Psi_k$ and the
-      orientation of the principle axis on the space spanned by $\hat{V}$
+      orientation of the principal axis on the space spanned by $\hat{V}$
       cover the truth in this simulation.  Dots correspond to true data generating
@@
 $\lambda_1/\lambda_2 = 1$ (cyan) since this value is on the edge of the
-parameter space and is not be covered by the 95\% posterior regions
+parameter space and is not covered by the 95\% posterior regions
 as constructed.  The slight under coverage for the other groups is likely due to the fact
@@
 
-We use this data to evaluate how well the shared subspace estimator
+We use these data to evaluate how well the shared subspace estimator
 performs when we fit the data using a shared subspace model of
@@
 the shared subspace estimator with $\hat{s} = rK$ will perform well,
-provided that we can identify a subspace, $\hat{V}\hat{V}^T$, that is
+provided that we can identify a subspace, $\hat{V}\hat{V}^T$ that is
 close to $\text{span}([U_1, ..., U_K])$. When
@@
 a larger shared subspace is warranted.  When $\hat{s}$ is too small,
-$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$ will be significantly smaller
+$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$ will be substantially smaller
 than one for at least some of the groups, regardless of $\hat{V}$
@@
 $\alpha_k =\frac{p}{n_k}$, to estimate $r$.  Although their estimator
-was derived for individual covariance estimation, we found that,
-Gavish and Donoho's estimator can also be applied to the pooled data
-to estimate $s$.  Specifically, we concatenate the data from all
-groups to create an $(\sum_k n_k) \times p$ dimensional matrix and
-apply their rank estimator to this matrix to choose $s$.
+was derived for individual covariance estimation, we found that Gavish
+and Donoho's estimator can also be applied to the pooled data to
+estimate $s$.  Specifically, we concatenate the data from all groups
+to create an $(\sum_k n_k) \times p$ dimensional matrix and apply
+their rank estimator to this matrix to choose $s$.
 
@@
 
-We fit these datasets using three different variants of the shared
+We estimated group-level covariance matrices from these datasets using three different variants of the shared
 subspace model.  For each of these fits we estimate $r$.  First, we
@@
 compare covariance matrices in a metabolomic analysis of \textit{Drosophila} data.  In
-this example, after controlling for mean effects due to fly age, gender and
+this example, after controlling for mean effects due to fly age, sex and
 genotype, we find little compelling evidence of differences in metabolite
@@
 and $D$ is an $(s-r)$-dimensional diagonal matrix.  We write
-$V = [V_1, V_2]$, with $V_1 \in \mathcal{V}_{p,r}$ as the basis for a
+$V = [V_1, V_2]$, with $V_1 \in \mathcal{V}_{p,r}$ as the basis for an
 $r$-dimensional shared subspace that explains the differences between
@@
 differences between groups may not manifest themselves in the first
-principle components.  For instance, in genetic analyses, there may be
+principal components.  For instance, in genetic analyses, there may be
 large between subject variability common to all groups which is
 unrelated to how those subjects are grouped. Differences between
-groups may manifest themselves in the smaller principle
+groups may manifest themselves in the smaller principal
 components. This formulation allows a complete pooling of several
-eigenvectors, yet a space to identify and compare relevant
+eigenvectors, in addition to a space to identify and compare relevant
 differences.  Using this framework we can easily visualize and compare the
@@
 In Figure \ref{fig:leukemia} we plot the results of our analysis.
-Panel \ref{fig:leukemiaRatio} shows that over $60\%$ of the variance
-can be explained by the  shared subspace for all
-groups, with over 80\% of the variance explained in  4 of the 6 groups.  Panel
-\ref{fig:leukemiaPosterior} reflects some significant differences in
-the posterior distribution of eigenvalues and eigenvectors between
-groups.  The $x$-axis corresponds to the orientation of the principle
-axis of $\Psi_k$ on $\hat{V}_1$ and the $y$-axis corresponds to the log ratio of
-the first two eigenvalues.  Several groups have significantly
-different orientations for the first principle component of the
-projected covariance matrix, as well as differences in the ratio of
-eigenvalues.  For instance, the subtype E2A-PBX1 has a larger
-eigenvalue ratio which reflects a more correlated distribution on this
-subspace.  The posterior regions for the TEL-AML1 and Hyperdip50
-groups are almost entirely overlapping, which suggests there is little
-detectable difference between their covariance structures.  Note that
-when the eigenvalue ratio is close to one (as is the case for the
-``other'' subtype), the distribution of expression values on the
-subspace is nearly spherical and thus the orientation of the primary
-eigenvector is at best weakly identifiable.  This is reflected in the
-wide posterior range of orientations of the first principle component.
+Panel \ref{fig:leukemiaRatio} shows that over $60\%$ of the estimated
+variation in the top $s$ eigenvectors of $\Sigma_k$ can be explained
+by the shared subspace for all groups, with over 80\% explained in 4
+of the 6 groups.  Panel \ref{fig:leukemiaPosterior} reflects some
+significant differences in the posterior distribution of eigenvalues
+and eigenvectors between groups.  The $x$-axis corresponds to the
+orientation of the principal axis of $\Psi_k$ on $\hat{V}_1$ and the
+$y$-axis corresponds to the log ratio of the first two eigenvalues.
+Several groups have significantly different orientations for the first
+principal component of the projected covariance matrix, as well as
+differences in the ratio of eigenvalues.  For instance, the subtype
+E2A-PBX1 has a larger eigenvalue ratio which reflects a more
+correlated distribution on this subspace.  The posterior regions for
+the TEL-AML1 and Hyperdip50 groups are almost entirely overlapping,
+which suggests there is little detectable difference between their
+covariance structures.  Note that when the eigenvalue ratio is close
+to one (as is the case for the ``other'' subtype), the distribution of
+expression values on the subspace is nearly spherical and thus the
+orientation of the primary eigenvector is at best weakly identifiable.
+This is reflected in the wide posterior range of orientations of the
+first principal component.
 
@@
     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/leukemiaRatio-07-01}
+        \includegraphics[width=\textwidth]{Figs/leukemiaRatio-07-07}
         \caption{$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$}
@@
     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/leukemiaPosterior-07-01}
+        \includegraphics[width=\textwidth]{Figs/leukemiaPosterior-07-07}
         \caption{95\% Posterior Regions }
@@
       \caption{a) Goodness of shared subspace fit for each of the
-        seven Leukemia groups.  Over 60\% of the variance is explained
-        on the shared subspace in each of the seven groups, with
-        more than 80\% explained in most groups.  b) 95\% posterior
-        regions for the eigenvalue ratio and primary eigenvector
-        orientation.  Regions for some pairs of groups are disjoint,
-        suggesting significant differences in the projected covariance
-        matrices.  For other groups (e.g. Hyperdip50 and TEL-AML1)
-        overlap in the posterior regions indicate that differences are
-        not detectable on this subspace. }
+        seven Leukemia groups.  The subspace explains over $60\%$ of
+        the estimated variation in the top $s$ eigenvectors of
+        $\Sigma_k$ in each of the seven groups, with more than 80\%
+        explained in most groups.  b) 95\% posterior regions for the
+        eigenvalue ratio and primary eigenvector orientation.  Regions
+        for some pairs of groups are disjoint, suggesting significant
+        differences in the projected covariance matrices.  For other
+        groups (e.g. Hyperdip50 and TEL-AML1) overlap in the posterior
+        regions indicate that differences are not detectable on this
+        subspace. }
 \label{fig:leukemia}
@@
 corner of the figure have positive loadings on the first column of $\hat{V}_1$
-but negative loadings on the second column of $\hat{V}_1$.  The principle axis
+but negative loadings on the second column of $\hat{V}_1$.  The principal axis
 for a group aligns in the direction of genes that exhibit the largest
 variability in that group.  Genes which lie in a direction orthogonal
-to the principle axis exhibit reduced variability, relative to the
+to the principal axis exhibit reduced variability, relative to the
 other groups. As an example, genes in the upper left corner of this
@@
 upper right corner of the plot (squares), since this cluster of genes lie in
-a direction nearly orthogonal to the principle axis.  In contrast, the
+a direction nearly orthogonal to the principal axis.  In contrast, the
 BCR-ABL group is aligned primarily with the second column of $\hat{V}_1$,
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{Figs/leukemia-biplot-05-13}
+    \includegraphics[width=0.5\textwidth]{Figs/leukemia-biplot-07-07}
     \qquad
@@
   subtypes and the loadings for each gene on the first two columns of
-  $\hat{V}$.  All gene's loadings are displayed in light gray, and the
+  $\hat{V}$.  The loadings for all of the genes are displayed in light gray, and the
   top 1\% of genes with with the largest magnitude loadings are
@@
 
-Next, we briefly discuss an example in which we could not readily identify
-significant covariance differences between groups.  Here, we apply
-shared subspace group covariance estimation to a metabolomic data
-analysis on fly aging \citep{Hoffman2014}.  We bin the data to include
-groups of flies less than 10 days (young), between 10 and 51 days
-(middle) and greater than 51 days (old), and further split by
-sex. The sample sizes range from 43 to 59 flies per group.  We
-analyze metabolomic data corresponding to metabolites with 3714
-mass-charge ratios.  After removing mean effects due to age and
-sex, we fit the shared subspace estimator and compare covariances
-between groups.  We identify a subspace that explains almost all of
-the variability in the first few components of $\Sigma_k$ for all
-groups (Figure \ref{fig:dmelanRatio}) but see little evidence for
-differences in metabolite covariances on this subspace, as evidenced
-by the large overlap in posterior distributions for each group (Figure
+Next, we briefly discuss an example in which we could not readily
+identify significant covariance differences between groups.  Here, we
+apply shared subspace group covariance estimation to a metabolomic
+data analysis on fly aging \citep{Hoffman2014}.  We bin the data to
+include groups of flies less than 10 days (young), between 10 and 51
+days (middle) and greater than 51 days (old), and further split by
+sex. The sample sizes range from 43 to 59 flies per group.  We analyze
+metabolomic data corresponding to metabolites with 3714 mass-charge
+ratios.  After removing mean effects due to age and sex, we fit the
+shared subspace estimator and compare covariances between groups.  We
+identify a subspace that explains almost all of the variability in the
+first few components of $\Sigma_k$ for all groups (Figure
+\ref{fig:dmelanRatio}) but see little evidence for differences in
+metabolite covariances on this subspace, as evidenced by the large
+overlap in posterior distributions for each group (Figure
 \ref{fig:dmelanPosterior}). This could be indicative of a large amount
 of variation that is common across all groups for the first $s$
-principle components. Differences in covariation across groups likely
-manifest themselves in the smaller principle components but are too
-small to detect given the sample sizes.
+principal components. Differences in covariation across groups, if
+there are any, likely manifest themselves in the smaller principal
+components that are too small to detect given the sample sizes.
 
@@
     \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/dmelanRatio-04-28}
+        \includegraphics[width=\textwidth]{Figs/dmelanRatio-07-07}
         \caption{$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$}
@@
     \begin{subfigure}[b]{0.45\textwidth}r
-        \includegraphics[width=\textwidth]{Figs/dmelanPosterior-07-01}
+        \includegraphics[width=\textwidth]{Figs/dmelanPosterior-07-07}
         \caption{95\% Posterior Regions}
@@
 example, with the biological data, the shared subspace highlights the
-full set of genes or metabolites that are correlated across group.
+full set of genes or metabolites that are correlated across groups.
 Differences between group covariance matrices can be understood in
-terms of differences in these sets of correlated molecules.  In the applied analysis, we demonstrated how
+terms of differences in these sets of correlated molecules.  In
+analyses of these datasets, we demonstrated how
 we can use these notions to visualize and contrast the posterior
@@
 once the rank of the space is specified.  We used maximum marginal
-likelihood to estimate $VV^T$ and then use MCMC to
+likelihood to estimate $VV^T$ and then used MCMC to
 infer $\Psi_k$.  By focusing on group differences for $\Psi_k$ on a
@@
   Girolami2011}.  It may also be possible, though computationally
-intensive, to jointly estimate the $s$ and $VV^T$ using for instance,
+intensive, to jointly estimate $s$ and $VV^T$ using for instance,
 a reversible-jump MCMC algorithm.
@@
 wide range of analyses, especially when the number of features is very
-large.
+large.  A repository for the replication code is available on GitHub \citep{FranksGit}.
 

### 2016-07-12 paper/arxiv/mscov.tex
Final pre-arxiv edits

--- 2016-07-09 paper/mscov.tex
@@
 \documentclass[12pt]{article}
-
+\pdfoutput=1
 \usepackage{amsmath,amssymb,amsthm}
@@
     (\href{mailto:amfranks@uw.edu}{amfranks@uw.edu}).  Peter D. Hoff
-    is a Professor at the Departments of Statistics and Biostatistics
+    is a Professor in the Departments of Statistics and Biostatistics
     at the University of Washington
@@
  Pathology, University of Washington), Dr.\ Jessica Hoffman
- (University of Alabama at Birmingham) and Dr. Julie Jossie (INRIA)
+ (University of Alabama at Birmingham) and Dr. Julie Josse (INRIA)
  for sharing data and ideas which contributed to framing of this
@@
   We develop a model-based method for evaluating heterogeneity among
-  several $p\times p$ covariance matrices in the big $p$, small $n$ setting.
+  several $p\times p$ covariance matrices in the large $p$, small $n$ setting.
   This is done by assuming a spiked covariance model for each group
@@
 \noindent\textbf{Keywords:} covariance estimation; spiked covariance
-model; Stiefel manifold; Grassmann manifold; big $p$, small $n$;
+model; Stiefel manifold; Grassmann manifold; large $p$, small $n$;
 high-dimensional data; empirical Bayes.

### 2016-07-12 helper.R
Flag for plotting log ratio

@@ -69,7 +69,10 @@ getAgeStat <- function(normsMat) {
 
 }
 
-posteriorPlot <- function(Osamps, OmegaSamps, s2samps, nsamps, groups, probRegion=0.95, hline=NULL, col=NULL, pch=NULL, lty=NULL, ymax=30, plotPoints=TRUE) {
+posteriorPlot <- function(Osamps, OmegaSamps, s2samps, nsamps, groups,
+                          probRegion=0.95, hline=NULL, col=NULL,
+                          pch=NULL, lty=NULL, ymax=30, logRatio=FALSE,
+                          plotPoints=TRUE, cex.axis=1.5) {
 
     ngroups <- length(groups)
     
@@ -83,9 +86,12 @@ posteriorPlot <- function(Osamps, OmegaSamps, s2samps, nsamps, groups, probRegio
         lty=rep(1, ngroups)
     }
     par(mar=c(5.1, 5.1, 4.1, 2.1))
+    ylab <- ifelse(logRatio,
+                   expression("log(" ~ lambda[1]/lambda[2] ~ ")"),
+                   expression(lambda[1]/lambda[2]))
     plot(0, 0, xlim=c(-pi/2, pi/2),
-         ylim=c(0, ymax), cex=0, xlab="Angle", ylab=expression(lambda[1]/lambda[2]), xaxt="n", cex.axis=1.5, cex.lab=1.5)
-    axis(1, at=seq(-pi/2, pi/2, by=pi/4), labels=expression(-pi/2, -pi/4, 0, pi/4, pi/2), cex.axis=1.5, cex.lab=1.5)
+         ylim=c(0, ymax), cex=0, xlab="Angle", ylab=ylab, xaxt="n", cex.axis=cex.axis, cex.lab=1.5)
+    axis(1, at=seq(-pi/2, pi/2, by=pi/4), labels=expression(-pi/2, -pi/4, 0, pi/4, pi/2), cex.axis=cex.axis, cex.lab=1.5)
 
     
     for(g in groups) {
@@ -102,9 +108,12 @@ posteriorPlot <- function(Osamps, OmegaSamps, s2samps, nsamps, groups, probRegio
         if(pmPoint[1] < 0)
             pmPoint <- pmPoint * c(-1, -1)
 
-        hp <- getHullPoints(nsamps, OmegaSamps[, g, ], Osamps[, , g, ])
+        hp <- getHullPoints(nsamps, OmegaSamps[, g, ], Osamps[, , g, ],
[…]

### 2016-07-12 simulationFigures.R
Prettify figures

@@ -1,5 +1,5 @@
 load("SdimLoss.Rdata")
-pdf("paper/Figs/LossVsDimension.pdf")
+pdf("paper/Figs/LossVsDimension.pdf", font="Times")
 par(mar=c(5.1, 5.1, 4.1, 2.1))
 plot(5:50, lossLst, type="l", lwd=3, ylim=c(0, 100), xlim=c(0, 50), ylab="Stein's Risk", xlab=expression(hat(s)), cex.axis=2, cex.lab=2)
 grid()
@@ -13,7 +13,7 @@ legend("topright", lty=2, lwd=3, col=c("blue", "red"),
                 bty='n', cex=2)
 dev.off()
 
-pdf(sprintf("paper/Figs/simRatio-s5.pdf", format(Sys.Date(), "%m-%d")))
+pdf(sprintf("paper/Figs/simRatio-s5.pdf", format(Sys.Date(), "%m-%d")), font="Times")
 load("sdimtest-5-1.RData")
 evalRatios <- with(resList[[1]],
     sapply(1:ngroups, function(k) {
@@ -23,14 +23,14 @@ evalRatios <- with(resList[[1]],
     (numer/denom) / correction
     }))
 
-par(mar=c(7.1, 4.5, 4.1, 2.1))
+par(mar=c(5.1, 4.5, 4.1, 2.1))
 barplot(evalRatios, ylim=c(0, 1), xlab="Group", ylab="Goodness of Fit", main="",
-        space=0.2, cex.axis=1.5, col="#646464", col.axis="#646464",
+        space=0.2, cex.axis=2, col="#646464", col.axis="black",
         col.main="#646464", border=NA,
-        names.arg=1:10, cex.names=2, cex.main=2, cex.lab=2, col.lab="#646464")
+        names.arg=1:10, cex.names=2, cex.main=1, cex.lab=2, col.lab="black")
 dev.off()
 
-pdf(sprintf("paper/Figs/simRatio-s20.pdf", format(Sys.Date(), "%m-%d")))
+pdf(sprintf("paper/Figs/simRatio-s20.pdf", format(Sys.Date(), "%m-%d")), font="Times")
 load("sdimtest-20-1.RData")
 evalRatios <- with(resList[[1]],
     sapply(1:ngroups, function(k) {
@@ -40,8 +40,8 @@ evalRatios <- with(resList[[1]],
     (numer/denom) / correction
     }))
 
[…]

### 2017-06-13 helper.R
OptimV2 for faster inference

@@ -964,7 +964,8 @@ optimV <- function(Slist, P, S, nvec, PhiList, PrecVec,
         
         V <- lineSearch(P, S, V, G, F, rho1, rho2, tauStart)
         Fcur <- F(V)
-        
+
+        Gprev <- G
         G <- 0
         for(k in 1:length(PhiList)) {
             G <- G + Slist[[k]] %*%  (V %*%  PhiList[[k]])  -
@@ -978,8 +979,79 @@ optimV <- function(Slist, P, S, nvec, PhiList, PrecVec,
 }
 
 
-## optimization algorithm based on Wen 2013
-lineSearch <- function(n, p, X, G, F, rho1, rho2, tauStart, maxIters=50) {
+optimV2 <- function(Slist, P, S, nvec, PhiList, PrecVec,
+                    Vinit=NULL, rho=0.1,
+                    maxIters=50, verbose=FALSE) {
+
+    if(is.null(Vinit)) 
+        V <- rustiefel(P, S)
+    else
+        V <- Vinit
+
+    ## Function to optimize
+    F <- function(V) {
+        obj <- 0
+        for(k in 1:length(PhiList)) {
+            obj <- obj +
+                1/2 * tr(  t(V) %*% Slist[[k]] %*% V %*% PhiList[[k]] ) -
+                1/2 * PrecVec[k] * tr(t(V) %*% Slist[[k]] %*% V)
+        }
+        obj
+    }
+
+    ## dF(X)/dX
+    dF <- function(V) {
+        G <- 0
+        for(k in 1:length(PhiList)) {
+
+            G <- G + Slist[[k]] %*%  (V %*%  PhiList[[k]])  -
+                PrecVec[k] * Slist[[k]] %*%  V
+        }
+        G
+    }
+
+
+    eta <- 0.2
+    iter <- 1
+    Vprev <- rustiefel(P, S)
+    Fprev <- 0
+    Fcur <- F(V)
+    Ccur <- Fcur
+    Gcur <- dF(V)
+    Gprev <- dF(Vprev)
+    Qcur <- 1
+
+    while(Fprev/Fcur < 1-1e-4 & iter < maxIters) {
+
+        Fprev <- Fcur
+        if(verbose) {
+            print(sprintf("Iteration %i: %f", iter, Fprev))
+        }
+
+        newV <- lineSearchBB(P, S, V, Vprev, Gcur, Gprev, F, rho, Ccur)
+
+        ## Update
[…]

### 2017-06-13 fit-dmelan.R
Updated EMFit function call, posterior plots

@@ -51,7 +51,7 @@ getRank(fit$residuals)
 ## Vinit <- svd(do.call(cbind, lapply(1:ngroups, function(k) svd(t(residualList[[k]]))$u[, 1:R])))$u[, 1:min(ngroups*R, S)]
 Vinit <- svd(do.call(cbind, lapply(1:ngroups, function(k) svd(t(residualList[[k]]))$u[, 1:S])))$u[, 1:S]
 PrecVec = rep(5, ngroups)
-EMFit <- subspaceEM(Slist, P=P, S=S, R=R, nvec=nvec, Vstart=Vinit, PrecVec=PrecVec, verbose=TRUE, rho1=1e-5, rho2=1-1e-5)
+EMFit <- subspaceEM(Slist, P=P, S=S, R=R, nvec=nvec, Vstart=Vinit, verbose=TRUE, rho1=1e-5, rho2=1-1e-5)
 Vinit <- EMFit$V
 
 ## In paper using, 04-28

### 2017-06-13 fit-leukemia.R

@@ -3,12 +3,24 @@ rm(list=ls())
 library(car)
 library(rstiefel)
 library(xtable)
+library(made4)
+library(coda)
+
 source("helper.R")
 source("fit-subspace.R")
 source("https://bioconductor.org/biocLite.R")
+
 ## biocLite("hgu95av2.db")
 ## library(hgu95av2.db)
 load("LeukemiaData/leukemia.RData")
+
+## Look at PCA for pooled data to evaluate differences in means
+YpooledDemeaned <- sweep(Ypooled, 2, colMeans(Ypooled), '-')
+indices <- which(typeVec %in% c("TEL-AML1", "T-ALL", "MLL"))
+plotarrays(YpooledDemeaned[indices, ], classvec=typeVec[indices],
+           star=FALSE, arraycol=c("red", "blue", "dark green"))
+plotarrays(YpooledDemeaned, classvec=typeVec,
+           star=FALSE)
 
 ## number of eigenvectors if pooled space
 S <- 14
@@ -24,17 +36,29 @@ for( k in 1:ngroups ) {
     Slist[[k]] <- t(residual) %*% residual
 }
 pooledResiduals <- do.call(rbind, residualList)
-plot(svd(residualList[[ngroups]]/sqrt(nvec[ngroups]))$d^2, pch=19, cex=0.75)
+plot(svd(residualList[[ngroups]]/sqrt(nvec[ngroups]))$d^2, pch=19, cex=0.75, ylim=c(0, 50))
 for(k in 1:6) {
     points(svd(residualList[[k]]/sqrt(nvec[k]))$d^2, col=k+1, pch=19, cex=0.75)
 }
 svd(residualList[[ngroups]]/sqrt(nvec[ngroups]))$d^2
 
-Vinit <- svd(do.call(cbind, lapply(1:ngroups, function(k) svd(t(residualList[[k]]))$u[, 1:R])))$u[, 1:min(ngroups*R, S)]
+Vinit <- svd(do.call(cbind, lapply(1:ngroups, function(k) svd(t(residualList[[k]]))$u[, 1:S])))$u[, 1:min(ngroups*R, S)]
+
+isoVar <- sapply(1:ngroups, function(i) median(apply(residualList[[i]], 2, var)))
+weightsList <- sapply(1:ngroups, function(i) {
[…]

### 2017-06-13 paper/JRSSB/response/response.tex
Paper edits and response

\documentclass{article}
\pdfoutput=1
\pdfminorversion=4
%\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{textcomp}
\usepackage{longtable}
\usepackage{multirow, booktabs}
\usepackage{natbib}
 \bibpunct{(}{)}{;}{a}{,}{,}
\usepackage{url}
\usepackage{hyperref}
\usepackage{amsmath, amsfonts, amssymb}
\usepackage{mathtools}
\usepackage{chemarrow}
\usepackage{subfigure}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{tikz}
\usepackage[margin=.75in]{geometry}

\renewcommand{\baselinestretch}{1.5}
\newcommand{\bl}[1]{{\mathbf #1}}
\newcommand{\bs}[1]{{\boldsymbol #1}}
\newcommand{\tr}{\text{tr}}
\newcommand{\etr}{\text{etr}}
\newcommand{\Exp}[1]{{\text{E}}[ \ensuremath{ #1 } ]  }

\newtheorem{theorem}{Theorem}

\newcommand{\edits}[1]{{\textsf{\textcolor{red}{#1}}}}

\newenvironment{tquotation}{\begin{quotation}\color{gray}}{\end{quotation}}
\newenvironment{myquotation}{\begin{quotation}\slshape}{\end{quotation}}
\newenvironment{resp}{\begin{quotation}\noindent\slshape}{\end{quotation}}
\newcommand{\todo}[1]{{\textsf{\textcolor{red}{TODO: #1}}}}
%\usepackage{amsthm}

\title{Shared Subspace Models for Multi-Group Covariance Estimation}

\begin{document}

\section{Reviewer BLAH}

\textbf{\noindent 1. Page 5 has $S_k = Y_k^TY_K$ while Page 10 has $S_k = Y_kY_k^T$}

\begin{resp}
We have fixedt his to correctly reflect the fact that $S_k = Y_k^TY_K$ as we have defined $Y_k$ and an $n \times p$ matrix.
\end{resp}

\noindent   \textbf{2. Page 8. It is not clear to me why $||Y_kV||_F/n_k$ is a consistent estimator for the sum
of the eigenvalues of $V^T\Sigma_kV$. The quantity  $||Y_kV||_F/n_k$  seems to be a consistent estimator of $||V^T\Sigma_kV||^2_F$ and thus $||Y_kV||_F/n_k$ converges to 0. A consistent estimator of the sum of the eigenvalues of $V^T\Sigma_kV$ is Tr($V^TY_k^TY_kV)/n_k$}

\begin{resp}
Indeed, $||Y_kV||^2_F/n_k$ is a consistent estimator for the sum of the first $s$ eigenvalues of $\Sigma_k$.  In the draft, we excluded the square and thank the reviewer for highlighting this error.  The following argument holds:

\begin{align*}
||Y_kV||_F^2/n_k &= \tr (V^TY_k^TY_k/n_kV)\\
&= \tr (V^T\hat{\Sigma}_kV)\\
&\to tr(V^T\Sigma_kV)\\
& = tr(\Psi_k + \sigma_k^2I)
\end{align*}

\end{resp}

\noindent \textbf{3. Page 11. Inference of $O_k$ deserves more discussion. In fact, I am a bit confused here. The matrix $O_k$ is not identifiable because $VO_k\Lambda_kO^T \tilde{V}\tilde{O_k}Λ_k\tilde{O}\tilde{V}^T$ where $\tilde{V} = VW$ and $\tilde{O}_k = W^TO_k$ for some orthonormal W.}

\begin{resp}
We thank the reviewer for highlighting this source of confusion.  As we noted in the paper, although $V$ is only identifiable up to right rotations, the matrix $VV^T$, which defines
the plane of variation shared by all groups, is identifiable. Of course, this implies that $O$ itself is only identifiable once we have fixed a basis $V$.  The particular basis for $V$ is usually determined by the particular initialization to the EM algorithm. This itself isn't a problem since we are primarily interested in relative differences between the groups on the shared subspace.  For instance, if we had fixed a different basis for the subspace in the leukemia example, then the left plot in Figure 7 would be rotated version of the current figure, but the top 1\% of genes with with the largest magnitude loadings would not change.  Similarly, we would still correctly identify the genes that are most variable in each leukemia group. \todo{We have clarified these points in the paper.}
\end{resp}

\noindent \textbf{4. Page 16. I don’t think concatenate the data from all groups is a good way for rank selection. Consider the case $r = 2, K = 2$ and $n_1 =\sqrt{n_2}$. The top two eigenvalues of
the first group are $\lambda_1+\sigma_1^2 = \lambda_2+\sigma_1^2 > \sigma_1^2$.  The top two eigenvalues of the second group are both $\sigma_2^2$  or perhaps very very close to $\sigma_2^2$. In other words, there is almost no signal in the second group. Combining the data from two groups will mask the signal in the first group and in this case it is impossible for Gavish and Donoho’s estimator to work. The correct way to do it in this case is to use Gavish and Donoho’s estimator only in the first group. Generally, rank selection in this problem is complicated. Depending on the situation, one may want to assign different weights to different groups while using Gavish and Donoho’s estimator. Here I just provide a counter-example. I hope the authors can do a thorough revision of this section.}

\begin{resp}
\end{resp}

\noindent \textbf{5. Page 18. The whole Section 5 is a bit trivial. The conclusion is obvious. More groups means more data, which certainly implies higher accuracy. Moreover, only the case when groups are identically distributed is analyzed. This is the case that everyone will have the right intuition. There seems no need to write an entire section discussing it.  Unless the authors can give a general analysis, I will suggest remove this section.}

\begin{resp}
We believe that this section still has merit, although we have reframed the section to emphasize the most important points.  The purpose of this section is not to point out that ``more groups mean more data'' nor to provide an analysis for the case of $iid$ groups.  Rather, section is meant to identify a benchmark on the accuracy of the shared subspace estimator.  In general, this is a very hard problem, as the subspace accuracy depends on both the eigenvalues and eigenvectors of $\psi_k$.  When the eigenvalues from each group are equal, we conjecture that $\tr(\hat{V}\hat{V}^TVV^T)/\sqrt(S) \geq $ with equality if and only if the eigenvectors from each group are the same.  When the eigenvectors are highly variable across groups, the accuracy of the shared subspac estimator increases dramatically.  Our simulations support this conjecture.

Finally, as noted by \citet{} estimating the accuracy of the sample eigenvectors in the spiked covariance model is a hard problem that is not entirely well studied.  In our case, although the ML solution is clearly related to the eigenvectors, we do not have access to a simple function relating the eigenvectors of each sample covariance matrix to the ML solution, and thus it is not obvious how we can use results from RMT. 
\end{resp}

\noindent \textbf{6. Since $\Sigma_k = V\Psi_kV^T + \sigma^2_kI$ direct calculation gives
$\frac{\sum_k n_k\Sigma_k}{\sum_k n_k} = V\Psi V^T + \sigma^2I$ where $\Psi = \frac{\sum_k n_k \psi_k}{\sum n_k}$ and $\sigma^2 = \sum_k \frac{n_k \sigma_k^2}{\sum n_K}$.  This suggest one can directly apply an SVD on the sample covariance $\hat{\Sigma} = \frac{\sum S_k}{\sum n_k}$ to obtain $\hat{V}$. This naive method ignores the difference among groups, but avoid estimation of other nuisance. When the difference between different groups are not very large, I suspect this naive method should perform better. I suggest the authors compare this naive alternative with their proposal on both simulated and real data.  It will be helpful for readers to be informed when to use the naive one and when to use the proposed EM algorithm.}

\begin{resp}
We thank the reviwer for the interesting suggestion.  The suggested comparisons are closely related to the results presented in row 2 of table 1, comparing the performance of different models when the groups are in fact iid.  When the groups are identical, , .  Most importantly, the reviewer highlights the importance of identifying a simple method for initializing the EM algorithm (see below).
\end{resp}

\textbf{7. I wonder if the EM algorithm can be stuck at local. I also wonder how the authors initialize the algorithm. In particular, does the naive estimator suggested in the last point serve as a good initializer?}

\begin{resp}

We are in fact using an initialization chosen in a very similar, but did not include the details in the first draft of the paper.  We now have a section in the paper on convergence of the EM algorithm and the importance of a reasonable initialization.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.5\textwidth]{../../Figs/mode-histogram}
        \label{fig:modeHist}
    \caption{ Subspace similarity ($\tr(\hat{V}\hat{V}^TVV^T)/s$)
      between EM estimates and the true shared subspace for 1000 random initializations (grey) and ``smart'' intialization (red line).  }
\end{figure}
\end{resp}

\section{}

The heterogeneity is imposed via $\Psi_k$’s and information is shared via V (of suitable dimensions).  The modeling framework is interesting, but it does not come to be as a “surprise” in any way or form. The authors propose an empirical Bayes framework to perform inference. Some “running” justifications via RMT is used in various places of the paper.  In summary, I find the paper interesting. However, the methods employed in the paper as well as the analysis comes across to me as rather straightforward. I do not see sufficient novelty in the paper in any of the areas of modeling, computation or theory to warrant a publication in JRSSB.  Some comments are provided below

\begin{resp}
\end{resp}

The estimation of the parameters carried out via EM comes across to me as standard. The sampling procedures, as well as the choice of priors seem to be (directly) motivated via earlier work, some of them done by a subset of the authors.  One of the main criticisms for EM type algorithms is that there is no rigorous justification of the convergence of this procedure – especially because the problem considered involves dealing with extremely non-convex optimization problems. What can be said about the quality of the estimates of V , $\phi$, etc that are obtained? Are they prone to poor local minima or do they reach good solutions with provable guarantees (possibly under assumptions on the data)? An important theme of current research at the interface of statistical computing and optimization is to understand formal guarantees on the EM procedure.  Some studies along these lines may greatly benefit the paper.

\begin{resp}
\todo{}
\end{resp}

The analysis in Section 5 is interesting but is not complete (agreed, some simulation results
are provided) for the joint estimation problem. Can the gains be quantified more formally
by appealing to RMT results — by accounting for the fact that several parameters are being
jointly estimated?

\begin{resp}
\end{resp}

The Gavish Donoho threshold applies for a specific class of models (way more special than
the classes studied in this paper) – I am not sure if their threshold can be justified rigorously
in the context of the models being studied in this paper.

\begin{resp}
The Gavish-Donoho estimator applies to the class of spiked covariance models (single group).  Since each covariance matrix is a spiked covariance matrix, in fact the Gavish-Donoho threshold applies to estimating the rank from each group individually.  The challenge generalizing the rank estimator to the shared subspace setting.  \todo{} 
\end{resp}

\section{editor}

In addition, the following points are based on my own non-expert reading of your work. If you decide to revise and resubmit, I would also like you to clarify these in the next version of the paper.

1. I think the problem is rather specific – so I wonder if your work is of generic enough interest for the mainstream of Series B readers.

\begin{resp}
\end{resp}

2. What happens if the spiked covariance model holds only approximately?

\begin{resp}
\end{resp}

3. In my first decision letter, I asked for the theoretical description of the properties of your method,but I am not sure if this has been done successfully. In particular, I cannot see theorem-type results quantifying the behaviour of your procedure.

\begin{resp}
\todo{BLAH}
\end{resp}

\end{document}

### 2017-06-13 paper/JRSSB/mscov.tex
Paper edits and response

--- 2016-07-12 paper/arxiv/mscov.tex
@@
-\documentclass[12pt]{article}
+\documentclass{statsoc}
 \pdfoutput=1
-\usepackage{amsmath,amssymb,amsthm}
+\pdfminorversion=4
+%\usepackage{amsmath,amssymb,amsthm}
 \usepackage{graphicx}
@@
 \usepackage{natbib}
+ \bibpunct{(}{)}{;}{a}{,}{,}
 \usepackage{url}
 \usepackage{hyperref}
-\usepackage{amsmath}
+\usepackage{amsmath, amsfonts, amssymb}
 \usepackage{mathtools}
 \usepackage{chemarrow}
-\usepackage{subcaption}
-\usepackage{amsmath}
-\usepackage{mathtools}
+\usepackage{subfigure}
 \usepackage{rotating}
@@
 
+\newtheorem{theorem}{Theorem}
+\newtheorem{conjecture}{Conjecture}
+
+\newcommand{\edits}[1]{{\textsf{\textcolor{red}{#1}}}}
+
+%\usepackage{amsthm}
+
+\title{Shared Subspace Models for Multi-Group Covariance Estimation}
+\author{Alexander M. Franks}
+\address{University of Washington, Seattle, Washington, USA}
+\author[Franks and Hoff]{Peter Hoff}
+\address{Duke University, Durham, North Carolina, USA}
+
 \begin{document}
 
-\title{Shared Subspace Models for Multi-Group Covariance Estimation
-%
-  \protect\thanks{Alexander M. Franks is a Moore/Sloan Data Science
-    and WRF Innovation in Data Science Postdoctoral Fellow
-    (\href{mailto:amfranks@uw.edu}{amfranks@uw.edu}).  Peter D. Hoff
-    is a Professor in the Departments of Statistics and Biostatistics
-    at the University of Washington
-    (\href{mailto:pdhoff@uw.edu}{pdfhoff@uw.edu}).
-%
-This work was partially supported 
- by the Washington Research Foundation Fund for Innovation in
- Data-Intensive Discovery, the Moore/Sloan Data Science Environments
- Project at the University of Washington and NSF grant DMS-1505136.
-%
- The authors are grateful to Dr.\ Daniel Promislow (Department of
- Pathology, University of Washington), Dr.\ Jessica Hoffman
- (University of Alabama at Birmingham) and Dr. Julie Josse (INRIA)
- for sharing data and ideas which contributed to framing of this
- paper.}}  \author{Alexander Franks and Peter Hoff} \date{\today}
-\maketitle 
 \begin{abstract}
-
-% From Hoff Although the covariance matrices corresponding to different
-% populations are unlikely to be exactly equal they can still exhibit a
-% high degree of similarity. For example, some pairs of variables may be
-% positively correlated across most groups, whereas the correlation
-% between other pairs may be consistently negative. In such cases much
-% of the similarity across covariance matrices can be described by
-% similarities in their principal axes, which are the axes that are
-% defined by the eigenvectors of the covariance matrices. Estimating the
-% degree of across-popu- lation eigenvector heterogeneity can be helpful
-% for a variety of estimation tasks.
 
@@
 in the covariance matrices across groups.  In particular, independent
-covariance estimation is equivalent to shared subspace
-estimation with a sufficiently large shared subspace.  In Section
-\ref{sec:app} we demonstrate the utility of a shared subspace model in
-an analysis of gene expression data from juvenile leukemia patients .
-Despite the large feature size ($p > 3000$) and small sample size
-($n < 100$ per group), we identify interpretable similarities and
-differences in gene covariances on a low dimensional subspace.
-Finally, we conclude with a brief exploration of heterogeneity in
-metabolomic data from a study on aging in the fly species
-\textit{Drosophila melanogaster}.
+covariance estimation is equivalent to shared subspace estimation with
+a sufficiently large shared subspace.  In Section \ref{sec:asymp} we
+use an asymptotic approximation to describe how shared subspace
+inference reduces bias when both $p$ and $n$ are large.  Finally, In
+Section \ref{sec:app} we demonstrate the utility of a shared subspace
+model in an analysis of gene expression data from juvenile leukemia
+patients .  Despite the large feature size ($p > 3000$) and small
+sample size ($n < 100$ per group), we identify interpretable
+similarities and differences in gene covariances on a low dimensional
+subspace.  Finally, we conclude with a brief exploration of
+heterogeneity in metabolomic data from a study on aging in the fly
+species \textit{Drosophila melanogaster}.
 
@@
     \centering
-    \begin{subfigure}[b]{0.25\textwidth}
-        \includegraphics[width=\textwidth]{Figs/3dplot}
-        \caption{Projection in $\mathbb{R}^3$}
+    \subfigure[Projection in $\mathbb{R}^3$]{
     \label{fig:3dplot}
-    \end{subfigure}
+    \includegraphics[width=0.25\textwidth]{../Figs/3dplot}}
 \quad
- %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-      %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.25\textwidth}
-        \includegraphics[width=\textwidth]{Figs/2d-scatter}
-        \caption{$Y_kV$}
+    \subfigure[$Y_kV$]{
     \label{fig:2dscatter}
-    \end{subfigure}
+    \includegraphics[width=0.25\textwidth]{../Figs/2d-scatter}}
 \quad 
-    \begin{subfigure}[b]{0.25\textwidth}
-        \includegraphics[width=\textwidth]{Figs/2d-scatter-orth}
-        \caption{$Y_kV_{\perp}$}
+\subfigure[$Y_kV_{\perp}$]{
 \label{fig:2dscatterorth}
-    \end{subfigure}
+\includegraphics[width=0.25\textwidth]{../Figs/2d-scatter-orth}}
 \caption{ Two groups of four-dimensional data (red and blue)
@@
        VV^T/\sigma^2_k\right)S_k/2)\\
-&= \frac{1}{2}\sum_k \tr\left(-M_k^{-1}V^T S_kV\right) + \text{tr}\left(VV^T S_k\right)/\sigma_k^2.
+%\nonumber &= \frac{1}{2}\sum_k \tr\left(-M_k^{-1}V^T S_kV\right) +
+%  \text{tr}\left(VV^T S_k\right)/\sigma_k^2\\
+&=\frac{1}{2}\sum_k \tr\left((\frac{1}{\sigma_k^2}I-M_k^{-1})V^T
+  S_kV\right).
+\label{eqn:likV}
 \end{align}
-
-% \begin{align}
-% p(V|...) &\propto \etr(\sum_{k=1}^K[O_k\Omega_kO^T_k]V^TS_k/(2\sigma^2_k)]V)\\
-% \end{align}
-
+%
 We maximize the marginal likelihood of $V$ with an EM algorithm, where
@@
 %
-$$p(M_k | V) \propto |M_k|^{(n_k + s + 1)/2}\etr(-(M_k^{-1}V^TS_kV)/2) $$
+$$p(M_k | V) \propto |M_k|^{-(n_k + s + 1)/2}\etr(-(M_k^{-1}V^TS_kV)/2) $$
 %
@@
 %
-$$p\left(\sigma^2 | V\right) \propto (\sigma_k^2)^{-n_k(p-s)/2-1}\etr\left(- (I-VV^T)S_k/[2\sigma_lk^2]\right)  $$
+$$p\left(\sigma^2 | V\right) \propto (\sigma_k^2)^{-n_k(p-s)/2-1}\etr\left(- (I-VV^T)S_k/[2\sigma_k^2]\right)  $$
 %
@@
 \end{enumerate}
-\item Compute $V_{t}$ = $\underset{V}{\text{arg } \text{max}}  \sum_k \text{tr}\left(-VE[M^{-1}|V_{(t-1)}]^T +
+\item Compute $V_{t}$ = $\underset{V}{\text{arg } \text{max}}  \sum_k \text{tr}\left(-VE[M^{-1}|V_{(t-1)}]V^T +
        E[\frac{1}{2\sigma_k^2}|V_{(t-1)}]VV^T\right)S_k/2)$ 
@@
 
+\paragraph{Initialization and Convergence:}
+\edits{The Stiefel manifold is compact and the marginal
+  likelihood is continuous, so the likelihood is bounded.  Thus, the EM
+  algorithm, which increases the likelihood at each iteration, will
+  converge to a stationary point \citep{WuEM}. However, maximizing the
+  marginal likelihood of the shared subspace model corresponds to a
+  non-convex optimization problem over the Grassmmanian manifold and
+  may converge to a sub-optimal local mode or stationary
+  point. Other work involving optimization on the Grassmmanian has
+  found convergence to non-optimal stationary values problematic and emphasized
+  the importance of good (e.g. $\sqrt{n}$-consistent) starting values
+  \citep{Cook2016}. Our empirical results on simulated data confirms
+  that randomly initialized starting values converge to sub-optimal
+  stationary values, and so in practice we initialize the algorithm at
+  a carefully chosen starting value.}
+
+\edits{Note that when the shared subspace model holds, the first $s$
+  eigenvectors, from \emph{any} of the groups can be used to construct a
+  $\sqrt{n}$-consistent estimator of $VV^T$.  In particular, if
+  $\hat{U}^{(k)}\hat{U}^{(k)^T}$ is the eigenprojection matrix for the
+  subspace spanned by the first $s$ eigenvectors of $S_k$ then it can
+  be shown that
+  $\sqrt{n} \text{ vec}(\hat{U}^{(k)}\hat{U}^{(k)^T} - VV^T)$
+  converges in distribution to a mean-zero normal
+  \citep{kollo2000}.  In the large $p$, small $n$ setting, such
+  classical asymptotic guarantees give little assurance that the
+  resulting estimators would be reasonable, but they nevertheless
+  suggest useful strategies for identifying starting value for the EM
+  algorithm.}
+
+\edits{In this work, we choose a subspace initialization strategy
+  based on sample eigenvectors of the data pooled from all groups.    Let
+  $ Z = \sum_k \pi_k \frac{Z_k}{\sigma_k}$ where $Z_k$ is a mean-zero
+  normal with covariance $\Sigma_k$ and $\pi_k = n_k / \sum_k
+  n_k$. Then $Z$ is a
+  mixture of mean-zero normal distributions with covariance 
+  \begin{align*}
+\Sigma_Z &= \sum_k \frac{\pi_k}{\sigma^2_k} \Sigma_k\\
+  &= V^T( \sum_k \frac{\pi_k}{\sigma_k^2} \Psi_k)V + I,
+  \end{align*} Clearly, the first $s$
+  eigenvectors of $\Sigma_Z$ span the shared subspace, $VV^T$.  This suggests
+  that we can estimate the shared subspace using the the scaled and pooled
+  data, $Y_{\text{pool}} = [\frac{1}{\sigma_1}Y_1;
+  \frac{1}{\sigma_2}Y_2;...;  \frac{1}{\sigma_k}Y_k]$, where
+  $Y_{\text{pool}}$ has dimension $\sum_k n_k \times p$.  
+  We use $\hat{U}_{\text{pool}}\hat{U}_{\text{pool}}^T$  as the
+  initial value for subspace estimation algorigthm where 
+  $\hat{U}_{\text{pool}}$ are the first $s$ eigenvectors of
+  $S_{\text{pool}} = Y_{\text{pool}}^TY_{\text{pool}}$ to
+  If we
+  treat $Y_{\text{pool}}$ as an iid sample from the mixture
+  distribution $Z$,  then it is known that  $\hat{U}_{\text{pool}}\hat{U}_{\text{pool}}^T$ is not
+  consistent when both $n$ and $p$ growing at the same rate.}
+\edits{For any $p$-vector $\eta$, the asymptotic bias of
+  $\eta^T\hat{U}_{\text{pool}}\hat{U}_{\text{pool}}^T\eta$ is well characterized as a function of the eigenvalues of $\Sigma_Z$\citep{Mestre2008}.  If either the eigenvalues of $\sum_k
+  \frac{\pi_k}{\sigma_k^2} \Psi_k$ or the total sample size $\sum_k n_k$
+  are large, $\hat{U}_{\text{pool}}\hat{U}_{\text{pool}}^T$ will
+  accurately estimate the shared subspace and likelihood based
+  optimization my have little added benefit.  When either the eigenvalues
+  or sample size are small to moderate in magnitude,
+  $\hat{U}_{\text{pool}}\hat{U}_{\text{pool}}^T$ is a useful starting value for the EM algorithm.}
+  % An asymptotically lower variance estimator can then be constructed
+  % by combining consistent estimators from each of the $K$ groups
+  % (e.g. by weighting each of the sample eigenprojection by the inverse
+  % Fisher Information).
+
+ % In this paper we consider initializations
+  % based on a weighted average of the eigenprojection matrices
+  % associated with each eigenvalue from each group:
+  % $ P = \frac{\sum_k \sum_i w_{ik} U_i^{(k)}U_i^{(k)^T}}{\sum_k\sum_i
+  %   w_{ik}}$.  In general, $P \notin \mathcal{G}_{p, s}$, i.e. P is
+  % not a projection matrix.  Thus, we choose the starting value to be
+  % the closest projection matrix
+  % $\hat{V}_{\text{init}}\hat{V}_{\text{init}}^T$ to $P$.
+  % $\hat{V}_{init}$ is the first $s$ eigenvectors of $P$.
+  % $\frac{\ell_i^{(k)}}{\sigma^2}$
+
+% \begin{figure}[t]
+%     \centering
+%     \includegraphics[width=0.5\textwidth]{../Figs/mode-histogram}
+%         \label{fig:modeHist}
+%     \caption{ Subspace similarity ($\tr(\hat{V}\hat{V}^TVV^T)/s$)
+%       between EM estimates and the true shared subspace for 1000 random initializations (grey) and ``smart'' intialization (red line).  }
+% \end{figure}
+
 \paragraph{Evaluating Goodness of Fit:}
 
-If $V$ is a basis for a shared subspace, then for each group, most
-of the non-isotropic variation in $Y_k$ should be preserved when
-projecting the data onto this space.  To characterize the extent to
-which this is true for different groups, we propose a simple estimator
-for the proportion of ``signal'' variance that lies on a given
-subspace.  Specifically, we use the following statistic for the ratio
-of the sum of the first $s$ eigenvalues of
-$V^T \Sigma_k V$ to the sum of the first $s$ eigenvalues
-of $\Sigma_k$:
+\edits{Tests for evaluating whether eigenvectors from multiple groups
+  span a common subspace were explored extensively by
+  \citet{schott1991some}.  These tests can be useful for assessing
+  whether a shared subspace model is appropriate, but cannot
+  be used to test whether a particular subspace explains
+  variation across groups.  These results are also based
+  on classical asymptotics and are thus less accurate when
+  $n \ll p$}
+
+Our goodness of fit measure if based on the fact that when $V$ is a
+basis for a shared subspace, then for each group, most of the
+non-isotropic variation in $Y_k$ should be preserved when projecting
+the data onto this space.  To characterize the extent to which this is
+true for different groups, we propose a simple estimator for the
+proportion of ``signal'' variance that lies on a given subspace.
+Specifically, we use the following statistic for the ratio of the sum
+of the first $s$ eigenvalues of $V^T \Sigma_k V$ to the sum of the
+first $s$ eigenvalues of $\Sigma_k$:
 %
 \begin{equation}
-\gamma(Y_k: V, \sigma_k^2) = \frac{||Y_kV||_F/n_k}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
-  ||Y_k\widetilde{V}||_F/n_k - \sigma_k^2ps/n_k}
+\gamma(Y_k: V, \sigma_k^2) = \frac{||Y_kV||^2_F/n_k}{\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
+  ||Y_k\widetilde{V}||^2_F/n_k - B_k }
 \label{eqn:ratio}
@@
 %
-\noindent where $||\cdot||_F$ is the Frobenius norm. Our motivation for
-this statistic is the fact that if
-$\Sigma_k = V\Psi_kV^T + \sigma_k^2I$ then
-$\gamma(Y_k: V, \sigma_k^2) \approx 1$.  We establish
-this in an asymptotic regime where $s$ is fixed, $p$,
-$n_k \rightarrow \infty$ and $p/n_k = \alpha_k$ constant.
-
-First, since $s$ is fixed and $n_k$ is growing, the numerator,
-$||Y_kV||_F/n_k$, is a consistent estimator for the sum of the
-eigenvalues of $V^T\Sigma_kV$.  The denominator,
+\noindent where $||\cdot||_F$ is the Frobenius norm and $B_k$ is a
+bias correction where $B_k = \sigma_k^2 p / n_k \sum_k
+\left(\frac{m_i^{(k)}}{m_i^{(k)}-\sigma^2_k}\right)$ with $m_i^{(k)}$
+the positive solution to the quadratic equation 
+\begin{equation}
+(m_i^{(k)})^2 + m_i^{(k)}(\sigma^2_kp/n_k - \sigma_k^2 - \hat{\lambda}_i^{(k)}) -
+\hat{\lambda}_i^{(k)}\sigma_k^2 = 0.
+\label{eqn:quad}
+\end{equation}
+and $\hat{\lambda}_i^{(k)}$ is the $i$-th eigenvalue of $S_k/n_k$.
+\edits{
+\begin{theorem}
+Assume $p/n_k \to \alpha_k$ and $s$ is fixed. If
+$\Sigma_k = V\Psi_kV^T + \sigma_k^2I$, then $\gamma(Y_k: V, \sigma_k^2)
+\to 1$ as $n_k, p \to \infty$.  
+\end{theorem}
+}
+\edits{
+\begin{proof}
+Since $s$ is fixed and $n_k$ is growing, the numerator,
+$||Y_kV||^2_F/n_k$, is a consistent estimator for the sum of the
+eigenvalues of $V^T\Sigma_kV$.  In the denominator,
 $\underset{\widetilde{V} \in \mathcal{V}_{p,
-    s}}{\text{max}}||Y_k\widetilde{V}||_F/n_k$
+    s}}{\text{max}}||Y_k\widetilde{V}||^2_F/n_k$
 is equivalent to the sum of the first $s$ eigenvalues of the sample
-covariance matrix $S_k/n_k$.  With $\hat{\lambda}^{(k)}_i$ the
-$i$-th eigenvalue of $S_k/n_k$, \citet{Baik2006} demonstrated that
-asymptotically as $p, n_k \rightarrow \infty$ with $p/n_k = \alpha_k$
-fixed
+covariance matrix $S_k/n_k$.  \citet{Baik2006} and others have demonstrated that
+asymptotically as $p, n_k \rightarrow \infty$ and $p/n_k = \alpha_k$, $\hat{\lambda}^{(k)}_i$ is
+positively biased.  Specifically,
 %
 \begin{eqnarray}
-\nonumber \hat{\lambda}^{(k)}_i &\rightarrow& \lambda^{(k)}_i\left(1 +
+\hat{\lambda}^{(k)}_i &\rightarrow& \lambda^{(k)}_i\left(1 +
                                     \frac{\sigma_k^2\alpha_k}{\lambda^{(k)}_i
-                                    - 1}\right)\\
-& \approx& \lambda^{(k)} + \sigma^2_k\frac{p}{n_k}
+                                    - \sigma^2_k}\right) 
 \label{eqn:lamAsymp}
@@
 %
-\noindent where the approximation in the second line is due to the
-fact that $\lambda_k/(\lambda_k - 1) \approx 1$ for $\lambda_k$ large.
-Thus,
-$\sum_i^s \hat{\lambda}^{(k)}_i \approx \sum_i^s \lambda^{(k)}_i +
-\sigma_k^2ps/n_k$
-and
-$\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
-||Y_k\widetilde{V}||_F/n_k - \hat{\sigma}_k^2ps/n_k$
-is a reasonable approximation to the sum of the first $s$ eigenvalues
-of $\Sigma_k$.  Alternatively, when $\lambda_k$ is small, a better
-approximation may be obtained by solving the quadratic equation
-implied by the first line of Equation \ref{eqn:lamAsymp}.
-
-As a consequence, the proposed statistic will be close to one for all
-groups when $VV^T$ is a shared subspace for the data and smaller if
+\noindent Replacing $\lambda_i^{(k)}$ by $m_i^{(k)}$ and assuming equality in \ref{eqn:lamAsymp} yields the quadratic equation \ref{eqn:quad}.  The
+solution, $m_i^{(k)}$, is an asymptotically (in $n$ and $p$) unbiased
+estimator of $\lambda_i^{(k)}$ and
+\begin{equation}
+\underset{\widetilde{V} \in \mathcal{V}_{p, s}}{\text{max}}
+  ||Y_k\widetilde{V}||^2_F/n_k - B_k  \to \sum_{i}^s \lambda_i^{(k)} 
+\end{equation}
+As such, when the shared subspace model holds both the
+numerator and denomiator of the goodness of fit statistic converge
+almost surely to $\sum_{i=1}^{s}
+\lambda_i^{(k)}$.  Therefore $\gamma(Y_k: V, \sigma_k^2)
+\to 1$.
+\end{proof}
+}
+% Our motivation for
+% this statistic is the fact that if
+% $\Sigma_k = V\Psi_kV^T + \sigma_k^2I$ then
+% $\gamma(Y_k: V, \sigma_k^2) \approx 1$.  We establish
+% this in an asymptotic regime where $s$ is fixed, $p$,
+% $n_k \rightarrow \infty$ and $p/n_k = \alpha_k$ constant.
+
+The goodness of fit statistic will be close to one for all
+groups when $VV^T$ is a shared subspace for the data and typically smaller if
 not.  The metric provides a useful indicator of which groups can be
@@
 practice, we estimate a shared subspace $\hat{V}$ and the isotropic
-variances $\hat{\sigma}_k^2$ using EM and compute the plug-in
-estimate $\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2)$.  When this
-statistic is small for some groups, it may suggest that the rank $s$
-of the inferred subspace needs to be larger to capture the variation
-in all groups. We investigate this in Section \ref{sec:simulation}, by
-computing the goodness of fit statistic for inferred subspaces of
-different dimensions on a single dataset. In Section \ref{sec:app}, we
-compute the estimates for subspaces inferred with real biological
-data.
-
-%% Problem here! Can't compare different sample sizes??
+variances $\hat{\sigma}_k^2$ using EM and compute the plug-in estimate
+$\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2)$.  When this statistic is
+small for some groups, it may suggest that the rank $s$ of the
+inferred subspace needs to be larger to capture the variation in all
+groups. If $\gamma(Y_k: \hat{V}, \hat{\sigma}_k^2)$ is substantially
+larger than 1 for a particular group, it suggests that the inferred subspace
+is too similar to the sample principal components from group $k$.  We
+investigate these issues in Section \ref{sec:simulation}, by computing the
+goodness of fit statistic for inferred subspaces of different
+dimensions on a single dataset. In Section \ref{sec:app}, we compute
+the estimates for subspaces inferred with real biological data.
 
@@
 The likelihood for $\Sigma_k$ given the sufficient statistic
-$S_k = Y_kY_k^T$ is given in Equation \ref{eqn:lik}.  For the
+$S_k = Y_k^TY_k$ is given in Equation \ref{eqn:lik}.  For the
 spiked PCA formulation, we must rewrite this likelihood in terms of $V$, $O_k$,
@@
     \centering
-%    \begin{subfigure}[b]{0.45\textwidth}
-    \includegraphics[width=0.5\textwidth]{Figs/posteriorRegions-chull}
-        \caption{95\% posterior regions}
+    \includegraphics[width=0.5\textwidth]{../Figs/posteriorRegions-chull}
         \label{fig:simPosterior}
-%    \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-      %(or a blank line to force the subfigure onto a new line)
-    % \begin{subfigure}[b]{0.4\textwidth}
-    %     \includegraphics[width=\textwidth]{Figs/simRatio-05-19}
-    %     \caption{Ratio}
-    %     \label{fig:simRatio}
-    % \end{subfigure}
     \caption{a) 95\% posterior regions for the log of the ratio of
@@
     \centering
-    \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=\textwidth]{Figs/LossVsDimension}
-        \caption{Stein's risk vs $\hat{s}$}
+    \subfigure[Stein's risk vs $\hat{s}$]{
       \label{fig:sdimension}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=\textwidth]{Figs/simRatio-s5}
-        \caption{$\hat{s}$ = 5}
+      \includegraphics[width=0.3\textwidth]{../Figs/LossVsDimension}}
+    \subfigure[$\hat{s}$ = 5]{
       \label{fig:ratio-s5}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=\textwidth]{Figs/simRatio-s20}
-        \caption{$\hat{s}$ = 20}
+      \includegraphics[width=0.3\textwidth]{../Figs/simRatio-s5}}
+    \subfigure[$\hat{s}$ = 20]{
         \label{fig:ratio-s20}
-    \end{subfigure}
+        \includegraphics[width=0.3\textwidth]{../Figs/simRatio-s20}}
       \caption{a) Stein's risk as a function of the
@@
 \begin{table}
-\begin{center}
-  \begin{tabular}{ l  l | c | c | c |}
-    \multicolumn{2}{c}{} & \multicolumn{3}{c}{\textbf{Inferential Model}} \\
-  \multicolumn{2}{c|}{}  & Adaptive & $\hat{\Sigma}_k=\hat{\Sigma}$
-                                                           & $\hat{s} = p$ \\  \cline{2-5}
-    \multirow{3}{*}{\rotatebox[origin=c]{90}{\textbf{Data Model}}} 
-& $s=r=2$ & 0.8 (0.7, 0.9) & 2.1 (1.7, 2.6) & 3.0 (2.9, 3.2) \\ \cline{2-5}
-   &   $s=r=2$, $\Sigma_k = \Sigma$ & 0.8 (0.7, 0.9) & 0.7 (0.6, 0.8) & 3.0 (2.9, 3.2)\\ \cline{2-5}
-   &  $s=p=200$ & 7.1 (6.2, 8.0) & 138.2 (119, 153) & 3.0 (2.9, 3.2) \\ \cline{2-5}
-  \end{tabular}
-  \caption[Table caption text]{Stein's risk (and 95\% loss intervals)
+  \caption[abcd]{Stein's risk (and 95\% loss intervals)
     for different inferential models and data generating models with
-    varying degrees of between-group covariance similarity.  For each
-    of $K=10$ groups, we simulate data from three different types of
-    shared subspace models.  For each of these models, $p=200$, $r=2$,
+    varying degrees of between-group covariance similarity.
+For each of $K=10$ groups, we simulate data from three different types of shared subspace models.
+For each of these models, $p=200$, $r=2$,
     $\sigma_k^2=1$ and $n_k=50$.  We also fit the data using three
@@
     risk (diagonal) but the adaptive estimator performs well relative
-    to the alternative misspecified model.}
-\label{table:groupLoss}
-\end{center}
+    to the alternative misspecified model. \label{table:groupLoss}}
+\centering
+ \begin{tabular}{ l  l | c | c | c |}
+    \multicolumn{2}{c}{} & \multicolumn{3}{c}{\textbf{Inferential Model}} \\
+  \multicolumn{2}{c|}{}  & Adaptive & $\hat{\Sigma}_k=\hat{\Sigma}$
+                                                           & $\hat{s} = p$ \\  \cline{2-5}
+    \multirow{3}{*}{\rotatebox[origin=c]{90}{\textbf{Data Model}}} 
+& $s=r=2$ & 0.8 (0.7, 0.9) & 2.1 (1.7, 2.6) & 3.0 (2.9, 3.2) \\ %\cline{2-5}
+   &   $s=r=2$, $\Sigma_k = \Sigma$ & 0.8 (0.7, 0.9) & 0.7 (0.6, 0.8) & 3.0 (2.9, 3.2)\\ %\cline{2-5}
+   &  $s=p=200$ & 7.1 (6.2, 8.0) & 138.2 (119, 153) & 3.0 (2.9, 3.2) \\ %\cline{2-5}
+  \end{tabular}
 \end{table}
@@
 
+\section{Reduction of Asymptotic Bias Via Pooling}
+\label{sec:asymp}
+Recently, there has been an interest in the asymptotic behavior of
+PCA-based covariance estimators in the setting in which
+$p, n \to \infty$ with $p/n=\alpha$ fixed.  Specifically, in the
+spiked covariance model it is known that when $p$ and $n$ are both
+large, the leading eigenvalues of the sample covariance matrix are
+positively biased and the empirical eigenvectors form a non-zero angle
+with the true eigenvectors \citep{Baik2006, Paul2007}.  Although this
+fact also implies that the shared subspace estimators are biased, a
+major advantage of shared subspace inference over independent
+estimation of multiple covariance matrices is that we reduce the
+asymptotic bias by pooling information across groups.  The bias
+reduction appears to be especially large when there is significant heterogeneity
+in the first $s$ eigenvectors of the projected covariance matrices.
+
+Throughout section we assume $K$ groups of data each with $n_k = n$
+observations per group and $s$ a fixed constant.  First, note that if
+$\hat{V}\hat{V}^T$ corresponds to the true shared subspace, then
+estimates $\hat{\psi}_k$ derived using the methods presented in
+Section \ref{sec:bayes} will consistently estimate $\psi_k$ as
+$n \to \infty$ regardless of whether $p$ increases as well because
+$Y_kV$ has a fixed number of columns.  For this reason, we focus
+explicitly on the accuracy of $\hat{V}\hat{V}^T$ (derived using the
+maximum marginal likelihood algorithm presented in Section
+\ref{sec:em}) as a function of the number of groups $K$ when both $p$
+and $n$ are of the same order of magnitude and much larger than $s$.
+As an accuracy metric, we study the behavior of
+$\tr(\hat{V}\hat{V}^TVV^T)/s$ which is bounded by zero and one
+and achieves a maximum of one if and only if $\hat{V}\hat{V}^T$
+corresponds to the true shared subspace.
+
+\begin{conjecture}
+\edits{Assume that the first $s$ eigenvalues from each of $K$ groups are
+  identical with $\lambda_i > \sigma^2(1 + \sqrt{\alpha})$.  Then,
+ for $p/n \to \alpha$ and $p, n
+  \to \infty$,  $\tr(\hat{V}\hat{V}^TVV^T)/s \to \xi$ with  
+\begin{equation}1 > \xi \geq  \frac{1}{s}\sum_{i=1}^s  \left(1-\frac{\alpha}{K(\lambda_i - 1)^2}\right) /\left(1 +
+    \frac{\alpha}{K(\lambda_i - 1)}\right).
+\label{eqn:asympBound}
+\end{equation}  }
+\end{conjecture}
+
+\edits{% Equality is achieved when the
+  % eigenvectors from each group are also identical (e.g. the data from each group is
+  % identically distributed).
+  We prove that the lower bound in \ref{eqn:asympBound} is
+in fact achieved when $Y_k$ are identically distributed and show in simulation
+that the subspace accuracy exceeds this bound when there is variation
+in the eigenvectors across groups. }
+In the case of iid groups, let the covariance matrix $\Sigma_k=\Sigma$ have the shared-subspace form given in
+Equation \ref{eqn:sspsi} and without loss of
+generality let $\psi_k = \psi$ be a diagonal matrix (e.g assume
+the columns of $V$ align with the eigenvectors of $\Sigma$).  In this
+case, the complete data likelihood of $V$ (Equation
+\ref{eqn:likV}) can be rewritten as
+%
+\begin{align*}
+\ell(V) &=\frac{1}{2}\sum_k \tr\left((\frac{1}{\sigma^2}I-M^{-1})V^T
+  S_kV\right)\\
+&=\frac{1}{2}\tr\left(DV^T(\sum_k S_k)V\right).
+\end{align*}
+%
+
+\noindent where $\sum_{k=1}^K S_k \sim \text{Wish}(\Sigma, Kn)$.
+Since $\psi$ is diagonal and $\sigma^2=1$, 
+$M = \sigma^2(\psi + I)$ is diagonal and thus
+$D = (\frac{1}{\sigma^2}I-M^{-1})$ is also diagonal with entries
+$0 < d_i < 1$ of decreasing magnitude.
+
+Then, the solution to
+$$\hat{V}^{(k)} = \underset{\widetilde{V} \in \mathcal{V}_{p,
+    s}}{\text{argmax }} \tr\left(D\widetilde{V}^T
+ \sum_k( S_k)\widetilde{V}\right).$$
+has $\hat{V}^{(k)}$ equal to the first $s$ eigenvectors of $\sum_k S_k$.  This is
+maximized when the columns of $V$ match the first empirical eigenvectors of
+$\sum_k S_k$ and has a maximum of $\sum_{i=1}^r d_i\ell_i$ where $\ell_i$ is the
+$i$th eigenvalue of $\sum_k S_k$.  Using a result from
+\citet{Paul2007}, it can be shown that as long as
+$\lambda_i > \sigma^2(1 + \sqrt{\alpha})$ where $\lambda_i$ is the $i$th
+eigenvalue of $\Sigma_k$, the asymptotic inner product between
+the $i$th sample eigenvector and the $i$th population eigenvector
+approaches a limit that is almost surely less than one
+$$|\langle\hat{V}_i, V_i\rangle| \to \sqrt{\left(1-\frac{\alpha}{K(\lambda_i - 1)^2}\right) /\left(1 +
+    \frac{\alpha}{K(\lambda_i - 1)}\right)} $$
+%
+% Note that in the likelihood specified in Equation \ref{eqn:likV} , the
+% sample covariance matrices, $S_k/n_k$, are sufficient statistics.
+% Since the sample covariance matrices are biased for the true
+% eigenvectors of $\Sigma_k$ when $p$ and $n$ increase at the same rate
+% then $\tr(\hat{V}\hat{V}^TVV^T)/s < 1$ almost surely \citep{}.
+
+\noindent As such, we can express asymptotic shared subspace
+accuracy for the identical groups model as
+\begin{align}
+\nonumber \tr(\hat{V}\hat{V}^{^T}VV^T)/s &= \frac{1}{s}\sum_{i=1}^s |\langle\hat{V}_i, V_i\rangle|^2\\
+&\to \frac{1}{s}\sum_{i=1}^s  \left(1-\frac{\alpha}{K(\lambda_i - 1)^2}\right) /\left(1 +
+    \frac{\alpha}{K(\lambda_i - 1)}\right).
+\label{eqn:lbound}
+\end{align}
+
+% Naturally, when the groups are identically distributed, $VV^T$ should
+% be estimated using the pooled data.  In this simple case, it is easy
+% to analytically express the improvement in subspace accuracy from
+% pooling data across groups.  The maximum likelihood estimate for
+% $VV^T$ for the pooled data can be expressed as
+% \begin{align*}
+% \hat{V} &= \underset{\widetilde{V} \in \mathcal{V}_{p,
+%     s}}{\text{argmax }} \tr\left(D\widetilde{V}^T
+%   (\sum_{k=1}^K S_k)\widetilde{V}\right)
+% \end{align*}
+% %
+
+% Thus, in this example we can express the accuracy of
+% $\hat{V}\hat{V}^T$ for the pooled estimator as:
+% %
+% \begin{align}
+% \nonumber \tr(\hat{V}\hat{V}^TVV^T)/s &= \sum_{i=1}^s |\langle\hat{V}_i, V_i\rangle|^2\\
+% &\approx\frac{1}{s}\sum_{i=1}^s  \left(1-\frac{\gamma}{K(\lambda_i - 1)^2}\right) /\left(1 +
+%     \frac{\gamma}{K(\lambda_i - 1)}\right).
+% \label{eqn:bound}
+% \end{align}
+
+Here, the accuracy of the estimate depends on $\alpha$, $K$ and the
+magnitude of the eigenvalues, with the bias naturally decreasing as the number
+of groups increases.  Most importantly, Equation
+\ref{eqn:lbound} provides a useful benchmark for understanding the
+bias of shared subspace estimates in the general setting in which
+$\psi_k$ varies across groups.  Our conjecture that the subspace
+accuracy is larger than the lower bound when the eigenvectors between
+groups are variable is consist with our simulation results.
+
+% This is supported by the results in Table
+% \ref{table:groupLoss} which show that when estimating groups of spiked
+% covariance matrices using the shared subspace estimator there is a
+% significant reduction in Stein's risk.
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.7\textwidth]{../Figs/asymptotics}
+    \caption{Subspace accuracy $\tr(\hat{V}\hat{V}^TVV^T)/s$ (solid)
+      and the asymptotics-based benchmark (dashed)
+      % $\frac{1}{s}\sum_{i=1}^s \left(1-\frac{\gamma}{K(\lambda_i -
+      %     1)^2}\right) /\left(1 + \frac{\gamma}{K(\lambda_i -
+      %     1)}\right)$
+      as a function of $K$.  When
+      $\lambda_1=\lambda_2$ (green), the assumptions used to derive
+      the benchmark (identically distributed groups) are met and thus the
+      subspace accuracy matches the benchmark.  However, when the
+      ratio $\lambda_1/\lambda_2$ is large, the subspace accuracy
+      metric can far exceed this benchmark if there is significant
+      variation in the eigenvectors of $\psi_k$ across groups.  Small
+      increases in accuracy over the benchmark are seen for moderately
+      anisotropic data (red) and large increases for highly
+      anisotropic data (blue).  }
+      \label{fig:asymptotics}
+\end{figure}
+
+In Figure \ref{fig:asymptotics} we depict the subspace accuracy metric
+$\tr(\hat{V}\hat{V}^TVV^T)/s$ and benchmark
+$\frac{1}{s}\sum_{i=1}^s \left(1-\frac{\alpha}{K(\lambda_i -
+    1)^2}\right) /\left(1 + \frac{\alpha}{K(\lambda_i - 1)}\right)$
+for simulated multi-group data generated under the shared subspace
+model with $s=2$, $n=50$, $p=200$ and three different sets of eigenvalues.  For each
+covariance matrix, the eigenvectors of $\psi_k$ were sampled uniformly
+from Stiefel manifold $\mathcal{V}_{2,2}$.  When $\psi_k$ is isotropic
+(green) the subspace similarity metric closely matches the benchmark
+since the assumptions used to derive this asymptotic result are met.
+However, when the eigenvectors of $\psi_k$ vary significantly across
+groups and $\lambda_1 \gg \lambda_2$, the subspace accuracy can far
+exceed this benchmark (blue).  Intuitively, when the first
+eigenvectors of two different groups are nearly orthogonal, each group
+provides a lot of information about orthogonal directions on $VV^T$
+and so the gains in accuracy exceed those that you would get by
+estimating the subspace from a single group with $K$ times the sample
+size.  In general the accuracy of shared subspace estimates depends on
+the variation in the eigenvectors of $\psi_k$ across groups as well as
+the magnitude of the eigenvalues and matrix dimensions $p$ and $n_k$.
+Although the shared subspace estimator improves on the accuracy of
+individually estimated covariance matrices, estimates can still be
+biased when $\alpha$ is very large or the eigenvalues of $\Sigma_k$
+are very small for all $k$.  In practice, one should estimate the
+approximate magnitude of the bias using the inferred eigenvalues of $\Sigma_k$.  When
+these inferred eigenvalues are significantly larger than
+$\hat{\sigma}_k^2(1+\sqrt{\alpha/K})$ the bias will likely be small.
+
+% \begin{align*}
+% \hat{V} &= \underset{\widetilde{V} \in \mathcal{V}_{p,
+%     s}}{\text{argmax }} \tr\left((I-M_k^{-1})^{1/2}\widetilde{V}^T
+%   (\sum_{k=1}^K S_k)\widetilde{V}(I-M_k^{-1})^{1/2}\right)
+% \end{align*}
+% Where $(I-M_k^{-1})^{1/2}$
+
+% \begin{align*}
+% Y &= ZV\psi_k^{1/2}V^T+E\\
+% Y^TY &= V\psi_k^{1/2}V^TZ^TZV\psi_k^{1/2}V^T + E^TE + \text{cross terms}
+% \end{align*}
+
+% \noindent where $Z$ and $E$ are $n$ by $p$ matrices of iid stanard
+% normals.
+
+% $$ \tr\left((I-M_k^{-1})\widetilde{V}^TVM_k^{1/2}V^TZ^TZVM_k^{1/2}V^T\widetilde{V}\right)$$
+%% \subsection{Consistency of $\Psi_k$ and $\sigma_k^2$ given $V$}
+
 \section{Analysis of Biological Data}
@@
     \centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/leukemiaRatio-07-07}
-        \caption{$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$}
+    \subfigure[$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$]{
         \label{fig:leukemiaRatio}
-    \end{subfigure}
+        \includegraphics[width=0.45\textwidth]{../Figs/leukemiaRatio-07-07}}
     ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
       %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/leukemiaPosterior-07-07}
-        \caption{95\% Posterior Regions }
+    \subfigure[95\% Posterior Regions]{
       \label{fig:leukemiaPosterior}
-      \end{subfigure}
+        \includegraphics[width=0.45\textwidth]{../Figs/leukemiaPosterior-07-07}}
       \caption{a) Goodness of shared subspace fit for each of the
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{Figs/leukemia-biplot-07-07}
+    \includegraphics[width=0.45\textwidth]{../Figs/leukemia-biplot-07-07}
     \qquad
-\raisebox{1.25\height}{
-\footnotesize
-\input{Figs/biplot.tab}
+\raisebox{1\height}{
+\scriptsize
+\input{../Figs/biplot.tab}
 }
@@
   quadrant they lie in.  Right) List of the gene's with the largest
-  loadings, grouped by quadrant. }
+  loadings, grouped by quadrant.}
 \label{fig:leukemiaBiplot}
 \end{figure}
-
-\paragraph{Analysis of Metabolomic Data:}
-
-Next, we briefly discuss an example in which we could not readily
-identify significant covariance differences between groups.  Here, we
-apply shared subspace group covariance estimation to a metabolomic
-data analysis on fly aging \citep{Hoffman2014}.  We bin the data to
-include groups of flies less than 10 days (young), between 10 and 51
-days (middle) and greater than 51 days (old), and further split by
-sex. The sample sizes range from 43 to 59 flies per group.  We analyze
-metabolomic data corresponding to metabolites with 3714 mass-charge
-ratios.  After removing mean effects due to age and sex, we fit the
-shared subspace estimator and compare covariances between groups.  We
-identify a subspace that explains almost all of the variability in the
-first few components of $\Sigma_k$ for all groups (Figure
-\ref{fig:dmelanRatio}) but see little evidence for differences in
-metabolite covariances on this subspace, as evidenced by the large
-overlap in posterior distributions for each group (Figure
-\ref{fig:dmelanPosterior}). This could be indicative of a large amount
-of variation that is common across all groups for the first $s$
-principal components. Differences in covariation across groups, if
-there are any, likely manifest themselves in the smaller principal
-components that are too small to detect given the sample sizes.
+\paragraph{Analysis of Metabolomic Data:} Next, we briefly discuss an
+example in which we could not readily identify significant covariance
+differences between groups.  Here, we apply shared subspace group
+covariance estimation to a metabolomic data analysis on fly aging
+\citep{Hoffman2014}.  We bin the data to include groups of flies less
+than 10 days (young), between 10 and 51 days (middle) and greater than
+51 days (old), and further split by sex. The sample sizes range from
+43 to 59 flies per group.  We analyze metabolomic data corresponding
+to metabolites with 3714 mass-charge ratios.  After removing mean
+effects due to age and sex, we fit the shared subspace estimator and
+compare covariances between groups.  We identify a subspace that
+explains almost all of the variability in the first few components of
+$\Sigma_k$ for all groups (Figure \ref{fig:dmelanRatio}) but see
+little evidence for differences in metabolite covariances on this
+subspace, as evidenced by the large overlap in posterior distributions
+for each group (Figure \ref{fig:dmelanPosterior}). This could be
+indicative of a large amount of variation that is common across all
+groups for the first $s$ principal components. Differences in
+covariation across groups, if there are any, likely manifest
+themselves in the smaller principal components that are too small to
+detect given the sample sizes.
 
@@
     \centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/dmelanRatio-07-07}
-        \caption{$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$}
+    \subfigure[$\gamma(Y_k: \hat{V},\hat{\sigma_k}^2)$]{
         \label{fig:dmelanRatio}
-      \end{subfigure}
+        \includegraphics[width=0.45\textwidth]{../Figs/dmelanRatio-07-07}}
     ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
       %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.45\textwidth}r
-        \includegraphics[width=\textwidth]{Figs/dmelanPosterior-07-07}
-        \caption{95\% Posterior Regions}
+    \subfigure[95\% Posterior Regions]{
       \label{fig:dmelanPosterior}
-    \end{subfigure}
+      \includegraphics[width=0.45\textwidth]{../Figs/dmelanPosterior-07-07}}
     \caption{a) Goodness of fit is close to one for all groups
@@
 
-% There are many different prior specifications for $\Psi_k$ and
-% $\sigma_k^2$ that we could consider for estimating posterior
-% uncertainty.  One simple approach would be to use the same prior
-% specification used in Section \ref{sec:em}.  That is, we can use
-% Jeffreys' prior for $\frac{1}{\sigma_k^2}$ and
-% $M_k^{-1} = (\Psi_k + I) ^{-1}/\sigma_k^2$.  In fact, it is easy to
-% show that the Gamma distribution for $\frac{1}{\sigma_k^2}$ and the
-% Wishart distribution for $M_k^{-1}$ are conjugate priors for the
-% respective quantities.  From posterior samples of $M_k$ and
-% $\sigma_k^2$ we can easily recover samples of $\Psi_k$.
-% Unfortunately, if $M_k^{-1}$ has a Wishart distribution, the implied
-% support of $\Psi_k$ is not constrained to the cone of positive
-% semi-definite matrices.  Direct prior specifications for $\Psi_k$ are
-% more appealing, but many obvious choices, like the inverse-Wishart
-% prior distribution for $\Psi_k$ do not lead to simple posterior
-% distributions.
-
 Fundamentally, our approach is quite general and can be integrated
@@
 
-\bibliographystyle{plainnat}
-\bibliography{refs.bib}
+%\pagebreak
+\bibliographystyle{chicago}
+\bibliography{refs}
 

### 2019-02-22 fit-leukemia-full.R
Full subspace fit

@@ -1,0 +1,474 @@
+rm(list=ls())
+
+library(car)
+library(rstiefel)
+library(xtable)
+library(made4)
+library(coda)
+library(colorspace)
+library(microbenchmark)
+library(cowplot)
+library(colorspace)
+library(ggdendro)
+library(tidyverse)
+
+source("subspace-functions.R")
+source("fit-subspace.R")
+source("leukemia_functions.R")
+
+load("LeukemiaData/leukemia.RData")
+
+## Look at PCA for pooled data to evaluate differences in means
+YpooledDemeaned <- sweep(Ypooled, 2, colMeans(Ypooled), '-')
+indices <- which(typeVec %in% c("TEL-AML1", "T-ALL", "MLL"))
+plotarrays(YpooledDemeaned[indices, ], axis1=8, axis2=9, classvec=typeVec[indices],
+           star=FALSE, arraycol=c("red", "blue", "dark green"))
+
+S <- 45
+## number of eigenvectors in group subspace
+R <- 45
+P <- ncol(Ypooled)
+
+residualList <- Slist <- list()
+for( k in 1:ngroups ) {
+    residual <- lm(Ylist[[k]] ~ 1)$residuals
+    residual <- t(lm(t(residual) ~ 1)$residuals)
+    residualList[[k]] <- residual
+    Slist[[k]] <- t(residual) %*% residual
+}
+pooledResiduals <- do.call(rbind, residualList)
+
+getRank(pooledResiduals)
+
+plot(svd(residualList[[ngroups]]/sqrt(nvec[ngroups]))$d^2, pch=19, cex=0.75, ylim=c(0, 50))
+for(k in 1:6) {
+    points(svd(residualList[[k]]/sqrt(nvec[k]))$d^2, col=k+1, pch=19, cex=0.75)
+}
+svd(residualList[[ngroups]]/sqrt(nvec[ngroups]))$d^2
+
+Vinit <- svd(do.call(cbind, lapply(1:ngroups, function(k) svd(t(residualList[[k]]))$u[, 1:min(S, nvec[k])])))$u[, 1:S]
+
+isoVar <- sapply(1:ngroups, function(i) median(apply(residualList[[i]], 2, var)))
+weightsList <- sapply(1:ngroups, function(i) {
+    evals <- svd(residualList[[i]]/sqrt(nvec[i]))$d^2
[…]

### 2019-02-22 paper/JRSSB/response/response.tex
Membership predictions

@@
 
-\section{Reviewer BLAH}
+\section*{Referee 1}
 
@@
 \begin{resp}
-We have fixedt his to correctly reflect the fact that $S_k = Y_k^TY_K$ as we have defined $Y_k$ and an $n \times p$ matrix.
+We have fixed this to correctly reflect the fact that $S_k = Y_k^TY_K$ as we have defined $Y_k$ and an $n \times p$ matrix.
 \end{resp}
@@
 In the draft, we excluded the square and thank the reviewer for highlighting this error.
-The following argument holds:
-
+Thus, $$||Y_kV||_F^2/n_k = \tr (V^TY_k^TY_k/n_kV)$$
+where $Y_k^TY_k/n_k \sim \text{Wish}(\frac{1}{n_k}\Sigma_k, n_k)$ and thus $V^TY_k^TY_k/n_kV \sim \text{Wish}(\frac{1}{n_k}V^T\Sigma_kV, n_k) \sim \text{Wish}(\frac{1}{n_k}(\Psi_k + \sigma_k^2I), n_k) $.
+Therefore, 
 \begin{align*}
 ||Y_kV||_F^2/n_k &= \tr (V^TY_k^TY_k/n_kV)\\
-&= \tr (V^T\hat{\Sigma}_kV)\\
 &\to tr(V^T\Sigma_kV)\\
@@
 \end{align*}
-
+for $n \to \infty$ and $s$ fixed (where the second line is by law of large numbers).
 \end{resp}
@@
 Similarly, we would still correctly identify the genes that are most variable in each leukemia group.
-\todo{We have clarified these points in the paper.}
+We have clarified these points in the paper in the section on inference for projected data covariance matrices.
 \end{resp}
@@
 \begin{resp}
+  The referee is correct that if there is a large difference in the noise variance $\sigma^2_k$ across groups then concatenating the data may not work well.
+In the problems presented in the paper, the noise variances were all very similar so this was not an issue.
+Nevertheless, we can address this issue by using Gavish and Donoho's estimator on the \emph{scaled} and pooled data, an $\sum n_k \times p$ matrix $Y_{\text{pool}} = [\frac{1}{\sigma_1}Y_1; \frac{1}{\sigma_2}Y_2;...;  \frac{1}{\sigma_k}Y_k]$.
+Since $\sigma^2_k$ is generally not known, we will typically rescale the data from each group using an estimate of $\sigma^2_k$.
+A robust estimator for the noise variance is described in \citet{Gavish2014} which is simply an appropriately scaled version of the median singular value of $Y_k$.
+Note that $Y_{\text{pool}}$ has covariance $V^T( \sum_k \frac{\pi_k}{\sigma_k^2} \Psi_k)V + I$ where $\pi_k = \frac{n_k}{\sum_i n_i}$.
+In other words, $Y_{\text{pool}} = X + Z$ where $V$ are the left eigenvectors of $X$ and $Z$ is the noise matrix with iid entries with mean zero and variance one.
+Gavish and Donoho's estimator assumes invariant white noise with finite fourth moments and so it can be applied appropriately to the scaled, pooled data.
+
+Standard model selection approaches can also be used to select the dimension of the shared subspace.
+This includes cross validation and information criteria like AIC and BIC.
+These will be computationally intensive since they require fitting the model multiple times for different values of $s$.
+We have revised this section to make note of these points.
+
 \end{resp}
@@
 \begin{resp}
-We believe that this section still has merit, although we have reframed the section to emphasize the most important points.
-The purpose of this section is not to point out that ``more groups mean more data'' nor to provide an analysis for the case of $iid$ groups.
-Rather, section is meant to identify a benchmark on the accuracy of the shared subspace estimator.
-In general, this is a very hard problem, as the subspace accuracy depends on both the eigenvalues and eigenvectors of $\psi_k$.
-When the eigenvalues from each group are equal, we conjecture that $\tr(\hat{V}\hat{V}^TVV^T)/\sqrt(S) \geq $ with equality if and only if the eigenvectors from each group are the same.
-When the eigenvectors are highly variable across groups, the accuracy of the shared subspac estimator increases dramatically.
-Our simulations support this conjecture.
+  We believe that this section has merit and have reframed the section to emphasize the most important points.
+The purpose of this section is not to point out that ``more groups mean more data'' and hence higher accuracy, nor to describe an analysis for the case of $iid$ groups.
+Rather, the section is meant to identify a benchmark on the accuracy of the shared subspace estimator.
+In general, this is a very hard problem, as the subspace accuracy depends on both the eigenvalues and eigenvectors of $\psi_k$ in a non-trivial way.
+In general we do not have access to an analytic expression of the MLE in terms of the eigenvectors and eigenvalues of $\Psi_k$.
+Thus, it is not obvious that we can apply RMT methods to identify exactly the asymptotic bias in the general case.
+However, when the eigenvalues from each group are equal, we conjecture that $\tr(\hat{V}\hat{V}^TVV^T)/\sqrt{s} \geq \frac{1}{s}\sum_{i=1}^s  \left(1-\frac{\alpha}{K(\lambda_i - 1)^2}\right) /\left(1 +
+    \frac{\alpha}{K(\lambda_i - 1)}\right)$.
+We prove that the equality is achieved when the eigenvectors from each group are identical and then show in simulation that the eigenvectors are highly variable across groups, the accuracy of the shared subspace estimator increases dramatically.
 
-Finally, as noted by \citet{} estimating the accuracy of the sample eigenvectors in the spiked covariance model is a hard problem that is not entirely well studied.
+% Finally, as noted by \citet{} estimating the accuracy of the sample eigenvectors in the spiked covariance model is a hard problem that is not entirely well studied.
 In our case, although the ML solution is clearly related to the eigenvectors, we do not have access to a simple function relating the eigenvectors of each sample covariance matrix to the ML solution, and thus it is not obvious how we can use results from RMT.
@@
 
-\begin{resp}
-We thank the reviwer for the interesting suggestion.
-The suggested comparisons are closely related to the results presented in row 2 of table 1, comparing the performance of different models when the groups are in fact iid.
-When the groups are identical, , .
-Most importantly, the reviewer highlights the importance of identifying a simple method for initializing the EM algorithm (see below).
-\end{resp}
-
 \textbf{7.
@@
 \begin{resp}
+We thank the referee for the interesting suggestion.
+% The suggested comparisons are closely related to the results presented in row 2 of table 1, comparing the performance of different models when the groups are in fact iid.
+Most importantly, the reviewer highlights the importance of identifying a simple method for initializing the EM algorithm.
+Empirical explorations show that the EM algorithm tends to converge to suboptimal local-minima when initialized randomly.
+We show in simulation that when we use the initial condition described above, the EM algorithm converges to a mode which is much closer to the truth (See an example in Fig 1 of this response).
 
-We are in fact using an initialization chosen in a very similar, but did not include the details in the first draft of the paper.
+We are in fact using an initialization chosen in a way that is very similar to the one proposed by this referee, but did not include the details in the first draft of the paper.
 We now have a section in the paper on convergence of the EM algorithm and the importance of a reasonable initialization.
+Specifically, we initialize our algorithm at $UU^T$ where $U$ are the eigenvectors of $Y_{\text{pool}}^T Y_{\text{pool}}$ and $Y_{\text{pool}}$ is the scaled and pooled data described in response to point $4$.
+The initialization is a robust estimator, even when the data is not normally distributed, but the estimate can be improved when using the likelihood-based estimator for normally distributed data.
+The initialization strategy and issues with convergence are described in more detail in a new subsection on the convergence of the EM algorithm.
 
@@
         \label{fig:modeHist}
-    \caption{ Subspace similarity ($\tr(\hat{V}\hat{V}^TVV^T)/s$)
-      between EM estimates and the true shared subspace for 1000 random initializations (grey) and ``smart'' intialization (red line).  }
+    \caption{ An example showing subspace similarity ($\tr(\hat{V}\hat{V}^TVV^T)/s$)
+      between EM estimates and the true shared subspace for 1000 random initializations (grey) and the initialization based on the eigenvectors of the scaled and pooled data (red line).  }
 \end{figure}
@@
 
-\section{}
+\section*{Referee 2}
 
-The heterogeneity is imposed via $\Psi_k$’s and information is shared via V (of suitable dimensions).
+% The heterogeneity is imposed via $\Psi_k$’s and information is shared via V (of suitable dimensions).
 The modeling framework is interesting, but it does not come to be as a “surprise” in any way or form.
@@
 
-\begin{resp}
-\end{resp}
+% \begin{resp}
+% \end{resp}
 
-The estimation of the parameters carried out via EM comes across to me as standard.
+\textbf{The estimation of the parameters carried out via EM comes across to me as standard.
 The sampling procedures, as well as the choice of priors seem to be (directly) motivated via earlier work, some of them done by a subset of the authors.
@@
 An important theme of current research at the interface of statistical computing and optimization is to understand formal guarantees on the EM procedure.
-Some studies along these lines may greatly benefit the paper.
+Some studies along these lines may greatly benefit the paper.}
 
 \begin{resp}
-\todo{}
+
+We have added a section  on convergence of the EM algorithm and intelligent choices for starting values.  The Stiefel manifold is compact and the marginal
+  likelihood is continuous, so the likelihood is bounded.  Thus, the EM
+  algorithm, which increases the likelihood at each iteration, will converge to a stationary point.
+Indeed the objective function is highly non-convex and thus intelligently initializing the EM algorithm is important.
+See the figure in response to point 7 by referee 1.
+% Unlike the EM algorithm applied to problems with missing data which grows in $n$ (e.g.
+mixture models) the population likelihood of $V$ converges to \todo{}
 \end{resp}
 
-The analysis in Section 5 is interesting but is not complete (agreed, some simulation results
+\textbf{The analysis in Section 5 is interesting but is not complete (agreed, some simulation results
 are provided) for the joint estimation problem. Can the gains be quantified more formally
 by appealing to RMT results — by accounting for the fact that several parameters are being
-jointly estimated?
+jointly estimated?}
 
 \begin{resp}
+We have reorganized the results in Section 5 to make the claims more clear.
+Specifically we provide a conjecture on the lower bound of the subspace accuracy, prove that bound is achieved when the data is identically distributed and show in simulation the bound is exceeded.
+Unfortunately, it is not clear how RMT results can be used to make a more precise claim about the accuracy of the shared subspace estimator in the most general setting.
+Existing RMT results describe the behavior of eigenvectors or eigenprojection matrix for small $n$, large $p$.
+In our problem, we have no analytic expression relating the MLE for the shared subspace estimator to the eigenvectors and eigenvalues.
 \end{resp}
 
-The Gavish Donoho threshold applies for a specific class of models (way more special than
+\textbf{The Gavish Donoho threshold applies for a specific class of models (way more special than
 the classes studied in this paper) – I am not sure if their threshold can be justified rigorously
-in the context of the models being studied in this paper.
+in the context of the models being studied in this paper.}
 
 \begin{resp}
-The Gavish-Donoho estimator applies to the class of spiked covariance models (single group).
-Since each covariance matrix is a spiked covariance matrix, in fact the Gavish-Donoho threshold applies to estimating the rank from each group individually.
-The challenge generalizing the rank estimator to the shared subspace setting.
-\todo{}
+In fact the Gavish-Donoho estimator applies to the \emph{broader} class of models than that studied here because they do not assume normality.
+Specifically, they focus on recovering a low rank matrix $X$ from a matrix $Y = X + \sigma Z$ where Z has iid mean zero entries with finite fourth moment.
+This setting applies to estimating the rank of each covariance matrix independently (since they are spiked covariance matrices).
+We can also apply the Gavish-Donoho applies to the scaled and pooled data as described in the paper and response to point 6 and 7 by referee 1.
 \end{resp}
 
-\section{editor}
-
-In addition, the following points are based on my own non-expert reading of your work. If you decide to revise and resubmit, I would also like you to clarify these in the next version of the paper.
-
-1. I think the problem is rather specific – so I wonder if your work is of generic enough interest for the mainstream of Series B readers.
+\section*{Associate Editor}
+\textbf{Two referees have considered your paper.
+They agree that your paper proposes an interesting approach to the problem of estimating the covariance matrix for different groups of observations in high-dimensions.
+The paper has some nice features.
+The proposed model can be estimated using the EM algorithm and an MCMC algorithm is proposed for posterior inference.
+In addition, some properties are investigated.
+However, both referees have concerns about convergence problems in the EM algorithm and there is little evidence of good mixing in the MCMC algorithm.} % The referees also questions whether the paper meets the level of novelty associated with JRSS B, the depth of the analysis in Section 5 (which is the main theoretical part) and the illustrations of the results.
+I agree with their concerns and I feel that the paper is currently not acceptable for JRSS B.
 
 \begin{resp}
+We have added a section to the paper discussing the convergence of the EM algorithm (see response to reviewers above).
+In addition, we provide trace plots and effective sample sizes for the log eigenvalue ratio and eigenvector angle plotted in Figure's 2, 5, and 7.
 \end{resp}
 
-2. What happens if the spiked covariance model holds only approximately?
+\begin{figure}[t]
+    \centering
+    \subfigure[Simulation (Figure 2).]{
+      \label{fig:sdimension}
+      \includegraphics[width=0.5\textwidth]{../../Figs/simulation-cov-trace.pdf}}\\
+    \subfigure[Leukemia data (Figure 5).]{
+      \label{fig:ratio-s5}
+      \includegraphics[width=0.5\textwidth]{../../Figs/leukemia-cov-trace.pdf}}\\
+    \subfigure[Metabolomics data (Figure 7)]{
+        \label{fig:ratio-s20}
+        \includegraphics[width=0.5\textwidth]{../../Figs/dmelan-cov-trace.pdf}}\\
+      \caption{Trace plot for log eigenvalue ratios and the first eigenvector angle, for all plots presented in the paper.
+Effective sample sizes are all large: for 1000 samples, it is over 500 for all parameters and nearly 1000 for most parameters, which reflects the fact that our samples are nearly independent.}
+\label{fig:dimensionPlots}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \subfigure[Simulation study (Figure 2).]{
+      \label{fig:sdimension}
+      \includegraphics[width=0.3\textwidth]{../../Figs/simulation-noisevar-trace.pdf}}
+    \subfigure[Leukemia data (Figure 5).]{
+      \label{fig:ratio-s5}
+      \includegraphics[width=0.3\textwidth]{../../Figs/leukemia-noisevar-trace.pdf}}
+    \subfigure[Metabolomics data (Figure 7).]{
+        \label{fig:ratio-s20}
+        \includegraphics[width=0.3\textwidth]{../../Figs/dmelan-noisevar-trace.pdf}}
+      \caption{Trace plot for $sigma^2_k$, for all datasets analyzed in the paper. Effective sample sizes are all close 1000, which reflects the fact that our samples are nearly independent.}
+\label{fig:dimensionPlots}
+\end{figure}
+
+\section*{Editor}
+
+\textbf{1. I think the problem is rather specific – so I wonder if your work is of generic enough interest for the mainstream of Series B readers.}
 
 \begin{resp}
+We believe that the problem is of sufficient generic interest for mainstream readers.
+First, understanding covariability of high dimensional data is of interest in several application areas.
+For example, in applications in the biological sciences, like those presented in the paper, the sample size is often limited by the number of available subjects and/or the expense of collecting samples, but the number of possible features for each subject is large (e.g.
+gene or protein expression for thousands of genes).
+In most of these problems the data can be naturally divided into multiple distinct groups (e.g.
+cases and controls).
+Further, it is clear that understanding second-moment information is essentially for understanding the mechanisms underlying the processes studied.
+
+From a methodological standpoint, shared subspace covariance estimation is closely related to many popular and widely studied statistical models.
+In particular, the ideas presented in our paper are relevant to sufficient dimension reduction, factor based methods and partial least squares as well as other methods for high dimensional covariance estimation (e.g.
+graphical lasso).
+Our paper examines the utility of exploring differences between groups of data on a common low dimensional space.
+This general idea may dovetail well with other common procedures for inferring patterns in high dimensional data.
+
 \end{resp}
 
-3.
-In my first decision letter, I asked for the theoretical description of the properties of your method,but I am not sure if this has been done successfully.
-In particular, I cannot see theorem-type results quantifying the behaviour of your procedure.
+\textbf{2. What happens if the spiked covariance model holds only approximately?}
 
 \begin{resp}
-\todo{BLAH}
+There are, broadly speaking, two ways in which the spiked covariance model may not hold.
+First, multivariate normality may hold, but the covariance matrix may not fit the ``low-rank + diagonal'' form.
+In other words, the $p \times p$ covariance matrix may have $p$ distinct eigenvalues (rather than $s + 1$).
+Even when the spiked covariance model does not hold, it is still a \emph{useful} model when $n < p$ setting because only a few eigenvectors and eigenvalues are accurately estimable.
+In particular, the distribution of $s+1st$ through $pth$ eigenvectors will be nearly indistinguishable from the distribution of eigenvectors associated with a single common eigenvalue of multiplicity (p-s) when these eigenvalues are very small relative to p/n.
+As is evident from the RMT results on eigenvector and eigenvalue estimation, as $n$ increase more eigenvalues/vectors are accurately estimable.
+The number of spikes is not to meant to reflect the number of \emph{true} unique eigenvalues, only the number of eigenvalues associated with estimable eigenvectors.
+
+Second, if the true distribution is very non-normal then the shared subspace model may not be appropriate, especially if the contours of the true data generating distribution are non-elliptical.
+However, we can still estimate a shared subspace using the eigenvectors of the scaled and pooled data.
+Under relatively weak conditions on the moments of the data generating distribution, we can use these eigenvectors to define a  $\sqrt{n}$-consistent estimator the space spanned by the first $s$ eigenvectors from each group (see \citep{Mestre2008}).
 \end{resp}
 
+\textbf{3.
+In my first decision letter, I asked for the theoretical description of the properties of your method, but I am not sure if this has been done successfully.
+In particular, I cannot see theorem-type results quantifying the behaviour of your procedure.}
+
+\begin{resp}
+We believe that after these revisions that the behavior of our method is well described.  In particular the paper now includes:
+\begin{itemize}
+\item Discussion of the convergence on the EM algorithm.  We describe $\sqrt{n}$-consistent initial values for the shared subspace estimator which improve convergence to more optimal modes.
+\item A theorem about the behavior of the goodness of fit estimator when $n << p$.  The goodness of fit test can be used to identify whether the inferred subspace is indeed reasonable.  
+\item A conjecture about the accuracy of the shared subspace estimator in the small n, large p setting.
+We provide a lower bound on the asymptotic bias of our estimator, prove that the lower bound holds exactly when the data from each groups are identically distributed and show in simulation that we exceed this bound.
+\end{itemize}
+\end{resp}
+
+\bibliographystyle{chicago}
+\bibliography{../refs}
+
 \end{document}

### 2019-02-22 paper/JRSSB/mscov.tex
Membership predictions

@@
 
+\ifdefined\revision
 \newcommand{\edits}[1]{{\textsf{\textcolor{red}{#1}}}}
+\else
+\newcommand{\edits}[1]{#1}
+\fi
 
@@
 \begin{equation}
-\Sigma_k = V\Psi_kV^T + \sigma^2_kI,
+\Sigma_k = \sigma^2_k(V\Psi_kV^T + I),
 \label{eqn:sspsi}
@@
 Note that for a shared subspace model,
-$V^T\Sigma_kV = \Psi_k + \sigma_k^2I$ is an anisotropic
+$V^T\Sigma_kV = \sigma_k^2(\Psi_k + I)$ is an anisotropic
 $s$-dimensional covariance matrix for the projected data, $Y_kV$.  In
@@
   \frac{1}{\sigma_2}Y_2;...;  \frac{1}{\sigma_k}Y_k]$, where
-  $Y_{\text{pool}}$ has dimension $\sum_k n_k \times p$.  
+  $Y_{\text{pool}}$ has dimension $(\sum_k n_k) \times p$.  
   We use $\hat{U}_{\text{pool}}\hat{U}_{\text{pool}}^T$  as the
-  initial value for subspace estimation algorigthm where 
+  initial value for subspace estimation algorithm where 
   $\hat{U}_{\text{pool}}$ are the first $s$ eigenvectors of
-  $S_{\text{pool}} = Y_{\text{pool}}^TY_{\text{pool}}$ to
+  $S_{\text{pool}} = Y_{\text{pool}}^TY_{\text{pool}}$.
   If we
@@
   consistent when both $n$ and $p$ growing at the same rate.}
-\edits{For any $p$-vector $\eta$, the asymptotic bias of
-  $\eta^T\hat{U}_{\text{pool}}\hat{U}_{\text{pool}}^T\eta$ is well characterized as a function of the eigenvalues of $\Sigma_Z$\citep{Mestre2008}.  If either the eigenvalues of $\sum_k
+\edits{For an arbitrary $p$-vector $\eta$, the asymptotic bias of
+  $\eta^T\hat{U}_{\text{pool}}\hat{U}_{\text{pool}}^T\eta$ is well
+  characterized as a function of the eigenvalues of $\Sigma_Z$ \citep{Mestre2008}.  If either the eigenvalues of $\sum_k
   \frac{\pi_k}{\sigma_k^2} \Psi_k$ or the total sample size $\sum_k n_k$
@@
   accurately estimate the shared subspace and likelihood based
-  optimization my have little added benefit.  When either the eigenvalues
-  or sample size are small to moderate in magnitude,
-  $\hat{U}_{\text{pool}}\hat{U}_{\text{pool}}^T$ is a useful starting value for the EM algorithm.}
+  optimization may not be necessary.  However, when either the eigenvalues are
+  small or the sample size is small the likelihood based analysis can
+  improve inference and $\hat{U}_{\text{pool}}\hat{U}_{\text{pool}}^T$ is a useful starting value for the EM algorithm.}
   % An asymptotically lower variance estimator can then be constructed
@@
 %     \caption{ Subspace similarity ($\tr(\hat{V}\hat{V}^TVV^T)/s$)
-%       between EM estimates and the true shared subspace for 1000 random initializations (grey) and ``smart'' intialization (red line).  }
+%       between EM estimates and the true shared subspace for 1000 random initializations (grey) and ``smart'' initialization (red line).  }
 % \end{figure}
@@
 As such, when the shared subspace model holds both the
-numerator and denomiator of the goodness of fit statistic converge
+numerator and denominator of the goodness of fit statistic converge
 almost surely to $\sum_{i=1}^{s}
@@
 is an $r \times r$ diagonal matrix of eigenvalues, and
-$O_k \in \mathcal{V}_{s,r}$ is the matrix of eigenvectors of $\Psi_k$.
+$O_k \in \mathcal{V}_{s,r}$ is the matrix of eigenvectors of
+$\Psi_k$.  
 For any individual group, this corresponds to the original spiked PCA
 model (Equation \ref{eqn:spiked}) with
-$U_k = VO_k \in \mathcal{V}_{p, r}$.  Differentiating the ranks $r$
+$U_k = VO_k \in \mathcal{V}_{p, r}$.  \edits{Note that the $V$ and $O_k$
+  are jointly unidentifiable because for any $s \times s$
+  orthonormal matrix $W, VO = VW^TWO = \tilde{V}\tilde{O}$. Once we fix a basis
+for the shared subspace, $O_k$ is identifiable.  As such, $O_k$
+should only be interpreted relative to the basis $V$, as determined by
+the EM algorithm described in Section \ref{sec:em}.}  Differentiating the ranks $r$
 and $s$ is helpful because it enables us to independently specify a
 subspace common to all groups and the possibly lower rank features on
-this space that are specific to individual groups.  Although our model
+this space that are specific to individual groups.
+
+Although our model
 is most useful when the covariance matrices are related across groups,
@@
 
+%\todo{Paragraph on implications if spiked covariance model does not hold?}
+
 \paragraph{Model Comparison and Rank Estimation:}
@@
 are fixed and known.  In practice this is not the case.  Prior to
-fitting a model we should estimate these quantities.
-\citet{Gavish2014} provide an asymptotically optimal (in mean squared
-error) singular value threshold for low rank matrix recovery with
-noisy data.  We use their rank estimator, which is a function of the
-median singular value of the data matrix and the ratio
-$\alpha_k =\frac{p}{n_k}$, to estimate $r$.  Although their estimator
-was derived for individual covariance estimation, we found that Gavish
-and Donoho's estimator can also be applied to the pooled data to
-estimate $s$.  Specifically, we concatenate the data from all groups
-to create an $(\sum_k n_k) \times p$ dimensional matrix and apply
-their rank estimator to this matrix to choose $s$.
-
-Using these rank estimators, we conduct a simulation which
+fitting a model we should estimate these quantities.  \edits{Standard
+  model selection methods can be applied to select the both $s$
+  and $r$.  Common approaches include cross validation and information
+  criteria like AIC and BIC.  However, these approaches are
+  computationally intensive since they require fitting the model for
+  each value of $s$ and $r$.}  Here, we estimate the model dimensions
+by applying an asymptotically optimal (in mean squared error) singular
+value threshold for low rank matrix recovery with noisy data
+\citep{Gavish2014}.  This rank estimator is a function of the median
+singular value of the data matrix and the ratio $\alpha_k = p/n_k$.
+\edits{Note that under the shared subspace model, the scaled and
+  pooled data described in section \ref{sec:em} can be expressed as
+  $Y_{\text{pooled}} = X + Z$ where $V$ are the left singular values
+  of $X$ and $Z$ is a noise matrix with zero mean and variance one.
+  This is the setting in which \citet{Gavish2014} develop a rank
+  estimation algorithm, and so it can be appropriately applied
+  to $Y_{\text{pooled}}$ to estimate $s$.}
+
+Using this rank estimation approach, we conduct a simulation which
 demonstrates the relative performance of shared subspace group

### 2019-02-22 paper/arxiv/mscov.tex
Membership predictions

@@
 \usepackage{chemarrow}
-\usepackage{subcaption}
+\usepackage{subfigure}
 \usepackage{amsmath}
@@
     (\href{mailto:amfranks@uw.edu}{amfranks@uw.edu}).  Peter D. Hoff
-    is a Professor in the Departments of Statistics and Biostatistics
-    at the University of Washington
-    (\href{mailto:pdhoff@uw.edu}{pdfhoff@uw.edu}).
+    is a Professor in the Department of Statistical Science at Duke University
+    (\href{mailto:peter.hoff@duke.edu}{peter.hoff@duke.edu}).
 %
@@
  Data-Intensive Discovery, the Moore/Sloan Data Science Environments
- Project at the University of Washington and NSF grant DMS-1505136.
+ Project at the University of Washington, NSF grant DMS-1505136, and NIH grant R03-CA211160-01.
 %
@@
 \maketitle
+
 \begin{abstract}
-
-% From Hoff Although the covariance matrices corresponding to different
-% populations are unlikely to be exactly equal they can still exhibit a
-% high degree of similarity. For example, some pairs of variables may be
-% positively correlated across most groups, whereas the correlation
-% between other pairs may be consistently negative. In such cases much
-% of the similarity across covariance matrices can be described by
-% similarities in their principal axes, which are the axes that are
-% defined by the eigenvectors of the covariance matrices. Estimating the
-% degree of across-popu- lation eigenvector heterogeneity can be helpful
-% for a variety of estimation tasks.
 
@@
     \centering
-    \begin{subfigure}[b]{0.25\textwidth}
-        \includegraphics[width=\textwidth]{Figs/3dplot}
-        \caption{Projection in $\mathbb{R}^3$}
+    \subfigure[Projection in $\mathbb{R}^3$]{
     \label{fig:3dplot}
-    \end{subfigure}
+    \includegraphics[width=0.25\textwidth]{Figs/3dplot}}
 \quad
- %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-      %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.25\textwidth}
-        \includegraphics[width=\textwidth]{Figs/2d-scatter}
-        \caption{$Y_kV$}
+    \subfigure[$Y_kV$]{
     \label{fig:2dscatter}
-    \end{subfigure}
+    \includegraphics[width=0.25\textwidth]{Figs/2d-scatter}}
 \quad 
-    \begin{subfigure}[b]{0.25\textwidth}
-        \includegraphics[width=\textwidth]{Figs/2d-scatter-orth}
-        \caption{$Y_kV_{\perp}$}
+\subfigure[$Y_kV_{\perp}$]{
 \label{fig:2dscatterorth}
-    \end{subfigure}
+\includegraphics[width=0.25\textwidth]{Figs/2d-scatter-orth}}
 \caption{ Two groups of four-dimensional data (red and blue)
@@
     \centering
-%    \begin{subfigure}[b]{0.45\textwidth}
     \includegraphics[width=0.5\textwidth]{Figs/posteriorRegions-chull}
-        \caption{95\% posterior regions}
         \label{fig:simPosterior}
-%    \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-      %(or a blank line to force the subfigure onto a new line)
-    % \begin{subfigure}[b]{0.4\textwidth}
-    %     \includegraphics[width=\textwidth]{Figs/simRatio-05-19}
-    %     \caption{Ratio}
-    %     \label{fig:simRatio}
-    % \end{subfigure}
     \caption{a) 95\% posterior regions for the log of the ratio of
@@
     \centering
-    \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=\textwidth]{Figs/LossVsDimension}
-        \caption{Stein's risk vs $\hat{s}$}
+    \subfigure[Stein's risk vs $\hat{s}$]{
       \label{fig:sdimension}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=\textwidth]{Figs/simRatio-s5}
-        \caption{$\hat{s}$ = 5}
+      \includegraphics[width=0.3\textwidth]{Figs/LossVsDimension}}
+    \subfigure[$\hat{s}$ = 5]{
       \label{fig:ratio-s5}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=\textwidth]{Figs/simRatio-s20}
-        \caption{$\hat{s}$ = 20}
+      \includegraphics[width=0.3\textwidth]{Figs/simRatio-s5}}
+    \subfigure[$\hat{s}$ = 20]{
         \label{fig:ratio-s20}
-    \end{subfigure}
+        \includegraphics[width=0.3\textwidth]{Figs/simRatio-s20}}
       \caption{a) Stein's risk as a function of the
@@
 
+\section{Reduction of Asymptotic Bias Via Pooling}
+\label{sec:asymp}
+Recently, there has been an interest in the asymptotic behavior of
+PCA-based covariance estimators in the setting in which
+$p, n \to \infty$ with $p/n=\gamma$ fixed.  Specifically, in the
+spiked covariance model it is known that when $p$ and $n$ are both
+large, the leading eigenvalues of the sample covariance matrix are
+positively biased and the empirical eigenvectors form a non-zero angle
+with the true eigenvectors \citep{Baik2006, Paul2007}.  Although this
+fact also implies that the shared subspace estimators are biased, a
+major advantage of shared subspace inference over independent
+estimation of multiple covariance matrices is that we reduce the
+asymptotic bias by pooling information across groups.  The bias
+reduction is especially large when there is significant heterogeneity
+in the first $s$ eigenvectors of the projected covariance matrices.
+
+Throughout section we assume $K$ groups of data each with $n_k = n$
+observations per group and $s$ a fixed constant.  First, note that if
+$\hat{V}\hat{V}^T$ corresponds to the true shared subspace, then
+estimates $\hat{\psi}_k$ derived using the methods presented in
+Section \ref{sec:bayes} will consistently estimate $\psi_k$ as
+$n \to \infty$ regardless of whether $p$ increases as well because
+$Y_kV$ has a fixed number of columns.  For this reason, we focus
+explicitly on the accuracy of $\hat{V}\hat{V}^T$ (derived using the
+maximum marginal likelihood algorithm presented in Section
+\ref{sec:em}) as a function of the number of groups $K$ when both $p$
+and $n$ are of the same order of magnitude and much larger than $s$.
+As an accuracy metric, we study the behavior of
+$\tr(\hat{V}\hat{V}^TVV^T)/s$.  This metric is bounded by zero and one
+and achieves a maximum of one if and only if $\hat{V}\hat{V}^T$
+corresponds to the true shared subspace.
+
+To start, we consider the simple case in which the data matrices $Y_k$
+are identically distributed with covariance $\Sigma_k=\Sigma$ of the
+shared-subspace form given in Equation \ref{eqn:sspsi} and
+$\sigma_k^2 = 1$.  Without loss of generality we can let
+$\psi_k = \psi$ be a diagonal matrix (e.g assume the columns of $V$
+align with the eigenvectors of $\Sigma$).  In this case, for a single
+group the complete data likelihood of $V$ (Equation \ref{eqn:likV})
+can be rewritten as
+%
+\begin{align*}
+\ell(V) &=\frac{1}{2}\tr\left((\frac{1}{\sigma^2}I-M^{-1})V^T
+  S_kV\right)\\
+&=\frac{1}{2}\tr\left(DV^TS_kV\right).
+\end{align*}
+%
+Since $\psi$ is diagonal and $\sigma^2=1$, 
+$M = \sigma^2(\psi + I)$ is diagonal and thus
+$D = (\frac{1}{\sigma^2}I-M^{-1})$ is also diagonal with entries
+$0 < d_i < 1$ of decreasing magnitude.  Then, the solution to
+$$\hat{V}^{(k)} = \underset{\widetilde{V} \in \mathcal{V}_{p,
+    s}}{\text{argmax }} \tr\left(D\widetilde{V}^T
+  S_k\widetilde{V}\right).$$
+has $\hat{V}^{(k)}$ equal to the first $s$ eigenvectors of $S_k$.  This is
+maximized when the columns of $V$ match the first empirical eigenvectors of
+$S_k$ and has a maximum of $\sum_{i=1}^r d_i\ell_i$ where $\ell_i$ is the
+$i$th eigenvalue of $S_k$.  As shown in \citet{Paul2007}, as long as
+$\lambda_i > 1 + \sqrt{\gamma}$ where $\lambda_i$ is the $i$th
+eigenvalue of $\Sigma_k$, the asymptotic inner product between
+the $i$th sample eigenvector and the $i$th population eigenvector
+approaches a limit that is almost surely less than one
+$$|\langle\hat{V}^{(k)}_i, V_i\rangle| \to \sqrt{\left(1-\frac{\gamma}{(\lambda_i - 1)^2}\right) /\left(1 +
+    \frac{\gamma}{(\lambda_i - 1)}\right)} $$
+%
+% Note that in the likelihood specified in Equation \ref{eqn:likV} , the
+% sample covariance matrices, $S_k/n_k$, are sufficient statistics.
+% Since the sample covariance matrices are biased for the true
+% eigenvectors of $\Sigma_k$ when $p$ and $n$ increase at the same rate
+% then $\tr(\hat{V}\hat{V}^TVV^T)/s < 1$ almost surely \citep{}.
+
+\noindent As such, for each independent estimate
+$\hat{V}^{(k)}\hat{V}^{(k)^T}$, we can approximate the shared subspace
+accuracy as
+\begin{align}
+\nonumber \tr(\hat{V}^{(k)}\hat{V}^{(k)^T}VV^T)/s &= \frac{1}{s}\sum_{i=1}^s |\langle\hat{V}^{(k)}_i, V_i\rangle|^2\\
+&\approx \frac{1}{s}\sum_{i=1}^s  \left(1-\frac{\gamma}{(\lambda_i - 1)^2}\right) /\left(1 +
+    \frac{\gamma}{(\lambda_i - 1)}\right).
+\label{eqn:lbound}
+\end{align}
+
+Naturally, when the groups are identically distributed, $VV^T$ should
+be estimated using the pooled data.  In this simple case, it is easy
+to analytically express the improvement in subspace accuracy from
+pooling data across groups.  The maximum likelihood estimate for
+$VV^T$ for the pooled data can be expressed as
+\begin{align*}
+\hat{V} &= \underset{\widetilde{V} \in \mathcal{V}_{p,
+    s}}{\text{argmax }} \tr\left(D\widetilde{V}^T
+  (\sum_{k=1}^K S_k)\widetilde{V}\right)
+\end{align*}
+%
+\noindent where $\sum_{k=1}^K S_k \sim \text{Wish}(\Sigma, Kn)$.
+Thus, in this example we can express the accuracy of
+$\hat{V}\hat{V}^T$ for the pooled estimator as:
+%
+\begin{align}
+\nonumber \tr(\hat{V}\hat{V}^TVV^T)/s &= \sum_{i=1}^s |\langle\hat{V}_i, V_i\rangle|^2\\
+&\approx\frac{1}{s}\sum_{i=1}^s  \left(1-\frac{\gamma}{K(\lambda_i - 1)^2}\right) /\left(1 +
+    \frac{\gamma}{K(\lambda_i - 1)}\right).
+\label{eqn:bound}
+\end{align}
+
+Here, the accuracy of the estimate depends on $\gamma$, $K$ and the
+magnitude of the eigenvalues, with the bias decreasing as the number
+of groups $K$ increases.  Empirically, in the fully general setting in
+which $\psi_k$ varies across groups we get bias reductions that scale
+similarly in $K$.  This is supported by the results in Table
+\ref{table:groupLoss} which show that when estimating groups of spiked
+covariance matrices using the shared subspace estimator there is a
+significant reduction in Stein's risk. Moreover, Equation
+\ref{eqn:bound} provides a useful benchmark for understanding the
+bias of shared subspace estimates.
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.7\textwidth]{Figs/asymptotics}
+    \caption{Subspace accuracy $\tr(\hat{V}\hat{V}^TVV^T)/s$ (solid)
+      and the asymptotics-based benchmark (dashed)
+      % $\frac{1}{s}\sum_{i=1}^s \left(1-\frac{\gamma}{K(\lambda_i -
+      %     1)^2}\right) /\left(1 + \frac{\gamma}{K(\lambda_i -
+      %     1)}\right)$
+      as a function of $K$.  When
+      $\lambda_1=\lambda_2$ (green), the assumptions used to derive
+      the benchmark (identically distributed groups) are met and thus the
+      subspace accuracy matches the benchmark.  However, when the
+      ratio $\lambda_1/\lambda_2$ is large, the subspace accuracy
+      metric can far exceed this benchmark if there is significant
+      variation in the eigenvectors of $\psi_k$ across groups.  Small
+      increases in accuracy over the benchmark are seen for moderately
+      anisotropic data (red) and large increases for highly
+      anisotropic data (blue).  }
+      \label{fig:asymptotics}
+\end{figure}
+
+In Figure \ref{fig:asymptotics} we depict the subspace accuracy metric
+$\tr(\hat{V}\hat{V}^TVV^T)/s$ and benchmark
+$\frac{1}{s}\sum_{i=1}^s \left(1-\frac{\gamma}{K(\lambda_i -
+    1)^2}\right) /\left(1 + \frac{\gamma}{K(\lambda_i - 1)}\right)$
+for simulated multi-group data generated under the shared subspace
+model with $s=2$, $n=50$, $p=200$ and three different sets of eigenvalues.  For each
+covariance matrix, the eigenvectors of $\psi_k$ were sampled uniformly
+from Stiefel manifold $\mathcal{V}_{2,2}$.  When $\psi_k$ is isotropic
+(green) the subspace similarity metric closely matches the benchmark
+since the assumptions used to derive this asymptotic result are met.
+However, when the eigenvectors of $\psi_k$ vary significantly across
+groups and $\lambda_1 \gg \lambda_2$, the subspace accuracy can far
+exceed this benchmark (blue).  Intuitively, when the first
+eigenvectors of two different groups are nearly orthogonal, each group
+provides a lot of information about orthogonal directions on $VV^T$
+and so the gains in accuracy exceed those that you would get by
+estimating the subspace from a single group with $K$ times the sample
+size.  In general the accuracy of shared subspace estimates depends on
+the variation in the eigenvectors of $\psi_k$ across groups as well as
+the magnitude of the eigenvalues and matrix dimensions $p$ and $n_k$.
+Although the shared subspace estimator improves on the accuracy of
+individually estimated covariance matrices, estimates can still be
+biased when $\gamma$ is very large or the eigenvalues of $\Sigma_k$
+are very small for all $k$.  In practice, one should estimate the
+approximate magnitude of the bias using the inferred eigenvalues of $\Sigma_k$.  When
+these inferred eigenvalues are significantly larger than
+$\hat{\sigma}_k^2(1+\sqrt{\gamma/K})$ the bias will likely be small.
+
+% \begin{align*}
+% \hat{V} &= \underset{\widetilde{V} \in \mathcal{V}_{p,
+%     s}}{\text{argmax }} \tr\left((I-M_k^{-1})^{1/2}\widetilde{V}^T
+%   (\sum_{k=1}^K S_k)\widetilde{V}(I-M_k^{-1})^{1/2}\right)
+% \end{align*}
+% Where $(I-M_k^{-1})^{1/2}$
+
+% \begin{align*}
+% Y &= ZV\psi_k^{1/2}V^T+E\\
+% Y^TY &= V\psi_k^{1/2}V^TZ^TZV\psi_k^{1/2}V^T + E^TE + \text{cross terms}
+% \end{align*}
+
+% \noindent where $Z$ and $E$ are $n$ by $p$ matrices of iid stanard
+% normals.
+
+% $$ \tr\left((I-M_k^{-1})\widetilde{V}^TVM_k^{1/2}V^TZ^TZVM_k^{1/2}V^T\widetilde{V}\right)$$
+%% \subsection{Consistency of $\Psi_k$ and $\sigma_k^2$ given $V$}
+
 \section{Analysis of Biological Data}
@@
     \centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/leukemiaRatio-07-07}
-        \caption{$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$}
+    \subfigure[$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$]{
         \label{fig:leukemiaRatio}
-    \end{subfigure}
+        \includegraphics[width=0.45\textwidth]{Figs/leukemiaRatio-07-07}}
     ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
       %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/leukemiaPosterior-07-07}
-        \caption{95\% Posterior Regions }
+    \subfigure[95\% Posterior Regions]{
       \label{fig:leukemiaPosterior}
-      \end{subfigure}
+        \includegraphics[width=0.45\textwidth]{Figs/leukemiaPosterior-07-07}}
       \caption{a) Goodness of shared subspace fit for each of the
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{Figs/leukemia-biplot-07-07}
+    \includegraphics[width=0.45\textwidth]{Figs/leukemia-biplot-07-07}
     \qquad
-\raisebox{1.25\height}{
-\footnotesize
+\raisebox{1\height}{
+\scriptsize
 \input{Figs/biplot.tab}
@@
     \centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{Figs/dmelanRatio-07-07}
-        \caption{$\gamma(Y_k: \hat{V}, \hat{\sigma_k}^2)$}
+    \subfigure[$\gamma(Y_k: \hat{V},\hat{\sigma_k}^2)$]{
         \label{fig:dmelanRatio}
-      \end{subfigure}
+        \includegraphics[width=0.45\textwidth]{Figs/dmelanRatio-07-07}}
     ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
       %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.45\textwidth}r
-        \includegraphics[width=\textwidth]{Figs/dmelanPosterior-07-07}
-        \caption{95\% Posterior Regions}
+    \subfigure[95\% Posterior Regions]{
       \label{fig:dmelanPosterior}
-    \end{subfigure}
+      \includegraphics[width=0.45\textwidth]{Figs/dmelanPosterior-07-07}}
     \caption{a) Goodness of fit is close to one for all groups
@@
 large.  A repository for the replication code is available on GitHub \citep{FranksGit}.
-
 \bibliographystyle{plainnat}

### 2019-02-22 Sdim-test.R

@@ -1,27 +1,8 @@
-#!/usr/bin/env Rscript
-
-#SBATCH -n 4  #Number of cores
-#SBATCH -N 1  #Number of cores 
-#SBATCH -t 15000  #Runtime in minutes
-#SBATCH -J ss3
-#SBATCH -o outfiles/ssg_%a.out #Standard output
-#SBATCH -e outfiles/ssg_%a.err #Standard error
-#SBATCH -p stats #Partition to submit to 
-#SBATCH --mem=10000  #Memory per node in MB (see also --mem-per-cpu)
-#SBATCH -a 1-100
-
-rm(list=ls())
-idx <- as.numeric(Sys.getenv("SLURM_ARRAY_TASK_ID"))
-
-## Need these for Rscript
-library("methods")
-library("utils")
-
 library(rstiefel)
 
 source("fit-subspace.R")
 source("generateData.R")
-source("helper.R")
+source("subspace-functions.R")
 
 ########################################################
 ############# Data Generation #########################

### 2019-02-22 asymptotics.R

@@ -127,3 +127,12 @@ lines(resMatList[[3]][1, ], lwd=3, col="dark green")
 lines(resMatList[[3]][4, ], lwd=3, col="dark green", lty=2)
 grid()
 legend("bottomright", legend=c("(5, 5)", "(5, 2)", "(50, 2)"), lwd=3, col=c("dark green", "red", "blue"), cex=1.5, bg="white", title=expression("("~lambda[1]~","~lambda[2]~")"))
+
+n <- 50
+S <- 2
+R <- S
+P <- 200
+
+adp <- function(n, e){ (1 - P/n / (e - 1)^2) / (1 + P/n / (e - 1)) }
+curve(adp(x, 10)/(adp(2*x, 10) + adp(2*x, 4.35))/2, from=50, 500)
+curve((adp(2*x, 10) + adp(2*x, 4.2))/2, from=50, 500, add=TRUE, col="blue")

### 2019-02-22 bayes-simulation.R

@@ -31,7 +31,7 @@ LambdaList[[5]] <- c(1000, 1000)
 
 Olist[[1]] <- matrix(c(sqrt(2)/2, sqrt(2)/2, -sqrt(2)/2, sqrt(2)/2), nrow=2)
 Olist[[2]] <- matrix(c(-sqrt(2)/2, sqrt(2)/2, sqrt(2)/2, sqrt(2)/2), nrow=2)
-Olist[[3]] <- matrix(c(-sqrt(2)/2, sqrt(2)/2, sqrt(2)/2, sqrt(2)/2), nrow=2)
+Olist[[3]] <- diag(2)## matrix(c(-sqrt(2)/2, sqrt(2)/2, sqrt(2)/2, sqrt(2)/2), nrow=2)
 Olist[[4]] <- diag(2)
 Olist[[5]] <- diag(2)
 
@@ -39,7 +39,7 @@ Opooled <- matrix(0, nrow=S, ncol=S)
 Opooled[1:2, 1:2] <- diag(2)
 if(R  < S) { 
     Opooled[(R+1):S, (R+1):S] <- rustiefel(S-R, S-R)
-    LambdaPooled <- rexp(S - R, 1/500)
+    LambdaPooled <- rexp(S - R, 1/50)
     for(k in 1:ngroups) {
         Ok <- matrix(0, nrow=S, ncol=S)
         Ok[1:R, 1:R] <- Olist[[k]]
@@ -71,7 +71,7 @@ dat$Ypooled <- Ypooled
 ngroups <- dat$ngroups
 nvec <- dat$nvec
 
-Vinit <- svd(do.call(cbind, lapply(1:ngroups, function(k) svd(t(dat$Ylist[[k]]))$u[, 1:S])))$u[, 1:S]
+Vinit <- svd(do.call(cbind, lapply(1:ngroups, function(k) svd(t(dat$Ylist[[k]]))$u[, 1:R])))$u[, 1:S]
 
 ## ulst <- lapply(1:ngroups, function(k) svd(t(dat$Ylist[[k]]))$u[, 1:S])
 
@@ -81,7 +81,7 @@ Vinit <- Vinit %*% rustiefel(S, S)
 
 
 microbenchmark(EMFit <- subspaceEM(dat$Ylist, P=P, S=S, R=R, nvec=dat$nvec,
-                                   Vstart=Vinit, verbose=TRUE, EM_iters=10, M_iters=100, stiefelAlgo=2), times=1)
[…]

### 2019-02-22 fit-leukemia.R

@@ -131,11 +131,21 @@ save(samples, initSS, file=sprintf("results/leukemiaBayes-%s.RData", format(Sys.
 
 regionColors <- c(4, 2:3, 1, 5:7)
 
+g1 <- 7
+g2 <- 2
+pmPsi6 <- getPostMeanPsi(samples$Osamps[, , g1 , ],
+                            samples$omegaSamps[, g1, ],
+                         samples$s2samps[g1, ], 100)
+pmPsi7 <- getPostMeanPsi(samples$Osamps[, , g2 , ],
+                            samples$omegaSamps[, g2, ],
+                         samples$s2samps[g2, ], 100)
+
+O <- svd(pmPsi6[1:R, 1:R] - pmPsi7[1:R, 1:R])$u[, 3:4]
 
 
 ## view data projected onto a 2D subspace
 unique(typeVec)
-g <- 7
+g <- 1
 
 
 O <- samples$Osamps[1:R, 1:R, g, 1000]
@@ -174,6 +184,9 @@ omegaSamps_proj[, , 1000]/(1-omegaSamps_proj[, , 1000])
 
 dev.off()
 
+
+samples$omegaSamps[, , 1000]/(1-samples$omegaSamps[, , 1000])
+samples$omegaSamps[, , 1000]/(1-samples$omegaSamps[, , 1000])
 
 pdf(sprintf("paper/Figs/leukemia-biplot-%i-%s.pdf", g, format(Sys.Date(), "%m-%d")), font="Times")
 
@@ -248,14 +261,16 @@ dev.off()
 
 symbol_names <- sapply(rownames(EMFit$V), function(x) {
     na.omit(select(hgu95av2.db, x, c("SYMBOL"))$SYMBOL)[1]
-})p
+})
+
 
 v1 <- Vstar[, view_indices[1]]
 v2 <- Vstar[, view_indices[2]]
 mag <- rowSums(Vstar[, view_indices]^2)
 names(mag) <- rownames(EMFit$V)
 
-gp_v1 <- find_go_groups(abs(v1), 0.1, go_name_length=60)
+gp_v1 <- find_go_groups(abs(v1), 0.05, go_name_length=60)
+gp_v1
 
 print(xtable(gp_v1[, c("Name", "Q-value", "Numer of Genes")], na="", quote=FALSE),
       include.rownames=TRUE, include.colnames=TRUE,
@@ -271,7 +286,18 @@ gp_v2
[…]

### 2019-02-22 fit-subspace.R

@@ -89,8 +89,14 @@ fitSubspace <- function(P, S, R, Q=S-R, Ylist, nvec, ngroups=length(Ylist),
                                   1, sum(nvec))
             
             ## Sample common O
-            O2 <- sampleO(YVpooled[, (R+1):(R+Q)], Ok[(R+1):(R+Q), (R+1):(R+Q)], 1,
+            O2 <- sampleO(YVpooled[, (R+1):(R+Q)], Olist[[k]][(R+1):(R+Q), (R+1):(R+Q)], 1,
                           Omega2)
+
+            ## order eigenvalues from largest to smallest
+            ord_omega <- order(Omega2, decreasing=TRUE)
+            Omega2 <- Omega2[ord_omega]
+            O2 <- O2[, ord_omega]
+            
         } else {
             Omega2 <- c()
             O2 <- matrix(nrow=0, ncol=0)
@@ -108,8 +114,8 @@ fitSubspace <- function(P, S, R, Q=S-R, Ylist, nvec, ngroups=length(Ylist),
 
                 Omega1 <- sampleOmega(YVlist[[k]], Olist[[k]][, 1:R],
                                          s2vec[k], nvec[k])
+                
 
-                OmegaList[[k]] <- c(Omega1, Omega2)
                 
             }
                 ## Sample O_k
@@ -117,10 +123,16 @@ fitSubspace <- function(P, S, R, Q=S-R, Ylist, nvec, ngroups=length(Ylist),
                 
                 O1 <- sampleO(YVlist[[k]][, 1:R], Olist[[k]][1:R, 1:R], s2vec[k],
                                   (OmegaList[[k]])[1:R])
-                                  
+
+                ## order eigenvalues from largest to smallest
+                ord_omega <- order(Omega1, decreasing=TRUE)
+                Omega1 <- Omega1[ord_omega]
+                O1 <- O1[, ord_omega]
+                
                 Ok <- as.matrix(bdiag(O1, O2))
             }
 
+            OmegaList[[k]] <- c(Omega1, Omega2)
             
             Olist[[k]] <- Ok
             
@@ -204,7 +216,7 @@ fitBayesianSpike <- function(P, R, SC, n, niters=100, nskip=1,
     }
     
     s2 <- init$s2
-    U <- init$U
+    U <- init$Upp
     omega <- init$omega
 
     SigInvInit <- getSigmaInv(P, init$U, init$omega, init$s2)

### 2019-02-22 initialization_simulations.R

@@ -4,10 +4,12 @@ rm(list=ls())
 library(microbenchmark)
 library(rstiefel)
 library(tidyverse)
+library(cowplot)
+library(colorspace)
 
 source("fit-subspace.R")
 source("generateData.R")
-source("helper.R")
+source("subspace-functions.R")
 
 ########################################################
 ############# Data Generation #########################
@@ -58,22 +60,30 @@ nvec <- dat$nvec
 
 Vinit <- svd(do.call(cbind, lapply(1:ngroups, function(k) svd(t(dat$Ylist[[k]]))$u[, 1:S])))$u[, 1:S]
 EMFit <- subspaceEM(dat$Ylist, P=P, S=S, R=R, nvec=dat$nvec,
-                    Vstart=Vinit, verbose=TRUE, max_EM_iters=10, stiefelAlgo=2)
+                    Vstart=Vinit, verbose=TRUE, EM_iters=100, M_iters=100, stiefelAlgo=2)
+
 svd_init <- tr((t(dat$V) %*%  EMFit$V) %*% (t(EMFit$V) %*% dat$V)) / S
 
 nruns <- 100
 similarity <- numeric(runs)
+
+Vinit_base <- svd(do.call(cbind, lapply(1:ngroups, function(k) svd(t(dat$Ylist[[k]]))$u[, 1:n])))$u[, 1:(n*ngroups)]
+
 for(i in nruns) {
 
     print(i)
-    Vinit <- rustiefel(P, S)
+    Vinit <- Vinit_base %*% rustiefel(n*ngroups, S)
     EMFit <- subspaceEM(dat$Ylist, P=P, S=S, R=R, nvec=dat$nvec,
-                        Vstart=Vinit, verbose=TRUE, max_EM_iters=10, stiefelAlgo=2)
+                    Vstart=Vinit, verbose=TRUE, EM_iters=100, M_iters=1000, stiefelAlgo=2)
     random_init <- tr((t(dat$V) %*%  EMFit$V) %*% (t(EMFit$V) %*% dat$V)) / S
     similarity[i] <- random_init
 }
 
-tibble(score=similarity) %>% ggplot() + geom_density(aes(x=score), fill=alpha("red", 0.5)) + geom_vline(xintercept=svd_init, size=2, col="blue")  +xlim(c(0, 1)) + xlab("Subspace Similarity") + ylab("") + theme_grey(base_size=16)
+load("results/similarity_scores.RData")
 
[…]

### 2019-02-22 ody-fit-generated.R

@@ -19,6 +19,7 @@ library("utils")
 
 library(rstiefel)
 
+scriptDir <- dirname(sys.frame(1)$ofile)
 source("fit-subspace.R")
 source("generateData.R")
 source("helper.R")
@@ -31,12 +32,14 @@ datType <- ((idx-1) %% 3) + 1
 ########################################################
 
 n <- 50
-S <- 2
-R <- 2
+S <- 5
+R <- 5
+Q <- 3
 P <- 200
 ngroups <- 10
 
 evals <- c(250, 25)
+evals <- c(125, 50, 250, 25, 4)
 
 niters <- 1000
 nwarmup <- niters/2
@@ -51,7 +54,7 @@ if(datType==1) {
         LambdaList[[k]] <- lam
     }
     
-    dat <- generateData(P=P, S=S, R=R, ngroups=ngroups,
+    dat <- generateData(P=P, S=S, R=2, ngroups=ngroups,
                         LambdaList=LambdaList, nvec=rep(n, ngroups))
 
     dat$genType <- "Shared subspace"
@@ -124,13 +127,23 @@ for(fitType in 1:1) {
         Slist <- dat$Slist
 
         Vinit <- svd(do.call(cbind, lapply(1:ngroups, function(k) svd(t(dat$Ylist[[k]]))$u[, 1:R])))$u[, 1:S]
+        Vinit <- rustiefel(P, S)
         EMFit <- subspaceEM(dat$Slist, P=P, S=S, R=R, nvec=dat$nvec,
-                            rho1=0.5, rho2=0.5, 
-                            Vstart=Vinit, verbose=FALSE, maxIters=100)
+                            PrecVec=rep(10, ngroups),
+                            rho1=0.1, rho2=0.9, 
+                            Vstart=Vinit, verbose=TRUE, maxIters=100)
         Vinit <- EMFit$V
+
+        A <- dat$Olist[[k]] %*% diag(dat$LambdaList[[k]]) %*% t(dat$Olist[[k]])
+        B <- solve(EMFit$PhiList[[k]])
+        cov2cor(R %*% B %*% t(R))
+        R <- t(dat$V) %*% EMFit$V
+        norm(t(EMFit$V) %*% dat$Slist[[k]] %*% EMFit$V, type="F")
+        
         ##cancor(svd(do.call(cbind, dat$Olist))$u, Vinit)$cor
         
-        OmegaList <- Ulist <- list()
+        
[…]

### 2019-02-22 rankTests.R

@@ -45,7 +45,7 @@ evals <- c(250, 25)
 niters <- 1000
 nwarmup <- niters/2
 
-## Subspace Shrinkage: Psi_k = VO_kLam_kO_k^TV^T
+## Subspace Shrinkage: Psni_k = VO_kLam_kO_k^TV^T
 LambdaList <- list()
 lam <- evals
 for( k in 1:ngroups) {

### 2019-02-22 scalability_simulation.R

@@ -1,8 +1,6 @@
 rm(list=ls())
 
 ## Need these for Rscript
-library(methods)
-library(utils)
 library(microbenchmark)
 library(rstiefel)
 library(tidyverse)
@@ -14,10 +12,10 @@ source("fit-subspace.R")
 source("generateData.R")
 source("subspace-functions.R")
 
+
 nbench_times <- 10
 Svec <- c(2, 10, 25, 50)
 Pvec <- seq(from=1000, to=10000, by=1000)
-
 
 ## Initialize sampler
 ngroups <- dat$ngroups
@@ -63,7 +61,14 @@ for(sindex in 1:length(Svec)) {
     }
 }
 
-## save(benchmark_array, file="benchmark-2-7-19-final.RData")
-dimnames(benchmark_array) = list(S = Svec, P = Pvec, Iter =1:nbench_times)
-as_tibble(reshape2::melt(benchmark_array)) %>% ggplot(aes(x=as.factor(P), y=value)) + geom_boxplot(aes(fill=factor(S))) + scale_fill_discrete_qualitative(palette="Set 2") +
-    ylab("Seconds")
+save(benchmark_array, file="results/benchmark_times.RData")
+
+pdf("paper/Figs/run_times.pdf", height=4, width=8)
+dimnames(benchmark_array) = list(S = Svec, P = Pvec, Iter = 1:nbench_times)
+as_tibble(reshape2::melt(benchmark_array)) %>% ggplot(aes(x=as.factor(P), y=value)) +
+    geom_jitter(aes(col=factor(S))) + scale_fill_discrete_qualitative(palette="Set 2") +
+    ylab("Seconds") + ylim(c(0, 300)) +
+    xlab("P") + scale_colour_discrete(name  = "S")
+
+dev.off()
+

### 2019-02-22 subspace-functions.R
Prediction functions

@@ -345,8 +345,13 @@ tgamma <- function(a, b, scale) {
 
 sampleOmega <- function(YV, O, s2, n) {
 
-    YVO <- YV %*% O
     R <- ncol(O)
+    
+    Rtilde <- min(n, R)
+    O <- O[, 1:Rtilde]
+    
+    YVO <- YV %*% O
+
     cvec <- diag(t(YVO) %*% YVO) /(2*s2)
 
     ## add divide by n for stability?
@@ -354,6 +359,9 @@ sampleOmega <- function(YV, O, s2, n) {
     g <- g/n
 
     omega <- 1 - g
+    
+    if(n < R)
+        omega <- c(omega, rep(0, R - n))
 
     omega
 
@@ -381,6 +389,7 @@ sampleO <- function(YV, O, s2, omega) {
 ## Ok is an S x R matrix,  V is a P x S and U is P x R
 ## B_k = Ok*omega_k*Ok^T
 ## A_k is sample covariance matrix for group k
+
 sampleV <- function(Slist, Ulist, s2vec, OmegaList, V, method="gibbs") {
 
     K <- length(Ulist)
@@ -607,13 +616,13 @@ makePrediction <- function(Y, V, Osamps, omegaSamps, s2samps, nvec,
     if(ngroups==1) {
         return(matrix(1, ncol=1, nrow=1))
     }
-    
+
     ##predictionMat <- matrix(0, nrow=length(genGroups), ncol=ngroups)
     ## Y <- Y / sum(Y^2)
     YV <- Y %*% V
     SigmaHat <- list()
         
-    probs <- matrix(0, nrow=numToAvg, ncol=ngroups)
+    ll_array <- array(0, dim=c(nrow(Y), numToAvg, ngroups))
     for(i in (nsamps-numToAvg+1):nsamps) {
             
         for(k  in 1:ngroups) {
@@ -623,35 +632,59 @@ makePrediction <- function(Y, V, Osamps, omegaSamps, s2samps, nvec,
             s2 <- s2samps[k, i]
             SigmaHat[[k]] <-  s2*(Ok %*% Lam %*% t(Ok) + diag(nrow(Ok)))
         }
-        
-        mp <- computeMembershipProbabilities(as.numeric(YV), SigmaHat)
-        probs[nsamps - i, ] <- mp
+
+        ll <- computeLL(YV, SigmaHat)
+        ll_array[, nsamps - i, ] <- ll
     }
-    probs
[…]
